---
title: "01 - ANES Data Inventory"
output: html_document
---

## Purpose

This document performs a read-only inventory of the ANES Time Series Cumulative Data File used in the dissertation. It verifies the dataset structure, election-year coverage, availability of the selected variables, and raw response distributions required for subsequent data preparation.

No analytical recoding, case exclusion, network estimation, or inferential analysis is performed in this document.

## Project Location Check
```{r}

```

```{r project-location}
getwd()
list.files()
```
## Project Setup

```{r project-setup}
project_root <- if (file.exists("ANES.Rproj")) {
  normalizePath(".", winslash = "/", mustWork = TRUE)
} else {
  normalizePath("..", winslash = "/", mustWork = TRUE)
}

project_file <- file.path(
  project_root,
  "ANES.Rproj"
)

project_root
file.exists(project_file)
```
## Official CDF File Check

```{r cdf-file-check}
cdf_path_config <- file.path(
  project_root,
  "config",
  "anes_cdf_path.txt"
)

stopifnot(
  file.exists(cdf_path_config)
)

cdf_file <- trimws(
  readLines(
    cdf_path_config,
    n = 1,
    warn = FALSE
  )
)

stopifnot(
  length(cdf_file) == 1,
  nzchar(cdf_file),
  file.exists(cdf_file)
)

cdf_file <- normalizePath(
  cdf_file,
  winslash = "/",
  mustWork = TRUE
)

expected_cdf_sha256 <-
  "45323c30faee1e9e67c2351c4302c31472b117050bef240d9339a264e2f3e078"

cdf_file_sha256 <- digest::digest(
  file = cdf_file,
  algo = "sha256",
  serialize = FALSE
)

cdf_inventory_input_audit <- data.frame(
  File = basename(cdf_file),
  Exists = file.exists(cdf_file),
  Size_bytes = file.info(cdf_file)$size,
  SHA256_matches = identical(
    cdf_file_sha256,
    expected_cdf_sha256
  )
)

cdf_inventory_input_audit

stopifnot(
  cdf_inventory_input_audit$Exists,
  cdf_inventory_input_audit$Size_bytes > 0,
  cdf_inventory_input_audit$SHA256_matches
)

"PASS: official CDF inventory input gate"
```

## Package Check

```{r package-check}
requireNamespace("haven", quietly = TRUE)
```
## Read ANES CDF

```{r read-cdf}
anes_raw <- haven::read_dta(cdf_file)

dim(anes_raw)
class(anes_raw)
```
## Core Variable Check

```{r core-variable-check}

# Five main election years used in the dissertation
main_years <- c(2004, 2012, 2016, 2020, 2024)

# Nine network nodes plus education
required_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223",
  "VCF0110"
)

# Check whether the key identifiers also exist
identifier_vars <- c(
  "Version",
  "VCF0004",
  "VCF0006",
  "VCF0006a"
)

all(identifier_vars %in% names(anes_raw))
all(required_vars %in% names(anes_raw))

# If anything is missing, print its name
setdiff(identifier_vars, names(anes_raw))
setdiff(required_vars, names(anes_raw))
```
## Version Field Check

```{r version-field-check}
grep(
  "version",
  names(anes_raw),
  value = TRUE,
  ignore.case = TRUE
)
```
## CDF Version

```{r cdf-version}
unique(anes_raw$Version)
```
## Main Wave Sample Sizes

```{r main-wave-sizes}
year_counts <- table(
  anes_raw$VCF0004,
  useNA = "ifany"
)

year_counts[as.character(main_years)]
```
## Raw Response Check: VCF0806 in 2004

```{r raw-vcf0806-2004}

# Select the raw VCF0806 values for the 2004 wave only
v0806_2004 <- anes_raw$VCF0806[
  anes_raw$VCF0004 == 2004
]

# Check variable type
class(v0806_2004)

# Check the variable label from the Stata file
attr(v0806_2004, "label")

# Check the labelled response codes
attr(v0806_2004, "labels")

# Count every raw value, including missing values
table(v0806_2004, useNA = "ifany")
```
## Raw Response Frequencies Across Main Waves: VCF0806

```{r raw-vcf0806-main-waves}

# Keep only the five main election years
v0806_main <- anes_raw[
  anes_raw$VCF0004 %in% main_years,
  c("VCF0004", "VCF0806")
]

# Show raw response frequencies for each year
table(
  v0806_main$VCF0004,
  v0806_main$VCF0806,
  useNA = "ifany"
)
```
## Raw Response Frequencies Across Main Waves: All Nine Nodes

```{r raw-all-nodes-main-waves}

# The nine network nodes
network_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223"
)

# Print the raw frequency table for each node
for (v in network_vars) {

  cat("\n\n========================================\n")
  cat("Variable:", v, "\n")
  cat("Variable label:", attr(anes_raw[[v]], "label"), "\n")
  cat("========================================\n")

  print(
    table(
      Year = anes_raw$VCF0004[
        anes_raw$VCF0004 %in% main_years
      ],
      RawValue = anes_raw[[v]][
        anes_raw$VCF0004 %in% main_years
      ],
      useNA = "ifany"
    )
  )
}
```
## Raw Value Labels: All Nine Nodes

```{r raw-value-labels-all-nodes}

for (v in network_vars) {

  cat("\n\n========================================\n")
  cat("Variable:", v, "\n")
  cat("Variable label:", attr(anes_raw[[v]], "label"), "\n")
  cat("========================================\n")

  print(attr(anes_raw[[v]], "labels"))
}
```
## 2024 Mode-Related Missingness Pattern Check

```{r 2024-missingness-pattern-check}

anes_2024 <- anes_raw[
  anes_raw$VCF0004 == 2024,
  c(
    "VCF0006",
    "VCF0806",
    "VCF0809",
    "VCF0888",
    "VCF0890",
    "VCF0894",
    "VCF9223"
  )
]

# Identify the special 2024 cases in each variable
flag_0806 <- anes_2024$VCF0806 == -1
flag_0809 <- anes_2024$VCF0809 == -1
flag_9223 <- anes_2024$VCF9223 == -1

flag_0888 <- is.na(anes_2024$VCF0888)
flag_0890 <- is.na(anes_2024$VCF0890)
flag_0894 <- is.na(anes_2024$VCF0894)

# Count each type
colSums(
  cbind(
    flag_0806,
    flag_0809,
    flag_9223,
    flag_0888,
    flag_0890,
    flag_0894
  )
)

# Check whether they identify exactly the same respondents
all(flag_0806 == flag_0809)
all(flag_0806 == flag_9223)
all(flag_0806 == flag_0888)
all(flag_0806 == flag_0890)
all(flag_0806 == flag_0894)
```
## Education Variable Inventory: VCF0110

```{r education-variable-inventory}

# Variable label
attr(anes_raw$VCF0110, "label")

# Value labels
attr(anes_raw$VCF0110, "labels")

# Raw frequencies across the five main election years
table(
  Year = anes_raw$VCF0004[
    anes_raw$VCF0004 %in% main_years
  ],
  RawValue = anes_raw$VCF0110[
    anes_raw$VCF0004 %in% main_years
  ],
  useNA = "ifany"
)
```
## Cross-Wave Respondent Identifier Inventory: VCF0006a

```{r cross-wave-id-inventory}

# Variable type
class(anes_raw$VCF0006a)

# Variable label
attr(anes_raw$VCF0006a, "label")

# Any labelled values
attr(anes_raw$VCF0006a, "labels")

# Number of non-missing IDs in each main election year
for (yr in main_years) {

  x <- anes_raw$VCF0006a[
    anes_raw$VCF0004 == yr
  ]

  cat(
    "\nYear:", yr,
    "\nRaw N:", length(x),
    "\nNon-missing VCF0006a:", sum(!is.na(x)),
    "\nUnique non-missing VCF0006a:", length(unique(x[!is.na(x)])),
    "\n"
  )
}
```
## Cross-Wave Respondent Overlap: Main Waves

```{r cross-wave-overlap-main-waves}

# Store the respondent IDs for each main election year
ids_by_year <- lapply(
  main_years,
  function(yr) {
    unique(
      anes_raw$VCF0006a[
        anes_raw$VCF0004 == yr &
        !is.na(anes_raw$VCF0006a)
      ]
    )
  }
)

names(ids_by_year) <- as.character(main_years)

# Create an empty overlap matrix
overlap_matrix <- matrix(
  0,
  nrow = length(main_years),
  ncol = length(main_years),
  dimnames = list(
    as.character(main_years),
    as.character(main_years)
  )
)

# Count respondents appearing in each pair of waves
for (i in seq_along(main_years)) {
  for (j in seq_along(main_years)) {

    overlap_matrix[i, j] <- length(
      intersect(
        ids_by_year[[i]],
        ids_by_year[[j]]
      )
    )
  }
}

overlap_matrix
```
### Interpretation

No respondent IDs were shared across the five main ANES CDF election-year samples according to VCF0006a. The selected wave samples therefore contain no detected cross-wave respondent overlap at the CDF case level.