---
title: "Node and Wave Selection Audit"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

find_project_root <- function(start = getwd()) {
  current <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(current, "ANES.Rproj"))) return(current)
    parent <- dirname(current)
    if (identical(parent, current)) stop("Cannot locate ANES.Rproj")
    current <- parent
  }
}

project_root <- find_project_root()
Sys.setenv(ANES_PROJECT_ROOT = project_root)
source(file.path(project_root, "R", "01A_node_and_wave_selection_engine.R"), local = knitr::knit_global())
table_dir <- file.path(project_root, "outputs", "tables", "supplement", "node_wave_selection")
figure_dir <- file.path(project_root, "outputs", "figures", "supplement", "node_wave_selection")
```

# Purpose and boundary

This audit reconstructs the documented evidence chain for the final nine-node set and the five primary election waves. It separates human substantive decisions from reproducible empirical checks. The code reproduces availability, measurement-map, complete-case, source-mapping and 2008 split-form evidence. It does not claim that an algorithm selected the research construct, and it does not estimate a network.

The surviving project record begins with a recorded pool of 16 candidate CDF variables. The exact original codebook search rules used to nominate those variables were not preserved. A separate retrospective audit now enumerates the full CDF and applies a deliberately broad reproducible screen. It repairs the provenance record but is not presented as the original or preregistered procedure.

# Retrospective candidate-pool completeness audit

```{r retrospective-audit}
retro_dir <- file.path(table_dir, "retrospective_candidate_audit")
retro_summary <- read.csv(
  file.path(retro_dir, "retrospective_candidate_audit_summary.csv"),
  check.names = FALSE
)
retro_survivors <- read.csv(
  file.path(retro_dir, "provisional_five_wave_survivors_32.csv"),
  check.names = FALSE
)
knitr::kable(retro_summary, row.names = FALSE)
knitr::kable(
  retro_survivors[, c(
    "variable", "label", "minimum_provisional_substantive_n",
    "minimum_provisional_category_count", "in_recorded_16",
    "retrospective_context"
  )],
  row.names = FALSE
)
```

The 32 fields are provisional raw-screen survivors, not 32 eligible network nodes. Official source mapping, wording, routing, scale, duplicate representation and the bounded substantive estimand require separate review.

# Candidate-node decisions

```{r candidate-node-decisions}
candidate_node_decisions <- read.csv(
  file.path(table_dir, "candidate_node_decisions.csv"),
  check.names = FALSE
)
knitr::kable(
  candidate_node_decisions[, c(
    "cdf_variable", "node_label", "construct_class", "final_role",
    "decision_status", "decision_reason"
  )],
  row.names = FALSE
)
```

# Final nine-node set

```{r final-nine}
final_nine <- read.csv(file.path(table_dir, "final_nine_node_map.csv"), check.names = FALSE)
knitr::kable(
  final_nine[, c(
    "final_order", "cdf_variable", "node_label", "primary_domain",
    "high_value_meaning", "measurement_caution"
  )],
  row.names = FALSE
)
```

## Current disposition of VCF9049

`VCF9049` is available in all five final waves and would cause little additional complete-case loss. Its surviving historical item-specific rationale is missing. The item is now retained outside the network by a current scope-and-measurement-boundary decision based on the combined provenance, category-harmonisation, bounded-scope and design-integrity considerations recorded in the decision register. This is not a reconstruction of the missing historical intention, and the item is not characterised as invalid or redundant.

# Candidate-set feasibility

```{r candidate-set-feasibility}
candidate_set_feasibility <- read.csv(
  file.path(table_dir, "candidate_node_set_feasibility.csv"),
  check.names = FALSE
)
knitr::kable(candidate_set_feasibility, row.names = FALSE)
```

# Wave decisions

```{r wave-decisions}
wave_decisions <- read.csv(file.path(table_dir, "wave_selection_decisions.csv"), check.names = FALSE)
knitr::kable(wave_decisions, row.names = FALSE)
```

```{r wave-selection-flow, echo=FALSE, out.width='90%'}
wave_selection_flow_file <- normalizePath(
  file.path(figure_dir, "wave_selection_flow.png"),
  winslash = "/",
  mustWork = TRUE
)
knitr::include_graphics(wave_selection_flow_file, rel_path = FALSE)
```

# Why 2008 was excluded

The 2008 overall matrix was technically estimable. The wave was not removed because of a failed matrix or an unfavourable network result. Four key nodes were administered through OLD/NEW split-form or branching designs, and the comparable cumulative fields produced a uniquely restricted OLD-form complete-case subset. The final exclusion is therefore a measurement-comparability and bounded-scope decision. The former 2008 sensitivity role was withdrawn; the diagnostics are retained only as historical design evidence.

```{r evidence-2008}
evidence_2008 <- read.csv(file.path(table_dir, "2008_exclusion_evidence.csv"), check.names = FALSE)
knitr::kable(evidence_2008, row.names = FALSE)
```

# Verified measurement map

```{r final-measurement-map}
final_measurement_map <- read.csv(
  file.path(table_dir, "final_five_wave_measurement_map.csv"),
  check.names = FALSE
)
knitr::kable(
  final_measurement_map[, c(
    "year", "final_order", "variable_name", "source_variable",
    "measurement_status", "cross_wave_comparability"
  )],
  row.names = FALSE
)
```

# Selection flow

```{r selection-flow, echo=FALSE, out.width='90%'}
selection_flow_file <- normalizePath(
  file.path(figure_dir, "node_selection_flow.png"),
  winslash = "/",
  mustWork = TRUE
)
knitr::include_graphics(selection_flow_file, rel_path = FALSE)
```

# Validation and disclosed evidence boundary

```{r validation}
validation <- read.csv(file.path(table_dir, "node_wave_selection_validation.csv"), check.names = FALSE)
status <- read.csv(file.path(table_dir, "node_wave_selection_status.csv"), check.names = FALSE)
knitr::kable(status, row.names = FALSE)
knitr::kable(validation, row.names = FALSE)
```

The current `VCF9049` decision is recorded and machine-checked. The exact historical rule that generated the recorded 16-item pool remains unrecovered. Submission readiness is therefore reported as `READY_WITH_DISCLOSED_RETROSPECTIVE_LIMITATION`, not as proof that the original screening history has been restored.

# Reproducibility boundary

No respondent-level ANES data are written by this audit. The aggregate tables can be shared, while the CDF and derived respondent-level RDS files remain outside the submission package. The full data-preparation notebooks and the bounded 2008 diagnostic code are listed in the Additional Materials code manifest.

The active project library matches the 164 records in `renv.lock`. The same module was also restored and rerun in a separate D-drive project library, including the retrospective audit, node/wave engine, validator and HTML render. This clean-library check applies to this module only and is not a claim that Steps 00-10 were all rerun independently.
