---
title: "03 - Data Cleaning"
output: html_document
---
## Project Root for Knitting

```{r project-root}
current_dir <- normalizePath(
  ".",
  winslash = "/",
  mustWork = TRUE
)

if (
  file.exists(
    file.path(
      current_dir,
      "ANES.Rproj"
    )
  )
) {
  project_root <- current_dir
} else if (
  basename(current_dir) == "notebooks" &&
    file.exists(
      file.path(
        dirname(current_dir),
        "ANES.Rproj"
      )
    )
) {
  project_root <- dirname(current_dir)
} else {
  stop(
    "Project root could not be identified."
  )
}

knitr::opts_knit$set(
  root.dir = project_root
)

project_root
```

## Purpose

This document cleans the extracted ANES data for the five approved election years: 2004, 2012, 2016, 2020, and 2024. Non-substantive response codes will be classified and converted to missing values using documented rules, while the extracted data remain unchanged. This stage does not apply complete-case filtering, education grouping, survey weighting, or network analysis.
## Input File

```{r input-file}
extracted_file <- file.path(
  "data",
  "extracted",
  "anes_cdf_20260205_extracted_2004_2012_2016_2020_2024.rds"
)

file.exists(extracted_file)
```
## Read Extracted Data

```{r read-extracted-data}
anes_extracted <- readRDS(extracted_file)

dim(anes_extracted)

names(anes_extracted)
```
## Define Network Variables

```{r define-network-variables}
network_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223"
)

network_vars

all(network_vars %in% names(anes_extracted))
```
## Raw Audit: VCF0806

```{r audit-vcf0806-raw}
table(
  Election_year = anes_extracted$VCF0004,
  Raw_value = anes_extracted$VCF0806,
  useNA = "ifany"
)
```
## Inspect Value Labels: VCF0806

```{r inspect-vcf0806-labels}
attr(
  anes_extracted$VCF0806,
  "label"
)

attr(
  anes_extracted$VCF0806,
  "labels"
)
```
## Clean VCF0806

```{r clean-vcf0806}
anes_clean <- anes_extracted

vcf0806_structural_missing <- (
  anes_clean$VCF0004 == 2024 &
  anes_clean$VCF0806 == -1
)

vcf0806_non_substantive <- (
  anes_clean$VCF0806 %in% c(0, 9)
)

sum(vcf0806_structural_missing)

sum(vcf0806_non_substantive)

anes_clean$VCF0806[
  vcf0806_structural_missing |
  vcf0806_non_substantive
] <- NA_real_

table(
  Election_year = anes_clean$VCF0004,
  Clean_value = anes_clean$VCF0806,
  useNA = "ifany"
)
```
## Raw Audit: VCF0809

```{r audit-vcf0809-raw}
table(
  Election_year = anes_extracted$VCF0004,
  Raw_value = anes_extracted$VCF0809,
  useNA = "ifany"
)

attr(
  anes_extracted$VCF0809,
  "label"
)

attr(
  anes_extracted$VCF0809,
  "labels"
)
```
## Clean VCF0809

```{r clean-vcf0809}
vcf0809_structural_missing <- (
  anes_clean$VCF0004 == 2024 &
  anes_clean$VCF0809 == -1
)

vcf0809_non_substantive <- (
  anes_clean$VCF0809 %in% c(0, 9)
)

sum(vcf0809_structural_missing)

sum(vcf0809_non_substantive)

anes_clean$VCF0809[
  vcf0809_structural_missing |
  vcf0809_non_substantive
] <- NA_real_

table(
  Election_year = anes_clean$VCF0004,
  Clean_value = anes_clean$VCF0809,
  useNA = "ifany"
)
```
## Raw Audit: VCF0838

```{r audit-vcf0838-raw}
table(
  Election_year = anes_extracted$VCF0004,
  Raw_value = anes_extracted$VCF0838,
  useNA = "ifany"
)

attr(
  anes_extracted$VCF0838,
  "label"
)

attr(
  anes_extracted$VCF0838,
  "labels"
)
```
## Clean VCF0838

```{r clean-vcf0838}
vcf0838_non_substantive <- (
  anes_clean$VCF0838 %in% c(0, 9)
)

sum(vcf0838_non_substantive)

anes_clean$VCF0838[
  vcf0838_non_substantive
] <- NA_real_

table(
  Election_year = anes_clean$VCF0004,
  Clean_value = anes_clean$VCF0838,
  useNA = "ifany"
)

range(
  anes_clean$VCF0838,
  na.rm = TRUE
)
```
## Raw Audit: VCF0839

```{r audit-vcf0839-raw}
table(
  Election_year = anes_extracted$VCF0004,
  Raw_value = anes_extracted$VCF0839,
  useNA = "ifany"
)

attr(
  anes_extracted$VCF0839,
  "label"
)

attr(
  anes_extracted$VCF0839,
  "labels"
)
```
## Clean VCF0839

```{r clean-vcf0839}
vcf0839_non_substantive <- (
  anes_clean$VCF0839 %in% c(0, 9)
)

sum(vcf0839_non_substantive)

anes_clean$VCF0839[
  vcf0839_non_substantive
] <- NA_real_

table(
  Election_year = anes_clean$VCF0004,
  Clean_value = anes_clean$VCF0839,
  useNA = "ifany"
)

range(
  anes_clean$VCF0839,
  na.rm = TRUE
)
```
## Raw Audit: VCF0879a

```{r audit-vcf0879a-raw}
table(
  Election_year = anes_extracted$VCF0004,
  Raw_value = anes_extracted$VCF0879a,
  useNA = "ifany"
)

attr(
  anes_extracted$VCF0879a,
  "label"
)

attr(
  anes_extracted$VCF0879a,
  "labels"
)
```
## Clean VCF0879a

```{r clean-vcf0879a}
vcf0879a_non_substantive <- (
  anes_clean$VCF0879a %in% c(8, 9)
)

sum(vcf0879a_non_substantive)

anes_clean$VCF0879a[
  vcf0879a_non_substantive
] <- NA_real_

table(
  Election_year = anes_clean$VCF0004,
  Clean_value = anes_clean$VCF0879a,
  useNA = "ifany"
)

sort(
  unique(
    anes_clean$VCF0879a[
      !is.na(anes_clean$VCF0879a)
    ]
  )
)
```
## Raw Audit: VCF0888

```{r audit-vcf0888-raw}
table(
  Election_year = anes_extracted$VCF0004,
  Raw_value = anes_extracted$VCF0888,
  useNA = "ifany"
)

attr(
  anes_extracted$VCF0888,
  "label"
)

attr(
  anes_extracted$VCF0888,
  "labels"
)
```
## Audit Existing NA: VCF0888

```{r audit-vcf0888-existing-na}
vcf0888_raw <- anes_extracted$VCF0888

sum(
  is.na(vcf0888_raw)
)

sum(
  haven::is_tagged_na(vcf0888_raw)
)

table(
  Election_year = anes_extracted$VCF0004[
    is.na(vcf0888_raw)
  ],
  NA_tag = haven::na_tag(
    vcf0888_raw[
      is.na(vcf0888_raw)
    ]
  ),
  useNA = "ifany"
)
```
## Compare Existing NA Cases: VCF0888

```{r compare-vcf0888-na-cases}
vcf0888_existing_na <- is.na(
  anes_extracted$VCF0888
)

vcf0806_2024_structural <- (
  anes_extracted$VCF0004 == 2024 &
  anes_extracted$VCF0806 == -1
)

vcf0809_2024_structural <- (
  anes_extracted$VCF0004 == 2024 &
  anes_extracted$VCF0809 == -1
)

identical(
  which(vcf0888_existing_na),
  which(vcf0806_2024_structural)
)

identical(
  which(vcf0888_existing_na),
  which(vcf0809_2024_structural)
)
```
## Clean VCF0888

```{r clean-vcf0888}
vcf0888_non_substantive <- (
  anes_clean$VCF0888 %in% c(8, 9)
)

sum(vcf0888_non_substantive)

anes_clean$VCF0888[
  vcf0888_non_substantive
] <- NA_real_

table(
  Election_year = anes_clean$VCF0004,
  Clean_value = anes_clean$VCF0888,
  useNA = "ifany"
)

sum(
  is.na(anes_clean$VCF0888)
)

sort(
  unique(
    anes_clean$VCF0888[
      !is.na(anes_clean$VCF0888)
    ]
  )
)
```
## Raw Audit: VCF0890

```{r audit-vcf0890-raw}
table(
  Election_year = anes_extracted$VCF0004,
  Raw_value = anes_extracted$VCF0890,
  useNA = "ifany"
)

attr(
  anes_extracted$VCF0890,
  "label"
)

attr(
  anes_extracted$VCF0890,
  "labels"
)
```
## Audit Existing NA: VCF0890

```{r audit-vcf0890-existing-na}
vcf0890_existing_na <- is.na(
  anes_extracted$VCF0890
)

sum(vcf0890_existing_na)

sum(
  haven::is_tagged_na(
    anes_extracted$VCF0890
  )
)

identical(
  which(vcf0890_existing_na),
  which(
    anes_extracted$VCF0004 == 2024 &
    anes_extracted$VCF0806 == -1
  )
)
```
## Clean VCF0890

```{r clean-vcf0890}
vcf0890_non_substantive <- (
  anes_clean$VCF0890 %in% c(8, 9)
)

sum(vcf0890_non_substantive)

anes_clean$VCF0890[
  vcf0890_non_substantive
] <- NA_real_

table(
  Election_year = anes_clean$VCF0004,
  Clean_value = anes_clean$VCF0890,
  useNA = "ifany"
)

sum(
  is.na(anes_clean$VCF0890)
)

sort(
  unique(
    anes_clean$VCF0890[
      !is.na(anes_clean$VCF0890)
    ]
  )
)
```
## Raw Audit: VCF0894

```{r audit-vcf0894-raw}
table(
  Election_year = anes_extracted$VCF0004,
  Raw_value = anes_extracted$VCF0894,
  useNA = "ifany"
)

attr(
  anes_extracted$VCF0894,
  "label"
)

attr(
  anes_extracted$VCF0894,
  "labels"
)
```
## Audit Existing NA: VCF0894

```{r audit-vcf0894-existing-na}
vcf0894_existing_na <- is.na(
  anes_extracted$VCF0894
)

sum(vcf0894_existing_na)

sum(
  haven::is_tagged_na(
    anes_extracted$VCF0894
  )
)

identical(
  which(vcf0894_existing_na),
  which(
    anes_extracted$VCF0004 == 2024 &
    anes_extracted$VCF0806 == -1
  )
)
```
## Clean VCF0894

```{r clean-vcf0894}
vcf0894_non_substantive <- (
  anes_clean$VCF0894 %in% c(8, 9)
)

sum(vcf0894_non_substantive)

anes_clean$VCF0894[
  vcf0894_non_substantive
] <- NA_real_

table(
  Election_year = anes_clean$VCF0004,
  Clean_value = anes_clean$VCF0894,
  useNA = "ifany"
)

sum(
  is.na(anes_clean$VCF0894)
)

sort(
  unique(
    anes_clean$VCF0894[
      !is.na(anes_clean$VCF0894)
    ]
  )
)
```
## Raw Audit: VCF9223

```{r audit-vcf9223-raw}
table(
  Election_year = anes_extracted$VCF0004,
  Raw_value = anes_extracted$VCF9223,
  useNA = "ifany"
)

attr(
  anes_extracted$VCF9223,
  "label"
)

attr(
  anes_extracted$VCF9223,
  "labels"
)
```
## Clean VCF9223

```{r clean-vcf9223}
vcf9223_structural_missing <- (
  anes_clean$VCF0004 == 2024 &
  anes_clean$VCF9223 == -1
)

vcf9223_non_substantive <- (
  anes_clean$VCF9223 %in% c(-9, -8)
)

sum(vcf9223_structural_missing)

sum(vcf9223_non_substantive)

anes_clean$VCF9223[
  vcf9223_structural_missing |
  vcf9223_non_substantive
] <- NA_real_

table(
  Election_year = anes_clean$VCF0004,
  Clean_value = anes_clean$VCF9223,
  useNA = "ifany"
)

sum(
  is.na(anes_clean$VCF9223)
)

sort(
  unique(
    anes_clean$VCF9223[
      !is.na(anes_clean$VCF9223)
    ]
  )
)
```
## Recode Six Node Directions

```{r recode-node-directions}
nodes_before_direction_recode <- as.data.frame(
  lapply(
    anes_clean[network_vars],
    as.numeric
  )
)

anes_clean[network_vars] <- nodes_before_direction_recode

anes_clean$VCF0806 <- ifelse(
  is.na(nodes_before_direction_recode$VCF0806),
  NA_real_,
  8 - as.numeric(anes_extracted$VCF0806)
)

anes_clean$VCF0809 <- ifelse(
  is.na(nodes_before_direction_recode$VCF0809),
  NA_real_,
  8 - as.numeric(anes_extracted$VCF0809)
)

anes_clean$VCF0879a <- ifelse(
  is.na(nodes_before_direction_recode$VCF0879a),
  NA_real_,
  6 - as.numeric(anes_extracted$VCF0879a)
)

anes_clean$VCF0888 <- ifelse(
  is.na(nodes_before_direction_recode$VCF0888),
  NA_real_,
  4 - as.numeric(anes_extracted$VCF0888)
)

anes_clean$VCF0890 <- ifelse(
  is.na(nodes_before_direction_recode$VCF0890),
  NA_real_,
  4 - as.numeric(anes_extracted$VCF0890)
)

anes_clean$VCF0894 <- ifelse(
  is.na(nodes_before_direction_recode$VCF0894),
  NA_real_,
  4 - as.numeric(anes_extracted$VCF0894)
)
```
## Audit Direction Recoding Frequencies

```{r audit-node-direction-frequencies}
reversed_nodes <- c(
  "VCF0806",
  "VCF0809",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894"
)

direction_frequency_audit <- lapply(
  reversed_nodes,
  function(variable_name) {
    table(
      Before = nodes_before_direction_recode[[variable_name]],
      After = anes_clean[[variable_name]],
      useNA = "ifany"
    )
  }
)

names(direction_frequency_audit) <- reversed_nodes

direction_frequency_audit
```
## Validate Node Directions

```{r validate-node-directions}
direction_validation <- data.frame(
  Node = network_vars,

  Required_action = c(
    "8 - original value",
    "8 - original value",
    "Unchanged",
    "Unchanged",
    "6 - original value",
    "4 - original value",
    "4 - original value",
    "4 - original value",
    "Unchanged"
  ),

  Value_mapping_correct = c(
    all(
      anes_clean$VCF0806[!is.na(anes_clean$VCF0806)] ==
        8 - as.numeric(
          anes_extracted$VCF0806[
            !is.na(anes_clean$VCF0806)
          ]
        )
    ),

    all(
      anes_clean$VCF0809[!is.na(anes_clean$VCF0809)] ==
        8 - as.numeric(
          anes_extracted$VCF0809[
            !is.na(anes_clean$VCF0809)
          ]
        )
    ),

    identical(
      anes_clean$VCF0838,
      nodes_before_direction_recode$VCF0838
    ),

    identical(
      anes_clean$VCF0839,
      nodes_before_direction_recode$VCF0839
    ),

    all(
      anes_clean$VCF0879a[!is.na(anes_clean$VCF0879a)] ==
        6 - as.numeric(
          anes_extracted$VCF0879a[
            !is.na(anes_clean$VCF0879a)
          ]
        )
    ),

    all(
      anes_clean$VCF0888[!is.na(anes_clean$VCF0888)] ==
        4 - as.numeric(
          anes_extracted$VCF0888[
            !is.na(anes_clean$VCF0888)
          ]
        )
    ),

    all(
      anes_clean$VCF0890[!is.na(anes_clean$VCF0890)] ==
        4 - as.numeric(
          anes_extracted$VCF0890[
            !is.na(anes_clean$VCF0890)
          ]
        )
    ),

    all(
      anes_clean$VCF0894[!is.na(anes_clean$VCF0894)] ==
        4 - as.numeric(
          anes_extracted$VCF0894[
            !is.na(anes_clean$VCF0894)
          ]
        )
    ),

    identical(
      anes_clean$VCF9223,
      nodes_before_direction_recode$VCF9223
    )
  ),

  Missing_pattern_unchanged = vapply(
    network_vars,
    function(variable_name) {
      identical(
        is.na(anes_clean[[variable_name]]),
        is.na(
          nodes_before_direction_recode[[variable_name]]
        )
      )
    },
    logical(1)
  ),

  Plain_numeric = vapply(
    anes_clean[network_vars],
    function(x) {
      is.numeric(x) &&
        is.null(attr(x, "labels"))
    },
    logical(1)
  )
)

direction_validation

stopifnot(
  all(direction_validation$Value_mapping_correct),
  all(direction_validation$Missing_pattern_unchanged),
  all(direction_validation$Plain_numeric)
)

"PASS: node direction recoding gate"
```
## Validate Clean Node Codes

```{r validate-clean-node-codes}
observed_clean_values <- lapply(
  anes_clean[network_vars],
  function(x) {
    sort(
      unique(
        as.numeric(
          x[!is.na(x)]
        )
      )
    )
  }
)

observed_clean_values
```
## Audit Missing Values by Year

```{r audit-missing-values-by-year}
missing_by_year <- sapply(
  network_vars,
  function(variable) {
    tapply(
      is.na(anes_clean[[variable]]),
      anes_clean$VCF0004,
      sum
    )
  }
)

missing_by_year
```
## Validate Unchanged Data Structure

```{r validate-unchanged-data-structure}
non_network_vars <- setdiff(
  names(anes_extracted),
  network_vars
)

dim(anes_clean)

identical(
  anes_clean[non_network_vars],
  anes_extracted[non_network_vars]
)
```
## Raw Audit: Education Variable VCF0110

```{r audit-vcf0110-raw}
table(
  Election_year = anes_extracted$VCF0004,
  Raw_value = anes_extracted$VCF0110,
  useNA = "ifany"
)

attr(
  anes_extracted$VCF0110,
  "label"
)

attr(
  anes_extracted$VCF0110,
  "labels"
)
```
## Clean Education Variable VCF0110

```{r clean-vcf0110}
vcf0110_non_substantive <- (
  anes_clean$VCF0110 == 0
)

sum(vcf0110_non_substantive)

anes_clean$VCF0110[
  vcf0110_non_substantive
] <- NA_real_

table(
  Election_year = anes_clean$VCF0004,
  Clean_value = anes_clean$VCF0110,
  useNA = "ifany"
)

sort(
  unique(
    as.numeric(
      anes_clean$VCF0110[
        !is.na(anes_clean$VCF0110)
      ]
    )
  )
)
```
## Validate Intended Variable Changes

```{r validate-intended-variable-changes}
cleaned_vars <- c(
  network_vars,
  "VCF0110"
)

unchanged_vars <- setdiff(
  names(anes_extracted),
  cleaned_vars
)

dim(anes_clean)

identical(
  anes_clean[unchanged_vars],
  anes_extracted[unchanged_vars]
)
```
## Prepare Cleaned Data Folder

```{r create-cleaned-folder}
if (!dir.exists("data/cleaned")) {
  dir.create(
    "data/cleaned",
    recursive = TRUE
  )
}

dir.exists("data/cleaned")
```
## Save Cleaned Data

```{r save-cleaned-data}
cleaned_dir <- file.path(
  project_root,
  "data",
  "cleaned"
)

dir.create(
  cleaned_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

cleaned_file <- file.path(
  cleaned_dir,
  "anes_cdf_20260205_cleaned_2004_2012_2016_2020_2024.rds"
)

saveRDS(
  anes_clean,
  cleaned_file
)

normalizePath(
  cleaned_file,
  winslash = "/"
)

file.exists(cleaned_file)
```

## Validate Cleaned Data File and SHA-256

```{r validate-cleaned-data-file}
anes_clean_check <- readRDS(
  cleaned_file
)

cleaned_file_sha256 <- digest::digest(
  file = cleaned_file,
  algo = "sha256",
  serialize = FALSE
)

cleaned_output_validation <- data.frame(
  File_exists = file.exists(cleaned_file),
  Round_trip_identical = identical(
    anes_clean,
    anes_clean_check
  ),
  Rows = nrow(anes_clean_check),
  Columns = ncol(anes_clean_check),
  SHA256 = cleaned_file_sha256
)

cleaned_output_validation

stopifnot(
  cleaned_output_validation$File_exists,
  cleaned_output_validation$Round_trip_identical,
  identical(
    dim(anes_clean_check),
    c(25197L, 18L)
  ),
  grepl(
    "^[0-9a-f]{64}$",
    cleaned_file_sha256
  )
)

"PASS: cleaned data write-read-hash gate"
```
