---
title: "04 Analysis-Ready Data"
output: html_document
---

## Purpose

This document prepares the analysis-ready ANES data for 2004, 2012, 2016, 2020, and 2024 from the validated cleaned dataset. It will apply the approved within-year nine-node complete-case rule while retaining identification, education, survey-mode, and survey-weight variables for later audits. No network model is estimated in this document.

## Confirm Cleaned Data Input

```{r confirm-cleaned-input}
current_dir <- normalizePath(
  ".",
  winslash = "/"
)

if (basename(current_dir) == "notebooks") {
  project_root <- dirname(current_dir)
} else {
  project_root <- current_dir
}

knitr::opts_knit$set(
  root.dir = project_root
)

cleaned_file <- file.path(
  project_root,
  "data",
  "cleaned",
  "anes_cdf_20260205_cleaned_2004_2012_2016_2020_2024.rds"
)

project_root
cleaned_file
file.exists(cleaned_file)
```
## Load Cleaned Data

```{r load-cleaned-data}
anes_cleaned <- readRDS(cleaned_file)

dim(anes_cleaned)

table(
  anes_cleaned$VCF0004,
  useNA = "ifany"
)
```
## Define Fixed Network Nodes

```{r define-network-nodes}
network_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223"
)

network_vars

length(network_vars)

all(
  network_vars %in% names(anes_cleaned)
)
```
## Audit Nine-Node Complete Cases

```{r audit-nine-node-complete-cases}
nine_node_complete <- complete.cases(
  anes_cleaned[network_vars]
)

complete_case_audit <- table(
  Election_year = anes_cleaned$VCF0004,
  Nine_node_complete = nine_node_complete
)

complete_case_audit

rowSums(complete_case_audit)
```
## Create Analysis-Ready Data

```{r create-analysis-ready-data}
anes_analysis_ready <- anes_cleaned[
  nine_node_complete,
]

dim(anes_analysis_ready)

table(
  anes_analysis_ready$VCF0004
)

all(
  complete.cases(
    anes_analysis_ready[network_vars]
  )
)
```
## Validate Analysis-Ready Membership After Direction Repair

```{r validate-analysis-ready-membership}
expected_year_counts <- c(
  "2004" = 789L,
  "2012" = 4353L,
  "2016" = 2709L,
  "2020" = 5390L,
  "2024" = 3525L
)

observed_year_counts <- table(
  as.numeric(
    anes_analysis_ready$VCF0004
  )
)

membership_key <- data.frame(
  Election_year = as.numeric(
    anes_analysis_ready$VCF0004
  ),
  Within_study_ID = as.numeric(
    anes_analysis_ready$VCF0006
  ),
  Cross_year_ID = as.numeric(
    anes_analysis_ready$VCF0006a
  )
)

membership_sha256 <- digest::digest(
  membership_key,
  algo = "sha256"
)

analysis_ready_object_sha256 <- digest::digest(
  anes_analysis_ready,
  algo = "sha256",
  serialize = TRUE
)

expected_analysis_ready_object_sha256 <-
  "78e5c1cff278a6640834113b769ef4251807ab30659b6effcfd6a14db58ada6c"

pre_repair_membership_sha256 <-
  "f16f3fe76315e03495595af62368124297c8e52fa475df80d28959393b65f334"

membership_hash_audit <- data.frame(
  Pre_repair_SHA256 = pre_repair_membership_sha256,
  Rebuilt_SHA256 = membership_sha256,
  Identical = identical(
    membership_sha256,
    pre_repair_membership_sha256
  )
)

analysis_ready_rebuild_validation <- data.frame(
  Check = c(
    "Dimensions remain 16766 by 18",
    "Election-year sample sizes are unchanged",
    "All nine nodes are complete",
    "Respondent membership and row order are unchanged",
    "Complete analysis-ready object matches the authorised object hash"
  ),
  Passed = c(
    identical(
      dim(anes_analysis_ready),
      c(16766L, 18L)
    ),
    identical(
      names(observed_year_counts),
      names(expected_year_counts)
    ) &&
      identical(
        as.integer(observed_year_counts),
        unname(expected_year_counts)
      ),
    all(
      complete.cases(
        anes_analysis_ready[network_vars]
      )
    ),
    identical(
      membership_sha256,
      pre_repair_membership_sha256
    ),
    identical(
      analysis_ready_object_sha256,
      expected_analysis_ready_object_sha256
    )
  )
)

observed_year_counts

membership_hash_audit

analysis_ready_rebuild_validation

stopifnot(
  all(analysis_ready_rebuild_validation$Passed)
)

"PASS: analysis-ready membership invariance gate"
```
## Save Analysis-Ready Data

```{r save-analysis-ready-data}
analysis_ready_dir <- file.path(
  project_root,
  "data",
  "analysis_ready"
)

dir.create(
  analysis_ready_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

analysis_ready_file <- file.path(
  analysis_ready_dir,
  "anes_cdf_20260205_analysis_ready_nine_node_complete_case_2004_2012_2016_2020_2024.rds"
)

saveRDS(
  anes_analysis_ready,
  analysis_ready_file,
  version = 3,
  compress = "gzip"
)

normalizePath(
  analysis_ready_file,
  winslash = "/"
)

file.exists(analysis_ready_file)
```
## Validate Analysis-Ready Data File

```{r validate-analysis-ready-data-file}
anes_analysis_ready_check <- readRDS(
  analysis_ready_file
)

analysis_ready_reload_identical <- identical(
  anes_analysis_ready,
  anes_analysis_ready_check
)

analysis_ready_reload_hash_matches <- identical(
  digest::digest(
    anes_analysis_ready_check,
    algo = "sha256",
    serialize = TRUE
  ),
  expected_analysis_ready_object_sha256
)

stopifnot(
  analysis_ready_reload_identical,
  analysis_ready_reload_hash_matches
)

data.frame(
  Check = c(
    "Reloaded object is identical to the in-memory object",
    "Reloaded object matches the authorised object hash"
  ),
  Passed = c(
    analysis_ready_reload_identical,
    analysis_ready_reload_hash_matches
  )
)

dim(anes_analysis_ready_check)
```
## Inspect Cross-Year Respondent ID

```{r inspect-cross-year-respondent-id}
class(
  anes_analysis_ready$VCF0006a
)

attr(
  anes_analysis_ready$VCF0006a,
  "label"
)

attr(
  anes_analysis_ready$VCF0006a,
  "labels"
)

summary(
  anes_analysis_ready$VCF0006a
)

table(
  Election_year = anes_analysis_ready$VCF0004,
  ID_is_NA = is.na(anes_analysis_ready$VCF0006a)
)
```
## Check Within-Year ID Uniqueness

```{r check-within-year-id-uniqueness}
id_year_pairs <- data.frame(
  Election_year = anes_analysis_ready$VCF0004,
  Cross_year_ID = anes_analysis_ready$VCF0006a
)

duplicate_within_year <- (
  duplicated(id_year_pairs) |
  duplicated(id_year_pairs, fromLast = TRUE)
)

table(
  Election_year = id_year_pairs$Election_year,
  Duplicate_within_year = duplicate_within_year
)

sum(duplicate_within_year)
```
## Audit Respondent Participation Across Years

```{r audit-respondent-participation}
respondent_wave_counts <- as.integer(
  table(id_year_pairs$Cross_year_ID)
)

wave_count_audit <- table(
  Number_of_years = respondent_wave_counts
)

wave_count_audit

length(respondent_wave_counts)

sum(respondent_wave_counts > 1)

sum(respondent_wave_counts)
```
## Pairwise Respondent Overlap Audit

```{r pairwise-respondent-overlap}
year_id_list <- split(
  id_year_pairs$Cross_year_ID,
  id_year_pairs$Election_year
)

pairwise_overlap <- sapply(
  year_id_list,
  function(column_ids) {
    sapply(
      year_id_list,
      function(row_ids) {
        length(
          intersect(column_ids, row_ids)
        )
      }
    )
  }
)

pairwise_overlap

all(
  pairwise_overlap[
    row(pairwise_overlap) != col(pairwise_overlap)
  ] == 0
)
```
## Verify Retained Audit Variables

```{r verify-retained-audit-variables}
retained_audit_vars <- c(
  "Version",
  "VCF0004",
  "VCF0006",
  "VCF0006a",
  "VCF0009x",
  "VCF0009y",
  "VCF0009z",
  "VCF0017",
  "VCF0110"
)

retained_audit_vars

all(
  retained_audit_vars %in%
    names(anes_analysis_ready)
)

identical(
  names(anes_analysis_ready),
  names(anes_cleaned)
)
```
## Audit Education Availability

```{r audit-education-availability}
education_missing_audit <- table(
  Election_year = anes_analysis_ready$VCF0004,
  Education_is_NA = is.na(
    anes_analysis_ready$VCF0110
  )
)

education_missing_audit

rowSums(education_missing_audit)
```
## Build Sample Flow Audit

```{r build-sample-flow-audit}
sample_flow_audit <- data.frame(
  Election_year = as.numeric(
    rownames(complete_case_audit)
  ),
  Cleaned_n = rowSums(
    complete_case_audit
  ),
  Excluded_missing_nine_node_n =
    complete_case_audit[, "FALSE"],
  Analysis_ready_n =
    complete_case_audit[, "TRUE"],
  Education_missing_n =
    education_missing_audit[, "TRUE"]
)

sample_flow_audit

all(
  sample_flow_audit$Cleaned_n ==
    sample_flow_audit$Excluded_missing_nine_node_n +
    sample_flow_audit$Analysis_ready_n
)
```
## Save Sample Flow Audit

```{r save-sample-flow-audit}
audit_dir <- file.path(
  project_root,
  "data",
  "audit"
)

dir.create(
  audit_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

sample_flow_file <- file.path(
  audit_dir,
  "anes_cdf_20260205_sample_flow_nine_node_complete_case_2004_2012_2016_2020_2024.csv"
)

write.csv(
  sample_flow_audit,
  sample_flow_file,
  row.names = FALSE
)

normalizePath(
  sample_flow_file,
  winslash = "/"
)

file.exists(sample_flow_file)
```
## Validate Sample Flow Audit File

```{r validate-sample-flow-audit}
sample_flow_check <- read.csv(
  sample_flow_file
)

sample_flow_check

identical(
  names(sample_flow_audit),
  names(sample_flow_check)
)

all(
  sample_flow_audit ==
    sample_flow_check
)
```
## Save Pairwise Respondent Overlap Audit

```{r save-pairwise-overlap-audit}
pairwise_overlap_audit <- data.frame(
  Election_year = as.numeric(
    rownames(pairwise_overlap)
  ),
  pairwise_overlap,
  row.names = NULL,
  check.names = FALSE
)

pairwise_overlap_audit

overlap_audit_file <- file.path(
  audit_dir,
  "anes_cdf_20260205_analysis_ready_pairwise_respondent_overlap_2004_2012_2016_2020_2024.csv"
)

write.csv(
  pairwise_overlap_audit,
  overlap_audit_file,
  row.names = FALSE
)

normalizePath(
  overlap_audit_file,
  winslash = "/"
)

file.exists(overlap_audit_file)
```
## Validate Pairwise Overlap Audit File

```{r validate-pairwise-overlap-audit}
pairwise_overlap_check <- read.csv(
  overlap_audit_file,
  check.names = FALSE
)

pairwise_overlap_check

identical(
  names(pairwise_overlap_audit),
  names(pairwise_overlap_check)
)

all(
  pairwise_overlap_audit ==
    pairwise_overlap_check
)
```
