---
title: "05 Preliminary Data Suitability Audit"
output: html_document
---

## Purpose

This document conducts a descriptive, non-model-based audit of category support in the fixed nine-node complete-case ANES samples for 2004, 2012, 2016, 2020, and 2024. It does not estimate correlations or network models, conduct weighted analysis, or interpret substantive results.

## Confirm Analysis-Ready Data Input

```{r confirm-analysis-ready-input}
current_dir <- normalizePath(
  ".",
  winslash = "/"
)

if (basename(current_dir) == "notebooks") {
  project_root <- dirname(current_dir)
} else {
  project_root <- current_dir
}

knitr::opts_knit$set(
  root.dir = project_root
)

analysis_ready_file <- file.path(
  project_root,
  "data",
  "analysis_ready",
  "anes_cdf_20260205_analysis_ready_nine_node_complete_case_2004_2012_2016_2020_2024.rds"
)

project_root
analysis_ready_file
file.exists(analysis_ready_file)
```
## Load Analysis-Ready Data

```{r load-analysis-ready-data}
anes_analysis_ready <- readRDS(
  analysis_ready_file
)

dim(anes_analysis_ready)

table(
  anes_analysis_ready$VCF0004,
  useNA = "ifany"
)
```
## Define Fixed Network Nodes

```{r define-fixed-network-nodes}
network_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223"
)

network_vars

length(network_vars)

all(
  network_vars %in%
    names(anes_analysis_ready)
)

all(
  complete.cases(
    anes_analysis_ready[network_vars]
  )
)
```
## Category Audit: VCF0806

```{r category-audit-vcf0806}
vcf0806_category_counts <- table(
  Election_year = anes_analysis_ready$VCF0004,
  Category = anes_analysis_ready$VCF0806
)

vcf0806_category_counts

rowSums(vcf0806_category_counts)

apply(
  vcf0806_category_counts,
  1,
  min
)
```
## Category Audit: VCF0809

```{r category-audit-vcf0809}
vcf0809_category_counts <- table(
  Election_year = anes_analysis_ready$VCF0004,
  Category = anes_analysis_ready$VCF0809
)

vcf0809_category_counts

rowSums(vcf0809_category_counts)

apply(
  vcf0809_category_counts,
  1,
  min
)
```
## Category Audit: VCF0838

```{r category-audit-vcf0838}
vcf0838_category_counts <- table(
  Election_year = anes_analysis_ready$VCF0004,
  Category = anes_analysis_ready$VCF0838
)

vcf0838_category_counts

rowSums(vcf0838_category_counts)

apply(
  vcf0838_category_counts,
  1,
  min
)
```
## Category Audit: VCF0839

```{r category-audit-vcf0839}
vcf0839_category_counts <- table(
  Election_year = anes_analysis_ready$VCF0004,
  Category = anes_analysis_ready$VCF0839
)

vcf0839_category_counts

rowSums(vcf0839_category_counts)

apply(
  vcf0839_category_counts,
  1,
  min
)
```

## Category Audit: VCF0879a

```{r category-audit-vcf0879a}
vcf0879a_category_counts <- table(
  Election_year = anes_analysis_ready$VCF0004,
  Category = anes_analysis_ready$VCF0879a
)

vcf0879a_category_counts

rowSums(vcf0879a_category_counts)

apply(
  vcf0879a_category_counts,
  1,
  min
)
```

## Category Audit: VCF0888

```{r category-audit-vcf0888}
vcf0888_category_counts <- table(
  Election_year = anes_analysis_ready$VCF0004,
  Category = anes_analysis_ready$VCF0888
)

vcf0888_category_counts

rowSums(vcf0888_category_counts)

apply(
  vcf0888_category_counts,
  1,
  min
)
```

## Category Audit: VCF0890

```{r category-audit-vcf0890}
vcf0890_category_counts <- table(
  Election_year = anes_analysis_ready$VCF0004,
  Category = anes_analysis_ready$VCF0890
)

vcf0890_category_counts

rowSums(vcf0890_category_counts)

apply(
  vcf0890_category_counts,
  1,
  min
)
```

## Category Audit: VCF0894

```{r category-audit-vcf0894}
vcf0894_category_counts <- table(
  Election_year = anes_analysis_ready$VCF0004,
  Category = anes_analysis_ready$VCF0894
)

vcf0894_category_counts

rowSums(vcf0894_category_counts)

apply(
  vcf0894_category_counts,
  1,
  min
)
```

## Category Audit: VCF9223

```{r category-audit-vcf9223}
vcf9223_category_counts <- table(
  Election_year = anes_analysis_ready$VCF0004,
  Category = anes_analysis_ready$VCF9223
)

vcf9223_category_counts

rowSums(vcf9223_category_counts)

apply(
  vcf9223_category_counts,
  1,
  min
)
```
## Analysis-Ready Validation Gate

```{r analysis-ready-validation-gate}
expected_year_counts <- c(
  "2004" = 789L,
  "2012" = 4353L,
  "2016" = 2709L,
  "2020" = 5390L,
  "2024" = 3525L
)

observed_year_counts <- table(
  anes_analysis_ready$VCF0004
)

stopifnot(
  identical(
    dim(anes_analysis_ready),
    c(16766L, 18L)
  ),
  identical(
    names(observed_year_counts),
    names(expected_year_counts)
  ),
  identical(
    as.integer(observed_year_counts),
    unname(expected_year_counts)
  ),
  length(network_vars) == 9L,
  !anyDuplicated(network_vars),
  all(
    network_vars %in%
      names(anes_analysis_ready)
  ),
  all(
    complete.cases(
      anes_analysis_ready[network_vars]
    )
  )
)

"PASS: analysis-ready validation gate"
```
## Category Validation Gate

```{r category-validation-gate}
expected_categories <- list(
  VCF0806 = 1:7,
  VCF0809 = 1:7,
  VCF0838 = 1:4,
  VCF0839 = 1:7,
  VCF0879a = c(1, 3, 5),
  VCF0888 = 1:3,
  VCF0890 = 1:3,
  VCF0894 = 1:3,
  VCF9223 = 1:4
)

stopifnot(
  identical(
    names(expected_categories),
    network_vars
  )
)

for (variable_name in network_vars) {
  category_counts <- table(
    Election_year = factor(
      as.numeric(
        anes_analysis_ready$VCF0004
      ),
      levels = as.integer(
        names(expected_year_counts)
      )
    ),
    Category = factor(
      as.numeric(
        anes_analysis_ready[[variable_name]]
      ),
      levels = expected_categories[[variable_name]]
    )
  )

  stopifnot(
    identical(
      as.integer(rowSums(category_counts)),
      unname(expected_year_counts)
    ),
    all(category_counts > 0)
  )
}

"PASS: category validation gate"
```
## Minimum Category Support Summary

```{r minimum-category-support-summary}
minimum_category_count <- matrix(
  NA_integer_,
  nrow = length(expected_year_counts),
  ncol = length(network_vars),
  dimnames = list(
    Election_year = names(expected_year_counts),
    Node = network_vars
  )
)

minimum_category_percent <- matrix(
  NA_real_,
  nrow = length(expected_year_counts),
  ncol = length(network_vars),
  dimnames = list(
    Election_year = names(expected_year_counts),
    Node = network_vars
  )
)

for (year_name in names(expected_year_counts)) {
  for (variable_name in network_vars) {
    category_counts <- table(
      factor(
        as.numeric(
          anes_analysis_ready[
            anes_analysis_ready$VCF0004 ==
              as.numeric(year_name),
            variable_name
          ]
        ),
        levels = expected_categories[[variable_name]]
      )
    )

    minimum_category_count[
      year_name,
      variable_name
    ] <- min(category_counts)

    minimum_category_percent[
      year_name,
      variable_name
    ] <- round(
      100 * min(category_counts) /
        sum(category_counts),
      2
    )
  }
}

minimum_category_count

minimum_category_percent

min(minimum_category_count)

min(minimum_category_percent)
```
## Pairwise Contingency-Table Cell Audit

```{r pairwise-contingency-cell-audit}
node_pairs <- combn(
  network_vars,
  2,
  simplify = FALSE
)

pairwise_cell_audit <- data.frame()

for (year_name in names(expected_year_counts)) {
  year_data <- anes_analysis_ready[
    anes_analysis_ready$VCF0004 ==
      as.numeric(year_name),
    network_vars
  ]

  for (node_pair in node_pairs) {
    cell_counts <- table(
      factor(
        as.numeric(year_data[[node_pair[1]]]),
        levels = expected_categories[[node_pair[1]]]
      ),
      factor(
        as.numeric(year_data[[node_pair[2]]]),
        levels = expected_categories[[node_pair[2]]]
      )
    )

    pairwise_cell_audit <- rbind(
      pairwise_cell_audit,
      data.frame(
        Election_year = as.integer(year_name),
        Node_1 = node_pair[1],
        Node_2 = node_pair[2],
        Pair_N = sum(cell_counts),
        Total_cells = length(cell_counts),
        Cells_0 = sum(cell_counts == 0),
        Cells_1_4 = sum(
          cell_counts >= 1 &
            cell_counts <= 4
        ),
        Cells_5_9 = sum(
          cell_counts >= 5 &
            cell_counts <= 9
        ),
        Cells_10_plus = sum(
          cell_counts >= 10
        ),
        Minimum_cell = min(cell_counts)
      )
    )
  }
}

pairwise_cell_audit$Pair_with_zero <- as.integer(
  pairwise_cell_audit$Cells_0 > 0
)

pairwise_cell_audit$Pair_with_1_4 <- as.integer(
  pairwise_cell_audit$Cells_1_4 > 0
)

pairwise_cell_audit$Pair_with_5_9 <- as.integer(
  pairwise_cell_audit$Cells_5_9 > 0
)

stopifnot(
  nrow(pairwise_cell_audit) == 180L,
  all(
    pairwise_cell_audit$Pair_N ==
      expected_year_counts[
        as.character(
          pairwise_cell_audit$Election_year
        )
      ]
  ),
  all(
    pairwise_cell_audit$Total_cells ==
      pairwise_cell_audit$Cells_0 +
      pairwise_cell_audit$Cells_1_4 +
      pairwise_cell_audit$Cells_5_9 +
      pairwise_cell_audit$Cells_10_plus
  )
)

pairwise_cell_summary <- aggregate(
  pairwise_cell_audit[
    c(
      "Cells_0",
      "Cells_1_4",
      "Cells_5_9",
      "Cells_10_plus",
      "Pair_with_zero",
      "Pair_with_1_4",
      "Pair_with_5_9"
    )
  ],
  by = list(
    Election_year =
      pairwise_cell_audit$Election_year
  ),
  FUN = sum
)

pairwise_cell_summary$Total_pairs <- 36L

pairwise_cell_summary <- pairwise_cell_summary[
  c(
    "Election_year",
    "Total_pairs",
    "Cells_0",
    "Cells_1_4",
    "Cells_5_9",
    "Cells_10_plus",
    "Pair_with_zero",
    "Pair_with_1_4",
    "Pair_with_5_9"
  )
]

low_frequency_pairs <- pairwise_cell_audit[
  pairwise_cell_audit$Cells_0 > 0 |
    pairwise_cell_audit$Cells_1_4 > 0,
  c(
    "Election_year",
    "Node_1",
    "Node_2",
    "Pair_N",
    "Cells_0",
    "Cells_1_4",
    "Minimum_cell"
  )
]

pairwise_cell_summary

low_frequency_pairs

"PASS: pairwise cell-count audit"
```
## 2004 Sparse-Cell Precision Warning

Passing the structural input checks does not demonstrate that every estimated edge will be precise. The 2004 sample contains zero and very low-frequency cells in several pairwise contingency tables. The year remains eligible for subsequent diagnostic estimation, but its edge-weight accuracy must be assessed carefully before substantive interpretation.

```{r record-2004-sparse-cell-warning}
pairwise_summary_2004 <- pairwise_cell_summary[
  pairwise_cell_summary$Election_year == 2004,
]

pairwise_details_2004 <- pairwise_cell_audit[
  pairwise_cell_audit$Election_year == 2004,
]

sparse_cell_warning_2004 <- data.frame(
  Election_year = 2004L,
  Total_node_pairs = pairwise_summary_2004$Total_pairs,
  Pairs_with_zero_cells =
    pairwise_summary_2004$Pair_with_zero,
  Pairs_with_cells_1_to_4 =
    pairwise_summary_2004$Pair_with_1_4,
  Minimum_cell_count = min(
    pairwise_details_2004$Minimum_cell
  ),
  Status = "PASS WITH SPARSE-CELL WARNING"
)

sparse_cell_warning_2004

stopifnot(
  nrow(pairwise_details_2004) == 36L,
  sparse_cell_warning_2004$Total_node_pairs == 36L,
  sparse_cell_warning_2004$Pairs_with_zero_cells == 4L,
  sparse_cell_warning_2004$Pairs_with_cells_1_to_4 == 19L,
  sparse_cell_warning_2004$Minimum_cell_count == 0L
)

"PASS WITH WARNING: 2004 contains sparse pairwise cells"
```
## Save Pairwise Cell Audit

```{r save-pairwise-cell-audit}
pairwise_cell_audit_file <- file.path(
  project_root,
  "data",
  "audit",
  "anes_cdf_20260205_analysis_ready_pairwise_cell_audit_2004_2012_2016_2020_2024.csv"
)

write.csv(
  pairwise_cell_audit,
  pairwise_cell_audit_file,
  row.names = FALSE
)

normalizePath(
  pairwise_cell_audit_file,
  winslash = "/"
)

file.exists(pairwise_cell_audit_file)
```
## Validate Pairwise Cell Audit File

```{r validate-pairwise-cell-audit}
pairwise_cell_audit_check <- read.csv(
  pairwise_cell_audit_file,
  check.names = FALSE
)

dim(pairwise_cell_audit_check)

identical(
  names(pairwise_cell_audit),
  names(pairwise_cell_audit_check)
)

isTRUE(
  all.equal(
    pairwise_cell_audit,
    pairwise_cell_audit_check,
    check.attributes = FALSE
  )
)
```