---
title: "06 Polychoric Input Diagnostics"
output: html_document
---

## Purpose

This notebook evaluates category support and whether the fixed nine-node complete-case ANES data for 2004, 2012, 2016, 2020, and 2024 can produce mathematically usable polychoric correlation matrices. It records estimation warnings, correlation-matrix checks, eigenvalues, and positive-definiteness results. It does not estimate a GGM, apply matrix repairs, use survey weights, or interpret substantive results.

## Check Required Diagnostic Packages

```{r check-diagnostic-packages}
package_check <- data.frame(
  Package = c("qgraph", "lavaan"),
  Installed = c(
    requireNamespace("qgraph", quietly = TRUE),
    requireNamespace("lavaan", quietly = TRUE)
  )
)

package_check
```
## Record Diagnostic Package Versions

```{r record-diagnostic-package-versions}
diagnostic_package_versions <- data.frame(
  Package = c("qgraph", "lavaan"),
  Version = c(
    as.character(packageVersion("qgraph")),
    as.character(packageVersion("lavaan"))
  )
)

diagnostic_package_versions
```
## Confirm Analysis-Ready Data Input

```{r confirm-analysis-ready-input}
current_dir <- normalizePath(
  getwd(),
  winslash = "/"
)

if (basename(current_dir) == "notebooks") {
  project_root <- dirname(current_dir)
} else {
  project_root <- current_dir
}

knitr::opts_knit$set(
  root.dir = project_root
)

analysis_ready_file <- file.path(
  project_root,
  "data",
  "analysis_ready",
  "anes_cdf_20260205_analysis_ready_nine_node_complete_case_2004_2012_2016_2020_2024.rds"
)

project_root

analysis_ready_file

file.exists(analysis_ready_file)
```
## Load Analysis-Ready Data

```{r load-analysis-ready-data}
anes_analysis_ready <- readRDS(
  analysis_ready_file
)

dim(anes_analysis_ready)

table(
  anes_analysis_ready$VCF0004
)
```

## Define Fixed Network Specification

```{r define-network-specification}
network_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223"
)

expected_year_counts <- c(
  "2004" = 789,
  "2012" = 4353,
  "2016" = 2709,
  "2020" = 5390,
  "2024" = 3525
)

expected_categories <- list(
  VCF0806  = c(1, 2, 3, 4, 5, 6, 7),
  VCF0809  = c(1, 2, 3, 4, 5, 6, 7),
  VCF0838  = c(1, 2, 3, 4),
  VCF0839  = c(1, 2, 3, 4, 5, 6, 7),
  VCF0879a = c(1, 3, 5),
  VCF0888  = c(1, 2, 3),
  VCF0890  = c(1, 2, 3),
  VCF0894  = c(1, 2, 3),
  VCF9223  = c(1, 2, 3, 4)
)

network_vars

expected_year_counts

identical(
  names(expected_categories),
  network_vars
)
```
## Validate Corrected Node-Direction Input

```{r validate-corrected-node-direction-input}
expected_analysis_ready_object_sha256 <-
  "78e5c1cff278a6640834113b769ef4251807ab30659b6effcfd6a14db58ada6c"

analysis_ready_file_sha256 <- digest::digest(
  file = analysis_ready_file,
  algo = "sha256",
  serialize = FALSE
)

analysis_ready_object_sha256 <- digest::digest(
  anes_analysis_ready,
  algo = "sha256",
  serialize = TRUE
)

reversed_nodes <- c(
  "VCF0806",
  "VCF0809",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894"
)

unchanged_direction_nodes <- c(
  "VCF0838",
  "VCF0839",
  "VCF9223"
)

node_direction_specification <- data.frame(
  Node = network_vars,

  Direction_processing = c(
    "Reversed in 03",
    "Reversed in 03",
    "Unchanged",
    "Unchanged",
    "Reversed in 03",
    "Reversed in 03",
    "Reversed in 03",
    "Reversed in 03",
    "Unchanged"
  ),

  Higher_values_mean = c(
    "Greater support for government health insurance",
    "Greater government responsibility for jobs and living standards",
    "More permissive abortion policy",
    "More government services and spending",
    "Greater support for increasing immigration",
    "Greater support for increased federal crime spending",
    "Greater support for increased federal public-school spending",
    "Greater support for increased federal welfare spending",
    "Lower perceived likelihood that immigration takes jobs"
  )
)

direction_input_validation <- data.frame(
  Check = c(
    "Corrected analysis-ready object SHA-256 matches",
    "Direction table follows the fixed node order",
    "Six reversed nodes are registered correctly",
    "Three unchanged nodes are registered correctly",
    "All node columns are plain numeric"
  ),

  Passed = c(
    identical(
      analysis_ready_object_sha256,
      expected_analysis_ready_object_sha256
    ),
    identical(
      node_direction_specification$Node,
      network_vars
    ),
    identical(
      node_direction_specification$Node[
        node_direction_specification$Direction_processing ==
          "Reversed in 03"
      ],
      reversed_nodes
    ),
    identical(
      node_direction_specification$Node[
        node_direction_specification$Direction_processing ==
          "Unchanged"
      ],
      unchanged_direction_nodes
    ),
    all(
      vapply(
        anes_analysis_ready[network_vars],
        function(x) {
          is.numeric(x) &&
            is.null(attr(x, "labels"))
        },
        logical(1)
      )
    )
  )
)

data.frame(
  Hash_type = c(
    "Compressed RDS file",
    "Deserialised analysis-ready object",
    "Expected deserialised object"
  ),
  SHA256 = c(
    analysis_ready_file_sha256,
    analysis_ready_object_sha256,
    expected_analysis_ready_object_sha256
  ),
  stringsAsFactors = FALSE
)

node_direction_specification

direction_input_validation

stopifnot(
  all(direction_input_validation$Passed)
)

"PASS: corrected node-direction input gate"
```
## Validate Analysis-Ready Categories

```{r validate-analysis-ready-categories}
category_match <- matrix(
  NA,
  nrow = length(expected_year_counts),
  ncol = length(network_vars),
  dimnames = list(
    Election_year = names(expected_year_counts),
    Node = network_vars
  )
)

for (year_name in names(expected_year_counts)) {
  for (variable_name in network_vars) {
    year_rows <- (
      anes_analysis_ready$VCF0004 ==
        as.numeric(year_name)
    )

    observed_categories <- sort(
      unique(
        as.numeric(
          anes_analysis_ready[[variable_name]][year_rows]
        )
      )
    )

    category_match[
      year_name,
      variable_name
    ] <- identical(
      observed_categories,
      as.numeric(
        expected_categories[[variable_name]]
      )
    )
  }
}

category_match

year_counts <- table(
  anes_analysis_ready$VCF0004
)

stopifnot(
  identical(
    dim(anes_analysis_ready),
    c(16766L, 18L)
  ),
  identical(
    names(year_counts),
    names(expected_year_counts)
  ),
  all(
    as.numeric(year_counts) ==
      unname(expected_year_counts)
  ),
  all(
    network_vars %in%
      names(anes_analysis_ready)
  ),
  all(
    complete.cases(
      anes_analysis_ready[network_vars]
    )
  ),
  all(category_match)
)

"PASS: analysis-ready category validation gate"
```
## Create Ordered-Factor Node Data

```{r create-ordered-node-data}
anes_ordinal <- anes_analysis_ready

for (variable_name in network_vars) {
  anes_ordinal[[variable_name]] <- ordered(
    as.numeric(
      anes_analysis_ready[[variable_name]]
    ),
    levels = expected_categories[[variable_name]]
  )
}

lapply(
  anes_ordinal[network_vars],
  levels
)
```

## Validate Ordered-Factor Conversion

```{r validate-ordered-node-data}
ordered_validation <- data.frame(
  Node = network_vars,
  Is_ordered = vapply(
    anes_ordinal[network_vars],
    is.ordered,
    logical(1)
  ),
  Levels_match = vapply(
    network_vars,
    function(variable_name) {
      identical(
        as.numeric(
          levels(
            anes_ordinal[[variable_name]]
          )
        ),
        as.numeric(
          expected_categories[[variable_name]]
        )
      )
    },
    logical(1)
  ),
  Missing_n = vapply(
    network_vars,
    function(variable_name) {
      sum(
        is.na(
          anes_ordinal[[variable_name]]
        )
      )
    },
    numeric(1)
  )
)

ordered_validation

stopifnot(
  identical(
    dim(anes_ordinal),
    dim(anes_analysis_ready)
  ),
  identical(
    anes_ordinal$VCF0004,
    anes_analysis_ready$VCF0004
  ),
  all(ordered_validation$Is_ordered),
  all(ordered_validation$Levels_match),
  all(ordered_validation$Missing_n == 0)
)

"PASS: ordered-factor conversion gate"
```

## Inspect Polychoric Function Interfaces

```{r inspect-polychoric-interfaces}
qgraph_cor_auto_arguments <- names(
  formals(qgraph::cor_auto)
)

lavaan_lavcor_arguments <- names(
  formals(lavaan::lavCor)
)

qgraph_cor_auto_arguments

lavaan_lavcor_arguments
```
## Prepare 2004 Ordered Pilot Data

```{r prepare-2004-ordered-pilot}
year_2004_rows <- (
  anes_ordinal$VCF0004 == 2004
)

pilot_data_2004 <- as.data.frame(
  anes_ordinal[
    year_2004_rows,
    network_vars
  ]
)

dim(pilot_data_2004)

vapply(
  pilot_data_2004,
  is.ordered,
  logical(1)
)

all(
  complete.cases(pilot_data_2004)
)

stopifnot(
  identical(
    dim(pilot_data_2004),
    c(789L, 9L)
  ),
  identical(
    names(pilot_data_2004),
    network_vars
  ),
  all(
    vapply(
      pilot_data_2004,
      is.ordered,
      logical(1)
    )
  ),
  all(
    complete.cases(pilot_data_2004)
  )
)

"PASS: 2004 ordered pilot data gate"
```
## Estimate 2004 Polychoric Correlation Matrix

The qgraph call below records qgraph-level warnings. Because `qgraph::cor_auto()` suppresses warnings from its internal `lavaan::lavCor()` call, the following direct lavaan reproduction is used for the formal estimation-warning audit.

```{r estimate-2004-polychoric-qgraph}
qgraph_warnings_2004 <- character(0)

polychoric_cor_2004 <- withCallingHandlers(
  qgraph::cor_auto(
    data = pilot_data_2004,
    detectOrdinal = TRUE,
    ordinalLevelMax = 7,
    forcePD = FALSE,
    missing = "listwise",
    verbose = FALSE
  ),
  warning = function(warning_condition) {
    qgraph_warnings_2004 <<- c(
      qgraph_warnings_2004,
      conditionMessage(warning_condition)
    )
    invokeRestart("muffleWarning")
  }
)

polychoric_cor_2004 <- unclass(
  polychoric_cor_2004
)

qgraph_warnings_2004
class(polychoric_cor_2004)
dim(polychoric_cor_2004)
round(polychoric_cor_2004, 3)
```
## Reproduce 2004 Matrix with lavaan

```{r reproduce-2004-polychoric-lavaan}
lavaan_warnings_2004 <- character(0)

lavaan_cor_2004 <- withCallingHandlers(
  lavaan::lavCor(
    object = pilot_data_2004,
    ordered = network_vars,
    missing = "listwise",
    estimator = "two.step",
    meanstructure = FALSE,
    cor_smooth = FALSE,
    output = "cor"
  ),
  warning = function(warning_condition) {
    lavaan_warnings_2004 <<- c(
      lavaan_warnings_2004,
      conditionMessage(warning_condition)
    )
    invokeRestart("muffleWarning")
  }
)

lavaan_cor_2004 <- unclass(
  lavaan_cor_2004
)

lavaan_warnings_2004
class(lavaan_cor_2004)
dim(lavaan_cor_2004)
```
## Compare 2004 Polychoric Implementations

```{r compare-2004-polychoric-implementations}
implementation_comparison_2004 <- data.frame(
  Maximum_absolute_difference = max(
    abs(
      polychoric_cor_2004 -
        lavaan_cor_2004
    )
  ),
  Equal_within_tolerance = all(
    abs(
      polychoric_cor_2004 -
        lavaan_cor_2004
    ) < 1e-12
  )
)

implementation_comparison_2004

stopifnot(
  identical(
    dim(polychoric_cor_2004),
    c(9L, 9L)
  ),
  identical(
    dim(polychoric_cor_2004),
    dim(lavaan_cor_2004)
  ),
  identical(
    dimnames(polychoric_cor_2004),
    dimnames(lavaan_cor_2004)
  ),
  implementation_comparison_2004$
    Maximum_absolute_difference < 1e-12,
  isTRUE(
    implementation_comparison_2004$
      Equal_within_tolerance
  )
)

"PASS: 2004 polychoric implementation comparison gate"
```
## Audit 2004 Estimation Warnings

```{r audit-2004-polychoric-warnings}
warning_audit_2004 <- data.frame(
  Source = c(
    "qgraph level",
    "direct lavaan"
  ),
  Warning_n = c(
    length(qgraph_warnings_2004),
    length(lavaan_warnings_2004)
  )
)

warning_audit_2004

qgraph_warnings_2004
lavaan_warnings_2004

if (any(warning_audit_2004$Warning_n > 0)) {
  stop(
    "FAIL: review the recorded 2004 estimation warnings"
  )
}

"PASS: 2004 polychoric warning gate"
```
## Validate 2004 Correlation Matrix Structure

```{r validate-2004-polychoric-matrix}
off_diagonal_2004 <- polychoric_cor_2004[
  lower.tri(polychoric_cor_2004)
]

matrix_structure_audit_2004 <- data.frame(
  Check = c(
    "Dimension is 9 by 9",
    "Node order is correct",
    "All values are finite",
    "Matrix is symmetric",
    "Diagonal equals one",
    "Values are within correlation range",
    "No perfect off-diagonal correlation"
  ),
  Passed = c(
    identical(
      dim(polychoric_cor_2004),
      c(9L, 9L)
    ),
    identical(
      rownames(polychoric_cor_2004),
      network_vars
    ) &&
      identical(
        colnames(polychoric_cor_2004),
        network_vars
      ),
    all(
      is.finite(polychoric_cor_2004)
    ),
    max(
      abs(
        polychoric_cor_2004 -
          t(polychoric_cor_2004)
      )
    ) < 1e-12,
    max(
      abs(
        diag(polychoric_cor_2004) - 1
      )
    ) < 1e-12,
    all(
      polychoric_cor_2004 >= -1 - 1e-12 &
        polychoric_cor_2004 <= 1 + 1e-12
    ),
    all(
      abs(off_diagonal_2004) < 1
    )
  )
)

matrix_numeric_diagnostics_2004 <- data.frame(
  Maximum_asymmetry = max(
    abs(
      polychoric_cor_2004 -
        t(polychoric_cor_2004)
    )
  ),
  Maximum_diagonal_deviation = max(
    abs(
      diag(polychoric_cor_2004) - 1
    )
  ),
  Minimum_off_diagonal = min(
    off_diagonal_2004
  ),
  Maximum_off_diagonal = max(
    off_diagonal_2004
  )
)

matrix_structure_audit_2004
matrix_numeric_diagnostics_2004

stopifnot(
  all(matrix_structure_audit_2004$Passed)
)

"PASS: 2004 polychoric matrix structure gate"
```
## Audit 2004 Eigenvalues and Positive Definiteness

```{r audit-2004-polychoric-eigenvalues}
eigenvalues_2004 <- eigen(
  polychoric_cor_2004,
  symmetric = TRUE,
  only.values = TRUE
)$values

eigenvalue_table_2004 <- data.frame(
  Eigenvalue_number = seq_along(
    eigenvalues_2004
  ),
  Eigenvalue = eigenvalues_2004
)

positive_definite_2004 <- all(
  eigenvalues_2004 > 0
)

positive_definiteness_audit_2004 <- data.frame(
  Minimum_eigenvalue = min(
    eigenvalues_2004
  ),
  Maximum_eigenvalue = max(
    eigenvalues_2004
  ),
  Condition_number = max(
    eigenvalues_2004
  ) / min(
    eigenvalues_2004
  ),
  Positive_definite = positive_definite_2004
)

eigenvalue_table_2004
positive_definiteness_audit_2004

if (!positive_definite_2004) {
  stop(
    "FAIL: 2004 polychoric matrix is not positive definite"
  )
}

"PASS: 2004 positive-definiteness gate"
```
## Prepare Ordered Data by Election Year

```{r prepare-ordered-data-by-year}
analysis_years <- as.numeric(
  names(expected_year_counts)
)

ordinal_data_by_year <- lapply(
  analysis_years,
  function(election_year) {
    as.data.frame(
      anes_ordinal[
        anes_ordinal$VCF0004 == election_year,
        network_vars
      ]
    )
  }
)

names(ordinal_data_by_year) <- names(
  expected_year_counts
)

year_input_audit <- data.frame(
  Election_year = analysis_years,
  Rows = vapply(
    ordinal_data_by_year,
    nrow,
    integer(1)
  ),
  Nodes = vapply(
    ordinal_data_by_year,
    ncol,
    integer(1)
  ),
  Complete = vapply(
    ordinal_data_by_year,
    function(year_data) {
      all(complete.cases(year_data))
    },
    logical(1)
  ),
  All_nodes_ordered = vapply(
    ordinal_data_by_year,
    function(year_data) {
      all(
        vapply(
          year_data,
          is.ordered,
          logical(1)
        )
      )
    },
    logical(1)
  )
)

year_input_audit

stopifnot(
  all(
    year_input_audit$Rows ==
      as.numeric(expected_year_counts)
  ),
  all(year_input_audit$Nodes == 9),
  all(year_input_audit$Complete),
  all(year_input_audit$All_nodes_ordered)
)

"PASS: all-year ordered input gate"
```
## Estimate Polychoric Matrices by Election Year

```{r estimate-polychoric-matrices-by-year}
polychoric_cor_by_year <- setNames(
  vector(
    "list",
    length(expected_year_counts)
  ),
  names(expected_year_counts)
)

polychoric_warnings_by_year <- setNames(
  vector(
    "list",
    length(expected_year_counts)
  ),
  names(expected_year_counts)
)

for (year_name in names(expected_year_counts)) {
  year_warnings <- character(0)

  year_matrix <- withCallingHandlers(
    lavaan::lavCor(
      object = ordinal_data_by_year[[year_name]],
      ordered = network_vars,
      missing = "listwise",
      estimator = "two.step",
      meanstructure = FALSE,
      cor_smooth = FALSE,
      output = "cor"
    ),
    warning = function(warning_condition) {
      year_warnings <<- c(
        year_warnings,
        conditionMessage(warning_condition)
      )
      invokeRestart("muffleWarning")
    }
  )

  polychoric_cor_by_year[[year_name]] <- unclass(
    year_matrix
  )

  polychoric_warnings_by_year[[year_name]] <- year_warnings
}

all_year_warning_audit <- data.frame(
  Election_year = analysis_years,
  Warning_n = vapply(
    polychoric_warnings_by_year,
    length,
    integer(1)
  )
)

pilot_reproduction_difference <- max(
  abs(
    polychoric_cor_by_year[["2004"]] -
      polychoric_cor_2004
  )
)

all_year_warning_audit
polychoric_warnings_by_year
pilot_reproduction_difference

if (any(all_year_warning_audit$Warning_n > 0)) {
  stop(
    "FAIL: review the recorded all-year polychoric warnings"
  )
}

stopifnot(
  pilot_reproduction_difference < 1e-12
)

"PASS: all-year polychoric estimation gate"
```
## Validate Polychoric Matrices by Election Year

```{r validate-polychoric-matrices-by-year}
polychoric_matrix_audit <- do.call(
  rbind,
  lapply(
    names(polychoric_cor_by_year),
    function(year_name) {
      year_matrix <- polychoric_cor_by_year[[year_name]]

      off_diagonal <- year_matrix[
        lower.tri(year_matrix)
      ]

      year_eigenvalues <- eigen(
        year_matrix,
        symmetric = TRUE,
        only.values = TRUE
      )$values

      data.frame(
        Election_year = as.numeric(year_name),
        Dimension_ok = identical(
          dim(year_matrix),
          c(9L, 9L)
        ),
        Node_order_ok =
          identical(
            rownames(year_matrix),
            network_vars
          ) &&
          identical(
            colnames(year_matrix),
            network_vars
          ),
        All_values_finite = all(
          is.finite(year_matrix)
        ),
        Maximum_asymmetry = max(
          abs(
            year_matrix -
              t(year_matrix)
          )
        ),
        Maximum_diagonal_deviation = max(
          abs(
            diag(year_matrix) - 1
          )
        ),
        Minimum_off_diagonal = min(
          off_diagonal
        ),
        Maximum_off_diagonal = max(
          off_diagonal
        ),
        Minimum_eigenvalue = min(
          year_eigenvalues
        ),
        Maximum_eigenvalue = max(
          year_eigenvalues
        ),
        Condition_number = max(
          year_eigenvalues
        ) / min(
          year_eigenvalues
        ),
        Positive_definite = all(
          year_eigenvalues > 0
        )
      )
    }
  )
)

rownames(polychoric_matrix_audit) <- NULL

polychoric_matrix_audit

stopifnot(
  all(polychoric_matrix_audit$Dimension_ok),
  all(polychoric_matrix_audit$Node_order_ok),
  all(polychoric_matrix_audit$All_values_finite),
  all(
    polychoric_matrix_audit$
      Maximum_asymmetry < 1e-12
  ),
  all(
    polychoric_matrix_audit$
      Maximum_diagonal_deviation < 1e-12
  ),
  all(
    polychoric_matrix_audit$
      Minimum_off_diagonal > -1
  ),
  all(
    polychoric_matrix_audit$
      Maximum_off_diagonal < 1
  ),
  all(
    polychoric_matrix_audit$
      Positive_definite
  )
)

"PASS: all-year polychoric matrix diagnostic gate"
```
## Build Polychoric Diagnostic Outputs

```{r build-polychoric-diagnostic-outputs}
stopifnot(
  identical(
    year_input_audit$Election_year,
    all_year_warning_audit$Election_year
  ),
  identical(
    year_input_audit$Election_year,
    polychoric_matrix_audit$Election_year
  )
)

polychoric_input_audit <- cbind(
  year_input_audit,
  Warning_n = all_year_warning_audit$Warning_n,
  polychoric_matrix_audit[
    names(polychoric_matrix_audit) !=
      "Election_year"
  ]
)

polychoric_diagnostic_bundle <- list(
  Analysis_ready_file = basename(
    analysis_ready_file
  ),
  Analysis_ready_file_SHA256 =
    analysis_ready_file_sha256,
  Analysis_ready_object_SHA256 =
    analysis_ready_object_sha256,
  Expected_analysis_ready_object_SHA256 =
    expected_analysis_ready_object_sha256,
  Node_direction_specification =
    node_direction_specification,
  Direction_input_validation =
    direction_input_validation,
  Election_years = analysis_years,
  Expected_year_counts =
    expected_year_counts,
  Network_vars = network_vars,
  Package_versions =
    diagnostic_package_versions,
  Warning_audit =
    all_year_warning_audit,
  Warnings =
    polychoric_warnings_by_year,
  Matrix_audit =
    polychoric_matrix_audit,
  Matrices =
    polychoric_cor_by_year,
  R_version = R.version.string
)

polychoric_input_audit

names(polychoric_diagnostic_bundle)
```
## Save Polychoric Diagnostic Outputs

```{r save-polychoric-diagnostic-outputs}
audit_dir <- file.path(
  project_root,
  "data",
  "audit"
)

dir.create(
  audit_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

polychoric_bundle_file <- file.path(
  audit_dir,
  "anes_cdf_20260205_nine_node_complete_case_polychoric_diagnostics_2004_2012_2016_2020_2024.rds"
)

polychoric_audit_file <- file.path(
  audit_dir,
  "anes_cdf_20260205_nine_node_complete_case_polychoric_diagnostics_2004_2012_2016_2020_2024.csv"
)

saveRDS(
  polychoric_diagnostic_bundle,
  polychoric_bundle_file
)

write.csv(
  polychoric_input_audit,
  polychoric_audit_file,
  row.names = FALSE
)

normalizePath(
  polychoric_bundle_file,
  winslash = "/"
)

normalizePath(
  polychoric_audit_file,
  winslash = "/"
)

file.exists(polychoric_bundle_file)
file.exists(polychoric_audit_file)
```
## Validate Saved Polychoric Diagnostic Outputs

```{r validate-saved-polychoric-outputs}
polychoric_bundle_check <- readRDS(
  polychoric_bundle_file
)

polychoric_audit_check <- read.csv(
  polychoric_audit_file,
  check.names = FALSE
)

bundle_identical <- identical(
  polychoric_diagnostic_bundle,
  polychoric_bundle_check
)

audit_equal <- isTRUE(
  all.equal(
    polychoric_input_audit,
    polychoric_audit_check,
    check.attributes = FALSE,
    tolerance = 1e-12
  )
)

saved_output_validation <- data.frame(
  Check = c(
    "RDS bundle is identical",
    "CSV audit is equal",
    "Corrected input object SHA-256 is preserved",
    "Node-direction specification is preserved",
    "Direction validation gate is preserved",
    "Five matrices are present",
    "Node order is preserved",
    "No estimation warnings",
    "All matrices are positive definite"
  ),

  Passed = c(
    bundle_identical,
    audit_equal,

    identical(
      polychoric_bundle_check$
        Analysis_ready_object_SHA256,
      expected_analysis_ready_object_sha256
    ) &&
      identical(
        polychoric_bundle_check$
          Expected_analysis_ready_object_SHA256,
        expected_analysis_ready_object_sha256
      ),

    identical(
      polychoric_bundle_check$
        Node_direction_specification,
      node_direction_specification
    ),

    identical(
      polychoric_bundle_check$
        Direction_input_validation,
      direction_input_validation
    ) &&
      all(
        polychoric_bundle_check$
          Direction_input_validation$Passed
      ),

    identical(
      names(
        polychoric_bundle_check$Matrices
      ),
      names(expected_year_counts)
    ),

    all(
      vapply(
        polychoric_bundle_check$Matrices,
        function(year_matrix) {
          identical(
            colnames(year_matrix),
            network_vars
          )
        },
        logical(1)
      )
    ),

    all(
      polychoric_bundle_check$
        Warning_audit$Warning_n == 0
    ),

    all(
      polychoric_bundle_check$
        Matrix_audit$Positive_definite
    )
  )
)

saved_output_validation

stopifnot(
  all(saved_output_validation$Passed)
)

"PASS: 06 polychoric diagnostic output gate"
```
## Formal Pre-Model Validation Gate

```{r formal-pre-model-validation-gate}
formal_pre_model_gate <- data.frame(
  Check = c(
    "Analysis-ready input exists",
    "Corrected node-direction input is validated",
    "Approved years and sample sizes match",
    "Nine nodes are complete",
    "Observed categories match specification",
    "Ordered-factor conversion is valid",
    "All-year estimation is warning-free",
    "All matrices pass mathematical checks",
    "Diagnostic output files exist",
    "Saved outputs passed round-trip validation"
  ),

  Passed = c(
    file.exists(analysis_ready_file),

    all(
      direction_input_validation$Passed
    ),

    identical(
      names(year_counts),
      names(expected_year_counts)
    ) &&
      all(
        as.numeric(year_counts) ==
          unname(expected_year_counts)
      ),

    length(network_vars) == 9 &&
      all(
        complete.cases(
          anes_analysis_ready[network_vars]
        )
      ),

    all(category_match),

    all(
      ordered_validation$Is_ordered &
        ordered_validation$Levels_match &
        ordered_validation$Missing_n == 0
    ),

    all(
      all_year_warning_audit$Warning_n == 0
    ),

    all(
      polychoric_matrix_audit$Dimension_ok &
        polychoric_matrix_audit$Node_order_ok &
        polychoric_matrix_audit$All_values_finite &
        polychoric_matrix_audit$
          Maximum_asymmetry < 1e-12 &
        polychoric_matrix_audit$
          Maximum_diagonal_deviation < 1e-12 &
        polychoric_matrix_audit$
          Positive_definite
    ),

    file.exists(polychoric_bundle_file) &&
      file.exists(polychoric_audit_file),

    all(saved_output_validation$Passed)
  )
)

formal_pre_model_gate

stopifnot(
  all(formal_pre_model_gate$Passed)
)

"PASS: formal pre-model polychoric input gate"
```
