---
title: "06B D15 Survey-Weight Pilot - Stages 0-1 Source and Descriptive Audit"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

## Purpose and boundary

This notebook performs Stages 0 and 1 of the authorised D15 pilot. Stage 0 verifies the official CDF input, the frozen five-wave nine-node complete-case object, weight availability and exact case-level matching. Stage 1 implements the two prospectively approved candidates, `W_CDF` and `W_POST_COMPAT`, and reports only weight-distribution and weighted-versus-unweighted category-proportion diagnostics. It does not estimate weighted correlations or networks and does not run bootstrap, centrality, NCT, education-group analysis, formal five-wave results or political interpretation.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  error = FALSE,
  warning = TRUE,
  message = TRUE
)
```

## Project root

```{r project-root}
current_dir <- normalizePath(
  ".",
  winslash = "/"
)

if (
  file.exists(
    file.path(
      current_dir,
      "ANES.Rproj"
    )
  )
) {
  project_root <- current_dir
} else if (
  basename(current_dir) == "notebooks" &&
    file.exists(
      file.path(
        dirname(current_dir),
        "ANES.Rproj"
      )
    )
) {
  project_root <- dirname(current_dir)
} else {
  stop(
    "Project root could not be identified. Open the ANES_dissertation project before rendering."
  )
}

knitr::opts_knit$set(
  root.dir = project_root
)

project_root
```

## Package gate

Stages 0 and 1 use only packages already locked by the existing project. They do not install or update packages.

```{r package-gate}
required_packages <- c(
  "digest",
  "haven",
  "tidyselect"
)

package_available <- vapply(
  required_packages,
  requireNamespace,
  quietly = TRUE,
  FUN.VALUE = logical(1)
)

package_gate <- data.frame(
  Package = required_packages,
  Available = unname(package_available),
  Version = vapply(
    required_packages,
    function(package_name) {
      if (requireNamespace(package_name, quietly = TRUE)) {
        as.character(
          utils::packageVersion(package_name)
        )
      } else {
        NA_character_
      }
    },
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

package_gate

stopifnot(
  all(package_gate$Available)
)
```

## D15 authorisation gate

The notebook reads the current decision registry. It does not accept a manually edited switch inside this file.

```{r d15-authorisation-gate}
registry_config <- file.path(
  project_root,
  "config",
  "decision_registry_path.txt"
)

stopifnot(
  file.exists(registry_config)
)

registry_lines <- trimws(
  readLines(
    registry_config,
    warn = FALSE
  )
)

registry_lines <- registry_lines[
  nzchar(registry_lines)
]

stopifnot(
  length(registry_lines) == 1L,
  file.exists(registry_lines)
)

registry_path <- normalizePath(
  registry_lines,
  winslash = "/",
  mustWork = TRUE
)

decision_registry <- utils::read.csv(
  registry_path,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

required_registry_columns <- c(
  "decision_id",
  "decision_status",
  "implementation_status",
  "researcher_approval_status",
  "analysis_effect",
  "last_verified"
)

stopifnot(
  all(
    required_registry_columns %in%
      names(decision_registry)
  )
)

normalise_status <- function(value) {
  value <- toupper(
    trimws(
      as.character(value)
    )
  )
  gsub(
    "[ -]+",
    "_",
    value
  )
}

d15 <- decision_registry[
  decision_registry$decision_id == "D15",
  ,
  drop = FALSE
]

d15_authorised <-
  nrow(d15) == 1L &&
  normalise_status(d15$decision_status) ==
    "PILOT_AUTHORISED" &&
  normalise_status(d15$implementation_status) ==
    "STAGE0_COMPLETE_STAGE1_AUTHORISED" &&
  normalise_status(d15$researcher_approval_status) ==
    "APPROVED_FOR_BOUNDED_PILOT_ONLY" &&
  normalise_status(d15$analysis_effect) ==
    "BLOCKS_CORE_GGM"

d15_gate <- data.frame(
  Registry = registry_path,
  D15_rows = nrow(d15),
  Decision_status = if (nrow(d15) == 1L) {
    d15$decision_status
  } else {
    NA_character_
  },
  Implementation_status = if (nrow(d15) == 1L) {
    d15$implementation_status
  } else {
    NA_character_
  },
  Researcher_approval = if (nrow(d15) == 1L) {
    d15$researcher_approval_status
  } else {
    NA_character_
  },
  Core_GGM_still_blocked = if (nrow(d15) == 1L) {
    normalise_status(d15$analysis_effect) ==
      "BLOCKS_CORE_GGM"
  } else {
    NA
  },
  D15_pilot_authorised = d15_authorised,
  stringsAsFactors = FALSE
)

d15_gate

stopifnot(
  d15_authorised
)
```

## Input paths and immutable-input gates

```{r input-paths-and-hashes}
cdf_path_config <- file.path(
  project_root,
  "config",
  "anes_cdf_path.txt"
)

analysis_ready_file <- file.path(
  project_root,
  "data",
  "analysis_ready",
  paste0(
    "anes_cdf_20260205_analysis_ready_nine_node_",
    "complete_case_2004_2012_2016_2020_2024.rds"
  )
)

stopifnot(
  file.exists(cdf_path_config),
  file.exists(analysis_ready_file)
)

cdf_path_lines <- trimws(
  readLines(
    cdf_path_config,
    warn = FALSE
  )
)

cdf_path_lines <- cdf_path_lines[
  nzchar(cdf_path_lines)
]

stopifnot(
  length(cdf_path_lines) == 1L,
  file.exists(cdf_path_lines)
)

cdf_file <- normalizePath(
  cdf_path_lines,
  winslash = "/",
  mustWork = TRUE
)

expected_cdf_sha256 <-
  "45323c30faee1e9e67c2351c4302c31472b117050bef240d9339a264e2f3e078"

expected_analysis_ready_object_sha256 <-
  "78e5c1cff278a6640834113b769ef4251807ab30659b6effcfd6a14db58ada6c"

cdf_file_sha256 <- digest::digest(
  file = cdf_file,
  algo = "sha256",
  serialize = FALSE
)

anes_analysis_ready <- readRDS(
  analysis_ready_file
)

analysis_ready_file_sha256 <- digest::digest(
  file = analysis_ready_file,
  algo = "sha256",
  serialize = FALSE
)

analysis_ready_object_sha256 <- digest::digest(
  anes_analysis_ready,
  algo = "sha256",
  serialize = TRUE
)

input_identity_gate <- data.frame(
  Input = c(
    "Official CDF Stata file",
    "Frozen analysis-ready RDS file",
    "Frozen analysis-ready R object"
  ),
  Observed_SHA256 = c(
    cdf_file_sha256,
    analysis_ready_file_sha256,
    analysis_ready_object_sha256
  ),
  Expected_SHA256 = c(
    expected_cdf_sha256,
    NA_character_,
    expected_analysis_ready_object_sha256
  ),
  Passed = c(
    identical(
      cdf_file_sha256,
      expected_cdf_sha256
    ),
    file.info(analysis_ready_file)$size > 0,
    identical(
      analysis_ready_object_sha256,
      expected_analysis_ready_object_sha256
    )
  ),
  stringsAsFactors = FALSE
)

input_identity_gate

stopifnot(
  all(input_identity_gate$Passed)
)
```

## Frozen sample and node specification

```{r frozen-sample-gate}
expected_years <- c(
  2004L,
  2012L,
  2016L,
  2020L,
  2024L
)

expected_year_counts <- c(
  "2004" = 789L,
  "2012" = 4353L,
  "2016" = 2709L,
  "2020" = 5390L,
  "2024" = 3525L
)

network_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223"
)

required_analysis_ready_vars <- c(
  "VCF0004",
  "VCF0006",
  "VCF0006a",
  "VCF0009z",
  network_vars
)

observed_year_counts <- table(
  anes_analysis_ready$VCF0004
)

sample_gate <- data.frame(
  Check = c(
    "Analysis-ready object is a data frame",
    "Required identifiers, weight and nodes are present",
    "Dimensions are 16766 by 18",
    "Five wave counts match the frozen sample",
    "All nine nodes are complete",
    "VCF9999 is not silently assumed to be in the RDS"
  ),
  Passed = c(
    is.data.frame(anes_analysis_ready),
    all(
      required_analysis_ready_vars %in%
        names(anes_analysis_ready)
    ),
    identical(
      dim(anes_analysis_ready),
      c(16766L, 18L)
    ),
    identical(
      names(observed_year_counts),
      names(expected_year_counts)
    ) &&
      identical(
        as.integer(observed_year_counts),
        unname(expected_year_counts)
      ),
    all(
      stats::complete.cases(
        anes_analysis_ready[
          network_vars
        ]
      )
    ),
    !"VCF9999" %in%
      names(anes_analysis_ready)
  ),
  stringsAsFactors = FALSE
)

sample_gate

stopifnot(
  all(sample_gate$Passed)
)
```

## Read only the required weight fields from the official CDF

The original CDF is read but never modified. `VCF9999` is joined in memory only for this pilot.

```{r read-official-weight-fields}
raw_weight_vars <- c(
  "VCF0004",
  "VCF0006",
  "VCF0006a",
  "VCF0009z",
  "VCF9999"
)

anes_weight_source <- haven::read_dta(
  cdf_file,
  col_select = tidyselect::all_of(
    raw_weight_vars
  )
)

stopifnot(
  identical(
    names(anes_weight_source),
    raw_weight_vars
  )
)

raw_variable_metadata <- data.frame(
  Variable = raw_weight_vars,
  Stata_label = vapply(
    raw_weight_vars,
    function(variable_name) {
      variable_label <- attr(
        anes_weight_source[[variable_name]],
        "label"
      )
      if (is.null(variable_label)) {
        NA_character_
      } else {
        as.character(variable_label)
      }
    },
    FUN.VALUE = character(1)
  ),
  Class = vapply(
    raw_weight_vars,
    function(variable_name) {
      paste(
        class(
          anes_weight_source[[variable_name]]
        ),
        collapse = ";"
      )
    },
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

raw_variable_metadata
```

## Exact complete-case membership match

```{r exact-case-match}
source_in_scope <- anes_weight_source[
  anes_weight_source$VCF0004 %in%
    expected_years,
  ,
  drop = FALSE
]

source_key <- as.numeric(
  source_in_scope$VCF0006a
)

analysis_key <- as.numeric(
  anes_analysis_ready$VCF0006a
)

source_duplicate_keys <- sum(
  duplicated(source_key)
)

analysis_duplicate_keys <- sum(
  duplicated(analysis_key)
)

match_index <- match(
  analysis_key,
  source_key
)

unmatched_analysis_cases <- sum(
  is.na(match_index)
)

match_gate <- data.frame(
  Check = c(
    "No duplicate VCF0006a cross-year IDs in official source",
    "No duplicate VCF0006a cross-year IDs in analysis-ready data",
    "Every analysis-ready case matches the official source"
  ),
  Observed = c(
    source_duplicate_keys,
    analysis_duplicate_keys,
    unmatched_analysis_cases
  ),
  Expected = c(
    0L,
    0L,
    0L
  ),
  Passed = c(
    source_duplicate_keys == 0L,
    analysis_duplicate_keys == 0L,
    unmatched_analysis_cases == 0L
  ),
  stringsAsFactors = FALSE
)

match_gate

stopifnot(
  all(match_gate$Passed)
)

pilot_weight_data <- anes_analysis_ready

pilot_weight_data$VCF0009z_source <- as.numeric(
  source_in_scope$VCF0009z[
    match_index
  ]
)

pilot_weight_data$VCF9999_source <- as.numeric(
  source_in_scope$VCF9999[
    match_index
  ]
)

vcf0009z_absolute_difference <- abs(
  as.numeric(
    pilot_weight_data$VCF0009z
  ) -
    pilot_weight_data$VCF0009z_source
)

vcf0009z_match_gate <- data.frame(
  Matched_cases = nrow(pilot_weight_data),
  Exact_matches = sum(
    vcf0009z_absolute_difference == 0,
    na.rm = TRUE
  ),
  Missing_comparisons = sum(
    is.na(vcf0009z_absolute_difference)
  ),
  Maximum_absolute_difference = max(
    vcf0009z_absolute_difference,
    na.rm = TRUE
  ),
  Passed = all(
    !is.na(vcf0009z_absolute_difference) &
      vcf0009z_absolute_difference == 0
  ),
  stringsAsFactors = FALSE
)

vcf0009z_match_gate

stopifnot(
  vcf0009z_match_gate$Passed
)
```

## Official source map and observed availability

The source-variable map below is transcribed from the official February 2026 CDF variable codebook. `NO CDF SOURCE LISTED` is retained explicitly for 2012 `VCF9999`.

```{r source-map-and-availability}
source_map <- data.frame(
  Election_year = rep(
    expected_years,
    times = 2L
  ),
  Weight_variable = rep(
    c(
      "VCF0009z",
      "VCF9999"
    ),
    each = length(expected_years)
  ),
  Official_role = c(
    rep(
      "CDF full-sample weight",
      length(expected_years)
    ),
    rep(
      "CDF post-election full-sample weight",
      length(expected_years)
    )
  ),
  Official_source_variable = c(
    "V040101",
    "weight_full",
    "V160101",
    "V200010a",
    "V240107a",
    "V040102",
    "NO CDF SOURCE LISTED",
    "V160102",
    "V200010b",
    "V240107b"
  ),
  Source_record_status = c(
    rep(
      "VERIFIED IN OFFICIAL CDF CODEBOOK",
      length(expected_years)
    ),
    "VERIFIED IN OFFICIAL CDF CODEBOOK",
    "NO CDF SOURCE LISTED",
    rep(
      "VERIFIED IN OFFICIAL CDF CODEBOOK",
      3L
    )
  ),
  Official_document = rep(
    paste0(
      "ANES Time Series Cumulative Data File ",
      "Variable Codebook, 2026-02-05"
    ),
    length(expected_years) * 2L
  ),
  Official_url = rep(
    paste0(
      "https://electionstudies.org/wp-content/uploads/",
      "2026/02/anes_timeseries_cdf_codebook_var_20260205.pdf"
    ),
    length(expected_years) * 2L
  ),
  stringsAsFactors = FALSE
)

summarise_weight_availability <- function(
  data,
  variable_name
) {
  do.call(
    rbind,
    lapply(
      expected_years,
      function(year_value) {
        year_rows <- as.integer(
          data$VCF0004
        ) == year_value

        weight_values <- as.numeric(
          data[[variable_name]][
            year_rows
          ]
        )

        nonmissing_values <- weight_values[
          !is.na(weight_values)
        ]

        finite_values <- nonmissing_values[
          is.finite(nonmissing_values)
        ]

        data.frame(
          Election_year = year_value,
          Weight_variable = sub(
            "_source$",
            "",
            variable_name
          ),
          Analysis_ready_N = length(weight_values),
          Missing_N = sum(
            is.na(weight_values)
          ),
          Nonfinite_nonmissing_N = sum(
            !is.finite(nonmissing_values)
          ),
          Nonpositive_finite_N = sum(
            finite_values <= 0
          ),
          Positive_finite_N = sum(
            finite_values > 0
          ),
          Minimum_positive = if (
            any(finite_values > 0)
          ) {
            min(
              finite_values[
                finite_values > 0
              ]
            )
          } else {
            NA_real_
          },
          Maximum_positive = if (
            any(finite_values > 0)
          ) {
            max(
              finite_values[
                finite_values > 0
              ]
            )
          } else {
            NA_real_
          },
          All_complete_cases_have_positive_weight =
            length(weight_values) > 0L &&
            all(
              !is.na(weight_values) &
                is.finite(weight_values) &
                weight_values > 0
            ),
          stringsAsFactors = FALSE
        )
      }
    )
  )
}

observed_availability <- rbind(
  summarise_weight_availability(
    pilot_weight_data,
    "VCF0009z_source"
  ),
  summarise_weight_availability(
    pilot_weight_data,
    "VCF9999_source"
  )
)

source_map_key <- paste(
  source_map$Election_year,
  source_map$Weight_variable,
  sep = "::"
)

availability_key <- paste(
  observed_availability$Election_year,
  observed_availability$Weight_variable,
  sep = "::"
)

source_match <- match(
  source_map_key,
  availability_key
)

stopifnot(
  !anyNA(source_match)
)

d15_weight_source_audit <- cbind(
  source_map,
  observed_availability[
    source_match,
    setdiff(
      names(observed_availability),
      c(
        "Election_year",
        "Weight_variable"
      )
    ),
    drop = FALSE
  ]
)

d15_weight_source_audit$Present_in_analysis_ready_RDS <-
  d15_weight_source_audit$Weight_variable %in%
  names(anes_analysis_ready)

d15_weight_source_audit$Stage_1_status <- ifelse(
  d15_weight_source_audit$All_complete_cases_have_positive_weight,
  "TECHNICALLY AVAILABLE FOR COMPLETE-CASE PILOT",
  ifelse(
    d15_weight_source_audit$Weight_variable == "VCF9999" &
      d15_weight_source_audit$Election_year == 2012L,
    "2012 FALLBACK DECISION REQUIRED BEFORE STAGE 1",
    "NOT TECHNICALLY AVAILABLE"
  )
)

d15_weight_source_audit
```

## Stage 0 result and hard boundary

```{r stage0-result}
stage0_result <- data.frame(
  Check = c(
    "Official CDF identity confirmed",
    "Frozen analysis-ready object confirmed",
    "All 16766 complete cases matched to the official CDF",
    "Stored VCF0009z reproduced exactly from the official CDF",
    "VCF0009z positive for all complete cases in all five waves",
    "VCF9999 positive for all complete cases in 2004 2016 2020 2024",
    "VCF9999 available for 2012 complete cases",
    "The 2012 W_POST_COMPAT substitute rule was registered before Stage 1"
  ),
  Passed = c(
    identical(
      cdf_file_sha256,
      expected_cdf_sha256
    ),
    identical(
      analysis_ready_object_sha256,
      expected_analysis_ready_object_sha256
    ),
    unmatched_analysis_cases == 0L,
    vcf0009z_match_gate$Passed,
    all(
      d15_weight_source_audit$All_complete_cases_have_positive_weight[
        d15_weight_source_audit$Weight_variable == "VCF0009z"
      ]
    ),
    all(
      d15_weight_source_audit$All_complete_cases_have_positive_weight[
        d15_weight_source_audit$Weight_variable == "VCF9999" &
          d15_weight_source_audit$Election_year %in%
            c(2004L, 2016L, 2020L, 2024L)
      ]
    ),
    d15_weight_source_audit$All_complete_cases_have_positive_weight[
      d15_weight_source_audit$Weight_variable == "VCF9999" &
        d15_weight_source_audit$Election_year == 2012L
    ],
    TRUE
  ),
  Consequence = c(
    "Continue",
    "Continue",
    "Continue",
    "Continue",
    "VCF0009z remains a valid five-wave pilot candidate",
    "VCF9999 is available for four waves",
    paste0(
      "Use only the registered wave-specific W_POST_COMPAT rule; ",
      "do not claim VCF9999 exists in 2012"
    ),
    "Stage 1 may proceed within its descriptive-only boundary"
  ),
  stringsAsFactors = FALSE
)

stage0_result
```

## Save Stage 0 outputs

```{r save-stage0-outputs}
pilot_table_dir <- file.path(
  project_root,
  "outputs",
  "tables",
  "pilot"
)

pilot_log_dir <- file.path(
  project_root,
  "outputs",
  "logs"
)

dir.create(
  pilot_table_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

dir.create(
  pilot_log_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

source_audit_file <- file.path(
  pilot_table_dir,
  "d15_weight_source_audit.csv"
)

validation_file <- file.path(
  pilot_table_dir,
  "d15_stage0_validation.csv"
)

metadata_file <- file.path(
  pilot_table_dir,
  "d15_raw_weight_variable_metadata.csv"
)

session_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage0_sessionInfo.txt"
)

log_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage0_source_audit.log"
)

utils::write.csv(
  d15_weight_source_audit,
  source_audit_file,
  row.names = FALSE,
  na = "NA"
)

utils::write.csv(
  stage0_result,
  validation_file,
  row.names = FALSE,
  na = "NA"
)

utils::write.csv(
  raw_variable_metadata,
  metadata_file,
  row.names = FALSE,
  na = "NA"
)

writeLines(
  capture.output(
    utils::sessionInfo()
  ),
  session_file
)

stage0_log <- c(
  "D15 SURVEY-WEIGHT PILOT - STAGE 0 SOURCE AUDIT",
  paste0(
    "Timestamp: ",
    format(
      Sys.time(),
      "%Y-%m-%d %H:%M:%S %z"
    )
  ),
  paste0(
    "Decision registry: ",
    registry_path
  ),
  paste0(
    "Official CDF: ",
    cdf_file
  ),
  paste0(
    "CDF SHA-256: ",
    cdf_file_sha256
  ),
  paste0(
    "Analysis-ready file: ",
    normalizePath(
      analysis_ready_file,
      winslash = "/"
    )
  ),
  paste0(
    "Analysis-ready file SHA-256: ",
    analysis_ready_file_sha256
  ),
  paste0(
    "Analysis-ready object SHA-256: ",
    analysis_ready_object_sha256
  ),
  paste0(
    "Matched complete cases: ",
    nrow(pilot_weight_data)
  ),
  "No weighted correlation or network was estimated.",
  paste0(
    "Stage 1 candidate rule registered: W_CDF and W_POST_COMPAT; ",
    "D15 remains not frozen."
  ),
  paste0(
    "Source audit output: ",
    normalizePath(
      source_audit_file,
      winslash = "/",
      mustWork = FALSE
    )
  ),
  paste0(
    "Validation output: ",
    normalizePath(
      validation_file,
      winslash = "/",
      mustWork = FALSE
    )
  )
)

writeLines(
  stage0_log,
  log_file
)

data.frame(
  Output = c(
    source_audit_file,
    validation_file,
    metadata_file,
    session_file,
    log_file
  ),
  Exists = file.exists(
    c(
      source_audit_file,
      validation_file,
      metadata_file,
      session_file,
      log_file
    )
  ),
  stringsAsFactors = FALSE
)
```

## Stage 1 approval and candidate construction

Stage 0 has been independently reviewed. The researcher prospectively approved two candidates on 2026-08-29:

- `W_CDF`: `VCF0009z` in every wave;
- `W_POST_COMPAT`: `VCF9999` in 2004, 2016, 2020 and 2024, with official `VCF0009z` (`weight_full`) as an explicitly labelled wave-specific substitute in 2012.

This approval does not freeze D15 and does not permit choosing a candidate from political or network results.

```{r stage1-candidate-construction}
stage0_required_checks <- c(
  "Official CDF identity confirmed",
  "Frozen analysis-ready object confirmed",
  "All 16766 complete cases matched to the official CDF",
  "Stored VCF0009z reproduced exactly from the official CDF",
  "VCF0009z positive for all complete cases in all five waves",
  "VCF9999 positive for all complete cases in 2004 2016 2020 2024"
)

stopifnot(
  all(
    stage0_result$Passed[
      match(
        stage0_required_checks,
        stage0_result$Check
      )
    ]
  )
)

stage1_data <- pilot_weight_data

stage1_data$W_CDF <- as.numeric(
  stage1_data$VCF0009z_source
)

stage1_data$W_POST_COMPAT <- ifelse(
  as.integer(stage1_data$VCF0004) == 2012L,
  as.numeric(stage1_data$VCF0009z_source),
  as.numeric(stage1_data$VCF9999_source)
)

candidate_map <- data.frame(
  Candidate = rep(
    c("W_CDF", "W_POST_COMPAT"),
    each = length(expected_years)
  ),
  Election_year = rep(
    expected_years,
    times = 2L
  ),
  CDF_weight_variable = c(
    rep("VCF0009z", length(expected_years)),
    "VCF9999",
    "VCF0009z",
    "VCF9999",
    "VCF9999",
    "VCF9999"
  ),
  Official_source_variable = c(
    "V040101",
    "weight_full",
    "V160101",
    "V200010a",
    "V240107a",
    "V040102",
    "weight_full",
    "V160102",
    "V200010b",
    "V240107b"
  ),
  Candidate_role = c(
    rep(
      "Uniform five-wave CDF full-sample-weight candidate",
      length(expected_years)
    ),
    "Post-election full-sample weight",
    paste0(
      "Wave-specific official 2012 substitute: ",
      "CDF full-sample weight"
    ),
    rep(
      "Post-election full-sample weight",
      3L
    )
  ),
  Rescaling_applied = FALSE,
  stringsAsFactors = FALSE
)

candidate_map

candidate_values_are_valid <- vapply(
  c("W_CDF", "W_POST_COMPAT"),
  function(candidate_name) {
    candidate_values <- stage1_data[[candidate_name]]
    all(
      !is.na(candidate_values) &
        is.finite(candidate_values) &
        candidate_values > 0
    )
  },
  FUN.VALUE = logical(1)
)

stopifnot(
  all(candidate_values_are_valid)
)
```

## Stage 1 weight-distribution audit

Kish effective sample size is used only as a diagnostic of weight dispersion. It is not substituted for model `N` in this stage.

```{r stage1-weight-distribution-audit}
safe_quantile <- function(values, probability) {
  as.numeric(
    stats::quantile(
      values,
      probs = probability,
      names = FALSE,
      type = 7
    )
  )
}

summarise_weight_distribution <- function(
  data,
  candidate_name,
  year_value
) {
  year_rows <- as.integer(data$VCF0004) == year_value
  weights <- as.numeric(
    data[[candidate_name]][year_rows]
  )

  valid_weights <- weights[
    !is.na(weights) &
      is.finite(weights) &
      weights > 0
  ]

  kish_ess <- if (length(valid_weights) > 0L) {
    sum(valid_weights)^2 / sum(valid_weights^2)
  } else {
    NA_real_
  }

  data.frame(
    Candidate = candidate_name,
    Election_year = year_value,
    Complete_case_N = length(weights),
    Missing_N = sum(is.na(weights)),
    Nonfinite_nonmissing_N = sum(
      !is.na(weights) & !is.finite(weights)
    ),
    Nonpositive_finite_N = sum(
      !is.na(weights) &
        is.finite(weights) &
        weights <= 0
    ),
    Valid_positive_N = length(valid_weights),
    Minimum = min(valid_weights),
    P01 = safe_quantile(valid_weights, 0.01),
    P05 = safe_quantile(valid_weights, 0.05),
    Median = stats::median(valid_weights),
    Mean = mean(valid_weights),
    P95 = safe_quantile(valid_weights, 0.95),
    P99 = safe_quantile(valid_weights, 0.99),
    Maximum = max(valid_weights),
    Standard_deviation = stats::sd(valid_weights),
    Coefficient_of_variation =
      stats::sd(valid_weights) / mean(valid_weights),
    Sum_of_weights = sum(valid_weights),
    Kish_effective_N = kish_ess,
    Kish_ESS_over_N = kish_ess / length(weights),
    Maximum_over_median =
      max(valid_weights) / stats::median(valid_weights),
    Rescaling_applied = FALSE,
    stringsAsFactors = FALSE
  )
}

d15_weight_distribution_audit <- do.call(
  rbind,
  lapply(
    c("W_CDF", "W_POST_COMPAT"),
    function(candidate_name) {
      do.call(
        rbind,
        lapply(
          expected_years,
          function(year_value) {
            summarise_weight_distribution(
              stage1_data,
              candidate_name,
              year_value
            )
          }
        )
      )
    }
  )
)

distribution_map_key <- paste(
  d15_weight_distribution_audit$Candidate,
  d15_weight_distribution_audit$Election_year,
  sep = "::"
)

candidate_map_key <- paste(
  candidate_map$Candidate,
  candidate_map$Election_year,
  sep = "::"
)

distribution_map_match <- match(
  distribution_map_key,
  candidate_map_key
)

stopifnot(
  !anyNA(distribution_map_match)
)

d15_weight_distribution_audit <- cbind(
  d15_weight_distribution_audit,
  candidate_map[
    distribution_map_match,
    c(
      "CDF_weight_variable",
      "Official_source_variable",
      "Candidate_role"
    ),
    drop = FALSE
  ]
)

d15_weight_distribution_audit
```

## Stage 1 weighted and unweighted category proportions

The calculation retains the observed ordered categories. It does not collapse, relabel or recode a node.

```{r stage1-category-proportions}
ordered_category_labels <- function(values) {
  if (is.factor(values)) {
    as.character(levels(values))
  } else {
    unique_values <- unique(values)
    unique_values <- unique_values[!is.na(unique_values)]
    as.character(sort(unique_values))
  }
}

summarise_node_proportions <- function(
  data,
  candidate_name,
  year_value,
  node_name
) {
  year_rows <- as.integer(data$VCF0004) == year_value
  node_values <- data[[node_name]][year_rows]
  weights <- as.numeric(
    data[[candidate_name]][year_rows]
  )

  category_labels <- ordered_category_labels(node_values)

  do.call(
    rbind,
    lapply(
      seq_along(category_labels),
      function(category_index) {
        category_label <- category_labels[[category_index]]
        in_category <- as.character(node_values) == category_label

        unweighted_proportion <- mean(in_category)
        weighted_proportion <-
          sum(weights[in_category]) / sum(weights)

        data.frame(
          Candidate = candidate_name,
          Election_year = year_value,
          Node = node_name,
          Category_order = category_index,
          Category_value = category_label,
          Complete_case_N = length(node_values),
          Category_N = sum(in_category),
          Unweighted_proportion = unweighted_proportion,
          Weighted_proportion = weighted_proportion,
          Difference_percentage_points =
            100 * (
              weighted_proportion -
                unweighted_proportion
            ),
          Absolute_difference_percentage_points =
            100 * abs(
              weighted_proportion -
                unweighted_proportion
            ),
          Weight_sum = sum(weights),
          Category_weight_sum = sum(weights[in_category]),
          stringsAsFactors = FALSE
        )
      }
    )
  )
}

d15_weighted_category_proportions <- do.call(
  rbind,
  lapply(
    c("W_CDF", "W_POST_COMPAT"),
    function(candidate_name) {
      do.call(
        rbind,
        lapply(
          expected_years,
          function(year_value) {
            do.call(
              rbind,
              lapply(
                network_vars,
                function(node_name) {
                  summarise_node_proportions(
                    stage1_data,
                    candidate_name,
                    year_value,
                    node_name
                  )
                }
              )
            )
          }
        )
      )
    }
  )
)

d15_category_shift_summary <- stats::aggregate(
  Absolute_difference_percentage_points ~
    Candidate + Election_year + Node,
  data = d15_weighted_category_proportions,
  FUN = max
)

names(d15_category_shift_summary)[
  names(d15_category_shift_summary) ==
    "Absolute_difference_percentage_points"
] <- "Maximum_absolute_difference_percentage_points"

d15_weighted_category_proportions
d15_category_shift_summary
```

## Stage 1 validation

```{r stage1-validation}
proportion_sum_check <- stats::aggregate(
  cbind(
    Unweighted_proportion,
    Weighted_proportion
  ) ~ Candidate + Election_year + Node,
  data = d15_weighted_category_proportions,
  FUN = sum
)

candidate_unweighted_check <- stats::reshape(
  d15_weighted_category_proportions[
    ,
    c(
      "Candidate",
      "Election_year",
      "Node",
      "Category_value",
      "Unweighted_proportion"
    )
  ],
  idvar = c(
    "Election_year",
    "Node",
    "Category_value"
  ),
  timevar = "Candidate",
  direction = "wide"
)

unweighted_difference <- abs(
  candidate_unweighted_check$Unweighted_proportion.W_CDF -
    candidate_unweighted_check$Unweighted_proportion.W_POST_COMPAT
)

stage1_validation <- data.frame(
  Check = c(
    "Both approved candidates contain only finite positive weights",
    "Ten wave-by-candidate weight summaries were produced",
    "All unweighted node-category proportions sum to one",
    "All weighted node-category proportions sum to one",
    "Unweighted proportions agree exactly across candidates",
    "No node value was missing in the frozen complete-case sample",
    "No correlation or network object was estimated"
  ),
  Passed = c(
    all(candidate_values_are_valid),
    nrow(d15_weight_distribution_audit) == 10L,
    all(
      abs(
        proportion_sum_check$Unweighted_proportion - 1
      ) < 1e-12
    ),
    all(
      abs(
        proportion_sum_check$Weighted_proportion - 1
      ) < 1e-12
    ),
    all(
      !is.na(unweighted_difference) &
        unweighted_difference < 1e-12
    ),
    all(
      stats::complete.cases(
        stage1_data[network_vars]
      )
    ),
    TRUE
  ),
  Consequence = c(
    "Continue",
    "Continue",
    "Continue",
    "Continue",
    "Continue",
    "Continue",
    "Hard boundary: stop before weighted polychoric estimation"
  ),
  stringsAsFactors = FALSE
)

stage1_validation

stopifnot(
  all(stage1_validation$Passed)
)
```

## Save Stage 1 outputs

```{r save-stage1-outputs}
distribution_file <- file.path(
  pilot_table_dir,
  "d15_weight_distribution_audit.csv"
)

proportion_file <- file.path(
  pilot_table_dir,
  "d15_weighted_category_proportions.csv"
)

shift_summary_file <- file.path(
  pilot_table_dir,
  "d15_category_shift_summary.csv"
)

stage1_validation_file <- file.path(
  pilot_table_dir,
  "d15_stage1_validation.csv"
)

stage1_session_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage1_sessionInfo.txt"
)

stage1_log_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage1_descriptive_audit.log"
)

utils::write.csv(
  d15_weight_distribution_audit,
  distribution_file,
  row.names = FALSE,
  na = "NA"
)

utils::write.csv(
  d15_weighted_category_proportions,
  proportion_file,
  row.names = FALSE,
  na = "NA"
)

utils::write.csv(
  d15_category_shift_summary,
  shift_summary_file,
  row.names = FALSE,
  na = "NA"
)

utils::write.csv(
  stage1_validation,
  stage1_validation_file,
  row.names = FALSE,
  na = "NA"
)

writeLines(
  capture.output(
    utils::sessionInfo()
  ),
  stage1_session_file
)

stage1_log <- c(
  "D15 SURVEY-WEIGHT PILOT - STAGE 1 DESCRIPTIVE AUDIT",
  paste0(
    "Timestamp: ",
    format(Sys.time(), "%Y-%m-%d %H:%M:%S %z")
  ),
  paste0("Decision registry: ", registry_path),
  paste0("Official CDF SHA-256: ", cdf_file_sha256),
  paste0(
    "Analysis-ready object SHA-256: ",
    analysis_ready_object_sha256
  ),
  paste0(
    "Candidates: ",
    paste(unique(candidate_map$Candidate), collapse = ", ")
  ),
  paste0(
    "Maximum observed absolute category shift (percentage points): ",
    format(
      max(
        d15_weighted_category_proportions$
          Absolute_difference_percentage_points
      ),
      digits = 10
    )
  ),
  "No weight rescaling was applied.",
  "Kish effective N is recorded only as a weight-dispersion diagnostic.",
  "No weighted correlation or network was estimated.",
  "Stage 2 remains blocked pending its reproducibility dependency gate."
)

writeLines(
  stage1_log,
  stage1_log_file
)

stage1_output_files <- c(
  distribution_file,
  proportion_file,
  shift_summary_file,
  stage1_validation_file,
  stage1_session_file,
  stage1_log_file
)

data.frame(
  Output = stage1_output_files,
  Exists = file.exists(stage1_output_files),
  stringsAsFactors = FALSE
)

stopifnot(
  all(file.exists(stage1_output_files))
)
```

## Stage 2A — Reproducibility dependency gate

This gate verifies the approved software dependency and its lockfile record. It does not estimate any correlation or network.

```{r stage2a-dependency-gate}
stage1_validation_path <- file.path(
  pilot_table_dir,
  "d15_stage1_validation.csv"
)

stopifnot(
  file.exists(stage1_validation_path)
)

stage1_validation_from_disk <- utils::read.csv(
  stage1_validation_path,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

stage1_all_passed <-
  nrow(stage1_validation_from_disk) == 7L &&
  all(stage1_validation_from_disk$Passed)

installed_lavaan_version <- if (
  requireNamespace("lavaan", quietly = TRUE)
) {
  as.character(utils::packageVersion("lavaan"))
} else {
  NA_character_
}

installed_wcorr_version <- if (
  requireNamespace("wCorr", quietly = TRUE)
) {
  as.character(utils::packageVersion("wCorr"))
} else {
  NA_character_
}

lockfile_path <- file.path(
  project_root,
  "renv.lock"
)

stopifnot(
  file.exists(lockfile_path),
  requireNamespace("jsonlite", quietly = TRUE)
)

lockfile <- jsonlite::fromJSON(
  lockfile_path,
  simplifyVector = FALSE
)

locked_wcorr_version <- if (
  !is.null(lockfile$Packages$wCorr$Version)
) {
  as.character(lockfile$Packages$wCorr$Version)
} else {
  NA_character_
}

wcorr_formals <- if (
  requireNamespace("wCorr", quietly = TRUE)
) {
  names(formals(wCorr::weightedCorr))
} else {
  character(0)
}

lavaan_formals <- if (
  requireNamespace("lavaan", quietly = TRUE)
) {
  names(formals(lavaan::lavCor))
} else {
  character(0)
}

stage2_dependency_gate <- data.frame(
  Check = c(
    "All seven Stage 1 validation checks passed",
    "lavaan is installed at the frozen version 0.7-2",
    "wCorr is installed at the approved version 1.9.8",
    "wCorr 1.9.8 is recorded in renv.lock",
    "wCorr weightedCorr exposes required arguments",
    "lavaan lavCor exposes required weighted ordinal arguments",
    "No correlation or network was estimated in this gate"
  ),
  Passed = c(
    stage1_all_passed,
    identical(
  installed_lavaan_version,
  "0.7.2"
    ),
    identical(
      installed_wcorr_version,
      "1.9.8"
    ),
    identical(
      locked_wcorr_version,
      "1.9.8"
    ),
    all(
      c(
        "x",
        "y",
        "method",
        "weights",
        "ML",
        "fast"
      ) %in% wcorr_formals
    ),
    all(
      c(
        "object",
        "ordered",
        "sampling_weights",
        "estimator",
        "cor_smooth"
      ) %in% lavaan_formals
    ),
    TRUE
  ),
  Observed = c(
    paste(
      sum(stage1_validation_from_disk$Passed),
      "of",
      nrow(stage1_validation_from_disk)
    ),
    installed_lavaan_version,
    installed_wcorr_version,
    locked_wcorr_version,
    paste(wcorr_formals, collapse = "; "),
    paste(lavaan_formals, collapse = "; "),
    "Dependency inspection only"
  ),
  Consequence = c(
    "Continue if TRUE",
    "Continue if TRUE",
    "Continue if TRUE",
    "Continue if TRUE",
    "Continue if TRUE",
    "Continue if TRUE",
    "Hard stop before weighted polychoric estimation"
  ),
  stringsAsFactors = FALSE
)

stage2_dependency_gate

stage2_dependency_file <- file.path(
  pilot_table_dir,
  "d15_stage2_dependency_gate.csv"
)

stage2_dependency_log <- file.path(
  pilot_log_dir,
  "06B_D15_stage2_dependency_gate.log"
)

stage2_dependency_session <- file.path(
  pilot_log_dir,
  "06B_D15_stage2_dependency_sessionInfo.txt"
)

utils::write.csv(
  stage2_dependency_gate,
  stage2_dependency_file,
  row.names = FALSE,
  na = "NA"
)

writeLines(
  capture.output(
    utils::sessionInfo()
  ),
  stage2_dependency_session
)

writeLines(
  c(
    "D15 STAGE 2A REPRODUCIBILITY DEPENDENCY GATE",
    paste0(
      "Timestamp: ",
      format(
        Sys.time(),
        "%Y-%m-%d %H:%M:%S %z"
      )
    ),
    paste0(
      "lavaan installed version: ",
      installed_lavaan_version
    ),
    paste0(
      "wCorr installed version: ",
      installed_wcorr_version
    ),
    paste0(
      "wCorr locked version: ",
      locked_wcorr_version
    ),
    paste0(
      "All dependency checks passed: ",
      all(stage2_dependency_gate$Passed)
    ),
    "No weighted correlation or network was estimated."
  ),
  stage2_dependency_log
)

stopifnot(
  file.exists(stage2_dependency_file),
  file.exists(stage2_dependency_log),
  file.exists(stage2_dependency_session),
  all(stage2_dependency_gate$Passed)
)
```

## Stage 2B — 2004 weighted-polychoric implementation pilot

This bounded pilot compares two weighted-polychoric software implementations for the 2004 overall complete-case sample. It does not estimate a GGM or perform network interpretation.

```{r stage2b-dependency-and-input-gate}
stage1_validation_disk <- utils::read.csv(
  file.path(
    pilot_table_dir,
    "d15_stage1_validation.csv"
  ),
  stringsAsFactors = FALSE
)

lock <- jsonlite::fromJSON(
  file.path(project_root, "renv.lock"),
  simplifyVector = FALSE
)

stage2_dependency_gate <- data.frame(
  Check = c(
    "All seven Stage 1 checks passed",
    "wCorr 1.9.8 is installed",
    "wCorr 1.9.8 is recorded in renv.lock",
    "lavaan 0.7-2 is installed",
    "The 2004 complete-case sample contains 789 respondents",
    "Both approved weight candidates are finite and positive"
  ),
  Passed = c(
    nrow(stage1_validation_disk) == 7L &&
      all(stage1_validation_disk$Passed),
    requireNamespace("wCorr", quietly = TRUE) &&
      as.character(
        utils::packageVersion("wCorr")
      ) == "1.9.8",
    !is.null(lock$Packages$wCorr) &&
      lock$Packages$wCorr$Version == "1.9.8",
    requireNamespace("lavaan", quietly = TRUE) &&
      as.character(
        utils::packageVersion("lavaan")
      ) == "0.7.2",
    sum(
      as.integer(stage1_data$VCF0004) == 2004L
    ) == 789L,
    all(
      is.finite(
        stage1_data$W_CDF[
          as.integer(stage1_data$VCF0004) == 2004L
        ]
      ) &
        stage1_data$W_CDF[
          as.integer(stage1_data$VCF0004) == 2004L
        ] > 0
    ) &&
      all(
        is.finite(
          stage1_data$W_POST_COMPAT[
            as.integer(stage1_data$VCF0004) == 2004L
          ]
        ) &
          stage1_data$W_POST_COMPAT[
            as.integer(stage1_data$VCF0004) == 2004L
          ] > 0
      )
  ),
  stringsAsFactors = FALSE
)

stage2_dependency_gate

utils::write.csv(
  stage2_dependency_gate,
  file.path(
    pilot_table_dir,
    "d15_stage2_dependency_gate.csv"
  ),
  row.names = FALSE
)

stopifnot(
  all(stage2_dependency_gate$Passed)
)

stage2_year <- 2004L

stage2_candidates <- c(
  "W_CDF",
  "W_POST_COMPAT"
)

stage2_data <- stage1_data[
  as.integer(stage1_data$VCF0004) == stage2_year,
  ,
  drop = FALSE
]

stopifnot(
  nrow(stage2_data) == 789L,
  identical(
    network_vars,
    c(
      "VCF0806",
      "VCF0809",
      "VCF0838",
      "VCF0839",
      "VCF0879a",
      "VCF0888",
      "VCF0890",
      "VCF0894",
      "VCF9223"
    )
  )
)
```

```{r stage2b-helper-functions}
make_ordered_factor <- function(values) {
  if (is.factor(values)) {
    observed_levels <- levels(
      droplevels(values)
    )

    return(
      ordered(
        as.character(values),
        levels = observed_levels
      )
    )
  }

  numeric_values <- as.numeric(values)

  ordered(
    numeric_values,
    levels = sort(
      unique(
        numeric_values[
          !is.na(numeric_values)
        ]
      )
    )
  )
}

run_and_capture <- function(calculation) {
  warning_messages <- character(0)
  ordinary_messages <- character(0)
  error_message <- NA_character_

  value <- tryCatch(
    withCallingHandlers(
      calculation(),
      warning = function(condition) {
        warning_messages <<- c(
          warning_messages,
          conditionMessage(condition)
        )
        invokeRestart("muffleWarning")
      },
      message = function(condition) {
        ordinary_messages <<- c(
          ordinary_messages,
          conditionMessage(condition)
        )
        invokeRestart("muffleMessage")
      }
    ),
    error = function(condition) {
      error_message <<- conditionMessage(condition)
      NULL
    }
  )

  list(
    value = value,
    warnings = unique(warning_messages),
    messages = unique(ordinary_messages),
    error = error_message
  )
}

collapse_conditions <- function(values) {
  if (length(values) == 0L) {
    ""
  } else {
    paste(values, collapse = " | ")
  }
}

diagnose_matrix <- function(
  matrix_object,
  candidate,
  implementation
) {
  matrix_present <-
    is.matrix(matrix_object) &&
    identical(
      dim(matrix_object),
      c(9L, 9L)
    )

  finite <- matrix_present &&
    all(is.finite(matrix_object))

  maximum_asymmetry <- if (finite) {
    max(
      abs(
        matrix_object -
          t(matrix_object)
      )
    )
  } else {
    NA_real_
  }

  maximum_diagonal_difference <- if (finite) {
    max(
      abs(
        diag(matrix_object) - 1
      )
    )
  } else {
    NA_real_
  }

  minimum_eigenvalue <- if (
    finite &&
    maximum_asymmetry <= 1e-10
  ) {
    tryCatch(
      min(
        eigen(
          matrix_object,
          symmetric = TRUE,
          only.values = TRUE
        )$values
      ),
      error = function(condition) {
        NA_real_
      }
    )
  } else {
    NA_real_
  }

  off_diagonal <- if (finite) {
    matrix_object[
      upper.tri(matrix_object)
    ]
  } else {
    numeric(0)
  }

  data.frame(
    Election_year = stage2_year,
    Candidate = candidate,
    Implementation = implementation,
    Matrix_present = matrix_present,
    Expected_dimension = "9 x 9",
    All_values_finite = finite,
    Maximum_asymmetry = maximum_asymmetry,
    Maximum_diagonal_difference =
      maximum_diagonal_difference,
    Minimum_eigenvalue = minimum_eigenvalue,
    Positive_definite =
      !is.na(minimum_eigenvalue) &&
      minimum_eigenvalue > 0,
    Unique_off_diagonal_values =
      length(off_diagonal),
    Minimum_off_diagonal = if (
      length(off_diagonal) > 0L
    ) {
      min(off_diagonal)
    } else {
      NA_real_
    },
    Maximum_off_diagonal = if (
      length(off_diagonal) > 0L
    ) {
      max(off_diagonal)
    } else {
      NA_real_
    },
    stringsAsFactors = FALSE
  )
}
```

```{r stage2b-estimation}
stage2_results <- list()
stage2_condition_records <- list()
condition_index <- 1L

for (candidate_name in stage2_candidates) {
  node_data <- data.frame(
    lapply(
      stage2_data[network_vars],
      make_ordered_factor
    ),
    check.names = FALSE
  )

  names(node_data) <- network_vars

  analysis_weights <- as.numeric(
    stage2_data[[candidate_name]]
  )

  stopifnot(
    all(
      vapply(
        node_data,
        function(variable) {
          nlevels(variable) >= 2L
        },
        FUN.VALUE = logical(1)
      )
    ),
    all(
      is.finite(analysis_weights) &
        analysis_weights > 0
    )
  )

  lavaan_data <- node_data
  lavaan_data$analysis_weight <-
    analysis_weights

  lavaan_run <- run_and_capture(
    function() {
      lavaan::lavCor(
        object = lavaan_data,
        ordered = network_vars,
        sampling_weights = "analysis_weight",
        missing = "listwise",
        se = "none",
        test = "none",
        estimator = "two.step",
        baseline = FALSE,
        cor_smooth = FALSE,
        output = "cor"
      )
    }
  )

  lavaan_matrix <- if (
    is.null(lavaan_run$value)
  ) {
    NULL
  } else {
    as.matrix(lavaan_run$value)
  }

  if (!is.null(lavaan_matrix)) {
    dimnames(lavaan_matrix) <- list(
      network_vars,
      network_vars
    )
  }

  stage2_condition_records[[
    condition_index
  ]] <- data.frame(
    Election_year = stage2_year,
    Candidate = candidate_name,
    Implementation = "lavaan_lavCor",
    Pair = "FULL_MATRIX",
    Warning_count =
      length(lavaan_run$warnings),
    Warnings =
      collapse_conditions(
        lavaan_run$warnings
      ),
    Message_count =
      length(lavaan_run$messages),
    Messages =
      collapse_conditions(
        lavaan_run$messages
      ),
    Error = ifelse(
      is.na(lavaan_run$error),
      "",
      lavaan_run$error
    ),
    stringsAsFactors = FALSE
  )

  condition_index <- condition_index + 1L

  wcorr_matrix <- diag(
    1,
    nrow = length(network_vars),
    ncol = length(network_vars)
  )

  dimnames(wcorr_matrix) <- list(
    network_vars,
    network_vars
  )

  for (
    first_index in
      seq_len(length(network_vars) - 1L)
  ) {
    for (
      second_index in
        seq.int(
          first_index + 1L,
          length(network_vars)
        )
    ) {
      first_node <-
        network_vars[first_index]

      second_node <-
        network_vars[second_index]

      pair_run <- run_and_capture(
        function() {
          wCorr::weightedCorr(
            x = node_data[[first_node]],
            y = node_data[[second_node]],
            method = "Polychoric",
            weights = analysis_weights,
            ML = FALSE,
            fast = TRUE
          )
        }
      )

      pair_value <- if (
        is.null(pair_run$value)
      ) {
        NA_real_
      } else {
        as.numeric(pair_run$value)
      }

      wcorr_matrix[
        first_index,
        second_index
      ] <- pair_value

      wcorr_matrix[
        second_index,
        first_index
      ] <- pair_value

      stage2_condition_records[[
        condition_index
      ]] <- data.frame(
        Election_year = stage2_year,
        Candidate = candidate_name,
        Implementation =
          "wCorr_weightedCorr",
        Pair = paste(
          first_node,
          second_node,
          sep = " -- "
        ),
        Warning_count =
          length(pair_run$warnings),
        Warnings =
          collapse_conditions(
            pair_run$warnings
          ),
        Message_count =
          length(pair_run$messages),
        Messages =
          collapse_conditions(
            pair_run$messages
          ),
        Error = ifelse(
          is.na(pair_run$error),
          "",
          pair_run$error
        ),
        stringsAsFactors = FALSE
      )

      condition_index <-
        condition_index + 1L
    }
  }

  stage2_results[[candidate_name]] <- list(
    lavaan = lavaan_matrix,
    wCorr = wcorr_matrix
  )
}

stage2_conditions <- do.call(
  rbind,
  stage2_condition_records
)

stage2_conditions
```

```{r stage2b-diagnostics-and-comparison}
stage2_diagnostics <- do.call(
  rbind,
  lapply(
    stage2_candidates,
    function(candidate_name) {
      rbind(
        diagnose_matrix(
          stage2_results[[
            candidate_name
          ]]$lavaan,
          candidate_name,
          "lavaan_lavCor"
        ),
        diagnose_matrix(
          stage2_results[[
            candidate_name
          ]]$wCorr,
          candidate_name,
          "wCorr_weightedCorr"
        )
      )
    }
  )
)

compare_implementations <- function(
  candidate_name
) {
  lavaan_matrix <-
    stage2_results[[
      candidate_name
    ]]$lavaan

  wcorr_matrix <-
    stage2_results[[
      candidate_name
    ]]$wCorr

  matrices_available <-
    is.matrix(lavaan_matrix) &&
    is.matrix(wcorr_matrix) &&
    identical(
      dim(lavaan_matrix),
      c(9L, 9L)
    ) &&
    identical(
      dim(wcorr_matrix),
      c(9L, 9L)
    )

  if (!matrices_available) {
    return(
      data.frame(
        Election_year = stage2_year,
        Candidate = candidate_name,
        Compared_pairs = 0L,
        Correlation_between_estimates = NA_real_,
        Median_absolute_difference = NA_real_,
        Mean_absolute_difference = NA_real_,
        Maximum_absolute_difference = NA_real_,
        Maximum_difference_pair = NA_character_,
        Sign_agreement_proportion = NA_real_,
        stringsAsFactors = FALSE
      )
    )
  }

  pair_indices <- which(
    upper.tri(lavaan_matrix),
    arr.ind = TRUE
  )

  lavaan_values <-
    lavaan_matrix[pair_indices]

  wcorr_values <-
    wcorr_matrix[pair_indices]

  absolute_differences <- abs(
    lavaan_values -
      wcorr_values
  )

  maximum_index <- which.max(
    absolute_differences
  )

  pair_labels <- paste(
    rownames(lavaan_matrix)[
      pair_indices[, 1]
    ],
    colnames(lavaan_matrix)[
      pair_indices[, 2]
    ],
    sep = " -- "
  )

  data.frame(
    Election_year = stage2_year,
    Candidate = candidate_name,
    Compared_pairs =
      length(lavaan_values),
    Correlation_between_estimates =
      stats::cor(
        lavaan_values,
        wcorr_values
      ),
    Median_absolute_difference =
      stats::median(
        absolute_differences
      ),
    Mean_absolute_difference =
      mean(
        absolute_differences
      ),
    Maximum_absolute_difference =
      max(
        absolute_differences
      ),
    Maximum_difference_pair =
      pair_labels[maximum_index],
    Sign_agreement_proportion =
      mean(
        sign(lavaan_values) ==
          sign(wcorr_values)
      ),
    stringsAsFactors = FALSE
  )
}

stage2_comparison <- do.call(
  rbind,
  lapply(
    stage2_candidates,
    compare_implementations
  )
)

stage2_diagnostics
stage2_comparison
```

```{r stage2b-validation-and-save}
stage2_validation <- data.frame(
  Check = c(
    "All four weighted-polychoric matrices were returned",
    "All matrix entries are finite",
    "All matrices are symmetric within tolerance",
    "All matrix diagonals equal one within tolerance",
    "All four matrices are positive definite",
    "Both implementation comparisons contain 36 pairs",
    "No estimation error was captured",
    "All warnings and messages were preserved for review",
    "No smoothing repair retry or category collapse was used",
    "No GGM or other network model was estimated"
  ),
  Passed = c(
    all(stage2_diagnostics$Matrix_present),
    all(stage2_diagnostics$All_values_finite),
    all(
      stage2_diagnostics$Maximum_asymmetry <=
        1e-10
    ),
    all(
      stage2_diagnostics$
        Maximum_diagonal_difference <=
        1e-10
    ),
    all(stage2_diagnostics$Positive_definite),
    all(
      stage2_comparison$Compared_pairs ==
        36L
    ),
    all(
      stage2_conditions$Error == ""
    ),
    TRUE,
    TRUE,
    TRUE
  ),
  stringsAsFactors = FALSE
)

stage2_validation

stage2_model_dir <- file.path(
  project_root,
  "outputs",
  "models",
  "pilot"
)

dir.create(
  stage2_model_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

diagnostics_file <- file.path(
  pilot_table_dir,
  "d15_stage2b_2004_weighted_polychoric_diagnostics.csv"
)

comparison_file <- file.path(
  pilot_table_dir,
  "d15_stage2b_2004_implementation_comparison.csv"
)

conditions_file <- file.path(
  pilot_table_dir,
  "d15_stage2b_2004_conditions.csv"
)

validation_file_stage2 <- file.path(
  pilot_table_dir,
  "d15_stage2b_2004_validation.csv"
)

matrix_file <- file.path(
  stage2_model_dir,
  "d15_stage2b_2004_weighted_polychoric_matrices.rds"
)

stage2_log_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage2b_2004.log"
)

stage2_session_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage2b_2004_sessionInfo.txt"
)

utils::write.csv(
  stage2_diagnostics,
  diagnostics_file,
  row.names = FALSE
)

utils::write.csv(
  stage2_comparison,
  comparison_file,
  row.names = FALSE
)

utils::write.csv(
  stage2_conditions,
  conditions_file,
  row.names = FALSE
)

utils::write.csv(
  stage2_validation,
  validation_file_stage2,
  row.names = FALSE
)

saveRDS(
  stage2_results,
  matrix_file
)

writeLines(
  capture.output(
    utils::sessionInfo()
  ),
  stage2_session_file
)

writeLines(
  c(
    "D15 STAGE 2B — 2004 WEIGHTED-POLYCHORIC PILOT",
    paste0(
      "Timestamp: ",
      format(
        Sys.time(),
        "%Y-%m-%d %H:%M:%S %z"
      )
    ),
    paste0(
      "Validation checks passed: ",
      sum(stage2_validation$Passed),
      "/",
      nrow(stage2_validation)
    ),
    paste0(
      "Captured warning count: ",
      sum(stage2_conditions$Warning_count)
    ),
    paste0(
      "Captured error count: ",
      sum(stage2_conditions$Error != "")
    ),
    "No smoothing, matrix repair, retry, category collapse or network estimation was performed."
  ),
  stage2_log_file
)

stage2_output_files <- c(
  diagnostics_file,
  comparison_file,
  conditions_file,
  validation_file_stage2,
  matrix_file,
  stage2_log_file,
  stage2_session_file
)

data.frame(
  Output = stage2_output_files,
  Exists = file.exists(stage2_output_files),
  stringsAsFactors = FALSE
)

stopifnot(
  all(file.exists(stage2_output_files))
)
```

## Hard stop after the 2004 Stage 2B pilot

Stop here. Do not extend the weighted-polychoric calculation to the other four waves and do not estimate EBICglasso until the 2004 implementation comparison, matrix diagnostics, warnings and errors have been independently reviewed.
## Stage 2C — Five-wave weighted-polychoric implementation check

This stage extends the validated 2004 implementation to the five approved overall waves. It remains a correlation-matrix diagnostic and does not estimate a network.

```{r stage2c-five-wave-estimation}
stage2c_years <- c(
  2004L,
  2012L,
  2016L,
  2020L,
  2024L
)

stage2c_candidates <- c(
  "W_CDF",
  "W_POST_COMPAT"
)

run_stage2c_candidate <- function(
  year_value,
  candidate_name
) {
  current_data <- stage1_data[
    as.integer(stage1_data$VCF0004) ==
      year_value,
    ,
    drop = FALSE
  ]

  expected_n <- unname(
    expected_year_counts[
      as.character(year_value)
    ]
  )

  stopifnot(
    nrow(current_data) == expected_n
  )

  node_data <- data.frame(
    lapply(
      current_data[network_vars],
      make_ordered_factor
    ),
    check.names = FALSE
  )

  names(node_data) <- network_vars

  analysis_weights <- as.numeric(
    current_data[[candidate_name]]
  )

  stopifnot(
    all(
      is.finite(analysis_weights) &
        analysis_weights > 0
    ),
    all(
      vapply(
        node_data,
        function(variable) {
          nlevels(variable) >= 2L
        },
        FUN.VALUE = logical(1)
      )
    )
  )

  lavaan_data <- node_data
  lavaan_data$analysis_weight <-
    analysis_weights

  lavaan_run <- run_and_capture(
    function() {
      lavaan::lavCor(
        object = lavaan_data,
        ordered = network_vars,
        sampling_weights = "analysis_weight",
        missing = "listwise",
        se = "none",
        test = "none",
        estimator = "two.step",
        baseline = FALSE,
        cor_smooth = FALSE,
        output = "cor"
      )
    }
  )

  lavaan_matrix <- if (
    is.null(lavaan_run$value)
  ) {
    NULL
  } else {
    as.matrix(lavaan_run$value)
  }

  if (!is.null(lavaan_matrix)) {
    dimnames(lavaan_matrix) <- list(
      network_vars,
      network_vars
    )
  }

  condition_rows <- list(
    data.frame(
      Election_year = year_value,
      Candidate = candidate_name,
      Implementation = "lavaan_lavCor",
      Pair = "FULL_MATRIX",
      Warning_count =
        length(lavaan_run$warnings),
      Warnings =
        collapse_conditions(
          lavaan_run$warnings
        ),
      Message_count =
        length(lavaan_run$messages),
      Messages =
        collapse_conditions(
          lavaan_run$messages
        ),
      Error = ifelse(
        is.na(lavaan_run$error),
        "",
        lavaan_run$error
      ),
      stringsAsFactors = FALSE
    )
  )

  wcorr_matrix <- diag(
    1,
    nrow = length(network_vars),
    ncol = length(network_vars)
  )

  dimnames(wcorr_matrix) <- list(
    network_vars,
    network_vars
  )

  condition_index <- 2L

  for (
    first_index in
      seq_len(length(network_vars) - 1L)
  ) {
    for (
      second_index in
        seq.int(
          first_index + 1L,
          length(network_vars)
        )
    ) {
      first_node <-
        network_vars[first_index]

      second_node <-
        network_vars[second_index]

      pair_run <- run_and_capture(
        function() {
          wCorr::weightedCorr(
            x = node_data[[first_node]],
            y = node_data[[second_node]],
            method = "Polychoric",
            weights = analysis_weights,
            ML = FALSE,
            fast = TRUE
          )
        }
      )

      pair_value <- if (
        is.null(pair_run$value)
      ) {
        NA_real_
      } else {
        as.numeric(pair_run$value)
      }

      wcorr_matrix[
        first_index,
        second_index
      ] <- pair_value

      wcorr_matrix[
        second_index,
        first_index
      ] <- pair_value

      condition_rows[[
        condition_index
      ]] <- data.frame(
        Election_year = year_value,
        Candidate = candidate_name,
        Implementation =
          "wCorr_weightedCorr",
        Pair = paste(
          first_node,
          second_node,
          sep = " -- "
        ),
        Warning_count =
          length(pair_run$warnings),
        Warnings =
          collapse_conditions(
            pair_run$warnings
          ),
        Message_count =
          length(pair_run$messages),
        Messages =
          collapse_conditions(
            pair_run$messages
          ),
        Error = ifelse(
          is.na(pair_run$error),
          "",
          pair_run$error
        ),
        stringsAsFactors = FALSE
      )

      condition_index <-
        condition_index + 1L
    }
  }

  list(
    Election_year = year_value,
    Candidate = candidate_name,
    Sample_N = nrow(current_data),
    lavaan = lavaan_matrix,
    wCorr = wcorr_matrix,
    conditions = do.call(
      rbind,
      condition_rows
    )
  )
}

stage2c_grid <- expand.grid(
  Election_year = stage2c_years,
  Candidate = stage2c_candidates,
  stringsAsFactors = FALSE
)

stage2c_results <- Map(
  run_stage2c_candidate,
  stage2c_grid$Election_year,
  stage2c_grid$Candidate
)

names(stage2c_results) <- paste(
  stage2c_grid$Election_year,
  stage2c_grid$Candidate,
  sep = "__"
)

stage2c_conditions <- do.call(
  rbind,
  lapply(
    stage2c_results,
    function(result) {
      result$conditions
    }
  )
)

stage2c_conditions
```

```{r stage2c-diagnostics}
diagnose_stage2c_matrix <- function(
  result,
  implementation
) {
  matrix_object <- result[[implementation]]

  matrix_present <-
    is.matrix(matrix_object) &&
    identical(
      dim(matrix_object),
      c(9L, 9L)
    )

  finite <- matrix_present &&
    all(is.finite(matrix_object))

  maximum_asymmetry <- if (finite) {
    max(
      abs(
        matrix_object -
          t(matrix_object)
      )
    )
  } else {
    NA_real_
  }

  maximum_diagonal_difference <- if (finite) {
    max(
      abs(
        diag(matrix_object) - 1
      )
    )
  } else {
    NA_real_
  }

  minimum_eigenvalue <- if (
    finite &&
    maximum_asymmetry <= 1e-10
  ) {
    tryCatch(
      min(
        eigen(
          matrix_object,
          symmetric = TRUE,
          only.values = TRUE
        )$values
      ),
      error = function(condition) {
        NA_real_
      }
    )
  } else {
    NA_real_
  }

  data.frame(
    Election_year =
      result$Election_year,
    Candidate =
      result$Candidate,
    Sample_N =
      result$Sample_N,
    Implementation =
      implementation,
    Matrix_present =
      matrix_present,
    All_values_finite =
      finite,
    Maximum_asymmetry =
      maximum_asymmetry,
    Maximum_diagonal_difference =
      maximum_diagonal_difference,
    Minimum_eigenvalue =
      minimum_eigenvalue,
    Positive_definite =
      !is.na(minimum_eigenvalue) &&
      minimum_eigenvalue > 0,
    Unique_off_diagonal_values = if (
      finite
    ) {
      length(
        matrix_object[
          upper.tri(matrix_object)
        ]
      )
    } else {
      0L
    },
    stringsAsFactors = FALSE
  )
}

stage2c_diagnostics <- do.call(
  rbind,
  lapply(
    stage2c_results,
    function(result) {
      rbind(
        diagnose_stage2c_matrix(
          result,
          "lavaan"
        ),
        diagnose_stage2c_matrix(
          result,
          "wCorr"
        )
      )
    }
  )
)

compare_stage2c_implementations <- function(
  result
) {
  lavaan_matrix <- result$lavaan
  wcorr_matrix <- result$wCorr

  available <-
    is.matrix(lavaan_matrix) &&
    is.matrix(wcorr_matrix) &&
    identical(
      dim(lavaan_matrix),
      c(9L, 9L)
    ) &&
    identical(
      dim(wcorr_matrix),
      c(9L, 9L)
    )

  if (!available) {
    return(
      data.frame(
        Election_year =
          result$Election_year,
        Candidate =
          result$Candidate,
        Compared_pairs = 0L,
        Correlation_between_estimates =
          NA_real_,
        Median_absolute_difference =
          NA_real_,
        Mean_absolute_difference =
          NA_real_,
        Maximum_absolute_difference =
          NA_real_,
        Maximum_difference_pair =
          NA_character_,
        Sign_agreement_proportion =
          NA_real_,
        stringsAsFactors = FALSE
      )
    )
  }

  pair_indices <- which(
    upper.tri(lavaan_matrix),
    arr.ind = TRUE
  )

  lavaan_values <-
    lavaan_matrix[pair_indices]

  wcorr_values <-
    wcorr_matrix[pair_indices]

  absolute_differences <- abs(
    lavaan_values -
      wcorr_values
  )

  pair_labels <- paste(
    rownames(lavaan_matrix)[
      pair_indices[, 1]
    ],
    colnames(lavaan_matrix)[
      pair_indices[, 2]
    ],
    sep = " -- "
  )

  maximum_index <- which.max(
    absolute_differences
  )

  data.frame(
    Election_year =
      result$Election_year,
    Candidate =
      result$Candidate,
    Compared_pairs =
      length(lavaan_values),
    Correlation_between_estimates =
      stats::cor(
        lavaan_values,
        wcorr_values
      ),
    Median_absolute_difference =
      stats::median(
        absolute_differences
      ),
    Mean_absolute_difference =
      mean(
        absolute_differences
      ),
    Maximum_absolute_difference =
      max(
        absolute_differences
      ),
    Maximum_difference_pair =
      pair_labels[maximum_index],
    Sign_agreement_proportion =
      mean(
        sign(lavaan_values) ==
          sign(wcorr_values)
      ),
    stringsAsFactors = FALSE
  )
}

stage2c_comparison <- do.call(
  rbind,
  lapply(
    stage2c_results,
    compare_stage2c_implementations
  )
)

stage2c_diagnostics
stage2c_comparison
```

```{r stage2c-validation-and-save}
stage2c_validation <- data.frame(
  Check = c(
    "Twenty weighted-polychoric matrices were returned",
    "All matrix entries are finite",
    "All matrices are symmetric within tolerance",
    "All matrix diagonals equal one within tolerance",
    "All matrices are positive definite",
    "Ten implementation comparisons contain 36 pairs each",
    "No estimation error was captured",
    "No estimation warning was captured",
    "No smoothing repair retry or category collapse was used",
    "No GGM or other network model was estimated"
  ),
  Passed = c(
    nrow(stage2c_diagnostics) == 20L &&
      all(stage2c_diagnostics$Matrix_present),
    all(
      stage2c_diagnostics$All_values_finite
    ),
    all(
      stage2c_diagnostics$
        Maximum_asymmetry <=
        1e-10
    ),
    all(
      stage2c_diagnostics$
        Maximum_diagonal_difference <=
        1e-10
    ),
    all(
      stage2c_diagnostics$
        Positive_definite
    ),
    nrow(stage2c_comparison) == 10L &&
      all(
        stage2c_comparison$
          Compared_pairs == 36L
      ),
    all(
      stage2c_conditions$Error == ""
    ),
    sum(
      stage2c_conditions$Warning_count
    ) == 0L,
    TRUE,
    TRUE
  ),
  stringsAsFactors = FALSE
)

stage2c_validation

stage2c_model_dir <- file.path(
  project_root,
  "outputs",
  "models",
  "pilot"
)

dir.create(
  stage2c_model_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

stage2c_diagnostics_file <- file.path(
  pilot_table_dir,
  "d15_stage2c_five_wave_weighted_polychoric_diagnostics.csv"
)

stage2c_comparison_file <- file.path(
  pilot_table_dir,
  "d15_stage2c_five_wave_implementation_comparison.csv"
)

stage2c_conditions_file <- file.path(
  pilot_table_dir,
  "d15_stage2c_five_wave_conditions.csv"
)

stage2c_validation_file <- file.path(
  pilot_table_dir,
  "d15_stage2c_five_wave_validation.csv"
)

stage2c_matrix_file <- file.path(
  stage2c_model_dir,
  "d15_stage2c_five_wave_weighted_polychoric_matrices.rds"
)

stage2c_log_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage2c_five_wave.log"
)

stage2c_session_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage2c_five_wave_sessionInfo.txt"
)

utils::write.csv(
  stage2c_diagnostics,
  stage2c_diagnostics_file,
  row.names = FALSE
)

utils::write.csv(
  stage2c_comparison,
  stage2c_comparison_file,
  row.names = FALSE
)

utils::write.csv(
  stage2c_conditions,
  stage2c_conditions_file,
  row.names = FALSE
)

utils::write.csv(
  stage2c_validation,
  stage2c_validation_file,
  row.names = FALSE
)

saveRDS(
  stage2c_results,
  stage2c_matrix_file
)

writeLines(
  capture.output(
    utils::sessionInfo()
  ),
  stage2c_session_file
)

writeLines(
  c(
    "D15 STAGE 2C — FIVE-WAVE WEIGHTED-POLYCHORIC CHECK",
    paste0(
      "Timestamp: ",
      format(
        Sys.time(),
        "%Y-%m-%d %H:%M:%S %z"
      )
    ),
    paste0(
      "Matrices returned: ",
      sum(
        stage2c_diagnostics$
          Matrix_present
      ),
      "/20"
    ),
    paste0(
      "Warnings captured: ",
      sum(
        stage2c_conditions$
          Warning_count
      )
    ),
    paste0(
      "Errors captured: ",
      sum(
        stage2c_conditions$
          Error != ""
      )
    ),
    paste0(
      "Validation checks passed: ",
      sum(stage2c_validation$Passed),
      "/",
      nrow(stage2c_validation)
    ),
    "No GGM or other network model was estimated."
  ),
  stage2c_log_file
)

stage2c_output_files <- c(
  stage2c_diagnostics_file,
  stage2c_comparison_file,
  stage2c_conditions_file,
  stage2c_validation_file,
  stage2c_matrix_file,
  stage2c_log_file,
  stage2c_session_file
)

stopifnot(
  all(
    file.exists(stage2c_output_files)
  )
)

data.frame(
  Output = stage2c_output_files,
  Exists = file.exists(
    stage2c_output_files
  ),
  stringsAsFactors = FALSE
)
```

## Hard stop after Stage 2C
## Stage 3A — 2004 diagnostic point-network pilot

Stage 2C passed independent review. This bounded pilot compares the
unweighted, W_CDF and W_POST_COMPAT point networks for 2004 under an
identical EBICglasso specification. It does not select the final weight,
run bootstrap, calculate centrality, perform NCT, or interpret political
results.
```{r stage3a-2004-input-gate}
stage3a_required_packages <- c(
  "qgraph",
  "digest"
)

stage3a_package_gate <- data.frame(
  Package = stage3a_required_packages,
  Available = vapply(
    stage3a_required_packages,
    requireNamespace,
    quietly = TRUE,
    FUN.VALUE = logical(1)
  ),
  Version = vapply(
    stage3a_required_packages,
    function(package_name) {
      if (requireNamespace(package_name, quietly = TRUE)) {
        as.character(utils::packageVersion(package_name))
      } else {
        NA_character_
      }
    },
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

stage3a_package_gate

stopifnot(
  all(stage3a_package_gate$Available)
)

stage2c_validation_path <- file.path(
  pilot_table_dir,
  "d15_stage2c_five_wave_validation.csv"
)

stage2c_matrix_path <- file.path(
  stage2c_model_dir,
  "d15_stage2c_five_wave_weighted_polychoric_matrices.rds"
)

unweighted_matrix_path <- file.path(
  project_root,
  "data",
  "audit",
  paste0(
    "anes_cdf_20260205_nine_node_complete_case_",
    "polychoric_diagnostics_2004_2012_2016_2020_2024.rds"
  )
)

stage07_pilot_path <- file.path(
  project_root,
  "outputs",
  "models",
  "pilot",
  "2004_pilot_network_gamma050.rds"
)

stage3a_required_files <- c(
  stage2c_validation_path,
  stage2c_matrix_path,
  unweighted_matrix_path,
  stage07_pilot_path
)

stopifnot(
  all(file.exists(stage3a_required_files))
)

stage2c_validation_disk <- utils::read.csv(
  stage2c_validation_path,
  stringsAsFactors = FALSE
)

stopifnot(
  nrow(stage2c_validation_disk) == 10L,
  all(as.logical(stage2c_validation_disk$Passed))
)

stage2c_saved <- readRDS(
  stage2c_matrix_path
)

unweighted_bundle <- readRDS(
  unweighted_matrix_path
)

stage07_pilot <- readRDS(
  stage07_pilot_path
)

stopifnot(
  "2004" %in% names(unweighted_bundle$Matrices),
  "2004__W_CDF" %in% names(stage2c_saved),
  "2004__W_POST_COMPAT" %in% names(stage2c_saved),
  !is.null(stage07_pilot$graph)
)

stage3a_input_matrices <- list(
  U_RAWN = as.matrix(
    unweighted_bundle$Matrices[["2004"]]
  ),
  W_CDF_RAWN = as.matrix(
    stage2c_saved[["2004__W_CDF"]]$lavaan
  ),
  W_POST_COMPAT_RAWN = as.matrix(
    stage2c_saved[["2004__W_POST_COMPAT"]]$lavaan
  )
)

stage3a_input_checks <- do.call(
  rbind,
  lapply(
    names(stage3a_input_matrices),
    function(candidate_name) {
      current_matrix <-
        stage3a_input_matrices[[candidate_name]]

      data.frame(
        Candidate = candidate_name,
        Rows = nrow(current_matrix),
        Columns = ncol(current_matrix),
        All_finite = all(is.finite(current_matrix)),
        Symmetric = isSymmetric(
          current_matrix,
          tol = 1e-10
        ),
        Diagonal_one = max(
          abs(diag(current_matrix) - 1)
        ) <= 1e-10,
        Node_order_correct =
          identical(
            rownames(current_matrix),
            network_vars
          ) &&
          identical(
            colnames(current_matrix),
            network_vars
          ),
        Minimum_eigenvalue = min(
          eigen(
            current_matrix,
            symmetric = TRUE,
            only.values = TRUE
          )$values
        ),
        Input_SHA256 = digest::digest(
          current_matrix,
          algo = "sha256",
          serialize = TRUE
        ),
        stringsAsFactors = FALSE
      )
    }
  )
)

stage3a_input_checks

stopifnot(
  nrow(stage3a_input_checks) == 3L,
  all(stage3a_input_checks$Rows == 9L),
  all(stage3a_input_checks$Columns == 9L),
  all(stage3a_input_checks$All_finite),
  all(stage3a_input_checks$Symmetric),
  all(stage3a_input_checks$Diagonal_one),
  all(stage3a_input_checks$Node_order_correct),
  all(stage3a_input_checks$Minimum_eigenvalue > 0)
)
```

```{r stage3a-2004-point-network-estimation}
stage3a_settings <- data.frame(
  Election_year = 2004L,
  Sample_N = 789L,
  Correlation = "polychoric",
  Estimator = "EBICglasso",
  Gamma = 0.50,
  Nlambda = 100L,
  Lambda_min_ratio = 0.01,
  Refit = FALSE,
  Threshold = FALSE,
  Weight_N_treatment = "raw complete-case N",
  Purpose = "diagnostic point-network comparison only",
  stringsAsFactors = FALSE
)

stage3a_settings

run_stage3a_network <- function(
  correlation_matrix,
  candidate_name
) {
  model_run <- run_and_capture(
    function() {
      qgraph::EBICglasso(
        S = correlation_matrix,
        n = 789L,
        gamma = 0.50,
        nlambda = 100L,
        lambda.min.ratio = 0.01,
        refit = FALSE,
        threshold = FALSE
      )
    }
  )

  if (!is.null(model_run$value)) {
    dimnames(model_run$value) <- list(
      network_vars,
      network_vars
    )
  }

  list(
    Candidate = candidate_name,
    value = model_run$value,
    warnings = model_run$warnings,
    messages = model_run$messages,
    error = model_run$error
  )
}

stage3a_runs <- lapply(
  names(stage3a_input_matrices),
  function(candidate_name) {
    run_stage3a_network(
      correlation_matrix =
        stage3a_input_matrices[[candidate_name]],
      candidate_name = candidate_name
    )
  }
)

names(stage3a_runs) <-
  names(stage3a_input_matrices)

stage3a_conditions <- do.call(
  rbind,
  lapply(
    stage3a_runs,
    function(model_run) {
      data.frame(
        Election_year = 2004L,
        Candidate = model_run$Candidate,
        Warning_count =
          length(model_run$warnings),
        Warnings =
          collapse_conditions(model_run$warnings),
        Message_count =
          length(model_run$messages),
        Messages =
          collapse_conditions(model_run$messages),
        Error = if (
          is.na(model_run$error)
        ) {
          ""
        } else {
          model_run$error
        },
        stringsAsFactors = FALSE
      )
    }
  )
)

stage3a_conditions_file <- file.path(
  pilot_table_dir,
  "d15_stage3a_2004_conditions.csv"
)

utils::write.csv(
  stage3a_conditions,
  stage3a_conditions_file,
  row.names = FALSE
)

stage3a_success <- vapply(
  stage3a_runs,
  function(model_run) {
    is.matrix(model_run$value) &&
      identical(
        dim(model_run$value),
        c(9L, 9L)
      ) &&
      (
        is.na(model_run$error) ||
          model_run$error == ""
      )
  },
  FUN.VALUE = logical(1)
)

if (!all(stage3a_success)) {
  stop(
    paste(
      "Stage 3A network estimation failed.",
      "Inspect d15_stage3a_2004_conditions.csv."
    )
  )
}
```

```{r stage3a-2004-network-diagnostics}
extract_stage3a_edges <- function(network_matrix) {
  network_matrix[
    upper.tri(network_matrix)
  ]
}

stage3a_networks <- lapply(
  stage3a_runs,
  function(model_run) {
    model_run$value
  }
)

stage3a_network_diagnostics <- do.call(
  rbind,
  lapply(
    names(stage3a_networks),
    function(candidate_name) {
      current_network <-
        stage3a_networks[[candidate_name]]

      current_edges <-
        extract_stage3a_edges(
          current_network
        )

      retained <- abs(current_edges) > 1e-12

      current_conditions <-
        stage3a_conditions[
          stage3a_conditions$Candidate ==
            candidate_name,
          ,
          drop = FALSE
        ]

      data.frame(
        Election_year = 2004L,
        Candidate = candidate_name,
        Sample_N = 789L,
        Gamma = 0.50,
        Compared_edges = length(current_edges),
        Retained_edges = sum(retained),
        Density = sum(retained) /
          length(current_edges),
        Global_strength =
          sum(abs(current_edges)),
        Minimum_edge =
          min(current_edges),
        Maximum_edge =
          max(current_edges),
        All_values_finite =
          all(is.finite(current_network)),
        Maximum_asymmetry =
          max(
            abs(
              current_network -
                t(current_network)
            )
          ),
        Maximum_diagonal =
          max(abs(diag(current_network))),
        Warning_count =
          current_conditions$Warning_count,
        Error =
          current_conditions$Error,
        Network_SHA256 =
          digest::digest(
            current_network,
            algo = "sha256",
            serialize = TRUE
          ),
        stringsAsFactors = FALSE
      )
    }
  )
)

compare_stage3a_networks <- function(
  first_name,
  second_name
) {
  first_edges <- extract_stage3a_edges(
    stage3a_networks[[first_name]]
  )

  second_edges <- extract_stage3a_edges(
    stage3a_networks[[second_name]]
  )

  absolute_difference <-
    abs(first_edges - second_edges)

  first_retained <-
    abs(first_edges) > 1e-12

  second_retained <-
    abs(second_edges) > 1e-12

  retained_union <-
    first_retained | second_retained

  retained_intersection <-
    first_retained & second_retained

  sign_agreement_union <- if (
    sum(retained_union) > 0L
  ) {
    mean(
      sign(first_edges[retained_union]) ==
        sign(second_edges[retained_union])
    )
  } else {
    NA_real_
  }

  edge_jaccard <- if (
    sum(retained_union) > 0L
  ) {
    sum(retained_intersection) /
      sum(retained_union)
  } else {
    1
  }

  data.frame(
    Election_year = 2004L,
    First_network = first_name,
    Second_network = second_name,
    Compared_edges = length(first_edges),
    Edge_weight_correlation =
      stats::cor(
        first_edges,
        second_edges
      ),
    Median_absolute_edge_difference =
      stats::median(absolute_difference),
    Mean_absolute_edge_difference =
      mean(absolute_difference),
    Maximum_absolute_edge_difference =
      max(absolute_difference),
    Sign_agreement_all_pairs =
      mean(
        sign(first_edges) ==
          sign(second_edges)
      ),
    Sign_agreement_retained_union =
      sign_agreement_union,
    Sign_reversals_among_common_edges =
      sum(
        first_edges[retained_intersection] *
          second_edges[retained_intersection] <
          0
      ),
    First_retained_edges =
      sum(first_retained),
    Second_retained_edges =
      sum(second_retained),
    Common_retained_edges =
      sum(retained_intersection),
    Retained_edge_union =
      sum(retained_union),
    Retained_edge_Jaccard =
      edge_jaccard,
    First_global_strength =
      sum(abs(first_edges)),
    Second_global_strength =
      sum(abs(second_edges)),
    Global_strength_difference =
      sum(abs(second_edges)) -
      sum(abs(first_edges)),
    stringsAsFactors = FALSE
  )
}

stage3a_comparison_pairs <- list(
  c("U_RAWN", "W_CDF_RAWN"),
  c("U_RAWN", "W_POST_COMPAT_RAWN"),
  c("W_CDF_RAWN", "W_POST_COMPAT_RAWN")
)

stage3a_network_comparison <- do.call(
  rbind,
  lapply(
    stage3a_comparison_pairs,
    function(comparison_pair) {
      compare_stage3a_networks(
        first_name = comparison_pair[1],
        second_name = comparison_pair[2]
      )
    }
  )
)

stage07_reference_graph <- as.matrix(
  stage07_pilot$graph
)

stage07_reproduction_difference <- max(
  abs(
    stage3a_networks[["U_RAWN"]] -
      stage07_reference_graph
  )
)

stage3a_network_diagnostics
stage3a_network_comparison
stage07_reproduction_difference
```

```{r stage3a-2004-validation-and-save}
stage3a_warning_types <- unique(
  stage3a_conditions$Warnings[
    nzchar(stage3a_conditions$Warnings)
  ]
)

stage3a_only_expected_dense_warning <-
  all(
    stage3a_conditions$Warning_count == 1L
  ) &&
  length(stage3a_warning_types) == 1L &&
  isTRUE(
    grepl(
      "dense regularized network",
      tolower(stage3a_warning_types),
      fixed = TRUE
    )
  )

stage3a_validation <- data.frame(
  Check = c(
    "Stage 2C validation passed before Stage 3A",
    "Three 2004 input matrices passed validation",
    "Three point networks were returned",
    "All point-network entries are finite",
    "All point networks are symmetric",
    "All point-network diagonals equal zero",
    "Each point network contains 36 possible edges",
    "No estimation error was captured",
    "Only the documented common dense-network warning was captured",
    "Three prespecified network comparisons were produced",
    "Unweighted network reproduces the existing Step 07 pilot",
    "No bootstrap centrality NCT or political interpretation was run",
    "No final weight candidate was selected"
  ),
  Passed = c(
    all(
      as.logical(
        stage2c_validation_disk$Passed
      )
    ),
    nrow(stage3a_input_checks) == 3L &&
      all(stage3a_input_checks$All_finite) &&
      all(stage3a_input_checks$Symmetric) &&
      all(stage3a_input_checks$Diagonal_one) &&
      all(stage3a_input_checks$Node_order_correct),
    length(stage3a_networks) == 3L &&
      all(stage3a_success),
    all(
      stage3a_network_diagnostics$
        All_values_finite
    ),
    all(
      stage3a_network_diagnostics$
        Maximum_asymmetry <= 1e-10
    ),
    all(
      stage3a_network_diagnostics$
        Maximum_diagonal <= 1e-10
    ),
    all(
      stage3a_network_diagnostics$
        Compared_edges == 36L
    ),
    all(
      stage3a_conditions$Error == ""
    ),
    stage3a_only_expected_dense_warning,
    nrow(stage3a_network_comparison) == 3L &&
      all(
        stage3a_network_comparison$
          Compared_edges == 36L
      ),
    stage07_reproduction_difference <= 1e-10,
    TRUE,
    TRUE
  ),
  stringsAsFactors = FALSE
)

stage3a_validation

stage3a_diagnostics_file <- file.path(
  pilot_table_dir,
  "d15_stage3a_2004_point_network_diagnostics.csv"
)

stage3a_comparison_file <- file.path(
  pilot_table_dir,
  "d15_stage3a_2004_point_network_comparison.csv"
)

stage3a_validation_file <- file.path(
  pilot_table_dir,
  "d15_stage3a_2004_validation.csv"
)

stage3a_model_file <- file.path(
  stage2c_model_dir,
  "d15_stage3a_2004_point_networks.rds"
)

stage3a_log_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage3a_2004.log"
)

stage3a_session_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage3a_2004_sessionInfo.txt"
)

utils::write.csv(
  stage3a_network_diagnostics,
  stage3a_diagnostics_file,
  row.names = FALSE
)

utils::write.csv(
  stage3a_network_comparison,
  stage3a_comparison_file,
  row.names = FALSE
)

utils::write.csv(
  stage3a_validation,
  stage3a_validation_file,
  row.names = FALSE
)

saveRDS(
  list(
    Settings = stage3a_settings,
    Input_checks = stage3a_input_checks,
    Input_matrices = stage3a_input_matrices,
    Point_networks = stage3a_networks,
    Conditions = stage3a_conditions,
    Diagnostics = stage3a_network_diagnostics,
    Comparisons = stage3a_network_comparison,
    Step07_reproduction_difference =
      stage07_reproduction_difference
  ),
  stage3a_model_file
)

writeLines(
  capture.output(
    utils::sessionInfo()
  ),
  stage3a_session_file
)

stage3a_warning_classification <- if (
  stage3a_only_expected_dense_warning
) {
  "EXPECTED_COMMON_DENSE_NETWORK_WARNING"
} else {
  "UNEXPECTED_OR_INCONSISTENT_WARNING"
}

writeLines(
  c(
    "D15 STAGE 3A — 2004 DIAGNOSTIC POINT-NETWORK PILOT",
    paste0(
      "Timestamp: ",
      format(
        Sys.time(),
        "%Y-%m-%d %H:%M:%S %z"
      )
    ),
    "Networks requested: 3",
    paste0(
      "Networks returned: ",
      sum(stage3a_success),
      "/3"
    ),
    paste0(
      "Warnings captured: ",
      sum(
        stage3a_conditions$
          Warning_count
      )
    ),
    paste0(
      "Warning classification: ",
      stage3a_warning_classification
    ),
    paste0(
      "Errors captured: ",
      sum(
        stage3a_conditions$
          Error != ""
      )
    ),
    paste0(
      "Validation checks passed: ",
      sum(stage3a_validation$Passed),
      "/",
      nrow(stage3a_validation)
    ),
    paste0(
      "Maximum difference from existing Step 07 unweighted network: ",
      format(
        stage07_reproduction_difference,
        scientific = TRUE
      )
    ),
    "No bootstrap, centrality, NCT or political interpretation was run.",
    "No final weight candidate was selected."
  ),
  stage3a_log_file
)

stage3a_output_files <- c(
  stage3a_diagnostics_file,
  stage3a_comparison_file,
  stage3a_conditions_file,
  stage3a_validation_file,
  stage3a_model_file,
  stage3a_log_file,
  stage3a_session_file
)

stopifnot(
  all(
    file.exists(stage3a_output_files)
  )
)

data.frame(
  Output = stage3a_output_files,
  Exists = file.exists(stage3a_output_files),
  stringsAsFactors = FALSE
)

if (!all(stage3a_validation$Passed)) {
  stop(
    paste(
      "Stage 3A completed with a failed validation check.",
      "Inspect the saved validation, conditions and log files."
    )
  )
}
```
## Stage 3B — Five-wave diagnostic point-network extension

Stage 3A passed independent review. This stage applies the identical
diagnostic EBICglasso specification to the five approved overall waves.
It compares unweighted, W_CDF and W_POST_COMPAT point networks without
selecting a final weight candidate. No bootstrap, centrality, NCT,
education-group analysis or political interpretation is performed.

```{r stage3b-five-wave-input-gate}
stage3b_required_packages <- c(
  "qgraph",
  "digest"
)

stage3b_package_gate <- data.frame(
  Package = stage3b_required_packages,
  Available = vapply(
    stage3b_required_packages,
    requireNamespace,
    quietly = TRUE,
    FUN.VALUE = logical(1)
  ),
  Version = vapply(
    stage3b_required_packages,
    function(package_name) {
      if (requireNamespace(package_name, quietly = TRUE)) {
        as.character(
          utils::packageVersion(package_name)
        )
      } else {
        NA_character_
      }
    },
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

stage3b_package_gate

stopifnot(
  all(stage3b_package_gate$Available)
)

stage3b_years <- c(
  2004L,
  2012L,
  2016L,
  2020L,
  2024L
)

stage3b_candidates <- c(
  "U_RAWN",
  "W_CDF_RAWN",
  "W_POST_COMPAT_RAWN"
)

stage3a_validation_path <- file.path(
  pilot_table_dir,
  "d15_stage3a_2004_validation.csv"
)

stage3a_model_path <- file.path(
  stage2c_model_dir,
  "d15_stage3a_2004_point_networks.rds"
)

stage2c_matrix_path <- file.path(
  stage2c_model_dir,
  "d15_stage2c_five_wave_weighted_polychoric_matrices.rds"
)

unweighted_matrix_path <- file.path(
  project_root,
  "data",
  "audit",
  paste0(
    "anes_cdf_20260205_nine_node_complete_case_",
    "polychoric_diagnostics_2004_2012_2016_2020_2024.rds"
  )
)

stage3b_required_files <- c(
  stage3a_validation_path,
  stage3a_model_path,
  stage2c_matrix_path,
  unweighted_matrix_path
)

stopifnot(
  all(file.exists(stage3b_required_files))
)

stage3a_validation_disk <- utils::read.csv(
  stage3a_validation_path,
  stringsAsFactors = FALSE
)

stopifnot(
  nrow(stage3a_validation_disk) == 13L,
  all(as.logical(stage3a_validation_disk$Passed))
)

stage3a_saved <- readRDS(
  stage3a_model_path
)

stage2c_saved <- readRDS(
  stage2c_matrix_path
)

unweighted_bundle <- readRDS(
  unweighted_matrix_path
)

stopifnot(
  all(
    as.character(stage3b_years) %in%
      names(unweighted_bundle$Matrices)
  ),
  all(
    paste0(stage3b_years, "__W_CDF") %in%
      names(stage2c_saved)
  ),
  all(
    paste0(stage3b_years, "__W_POST_COMPAT") %in%
      names(stage2c_saved)
  ),
  all(
    stage3b_candidates %in%
      names(stage3a_saved$Point_networks)
  )
)

stage3b_input_map <- do.call(
  rbind,
  lapply(
    stage3b_years,
    function(year_value) {
      data.frame(
        Election_year = year_value,
        Candidate = stage3b_candidates,
        stringsAsFactors = FALSE
      )
    }
  )
)

stage3b_input_map$Sample_N <- as.integer(
  unname(
    expected_year_counts[
      as.character(
        stage3b_input_map$Election_year
      )
    ]
  )
)

stage3b_input_map$Key <- paste(
  stage3b_input_map$Election_year,
  stage3b_input_map$Candidate,
  sep = "__"
)

stopifnot(
  !anyNA(stage3b_input_map$Sample_N),
  nrow(stage3b_input_map) == 15L,
  !anyDuplicated(stage3b_input_map$Key)
)

stage3b_input_matrices <- setNames(
  lapply(
    seq_len(nrow(stage3b_input_map)),
    function(row_number) {
      year_value <-
        stage3b_input_map$Election_year[row_number]

      candidate_name <-
        stage3b_input_map$Candidate[row_number]

      if (candidate_name == "U_RAWN") {
        matrix_object <- unweighted_bundle$
          Matrices[[as.character(year_value)]]
      } else if (candidate_name == "W_CDF_RAWN") {
        matrix_object <- stage2c_saved[[
          paste0(year_value, "__W_CDF")
        ]]$lavaan
      } else if (
        candidate_name == "W_POST_COMPAT_RAWN"
      ) {
        matrix_object <- stage2c_saved[[
          paste0(year_value, "__W_POST_COMPAT")
        ]]$lavaan
      } else {
        stop(
          paste(
            "Unexpected Stage 3B candidate:",
            candidate_name
          )
        )
      }

      as.matrix(matrix_object)
    }
  ),
  stage3b_input_map$Key
)

stage3b_input_checks <- do.call(
  rbind,
  lapply(
    seq_len(nrow(stage3b_input_map)),
    function(row_number) {
      current_key <-
        stage3b_input_map$Key[row_number]

      current_matrix <-
        stage3b_input_matrices[[current_key]]

      data.frame(
        Election_year =
          stage3b_input_map$
            Election_year[row_number],
        Candidate =
          stage3b_input_map$
            Candidate[row_number],
        Key = current_key,
        Sample_N =
          stage3b_input_map$
            Sample_N[row_number],
        Rows = nrow(current_matrix),
        Columns = ncol(current_matrix),
        All_finite =
          all(is.finite(current_matrix)),
        Symmetric = isSymmetric(
          current_matrix,
          tol = 1e-10
        ),
        Diagonal_one =
          max(
            abs(diag(current_matrix) - 1)
          ) <= 1e-10,
        Node_order_correct =
          identical(
            rownames(current_matrix),
            network_vars
          ) &&
          identical(
            colnames(current_matrix),
            network_vars
          ),
        Minimum_eigenvalue = min(
          eigen(
            current_matrix,
            symmetric = TRUE,
            only.values = TRUE
          )$values
        ),
        Input_SHA256 = digest::digest(
          current_matrix,
          algo = "sha256",
          serialize = TRUE
        ),
        stringsAsFactors = FALSE
      )
    }
  )
)

stage3b_input_checks

stopifnot(
  nrow(stage3b_input_checks) == 15L,
  all(stage3b_input_checks$Rows == 9L),
  all(stage3b_input_checks$Columns == 9L),
  all(stage3b_input_checks$All_finite),
  all(stage3b_input_checks$Symmetric),
  all(stage3b_input_checks$Diagonal_one),
  all(stage3b_input_checks$Node_order_correct),
  all(stage3b_input_checks$Minimum_eigenvalue > 0)
)

stage3b_settings <- data.frame(
  Election_year =
    stage3b_input_map$Election_year,
  Candidate =
    stage3b_input_map$Candidate,
  Sample_N =
    stage3b_input_map$Sample_N,
  Correlation = "polychoric",
  Estimator = "EBICglasso",
  Gamma = 0.50,
  Nlambda = 100L,
  Lambda_min_ratio = 0.01,
  Refit = FALSE,
  Threshold = FALSE,
  Weight_N_treatment =
    "raw complete-case N",
  Purpose =
    "diagnostic point-network comparison only",
  stringsAsFactors = FALSE
)

stage3b_settings
```

```{r stage3b-five-wave-point-network-estimation}
run_stage3b_network <- function(
  correlation_matrix,
  year_value,
  candidate_name,
  sample_n
) {
  model_run <- run_and_capture(
    function() {
      qgraph::EBICglasso(
        S = correlation_matrix,
        n = sample_n,
        gamma = 0.50,
        nlambda = 100L,
        lambda.min.ratio = 0.01,
        refit = FALSE,
        threshold = FALSE
      )
    }
  )

  if (!is.null(model_run$value)) {
    dimnames(model_run$value) <- list(
      network_vars,
      network_vars
    )
  }

  list(
    Election_year = year_value,
    Candidate = candidate_name,
    Sample_N = sample_n,
    value = model_run$value,
    warnings = model_run$warnings,
    messages = model_run$messages,
    error = model_run$error
  )
}

stage3b_runs <- setNames(
  lapply(
    seq_len(nrow(stage3b_input_map)),
    function(row_number) {
      current_key <-
        stage3b_input_map$Key[row_number]

      run_stage3b_network(
        correlation_matrix =
          stage3b_input_matrices[[current_key]],
        year_value =
          stage3b_input_map$
            Election_year[row_number],
        candidate_name =
          stage3b_input_map$
            Candidate[row_number],
        sample_n =
          stage3b_input_map$
            Sample_N[row_number]
      )
    }
  ),
  stage3b_input_map$Key
)

stage3b_conditions <- do.call(
  rbind,
  lapply(
    stage3b_runs,
    function(model_run) {
      data.frame(
        Election_year =
          model_run$Election_year,
        Candidate =
          model_run$Candidate,
        Sample_N =
          model_run$Sample_N,
        Warning_count =
          length(model_run$warnings),
        Warnings =
          collapse_conditions(
            model_run$warnings
          ),
        Message_count =
          length(model_run$messages),
        Messages =
          collapse_conditions(
            model_run$messages
          ),
        Error = if (
          is.na(model_run$error)
        ) {
          ""
        } else {
          model_run$error
        },
        stringsAsFactors = FALSE
      )
    }
  )
)

stage3b_conditions_file <- file.path(
  pilot_table_dir,
  "d15_stage3b_five_wave_conditions.csv"
)

utils::write.csv(
  stage3b_conditions,
  stage3b_conditions_file,
  row.names = FALSE
)

stage3b_success <- vapply(
  stage3b_runs,
  function(model_run) {
    is.matrix(model_run$value) &&
      identical(
        dim(model_run$value),
        c(9L, 9L)
      ) &&
      (
        is.na(model_run$error) ||
          model_run$error == ""
      )
  },
  FUN.VALUE = logical(1)
)

if (!all(stage3b_success)) {
  stop(
    paste(
      "Stage 3B network estimation failed.",
      "Inspect d15_stage3b_five_wave_conditions.csv."
    )
  )
}

stage3b_networks <- lapply(
  stage3b_runs,
  function(model_run) {
    model_run$value
  }
)

all_stage3b_warning_messages <- unlist(
  lapply(
    stage3b_runs,
    function(model_run) {
      model_run$warnings
    }
  ),
  use.names = FALSE
)

stage3b_only_expected_warnings <- if (
  length(all_stage3b_warning_messages) == 0L
) {
  TRUE
} else {
  all(
    grepl(
      "dense regularized network",
      tolower(all_stage3b_warning_messages),
      fixed = TRUE
    )
  )
}

stage3b_warning_classification <- if (
  length(all_stage3b_warning_messages) == 0L
) {
  "NO_WARNINGS"
} else if (
  stage3b_only_expected_warnings
) {
  "EXPECTED_DENSE_NETWORK_WARNINGS_ONLY"
} else {
  "UNEXPECTED_WARNING_PRESENT"
}

stage3b_conditions
stage3b_warning_classification
```

```{r stage3b-five-wave-network-diagnostics}
extract_stage3b_edges <- function(network_matrix) {
  network_matrix[
    upper.tri(network_matrix)
  ]
}

stage3b_network_diagnostics <- do.call(
  rbind,
  lapply(
    seq_len(nrow(stage3b_input_map)),
    function(row_number) {
      current_key <-
        stage3b_input_map$Key[row_number]

      current_network <-
        stage3b_networks[[current_key]]

      current_edges <-
        extract_stage3b_edges(
          current_network
        )

      retained_edges <-
        abs(current_edges) > 1e-12

      current_conditions <-
        stage3b_conditions[
          stage3b_conditions$Election_year ==
            stage3b_input_map$
              Election_year[row_number] &
            stage3b_conditions$Candidate ==
              stage3b_input_map$
                Candidate[row_number],
          ,
          drop = FALSE
        ]

      current_input_check <-
        stage3b_input_checks[
          stage3b_input_checks$Key ==
            current_key,
          ,
          drop = FALSE
        ]

      data.frame(
        Election_year =
          stage3b_input_map$
            Election_year[row_number],
        Candidate =
          stage3b_input_map$
            Candidate[row_number],
        Sample_N =
          stage3b_input_map$
            Sample_N[row_number],
        Gamma = 0.50,
        Compared_edges =
          length(current_edges),
        Retained_edges =
          sum(retained_edges),
        Density =
          sum(retained_edges) /
          length(current_edges),
        Global_strength =
          sum(abs(current_edges)),
        Minimum_edge =
          min(current_edges),
        Maximum_edge =
          max(current_edges),
        All_values_finite =
          all(is.finite(current_network)),
        Maximum_asymmetry =
          max(
            abs(
              current_network -
                t(current_network)
            )
          ),
        Maximum_diagonal =
          max(abs(diag(current_network))),
        Input_minimum_eigenvalue =
          current_input_check$
            Minimum_eigenvalue,
        Warning_count =
          current_conditions$
            Warning_count,
        Error =
          current_conditions$Error,
        Input_SHA256 =
          current_input_check$
            Input_SHA256,
        Network_SHA256 =
          digest::digest(
            current_network,
            algo = "sha256",
            serialize = TRUE
          ),
        stringsAsFactors = FALSE
      )
    }
  )
)

compare_stage3b_networks <- function(
  year_value,
  first_name,
  second_name
) {
  first_key <- paste(
    year_value,
    first_name,
    sep = "__"
  )

  second_key <- paste(
    year_value,
    second_name,
    sep = "__"
  )

  first_edges <- extract_stage3b_edges(
    stage3b_networks[[first_key]]
  )

  second_edges <- extract_stage3b_edges(
    stage3b_networks[[second_key]]
  )

  absolute_difference <-
    abs(first_edges - second_edges)

  first_retained <-
    abs(first_edges) > 1e-12

  second_retained <-
    abs(second_edges) > 1e-12

  retained_union <-
    first_retained | second_retained

  retained_intersection <-
    first_retained & second_retained

  sign_agreement_union <- if (
    sum(retained_union) > 0L
  ) {
    mean(
      sign(first_edges[retained_union]) ==
        sign(second_edges[retained_union])
    )
  } else {
    NA_real_
  }

  edge_jaccard <- if (
    sum(retained_union) > 0L
  ) {
    sum(retained_intersection) /
      sum(retained_union)
  } else {
    1
  }

  data.frame(
    Election_year = year_value,
    First_network = first_name,
    Second_network = second_name,
    Compared_edges =
      length(first_edges),
    Edge_weight_correlation =
      stats::cor(
        first_edges,
        second_edges
      ),
    Median_absolute_edge_difference =
      stats::median(
        absolute_difference
      ),
    Mean_absolute_edge_difference =
      mean(absolute_difference),
    Maximum_absolute_edge_difference =
      max(absolute_difference),
    Sign_agreement_all_pairs =
      mean(
        sign(first_edges) ==
          sign(second_edges)
      ),
    Sign_agreement_retained_union =
      sign_agreement_union,
    Sign_reversals_among_common_edges =
      sum(
        first_edges[retained_intersection] *
          second_edges[retained_intersection] <
          0
      ),
    First_retained_edges =
      sum(first_retained),
    Second_retained_edges =
      sum(second_retained),
    Common_retained_edges =
      sum(retained_intersection),
    Retained_edge_union =
      sum(retained_union),
    Retained_edge_Jaccard =
      edge_jaccard,
    First_global_strength =
      sum(abs(first_edges)),
    Second_global_strength =
      sum(abs(second_edges)),
    Global_strength_difference =
      sum(abs(second_edges)) -
      sum(abs(first_edges)),
    stringsAsFactors = FALSE
  )
}

stage3b_comparison_design <- do.call(
  rbind,
  lapply(
    stage3b_years,
    function(year_value) {
      data.frame(
        Election_year = year_value,
        First_network = c(
          "U_RAWN",
          "U_RAWN",
          "W_CDF_RAWN"
        ),
        Second_network = c(
          "W_CDF_RAWN",
          "W_POST_COMPAT_RAWN",
          "W_POST_COMPAT_RAWN"
        ),
        stringsAsFactors = FALSE
      )
    }
  )
)

stage3b_network_comparison <- do.call(
  rbind,
  lapply(
    seq_len(
      nrow(stage3b_comparison_design)
    ),
    function(row_number) {
      compare_stage3b_networks(
        year_value =
          stage3b_comparison_design$
            Election_year[row_number],
        first_name =
          stage3b_comparison_design$
            First_network[row_number],
        second_name =
          stage3b_comparison_design$
            Second_network[row_number]
      )
    }
  )
)

stage3b_2004_reproduction <- data.frame(
  Candidate = stage3b_candidates,
  Maximum_absolute_difference = vapply(
    stage3b_candidates,
    function(candidate_name) {
      max(
        abs(
          stage3b_networks[[
            paste(
              2004L,
              candidate_name,
              sep = "__"
            )
          ]] -
            stage3a_saved$
              Point_networks[[
                candidate_name
              ]]
        )
      )
    },
    FUN.VALUE = numeric(1)
  ),
  stringsAsFactors = FALSE
)

stage3b_maximum_2004_reproduction_difference <-
  max(
    stage3b_2004_reproduction$
      Maximum_absolute_difference
  )

stage3b_network_diagnostics
stage3b_network_comparison
stage3b_2004_reproduction
stage3b_warning_classification
```

```{r stage3b-five-wave-validation-and-save}
stage3b_validation <- data.frame(
  Check = c(
    "Stage 3A validation passed before Stage 3B",
    "Fifteen five-wave input matrices passed validation",
    "Fifteen point networks were returned",
    "All point-network entries are finite",
    "All point networks are symmetric",
    "All point-network diagonals equal zero",
    "Each point network contains 36 possible edges",
    "No estimation error was captured",
    "No unexpected estimation warning was captured",
    "Fifteen network comparisons were produced",
    "The 2004 networks reproduce the Stage 3A results",
    "The EBICglasso specification was identical across all runs",
    "No bootstrap centrality NCT or political interpretation was run",
    "No final weight candidate was selected"
  ),
  Passed = c(
    all(
      as.logical(
        stage3a_validation_disk$Passed
      )
    ),
    nrow(stage3b_input_checks) == 15L &&
      all(stage3b_input_checks$All_finite) &&
      all(stage3b_input_checks$Symmetric) &&
      all(stage3b_input_checks$Diagonal_one) &&
      all(stage3b_input_checks$Node_order_correct) &&
      all(
        stage3b_input_checks$
          Minimum_eigenvalue > 0
      ),
    length(stage3b_networks) == 15L &&
      all(stage3b_success),
    all(
      stage3b_network_diagnostics$
        All_values_finite
    ),
    all(
      stage3b_network_diagnostics$
        Maximum_asymmetry <= 1e-10
    ),
    all(
      stage3b_network_diagnostics$
        Maximum_diagonal <= 1e-10
    ),
    all(
      stage3b_network_diagnostics$
        Compared_edges == 36L
    ),
    all(stage3b_conditions$Error == ""),
    stage3b_only_expected_warnings,
    nrow(stage3b_network_comparison) == 15L &&
      all(
        stage3b_network_comparison$
          Compared_edges == 36L
      ),
    stage3b_maximum_2004_reproduction_difference <=
      1e-10,
    all(stage3b_settings$Gamma == 0.50) &&
      all(stage3b_settings$Nlambda == 100L) &&
      all(
        stage3b_settings$
          Lambda_min_ratio == 0.01
      ) &&
      all(!stage3b_settings$Refit) &&
      all(!stage3b_settings$Threshold),
    TRUE,
    TRUE
  ),
  stringsAsFactors = FALSE
)

stage3b_validation

stage3b_diagnostics_file <- file.path(
  pilot_table_dir,
  "d15_stage3b_five_wave_point_network_diagnostics.csv"
)

stage3b_comparison_file <- file.path(
  pilot_table_dir,
  "d15_stage3b_five_wave_point_network_comparison.csv"
)

stage3b_validation_file <- file.path(
  pilot_table_dir,
  "d15_stage3b_five_wave_validation.csv"
)

stage3b_model_file <- file.path(
  stage2c_model_dir,
  "d15_stage3b_five_wave_point_networks.rds"
)

stage3b_log_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage3b_five_wave.log"
)

stage3b_session_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage3b_five_wave_sessionInfo.txt"
)

utils::write.csv(
  stage3b_network_diagnostics,
  stage3b_diagnostics_file,
  row.names = FALSE
)

utils::write.csv(
  stage3b_network_comparison,
  stage3b_comparison_file,
  row.names = FALSE
)

utils::write.csv(
  stage3b_validation,
  stage3b_validation_file,
  row.names = FALSE
)

saveRDS(
  list(
    Settings = stage3b_settings,
    Input_checks = stage3b_input_checks,
    Input_matrices = stage3b_input_matrices,
    Point_networks = stage3b_networks,
    Conditions = stage3b_conditions,
    Diagnostics = stage3b_network_diagnostics,
    Comparisons = stage3b_network_comparison,
    Stage3A_reproduction =
      stage3b_2004_reproduction,
    Warning_classification =
      stage3b_warning_classification
  ),
  stage3b_model_file
)

writeLines(
  capture.output(
    utils::sessionInfo()
  ),
  stage3b_session_file
)

writeLines(
  c(
    "D15 STAGE 3B — FIVE-WAVE DIAGNOSTIC POINT-NETWORK EXTENSION",
    paste0(
      "Timestamp: ",
      format(
        Sys.time(),
        "%Y-%m-%d %H:%M:%S %z"
      )
    ),
    "Networks requested: 15",
    paste0(
      "Networks returned: ",
      sum(stage3b_success),
      "/15"
    ),
    paste0(
      "Warnings captured: ",
      sum(
        stage3b_conditions$
          Warning_count
      )
    ),
    paste0(
      "Warning classification: ",
      stage3b_warning_classification
    ),
    paste0(
      "Errors captured: ",
      sum(
        stage3b_conditions$
          Error != ""
      )
    ),
    paste0(
      "Validation checks passed: ",
      sum(stage3b_validation$Passed),
      "/",
      nrow(stage3b_validation)
    ),
    paste0(
      "Maximum 2004 reproduction difference: ",
      format(
        stage3b_maximum_2004_reproduction_difference,
        scientific = TRUE
      )
    ),
    "No bootstrap, centrality, NCT or political interpretation was run.",
    "No final weight candidate was selected."
  ),
  stage3b_log_file
)

stage3b_output_files <- c(
  stage3b_diagnostics_file,
  stage3b_comparison_file,
  stage3b_conditions_file,
  stage3b_validation_file,
  stage3b_model_file,
  stage3b_log_file,
  stage3b_session_file
)

stopifnot(
  all(
    file.exists(stage3b_output_files)
  )
)

data.frame(
  Output = stage3b_output_files,
  Exists = file.exists(stage3b_output_files),
  stringsAsFactors = FALSE
)

if (!all(stage3b_validation$Passed)) {
  stop(
    paste(
      "Stage 3B completed with a failed validation check.",
      "Inspect the saved validation, conditions and log files."
    )
  )
}
```

## Stage 3C — 2020 W_POST_COMPAT lambda-boundary diagnostic

This bounded diagnostic investigates the lowest-lambda boundary message
observed for the 2020 W_POST_COMPAT network. It compares the registered
lambda grid with a one-decade extension while keeping gamma and all other
estimation settings fixed. The extended grid is not a new primary model
and will not be selected automatically.

```{r stage3c-input-and-specification-gate}
stage3b_validation_path <- file.path(
  pilot_table_dir,
  "d15_stage3b_five_wave_validation.csv"
)

stage3b_conditions_path <- file.path(
  pilot_table_dir,
  "d15_stage3b_five_wave_conditions.csv"
)

stage3b_model_path <- file.path(
  stage2c_model_dir,
  "d15_stage3b_five_wave_point_networks.rds"
)

stage3c_required_files <- c(
  stage3b_validation_path,
  stage3b_conditions_path,
  stage3b_model_path
)

stopifnot(
  all(file.exists(stage3c_required_files)),
  requireNamespace("qgraph", quietly = TRUE),
  requireNamespace("digest", quietly = TRUE)
)

stage3b_validation_disk <- utils::read.csv(
  stage3b_validation_path,
  stringsAsFactors = FALSE
)

stage3b_conditions_disk <- utils::read.csv(
  stage3b_conditions_path,
  stringsAsFactors = FALSE
)

stage3b_saved <- readRDS(
  stage3b_model_path
)

stopifnot(
  nrow(stage3b_validation_disk) == 14L,
  all(as.logical(stage3b_validation_disk$Passed)),
  "2020__W_POST_COMPAT_RAWN" %in%
    names(stage3b_saved$Input_matrices),
  "2020__W_POST_COMPAT_RAWN" %in%
    names(stage3b_saved$Point_networks)
)

stage3c_previous_condition <-
  stage3b_conditions_disk[
    stage3b_conditions_disk$Election_year == 2020L &
      stage3b_conditions_disk$Candidate ==
        "W_POST_COMPAT_RAWN",
    ,
    drop = FALSE
  ]

stage3c_previous_boundary_message <-
  nrow(stage3c_previous_condition) == 1L &&
  stage3c_previous_condition$Message_count == 1L &&
  grepl(
    "lowest lambda selected",
    tolower(stage3c_previous_condition$Messages),
    fixed = TRUE
  )

stopifnot(
  stage3c_previous_boundary_message
)

stage3c_input_matrix <- as.matrix(
  stage3b_saved$Input_matrices[[
    "2020__W_POST_COMPAT_RAWN"
  ]]
)

stage3c_reference_network <- as.matrix(
  stage3b_saved$Point_networks[[
    "2020__W_POST_COMPAT_RAWN"
  ]]
)

stage3c_input_gate <- data.frame(
  Check = c(
    "Input matrix is 9 by 9",
    "All input values are finite",
    "Input matrix is symmetric",
    "Input diagonal equals one",
    "Node order is correct",
    "Input matrix is positive definite",
    "Reference network is 9 by 9",
    "Previous lowest-lambda message is documented"
  ),
  Passed = c(
    identical(
      dim(stage3c_input_matrix),
      c(9L, 9L)
    ),
    all(is.finite(stage3c_input_matrix)),
    isSymmetric(
      stage3c_input_matrix,
      tol = 1e-10
    ),
    max(
      abs(
        diag(stage3c_input_matrix) - 1
      )
    ) <= 1e-10,
    identical(
      rownames(stage3c_input_matrix),
      network_vars
    ) &&
      identical(
        colnames(stage3c_input_matrix),
        network_vars
      ),
    min(
      eigen(
        stage3c_input_matrix,
        symmetric = TRUE,
        only.values = TRUE
      )$values
    ) > 0,
    identical(
      dim(stage3c_reference_network),
      c(9L, 9L)
    ),
    stage3c_previous_boundary_message
  ),
  stringsAsFactors = FALSE
)

stage3c_input_gate

stopifnot(
  all(stage3c_input_gate$Passed)
)

stage3c_specifications <- data.frame(
  Specification = c(
    "CURRENT_GRID",
    "EXTENDED_GRID"
  ),
  Lambda_min_ratio = c(
    0.01,
    0.001
  ),
  Nlambda = c(
    100L,
    150L
  ),
  Gamma = c(
    0.50,
    0.50
  ),
  Sample_N = c(
    5390L,
    5390L
  ),
  Refit = c(
    FALSE,
    FALSE
  ),
  Threshold = c(
    FALSE,
    FALSE
  ),
  Purpose = c(
    "Reproduce registered Stage 3B specification",
    "One-decade diagnostic lambda-path extension"
  ),
  stringsAsFactors = FALSE
)

stage3c_specifications
```

```{r stage3c-run-lambda-paths}
run_stage3c_path <- function(
  specification_name,
  lambda_min_ratio,
  nlambda_value
) {
  path_run <- run_and_capture(
    function() {
      qgraph::EBICglasso(
        S = stage3c_input_matrix,
        n = 5390L,
        gamma = 0.50,
        nlambda = nlambda_value,
        lambda.min.ratio =
          lambda_min_ratio,
        refit = FALSE,
        threshold = FALSE,
        returnAllResults = TRUE
      )
    }
  )

  list(
    Specification =
      specification_name,
    Lambda_min_ratio =
      lambda_min_ratio,
    Nlambda =
      nlambda_value,
    value =
      path_run$value,
    warnings =
      path_run$warnings,
    messages =
      path_run$messages,
    error =
      path_run$error
  )
}

stage3c_path_runs <- setNames(
  lapply(
    seq_len(
      nrow(stage3c_specifications)
    ),
    function(row_number) {
      run_stage3c_path(
        specification_name =
          stage3c_specifications$
            Specification[row_number],
        lambda_min_ratio =
          stage3c_specifications$
            Lambda_min_ratio[row_number],
        nlambda_value =
          stage3c_specifications$
            Nlambda[row_number]
      )
    }
  ),
  stage3c_specifications$Specification
)

stage3c_conditions <- do.call(
  rbind,
  lapply(
    stage3c_path_runs,
    function(path_run) {
      data.frame(
        Election_year = 2020L,
        Candidate =
          "W_POST_COMPAT_RAWN",
        Specification =
          path_run$Specification,
        Lambda_min_ratio =
          path_run$Lambda_min_ratio,
        Nlambda =
          path_run$Nlambda,
        Warning_count =
          length(path_run$warnings),
        Warnings =
          collapse_conditions(
            path_run$warnings
          ),
        Message_count =
          length(path_run$messages),
        Messages =
          collapse_conditions(
            path_run$messages
          ),
        Error = if (
          is.na(path_run$error)
        ) {
          ""
        } else {
          path_run$error
        },
        stringsAsFactors = FALSE
      )
    }
  )
)

stage3c_conditions_file <- file.path(
  pilot_table_dir,
  "d15_stage3c_2020_wpost_conditions.csv"
)

utils::write.csv(
  stage3c_conditions,
  stage3c_conditions_file,
  row.names = FALSE
)

stage3c_success <- vapply(
  stage3c_path_runs,
  function(path_run) {
    !is.null(path_run$value) &&
      is.list(path_run$value) &&
      all(
        c(
          "ebic",
          "lambda",
          "optnet"
        ) %in%
          names(path_run$value)
      ) &&
      (
        is.na(path_run$error) ||
          path_run$error == ""
      )
  },
  FUN.VALUE = logical(1)
)

if (!all(stage3c_success)) {
  stop(
    paste(
      "Stage 3C lambda-path estimation failed.",
      "Inspect d15_stage3c_2020_wpost_conditions.csv."
    )
  )
}

stage3c_selected_networks <- setNames(
  lapply(
    stage3c_path_runs,
    function(path_run) {
      selected_network <- as.matrix(
        path_run$value$optnet
      )

      dimnames(selected_network) <- list(
        network_vars,
        network_vars
      )

      selected_network
    }
  ),
  names(stage3c_path_runs)
)

all_stage3c_warning_messages <- unlist(
  lapply(
    stage3c_path_runs,
    function(path_run) {
      path_run$warnings
    }
  ),
  use.names = FALSE
)

all_stage3c_ordinary_messages <- unlist(
  lapply(
    stage3c_path_runs,
    function(path_run) {
      path_run$messages
    }
  ),
  use.names = FALSE
)

stage3c_only_expected_warnings <- if (
  length(all_stage3c_warning_messages) == 0L
) {
  TRUE
} else {
  all(
    grepl(
      "dense regularized network",
      tolower(
        all_stage3c_warning_messages
      ),
      fixed = TRUE
    )
  )
}

stage3c_only_expected_messages <- if (
  length(all_stage3c_ordinary_messages) == 0L
) {
  TRUE
} else {
  all(
    grepl(
      "lowest lambda selected",
      tolower(
        all_stage3c_ordinary_messages
      ),
      fixed = TRUE
    )
  )
}

stage3c_conditions
```

```{r stage3c-path-and-network-diagnostics}
extract_stage3c_edges <- function(network_matrix) {
  network_matrix[
    upper.tri(network_matrix)
  ]
}

stage3c_path_diagnostics <- do.call(
  rbind,
  lapply(
    names(stage3c_path_runs),
    function(specification_name) {
      path_run <-
        stage3c_path_runs[[
          specification_name
        ]]

      path_result <- path_run$value

      selected_index <- which.min(
        path_result$ebic
      )

      selected_lambda <-
        path_result$lambda[
          selected_index
        ]

      lambda_maximum <- max(
        path_result$lambda
      )

      selected_network <-
        stage3c_selected_networks[[
          specification_name
        ]]

      selected_edges <-
        extract_stage3c_edges(
          selected_network
        )

      current_condition <-
        stage3c_conditions[
          stage3c_conditions$Specification ==
            specification_name,
          ,
          drop = FALSE
        ]

      data.frame(
        Election_year = 2020L,
        Candidate =
          "W_POST_COMPAT_RAWN",
        Specification =
          specification_name,
        Lambda_min_ratio =
          path_run$Lambda_min_ratio,
        Nlambda =
          path_run$Nlambda,
        Selected_index =
          selected_index,
        Selected_lambda =
          selected_lambda,
        Lambda_maximum =
          lambda_maximum,
        Selected_lambda_ratio =
          selected_lambda /
          lambda_maximum,
        At_lower_boundary =
          selected_index == 1L,
        At_upper_boundary =
          selected_index ==
          length(path_result$lambda),
        Minimum_EBIC =
          min(path_result$ebic),
        EBIC_at_lower_boundary =
          path_result$ebic[1L],
        Lower_boundary_delta_EBIC =
          path_result$ebic[1L] -
          min(path_result$ebic),
        Selected_edges =
          sum(
            abs(selected_edges) >
              1e-12
          ),
        Density =
          sum(
            abs(selected_edges) >
              1e-12
          ) /
          length(selected_edges),
        Global_strength =
          sum(abs(selected_edges)),
        All_EBIC_values_finite =
          all(is.finite(path_result$ebic)),
        All_lambda_values_finite =
          all(is.finite(path_result$lambda)),
        Warning_count =
          current_condition$Warning_count,
        Message_count =
          current_condition$Message_count,
        Network_SHA256 =
          digest::digest(
            selected_network,
            algo = "sha256",
            serialize = TRUE
          ),
        stringsAsFactors = FALSE
      )
    }
  )
)

stage3c_full_path <- do.call(
  rbind,
  lapply(
    names(stage3c_path_runs),
    function(specification_name) {
      path_run <-
        stage3c_path_runs[[
          specification_name
        ]]

      path_result <- path_run$value

      selected_index <- which.min(
        path_result$ebic
      )

      data.frame(
        Election_year = 2020L,
        Candidate =
          "W_POST_COMPAT_RAWN",
        Specification =
          specification_name,
        Grid_index =
          seq_along(path_result$lambda),
        Lambda =
          path_result$lambda,
        Lambda_ratio =
          path_result$lambda /
          max(path_result$lambda),
        EBIC =
          path_result$ebic,
        Selected =
          seq_along(path_result$lambda) ==
          selected_index,
        stringsAsFactors = FALSE
      )
    }
  )
)

compare_stage3c_networks <- function(
  comparison_name,
  first_network,
  second_network
) {
  first_edges <-
    extract_stage3c_edges(
      first_network
    )

  second_edges <-
    extract_stage3c_edges(
      second_network
    )

  absolute_difference <-
    abs(first_edges - second_edges)

  first_retained <-
    abs(first_edges) > 1e-12

  second_retained <-
    abs(second_edges) > 1e-12

  retained_union <-
    first_retained | second_retained

  retained_intersection <-
    first_retained & second_retained

  data.frame(
    Election_year = 2020L,
    Candidate =
      "W_POST_COMPAT_RAWN",
    Comparison =
      comparison_name,
    Compared_edges =
      length(first_edges),
    Edge_weight_correlation =
      stats::cor(
        first_edges,
        second_edges
      ),
    Median_absolute_edge_difference =
      stats::median(
        absolute_difference
      ),
    Mean_absolute_edge_difference =
      mean(absolute_difference),
    Maximum_absolute_edge_difference =
      max(absolute_difference),
    First_retained_edges =
      sum(first_retained),
    Second_retained_edges =
      sum(second_retained),
    Common_retained_edges =
      sum(retained_intersection),
    Retained_edge_union =
      sum(retained_union),
    Retained_edge_Jaccard = if (
      sum(retained_union) > 0L
    ) {
      sum(retained_intersection) /
        sum(retained_union)
    } else {
      1
    },
    Sign_reversals_among_common_edges =
      sum(
        first_edges[retained_intersection] *
          second_edges[retained_intersection] <
          0
      ),
    First_global_strength =
      sum(abs(first_edges)),
    Second_global_strength =
      sum(abs(second_edges)),
    Global_strength_difference =
      sum(abs(second_edges)) -
      sum(abs(first_edges)),
    stringsAsFactors = FALSE
  )
}

stage3c_network_comparison <- rbind(
  compare_stage3c_networks(
    comparison_name =
      "STAGE3B_REFERENCE_VS_CURRENT_GRID",
    first_network =
      stage3c_reference_network,
    second_network =
      stage3c_selected_networks[[
        "CURRENT_GRID"
      ]]
  ),
  compare_stage3c_networks(
    comparison_name =
      "CURRENT_GRID_VS_EXTENDED_GRID",
    first_network =
      stage3c_selected_networks[[
        "CURRENT_GRID"
      ]],
    second_network =
      stage3c_selected_networks[[
        "EXTENDED_GRID"
      ]]
  )
)

stage3c_current_reproduction_difference <-
  max(
    abs(
      stage3c_reference_network -
        stage3c_selected_networks[[
          "CURRENT_GRID"
        ]]
    )
  )

stage3c_extended_row <-
  stage3c_path_diagnostics[
    stage3c_path_diagnostics$Specification ==
      "EXTENDED_GRID",
    ,
    drop = FALSE
  ]

stage3c_boundary_status <- if (
  stage3c_extended_row$At_lower_boundary
) {
  "EXTENDED_GRID_REMAINS_AT_LOWER_BOUNDARY"
} else if (
  stage3c_extended_row$At_upper_boundary
) {
  "EXTENDED_GRID_SELECTED_UPPER_BOUNDARY"
} else {
  "EXTENDED_GRID_BRACKETED_INTERNAL_OPTIMUM"
}

stage3c_path_diagnostics
stage3c_network_comparison
stage3c_boundary_status
```

```{r stage3c-validation-and-save}
stage3c_validation <- data.frame(
  Check = c(
    "Stage 3B validation passed before Stage 3C",
    "Previous 2020 lowest-lambda message was documented",
    "The 2020 W_POST_COMPAT input passed validation",
    "Two lambda-path objects were returned",
    "All EBIC and lambda values are finite",
    "No estimation error was captured",
    "Only documented dense-network warnings were captured",
    "Only documented lowest-lambda messages were captured",
    "Current-grid network reproduces the Stage 3B network",
    "Current grid reproduces the lower-boundary selection",
    "Two path-diagnostic records were produced",
    "The full path contains 250 prespecified grid points",
    "Gamma refit and threshold settings were held fixed",
    "No bootstrap centrality NCT or political interpretation was run",
    "No final weight or lambda specification was selected"
  ),
  Passed = c(
    all(
      as.logical(
        stage3b_validation_disk$Passed
      )
    ),
    stage3c_previous_boundary_message,
    all(stage3c_input_gate$Passed),
    length(stage3c_path_runs) == 2L &&
      all(stage3c_success),
    all(
      stage3c_path_diagnostics$
        All_EBIC_values_finite
    ) &&
      all(
        stage3c_path_diagnostics$
          All_lambda_values_finite
      ),
    all(stage3c_conditions$Error == ""),
    stage3c_only_expected_warnings,
    stage3c_only_expected_messages,
    stage3c_current_reproduction_difference <=
      1e-10,
    stage3c_path_diagnostics$
      At_lower_boundary[
        stage3c_path_diagnostics$
          Specification ==
          "CURRENT_GRID"
      ],
    nrow(stage3c_path_diagnostics) == 2L,
    nrow(stage3c_full_path) == 250L,
    all(
      stage3c_specifications$Gamma ==
        0.50
    ) &&
      all(!stage3c_specifications$Refit) &&
      all(!stage3c_specifications$Threshold),
    TRUE,
    TRUE
  ),
  stringsAsFactors = FALSE
)

stage3c_validation

stage3c_path_file <- file.path(
  pilot_table_dir,
  "d15_stage3c_2020_wpost_lambda_path.csv"
)

stage3c_diagnostics_file <- file.path(
  pilot_table_dir,
  "d15_stage3c_2020_wpost_lambda_diagnostics.csv"
)

stage3c_comparison_file <- file.path(
  pilot_table_dir,
  "d15_stage3c_2020_wpost_network_comparison.csv"
)

stage3c_validation_file <- file.path(
  pilot_table_dir,
  "d15_stage3c_2020_wpost_validation.csv"
)

stage3c_model_file <- file.path(
  stage2c_model_dir,
  "d15_stage3c_2020_wpost_lambda_paths.rds"
)

stage3c_log_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage3c_2020_wpost.log"
)

stage3c_session_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage3c_2020_wpost_sessionInfo.txt"
)

utils::write.csv(
  stage3c_full_path,
  stage3c_path_file,
  row.names = FALSE
)

utils::write.csv(
  stage3c_path_diagnostics,
  stage3c_diagnostics_file,
  row.names = FALSE
)

utils::write.csv(
  stage3c_network_comparison,
  stage3c_comparison_file,
  row.names = FALSE
)

utils::write.csv(
  stage3c_validation,
  stage3c_validation_file,
  row.names = FALSE
)

saveRDS(
  list(
    Specifications =
      stage3c_specifications,
    Input_gate =
      stage3c_input_gate,
    Input_matrix =
      stage3c_input_matrix,
    Stage3B_reference_network =
      stage3c_reference_network,
    Path_runs =
      stage3c_path_runs,
    Path_diagnostics =
      stage3c_path_diagnostics,
    Full_path =
      stage3c_full_path,
    Network_comparison =
      stage3c_network_comparison,
    Conditions =
      stage3c_conditions,
    Boundary_status =
      stage3c_boundary_status
  ),
  stage3c_model_file
)

writeLines(
  capture.output(
    utils::sessionInfo()
  ),
  stage3c_session_file
)

writeLines(
  c(
    "D15 STAGE 3C — 2020 W_POST_COMPAT LAMBDA-BOUNDARY DIAGNOSTIC",
    paste0(
      "Timestamp: ",
      format(
        Sys.time(),
        "%Y-%m-%d %H:%M:%S %z"
      )
    ),
    "Lambda paths requested: 2",
    paste0(
      "Lambda paths returned: ",
      sum(stage3c_success),
      "/2"
    ),
    paste0(
      "Warnings captured: ",
      sum(
        stage3c_conditions$
          Warning_count
      )
    ),
    paste0(
      "Messages captured: ",
      sum(
        stage3c_conditions$
          Message_count
      )
    ),
    paste0(
      "Errors captured: ",
      sum(
        stage3c_conditions$
          Error != ""
      )
    ),
    paste0(
      "Boundary status: ",
      stage3c_boundary_status
    ),
    paste0(
      "Validation checks passed: ",
      sum(stage3c_validation$Passed),
      "/",
      nrow(stage3c_validation)
    ),
    paste0(
      "Current-grid reproduction difference: ",
      format(
        stage3c_current_reproduction_difference,
        scientific = TRUE
      )
    ),
    "The extended grid is diagnostic only.",
    "No final weight or lambda specification was selected.",
    "No bootstrap, centrality, NCT or political interpretation was run."
  ),
  stage3c_log_file
)

stage3c_output_files <- c(
  stage3c_path_file,
  stage3c_diagnostics_file,
  stage3c_comparison_file,
  stage3c_conditions_file,
  stage3c_validation_file,
  stage3c_model_file,
  stage3c_log_file,
  stage3c_session_file
)

stopifnot(
  all(
    file.exists(stage3c_output_files)
  )
)

data.frame(
  Output = stage3c_output_files,
  Exists = file.exists(stage3c_output_files),
  stringsAsFactors = FALSE
)

if (!all(stage3c_validation$Passed)) {
  stop(
    paste(
      "Stage 3C completed with a failed validation check.",
      "Inspect the saved Stage 3C files."
    )
  )
}
```

## Stage 3C independent-review disposition

Stage 3C passed independent review. The extended lambda path bracketed
an internal optimum, while the current-grid and extended-grid networks
were practically identical. The registered lambda grid is retained for
the remaining diagnostic stress test. This decision does not select a
final weight specification or authorise formal network analysis.
## Stage 3D — Kish effective-sample-size stress test

This diagnostic compares each weighted raw-N network with an otherwise
identical network in which rounded Kish effective sample size is supplied
to EBICglasso as n. Kish ESS is not treated as a validated or selectable
primary sample-size definition. No categorical materiality threshold is
introduced after inspecting the results.

```{r stage3d-kish-ess-input-gate}
stage3d_required_packages <- c(
  "qgraph",
  "digest"
)

stage3d_package_gate <- data.frame(
  Package = stage3d_required_packages,
  Available = vapply(
    stage3d_required_packages,
    requireNamespace,
    quietly = TRUE,
    FUN.VALUE = logical(1)
  ),
  Version = vapply(
    stage3d_required_packages,
    function(package_name) {
      if (requireNamespace(package_name, quietly = TRUE)) {
        as.character(
          utils::packageVersion(package_name)
        )
      } else {
        NA_character_
      }
    },
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

stage3d_package_gate

stopifnot(
  all(stage3d_package_gate$Available)
)

stage3b_validation_path <- file.path(
  pilot_table_dir,
  "d15_stage3b_five_wave_validation.csv"
)

stage3c_validation_path <- file.path(
  pilot_table_dir,
  "d15_stage3c_2020_wpost_validation.csv"
)

stage3b_model_path <- file.path(
  stage2c_model_dir,
  "d15_stage3b_five_wave_point_networks.rds"
)

stage3d_weight_audit_path <- file.path(
  pilot_table_dir,
  "d15_weight_distribution_audit.csv"
)

stage3d_required_files <- c(
  stage3b_validation_path,
  stage3c_validation_path,
  stage3b_model_path,
  stage3d_weight_audit_path
)

stopifnot(
  all(file.exists(stage3d_required_files))
)

stage3b_validation_disk <- utils::read.csv(
  stage3b_validation_path,
  stringsAsFactors = FALSE
)

stage3c_validation_disk <- utils::read.csv(
  stage3c_validation_path,
  stringsAsFactors = FALSE
)

stage3b_saved <- readRDS(
  stage3b_model_path
)

stage3d_weight_audit <- utils::read.csv(
  stage3d_weight_audit_path,
  stringsAsFactors = FALSE
)

stage3d_weight_audit$Election_year <- as.integer(
  stage3d_weight_audit$Election_year
)

stage3d_weight_audit$Complete_case_N <- as.integer(
  stage3d_weight_audit$Complete_case_N
)

stage3d_weight_audit$Kish_effective_N <- as.numeric(
  stage3d_weight_audit$Kish_effective_N
)

stopifnot(
  nrow(stage3b_validation_disk) == 14L,
  all(as.logical(stage3b_validation_disk$Passed)),
  nrow(stage3c_validation_disk) == 15L,
  all(as.logical(stage3c_validation_disk$Passed))
)

stage3d_years <- c(
  2004L,
  2012L,
  2016L,
  2020L,
  2024L
)

stage3d_weight_candidates <- c(
  "W_CDF",
  "W_POST_COMPAT"
)

stage3d_input_map <- expand.grid(
  Election_year = stage3d_years,
  Weight_candidate = stage3d_weight_candidates,
  KEEP.OUT.ATTRS = FALSE,
  stringsAsFactors = FALSE
)

stage3d_input_map$Raw_candidate <- ifelse(
  stage3d_input_map$Weight_candidate == "W_CDF",
  "W_CDF_RAWN",
  "W_POST_COMPAT_RAWN"
)

stage3d_input_map$Kish_candidate <- ifelse(
  stage3d_input_map$Weight_candidate == "W_CDF",
  "W_CDF_KISH",
  "W_POST_COMPAT_KISH"
)

stage3d_input_map$Raw_key <- paste(
  stage3d_input_map$Election_year,
  stage3d_input_map$Raw_candidate,
  sep = "__"
)

stage3d_input_map$Kish_key <- paste(
  stage3d_input_map$Election_year,
  stage3d_input_map$Kish_candidate,
  sep = "__"
)

get_stage3d_raw_n <- function(
  year_value,
  candidate_name
) {
  selected_row <- stage3b_saved$Settings[
    stage3b_saved$Settings$Election_year == year_value &
      stage3b_saved$Settings$Candidate == candidate_name,
    ,
    drop = FALSE
  ]

  if (nrow(selected_row) != 1L) {
    stop(
      paste(
        "Stage 3D could not uniquely identify raw N for",
        year_value,
        candidate_name
      )
    )
  }

  as.numeric(selected_row$Sample_N)
}

get_stage3d_kish_ess <- function(
  year_value,
  weight_candidate
) {
  selected_row <- stage3d_weight_audit[
    stage3d_weight_audit$Election_year == year_value &
      stage3d_weight_audit$Candidate == weight_candidate,
    ,
    drop = FALSE
  ]

  if (nrow(selected_row) != 1L) {
    stop(
      paste(
        "Stage 3D could not uniquely identify Kish ESS for",
        year_value,
        weight_candidate
      )
    )
  }

  as.numeric(selected_row$Kish_effective_N)
}

get_stage3d_complete_case_n <- function(
  year_value,
  weight_candidate
) {
  selected_row <- stage3d_weight_audit[
    stage3d_weight_audit$Election_year == year_value &
      stage3d_weight_audit$Candidate == weight_candidate,
    ,
    drop = FALSE
  ]

  if (nrow(selected_row) != 1L) {
    stop(
      paste(
        "Stage 3D could not uniquely identify complete-case N for",
        year_value,
        weight_candidate
      )
    )
  }

  as.integer(selected_row$Complete_case_N)
}

stage3d_input_map$Raw_N <- vapply(
  seq_len(nrow(stage3d_input_map)),
  function(row_number) {
    get_stage3d_raw_n(
      stage3d_input_map$Election_year[row_number],
      stage3d_input_map$Raw_candidate[row_number]
    )
  },
  FUN.VALUE = numeric(1)
)

stage3d_input_map$Audit_complete_case_N <- vapply(
  seq_len(nrow(stage3d_input_map)),
  function(row_number) {
    get_stage3d_complete_case_n(
      stage3d_input_map$Election_year[row_number],
      stage3d_input_map$Weight_candidate[row_number]
    )
  },
  FUN.VALUE = integer(1)
)

stage3d_input_map$Kish_effective_N_exact <- vapply(
  seq_len(nrow(stage3d_input_map)),
  function(row_number) {
    get_stage3d_kish_ess(
      stage3d_input_map$Election_year[row_number],
      stage3d_input_map$Weight_candidate[row_number]
    )
  },
  FUN.VALUE = numeric(1)
)

stage3d_input_map$Kish_N_supplied <- as.integer(
  round(
    stage3d_input_map$Kish_effective_N_exact
  )
)

stage3d_input_map$Kish_ESS_over_raw_N <-
  stage3d_input_map$Kish_effective_N_exact /
  stage3d_input_map$Raw_N

stopifnot(
  nrow(stage3d_input_map) == 10L,
  !anyDuplicated(stage3d_input_map$Kish_key),
  all(stage3d_input_map$Raw_N ==
        stage3d_input_map$Audit_complete_case_N),
  all(is.finite(
    stage3d_input_map$Kish_effective_N_exact
  )),
  all(stage3d_input_map$Kish_N_supplied > 0L),
  all(stage3d_input_map$Kish_N_supplied <=
        stage3d_input_map$Raw_N),
  all(stage3d_input_map$Raw_key %in%
        names(stage3b_saved$Input_matrices)),
  all(stage3d_input_map$Raw_key %in%
        names(stage3b_saved$Point_networks))
)

stage3d_input_matrices <- setNames(
  lapply(
    stage3d_input_map$Raw_key,
    function(current_key) {
      as.matrix(
        stage3b_saved$Input_matrices[[current_key]]
      )
    }
  ),
  stage3d_input_map$Kish_key
)

stage3d_input_checks <- do.call(
  rbind,
  lapply(
    seq_len(nrow(stage3d_input_map)),
    function(row_number) {
      current_key <-
        stage3d_input_map$Kish_key[row_number]

      current_matrix <-
        stage3d_input_matrices[[current_key]]

      data.frame(
        Election_year =
          stage3d_input_map$Election_year[row_number],
        Weight_candidate =
          stage3d_input_map$Weight_candidate[row_number],
        Kish_candidate =
          stage3d_input_map$Kish_candidate[row_number],
        Raw_N =
          stage3d_input_map$Raw_N[row_number],
        Kish_effective_N_exact =
          stage3d_input_map$
            Kish_effective_N_exact[row_number],
        Kish_N_supplied =
          stage3d_input_map$Kish_N_supplied[row_number],
        Rows = nrow(current_matrix),
        Columns = ncol(current_matrix),
        All_finite =
          all(is.finite(current_matrix)),
        Symmetric = isSymmetric(
          current_matrix,
          tol = 1e-10
        ),
        Diagonal_one =
          max(
            abs(diag(current_matrix) - 1)
          ) <= 1e-10,
        Node_order_correct =
          identical(
            rownames(current_matrix),
            network_vars
          ) &&
          identical(
            colnames(current_matrix),
            network_vars
          ),
        Minimum_eigenvalue = min(
          eigen(
            current_matrix,
            symmetric = TRUE,
            only.values = TRUE
          )$values
        ),
        Input_SHA256 = digest::digest(
          current_matrix,
          algo = "sha256",
          serialize = TRUE
        ),
        stringsAsFactors = FALSE
      )
    }
  )
)

stage3d_input_checks

stopifnot(
  nrow(stage3d_input_checks) == 10L,
  all(stage3d_input_checks$Rows == 9L),
  all(stage3d_input_checks$Columns == 9L),
  all(stage3d_input_checks$All_finite),
  all(stage3d_input_checks$Symmetric),
  all(stage3d_input_checks$Diagonal_one),
  all(stage3d_input_checks$Node_order_correct),
  all(stage3d_input_checks$Minimum_eigenvalue > 0)
)

stage3d_settings <- data.frame(
  Election_year =
    stage3d_input_map$Election_year,
  Weight_candidate =
    stage3d_input_map$Weight_candidate,
  Raw_N =
    stage3d_input_map$Raw_N,
  Kish_effective_N_exact =
    stage3d_input_map$Kish_effective_N_exact,
  Kish_N_supplied =
    stage3d_input_map$Kish_N_supplied,
  Correlation = "weighted polychoric",
  Estimator = "EBICglasso",
  Gamma = 0.50,
  Nlambda = 100L,
  Lambda_min_ratio = 0.01,
  Refit = FALSE,
  Threshold = FALSE,
  Purpose =
    "diagnostic Kish-ESS stress test only",
  stringsAsFactors = FALSE
)

stage3d_settings
```

### 估计十张 Kish 网络

```{r stage3d-kish-ess-network-estimation}
run_stage3d_network <- function(
  correlation_matrix,
  year_value,
  weight_candidate,
  kish_candidate,
  raw_n,
  kish_ess_exact,
  kish_n
) {
  model_run <- run_and_capture(
    function() {
      qgraph::EBICglasso(
        S = correlation_matrix,
        n = kish_n,
        gamma = 0.50,
        nlambda = 100L,
        lambda.min.ratio = 0.01,
        refit = FALSE,
        threshold = FALSE
      )
    }
  )

  if (!is.null(model_run$value)) {
    dimnames(model_run$value) <- list(
      network_vars,
      network_vars
    )
  }

  list(
    Election_year = year_value,
    Weight_candidate = weight_candidate,
    Kish_candidate = kish_candidate,
    Raw_N = raw_n,
    Kish_effective_N_exact = kish_ess_exact,
    Kish_N_supplied = kish_n,
    value = model_run$value,
    warnings = model_run$warnings,
    messages = model_run$messages,
    error = model_run$error
  )
}

stage3d_runs <- setNames(
  lapply(
    seq_len(nrow(stage3d_input_map)),
    function(row_number) {
      current_key <-
        stage3d_input_map$Kish_key[row_number]

      run_stage3d_network(
        correlation_matrix =
          stage3d_input_matrices[[current_key]],
        year_value =
          stage3d_input_map$Election_year[row_number],
        weight_candidate =
          stage3d_input_map$Weight_candidate[row_number],
        kish_candidate =
          stage3d_input_map$Kish_candidate[row_number],
        raw_n =
          stage3d_input_map$Raw_N[row_number],
        kish_ess_exact =
          stage3d_input_map$
            Kish_effective_N_exact[row_number],
        kish_n =
          stage3d_input_map$Kish_N_supplied[row_number]
      )
    }
  ),
  stage3d_input_map$Kish_key
)

is_expected_stage3d_warning <- function(values) {
  if (length(values) == 0L) {
    logical(0)
  } else {
    grepl(
      "dense regularized network",
      tolower(values),
      fixed = TRUE
    )
  }
}

is_expected_stage3d_message <- function(values) {
  if (length(values) == 0L) {
    logical(0)
  } else {
    grepl(
      "network with lowest lambda selected",
      tolower(values),
      fixed = TRUE
    )
  }
}

stage3d_conditions <- do.call(
  rbind,
  lapply(
    stage3d_runs,
    function(model_run) {
      expected_warning_flags <-
        is_expected_stage3d_warning(
          model_run$warnings
        )

      expected_message_flags <-
        is_expected_stage3d_message(
          model_run$messages
        )

      data.frame(
        Election_year =
          model_run$Election_year,
        Weight_candidate =
          model_run$Weight_candidate,
        Kish_candidate =
          model_run$Kish_candidate,
        Raw_N =
          model_run$Raw_N,
        Kish_effective_N_exact =
          model_run$Kish_effective_N_exact,
        Kish_N_supplied =
          model_run$Kish_N_supplied,
        Warning_count =
          length(model_run$warnings),
        Warnings =
          collapse_conditions(
            model_run$warnings
          ),
        Unexpected_warning_count =
          sum(!expected_warning_flags),
        Message_count =
          length(model_run$messages),
        Messages =
          collapse_conditions(
            model_run$messages
          ),
        Lowest_lambda_message =
          any(expected_message_flags),
        Unexpected_message_count =
          sum(!expected_message_flags),
        Error = if (
          is.na(model_run$error)
        ) {
          ""
        } else {
          model_run$error
        },
        stringsAsFactors = FALSE
      )
    }
  )
)

stage3d_conditions_file <- file.path(
  pilot_table_dir,
  "d15_stage3d_kish_ess_conditions.csv"
)

utils::write.csv(
  stage3d_conditions,
  stage3d_conditions_file,
  row.names = FALSE
)

stage3d_success <- vapply(
  stage3d_runs,
  function(model_run) {
    is.matrix(model_run$value) &&
      identical(
        dim(model_run$value),
        c(9L, 9L)
      ) &&
      (
        is.na(model_run$error) ||
          model_run$error == ""
      )
  },
  FUN.VALUE = logical(1)
)

stage3d_conditions

if (!all(stage3d_success)) {
  stop(
    paste(
      "Stage 3D network estimation failed.",
      "Inspect d15_stage3d_kish_ess_conditions.csv."
    )
  )
}

stage3d_networks <- lapply(
  stage3d_runs,
  function(model_run) {
    model_run$value
  }
)
```

### 比较 raw N 与 Kish ESS

```{r stage3d-kish-ess-diagnostics}
extract_stage3d_edges <- function(network_matrix) {
  network_matrix[
    upper.tri(network_matrix)
  ]
}

stage3d_network_diagnostics <- do.call(
  rbind,
  lapply(
    seq_len(nrow(stage3d_input_map)),
    function(row_number) {
      kish_key <-
        stage3d_input_map$Kish_key[row_number]

      current_network <-
        stage3d_networks[[kish_key]]

      current_edges <-
        extract_stage3d_edges(
          current_network
        )

      current_conditions <-
        stage3d_conditions[
          stage3d_conditions$Election_year ==
            stage3d_input_map$
              Election_year[row_number] &
            stage3d_conditions$Weight_candidate ==
              stage3d_input_map$
                Weight_candidate[row_number],
          ,
          drop = FALSE
        ]

      current_input_check <-
        stage3d_input_checks[
          stage3d_input_checks$Election_year ==
            stage3d_input_map$
              Election_year[row_number] &
            stage3d_input_checks$Weight_candidate ==
              stage3d_input_map$
                Weight_candidate[row_number],
          ,
          drop = FALSE
        ]

      retained_edges <-
        abs(current_edges) > 1e-12

      data.frame(
        Election_year =
          stage3d_input_map$Election_year[row_number],
        Weight_candidate =
          stage3d_input_map$Weight_candidate[row_number],
        Kish_candidate =
          stage3d_input_map$Kish_candidate[row_number],
        Raw_N =
          stage3d_input_map$Raw_N[row_number],
        Kish_effective_N_exact =
          stage3d_input_map$
            Kish_effective_N_exact[row_number],
        Kish_N_supplied =
          stage3d_input_map$Kish_N_supplied[row_number],
        Kish_ESS_over_raw_N =
          stage3d_input_map$
            Kish_ESS_over_raw_N[row_number],
        Compared_edges =
          length(current_edges),
        Retained_edges =
          sum(retained_edges),
        Density =
          sum(retained_edges) /
          length(current_edges),
        Global_strength =
          sum(abs(current_edges)),
        Minimum_edge =
          min(current_edges),
        Maximum_edge =
          max(current_edges),
        All_values_finite =
          all(is.finite(current_network)),
        Maximum_asymmetry =
          max(
            abs(
              current_network -
                t(current_network)
            )
          ),
        Maximum_diagonal =
          max(abs(diag(current_network))),
        Input_minimum_eigenvalue =
          current_input_check$Minimum_eigenvalue,
        Warning_count =
          current_conditions$Warning_count,
        Lowest_lambda_message =
          current_conditions$Lowest_lambda_message,
        Error =
          current_conditions$Error,
        Input_SHA256 =
          current_input_check$Input_SHA256,
        Network_SHA256 = digest::digest(
          current_network,
          algo = "sha256",
          serialize = TRUE
        ),
        stringsAsFactors = FALSE
      )
    }
  )
)

compare_stage3d_rawn_and_kish <- function(
  row_number
) {
  year_value <-
    stage3d_input_map$Election_year[row_number]

  weight_candidate <-
    stage3d_input_map$Weight_candidate[row_number]

  raw_key <-
    stage3d_input_map$Raw_key[row_number]

  kish_key <-
    stage3d_input_map$Kish_key[row_number]

  raw_network <- as.matrix(
    stage3b_saved$Point_networks[[raw_key]]
  )

  kish_network <- as.matrix(
    stage3d_networks[[kish_key]]
  )

  raw_edges <-
    extract_stage3d_edges(raw_network)

  kish_edges <-
    extract_stage3d_edges(kish_network)

  absolute_difference <-
    abs(raw_edges - kish_edges)

  raw_retained <-
    abs(raw_edges) > 1e-12

  kish_retained <-
    abs(kish_edges) > 1e-12

  retained_union <-
    raw_retained | kish_retained

  retained_intersection <-
    raw_retained & kish_retained

  retained_jaccard <- if (
    sum(retained_union) > 0L
  ) {
    sum(retained_intersection) /
      sum(retained_union)
  } else {
    1
  }

  sign_agreement_union <- if (
    sum(retained_union) > 0L
  ) {
    mean(
      sign(raw_edges[retained_union]) ==
        sign(kish_edges[retained_union])
    )
  } else {
    NA_real_
  }

  raw_global_strength <-
    sum(abs(raw_edges))

  kish_global_strength <-
    sum(abs(kish_edges))

  data.frame(
    Election_year = year_value,
    Weight_candidate = weight_candidate,
    Raw_candidate =
      stage3d_input_map$Raw_candidate[row_number],
    Kish_candidate =
      stage3d_input_map$Kish_candidate[row_number],
    Raw_N =
      stage3d_input_map$Raw_N[row_number],
    Kish_effective_N_exact =
      stage3d_input_map$
        Kish_effective_N_exact[row_number],
    Kish_N_supplied =
      stage3d_input_map$Kish_N_supplied[row_number],
    Kish_ESS_over_raw_N =
      stage3d_input_map$
        Kish_ESS_over_raw_N[row_number],
    Compared_edges =
      length(raw_edges),
    Edge_weight_correlation =
      stats::cor(raw_edges, kish_edges),
    Median_absolute_edge_difference =
      stats::median(absolute_difference),
    Mean_absolute_edge_difference =
      mean(absolute_difference),
    Maximum_absolute_edge_difference =
      max(absolute_difference),
    Sign_agreement_all_pairs =
      mean(
        sign(raw_edges) ==
          sign(kish_edges)
      ),
    Sign_agreement_retained_union =
      sign_agreement_union,
    Sign_reversals_among_common_edges =
      sum(
        raw_edges[retained_intersection] *
          kish_edges[retained_intersection] < 0
      ),
    Raw_retained_edges =
      sum(raw_retained),
    Kish_retained_edges =
      sum(kish_retained),
    Common_retained_edges =
      sum(retained_intersection),
    Retained_edge_union =
      sum(retained_union),
    Retained_edge_Jaccard =
      retained_jaccard,
    Raw_global_strength =
      raw_global_strength,
    Kish_global_strength =
      kish_global_strength,
    Global_strength_difference =
      kish_global_strength -
      raw_global_strength,
    Global_strength_percent_difference =
      if (raw_global_strength == 0) {
        NA_real_
      } else {
        100 *
          (
            kish_global_strength -
              raw_global_strength
          ) /
          raw_global_strength
      },
    stringsAsFactors = FALSE
  )
}

stage3d_network_comparison <- do.call(
  rbind,
  lapply(
    seq_len(nrow(stage3d_input_map)),
    compare_stage3d_rawn_and_kish
  )
)

stage3d_network_diagnostics
stage3d_network_comparison
```

### 验证和保存

```{r stage3d-kish-ess-validation-and-save}
stage3d_validation <- data.frame(
  Check = c(
    "Stage 3B validation passed before Stage 3D",
    "Stage 3C validation passed before Stage 3D",
    "Ten weighted Kish-ESS candidate records were mapped",
    "Raw sample counts match the Stage 1 audit",
    "Exact Kish effective sample sizes are finite and valid",
    "Rounded Kish N values are positive and do not exceed raw N",
    "Ten weighted input matrices passed validation",
    "Ten Kish-ESS point networks were returned",
    "All Kish-ESS network entries are finite",
    "All Kish-ESS networks are symmetric with zero diagonals",
    "Each Kish-ESS network contains 36 possible edges",
    "No estimation error was captured",
    "No unexpected estimation warning was captured",
    "No unexpected ordinary message was captured",
    "Ten raw-N versus Kish-ESS comparisons were produced",
    "The EBICglasso specification was held fixed",
    "No bootstrap centrality NCT or political interpretation was run",
    "No final weight or EBIC sample-size definition was selected"
  ),
  Passed = c(
    all(as.logical(
      stage3b_validation_disk$Passed
    )),
    all(as.logical(
      stage3c_validation_disk$Passed
    )),
    nrow(stage3d_input_map) == 10L &&
      !anyDuplicated(stage3d_input_map$Kish_key),
    all(
      stage3d_input_map$Raw_N ==
        stage3d_input_map$Audit_complete_case_N
    ),
    all(is.finite(
      stage3d_input_map$Kish_effective_N_exact
    )) &&
      all(
        stage3d_input_map$Kish_effective_N_exact >
          0
      ),
    all(stage3d_input_map$Kish_N_supplied > 0L) &&
      all(
        stage3d_input_map$Kish_N_supplied <=
          stage3d_input_map$Raw_N
      ),
    nrow(stage3d_input_checks) == 10L &&
      all(stage3d_input_checks$All_finite) &&
      all(stage3d_input_checks$Symmetric) &&
      all(stage3d_input_checks$Diagonal_one) &&
      all(stage3d_input_checks$Node_order_correct) &&
      all(
        stage3d_input_checks$Minimum_eigenvalue >
          0
      ),
    length(stage3d_networks) == 10L &&
      all(stage3d_success),
    all(
      stage3d_network_diagnostics$
        All_values_finite
    ),
    all(
      stage3d_network_diagnostics$
        Maximum_asymmetry <= 1e-10
    ) &&
      all(
        stage3d_network_diagnostics$
          Maximum_diagonal <= 1e-10
      ),
    all(
      stage3d_network_diagnostics$
        Compared_edges == 36L
    ),
    all(stage3d_conditions$Error == ""),
    all(
      stage3d_conditions$
        Unexpected_warning_count == 0L
    ),
    all(
      stage3d_conditions$
        Unexpected_message_count == 0L
    ),
    nrow(stage3d_network_comparison) == 10L &&
      all(
        stage3d_network_comparison$
          Compared_edges == 36L
      ) &&
      all(is.finite(
        stage3d_network_comparison$
          Edge_weight_correlation
      )),
    all(stage3d_settings$Gamma == 0.50) &&
      all(stage3d_settings$Nlambda == 100L) &&
      all(
        stage3d_settings$
          Lambda_min_ratio == 0.01
      ) &&
      all(!stage3d_settings$Refit) &&
      all(!stage3d_settings$Threshold),
    TRUE,
    TRUE
  ),
  stringsAsFactors = FALSE
)

stage3d_validation

stage3d_diagnostics_file <- file.path(
  pilot_table_dir,
  "d15_stage3d_kish_ess_point_network_diagnostics.csv"
)

stage3d_comparison_file <- file.path(
  pilot_table_dir,
  "d15_stage3d_rawn_vs_kish_comparison.csv"
)

stage3d_validation_file <- file.path(
  pilot_table_dir,
  "d15_stage3d_kish_ess_validation.csv"
)

stage3d_model_file <- file.path(
  stage2c_model_dir,
  "d15_stage3d_kish_ess_point_networks.rds"
)

stage3d_log_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage3d_kish_ess.log"
)

stage3d_session_file <- file.path(
  pilot_log_dir,
  "06B_D15_stage3d_kish_ess_sessionInfo.txt"
)

utils::write.csv(
  stage3d_network_diagnostics,
  stage3d_diagnostics_file,
  row.names = FALSE
)

utils::write.csv(
  stage3d_network_comparison,
  stage3d_comparison_file,
  row.names = FALSE
)

utils::write.csv(
  stage3d_validation,
  stage3d_validation_file,
  row.names = FALSE
)

saveRDS(
  list(
    Settings = stage3d_settings,
    Input_map = stage3d_input_map,
    Input_checks = stage3d_input_checks,
    Input_matrices = stage3d_input_matrices,
    Kish_point_networks = stage3d_networks,
    Conditions = stage3d_conditions,
    Diagnostics = stage3d_network_diagnostics,
    RawN_vs_Kish_comparisons =
      stage3d_network_comparison
  ),
  stage3d_model_file
)

writeLines(
  capture.output(
    utils::sessionInfo()
  ),
  stage3d_session_file
)

writeLines(
  c(
    "D15 STAGE 3D — KISH EFFECTIVE-SAMPLE-SIZE STRESS TEST",
    paste0(
      "Timestamp: ",
      format(
        Sys.time(),
        "%Y-%m-%d %H:%M:%S %z"
      )
    ),
    "Kish-ESS networks requested: 10",
    paste0(
      "Kish-ESS networks returned: ",
      sum(stage3d_success),
      "/10"
    ),
    paste0(
      "Lowest-lambda messages recorded: ",
      sum(
        stage3d_conditions$
          Lowest_lambda_message
      )
    ),
    paste0(
      "Unexpected warnings recorded: ",
      sum(
        stage3d_conditions$
          Unexpected_warning_count
      )
    ),
    paste0(
      "Unexpected messages recorded: ",
      sum(
        stage3d_conditions$
          Unexpected_message_count
      )
    ),
    paste0(
      "Errors recorded: ",
      sum(stage3d_conditions$Error != "")
    ),
    paste0(
      "Validation checks passed: ",
      sum(stage3d_validation$Passed),
      "/",
      nrow(stage3d_validation)
    ),
    "Kish ESS was used only as a diagnostic EBIC n stress test.",
    "No categorical materiality threshold was applied.",
    "No bootstrap, centrality, NCT or political interpretation was run.",
    "No final weight or EBIC sample-size definition was selected."
  ),
  stage3d_log_file
)

stage3d_output_files <- c(
  stage3d_conditions_file,
  stage3d_diagnostics_file,
  stage3d_comparison_file,
  stage3d_validation_file,
  stage3d_model_file,
  stage3d_log_file,
  stage3d_session_file
)

stopifnot(
  all(file.exists(stage3d_output_files))
)

data.frame(
  Output = stage3d_output_files,
  Exists = file.exists(stage3d_output_files),
  stringsAsFactors = FALSE
)

if (!all(stage3d_validation$Passed)) {
  stop(
    paste(
      "Stage 3D completed with a failed validation check.",
      "Inspect the saved Stage 3D files."
    )
  )
}
```

## Hard stop after Stage 3D

Stop here. Do not select or freeze a final weight treatment, replace raw
sample N with Kish ESS, run bootstrap, centrality, NCT, education-group
analysis or formal five-wave analysis until the complete D15 pilot has
been independently reviewed.