---
title: "07 Primary Network Estimation and Stability"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

## Purpose and Current Status

This notebook is the formal location for Step 07 of the ANES dissertation workflow. A separately authorised, non-inferential 2004 pilot evaluated the computational behaviour of 10, 1,000 and 2,500 nonparametric edge-weight bootstrap repetitions. D15 and D21 now authorise the bounded five-wave primary point-network and edge-accuracy workflow recorded below. Centrality, case-dropping bootstrap, NCT, education-group analysis, sensitivity analysis and substantive political interpretation remain outside this execution.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  error = FALSE,
  warning = TRUE,
  message = TRUE
)
```

## Project Root

```{r project-root}
current_dir <- normalizePath(
  ".",
  winslash = "/"
)

if (
  file.exists(
    file.path(
      current_dir,
      "ANES.Rproj"
    )
  )
) {
  project_root <- current_dir
} else if (
  basename(current_dir) == "notebooks" &&
    file.exists(
      file.path(
        dirname(current_dir),
        "ANES.Rproj"
      )
    )
) {
  project_root <- dirname(current_dir)
} else {
  stop(
    "Project root could not be identified."
  )
}

knitr::opts_knit$set(
  root.dir = project_root
)

project_root
```

## Execution Authorisation

The authorisation state is read at render time from the single decision-registry path recorded in the project configuration. The notebook does not contain editable copies of the D15 or D21 status and does not accept an environment-variable override.

```{r decision-registry}
registry_config <- file.path(
  project_root,
  "config",
  "decision_registry_path.txt"
)

registry_path <- if (file.exists(registry_config)) {
  configured_lines <- trimws(
    readLines(
      registry_config,
      warn = FALSE
    )
  )
  configured_lines <- configured_lines[nzchar(configured_lines)]
  if (length(configured_lines) == 1L) {
    configured_lines
  } else {
    NA_character_
  }
} else {
  NA_character_
}

registry_read_error <- NA_character_
decision_registry <- tryCatch(
  {
    if (
      length(registry_path) != 1L ||
        is.na(registry_path) ||
        !file.exists(registry_path)
    ) {
      stop(
        "Configured decision registry is unavailable."
      )
    }
    utils::read.csv(
      registry_path,
      stringsAsFactors = FALSE,
      check.names = FALSE
    )
  },
  error = function(condition) {
    registry_read_error <<- conditionMessage(condition)
    data.frame()
  }
)

required_registry_columns <- c(
  "decision_id",
  "decision_status",
  "implementation_status",
  "researcher_approval_status",
  "supervisor_confirmation_status",
  "analysis_effect",
  "last_verified"
)

registry_schema_valid <- all(
  required_registry_columns %in%
    names(decision_registry)
)

normalise_status <- function(value) {
  value <- toupper(trimws(as.character(value)))
  gsub("[ -]+", "_", value)
}

decision_row <- function(decision_id) {
  if (!registry_schema_valid) {
    return(data.frame())
  }
  decision_registry[
    decision_registry$decision_id == decision_id,
    ,
    drop = FALSE
  ]
}

d15 <- decision_row("D15")
d21 <- decision_row("D21")

d15_unique <- nrow(d15) == 1L
d21_unique <- nrow(d21) == 1L

d15_passed <- d15_unique &&
  normalise_status(d15$researcher_approval_status) == "APPROVED" &&
  normalise_status(d15$supervisor_confirmation_status) %in%
    c(
      "CONFIRMED",
      "APPROVED",
      "NOT_REQUIRED",
      "DEFERRED_BY_RESEARCHER"
    ) &&
  normalise_status(d15$analysis_effect) != "BLOCKS_CORE_GGM"

d21_passed <- d21_unique &&
  normalise_status(d21$researcher_approval_status) == "APPROVED" &&
  normalise_status(d21$implementation_status) %in%
    c(
      "AUTHORISED",
      "AUTHORIZED",
      "COMPLETE_AND_VALIDATED"
    ) &&
  normalise_status(d21$analysis_effect) != "BLOCKS_CORE_GGM"

registry_value <- function(row, column) {
  if (
    nrow(row) == 1L &&
      column %in% names(row)
  ) {
    as.character(row[[column]])
  } else {
    NA_character_
  }
}

authorisation_status <- data.frame(
  Decision = c(
    "D15 survey-weight decision",
    "D21 formal Step-07 execution authority"
  ),
  Decision_status = c(
    registry_value(d15, "decision_status"),
    registry_value(d21, "decision_status")
  ),
  Implementation_status = c(
    registry_value(d15, "implementation_status"),
    registry_value(d21, "implementation_status")
  ),
  Researcher_approval = c(
    registry_value(d15, "researcher_approval_status"),
    registry_value(d21, "researcher_approval_status")
  ),
  Supervisor_confirmation = c(
    registry_value(d15, "supervisor_confirmation_status"),
    registry_value(d21, "supervisor_confirmation_status")
  ),
  Analysis_effect = c(
    registry_value(d15, "analysis_effect"),
    registry_value(d21, "analysis_effect")
  ),
  Last_verified = c(
    registry_value(d15, "last_verified"),
    registry_value(d21, "last_verified")
  ),
  Passed = c(
    d15_passed,
    d21_passed
  ),
  stringsAsFactors = FALSE
)

authorisation_status
```

```{r authorisation-gate}
authorisation_passed <- is.na(registry_read_error) &&
  registry_schema_valid &&
  d15_unique &&
  d21_unique &&
  all(authorisation_status$Passed)

data.frame(
  Registry = registry_path,
  Registry_read_error = registry_read_error,
  Required_schema_present = registry_schema_valid,
  D15_unique = d15_unique,
  D21_unique = d21_unique,
  Authorisation_passed = authorisation_passed,
  stringsAsFactors = FALSE
)
```

## Read-Only Technical Preflight

The preflight checks the repaired notebooks, the exact locked and installed `bootnet` version, the expected outputs from Steps 00–06, and the internal structure of the saved analysis-ready and polychoric-diagnostic objects. It does not read or modify the ANES raw file.

```{r preflight-paths}
cleaning_rmd <- file.path(
  project_root,
  "notebooks",
  "03_data_cleaning.Rmd"
)

analysis_ready_rmd <- file.path(
  project_root,
  "notebooks",
  "04_analysis_ready_data.Rmd"
)

renv_lockfile <- file.path(
  project_root,
  "renv.lock"
)

reproduction_report <- file.path(
  project_root,
  "outputs",
  "logs",
  "00_06_REPRODUCTION_CHECK.md"
)

html_files <- file.path(
  project_root,
  "notebooks",
  c(
    "00_environment.html",
    "01_data_inventory.html",
    "02_data_extraction.html",
    "03_data_cleaning.html",
    "04_analysis_ready_data.html",
    "05_preliminary_data_suitability_audit.html",
    "06-Polychoric-Input-Diagnostics.html"
  )
)

expected_output_files <- file.path(
  project_root,
  c(
    "data/extracted/anes_cdf_20260205_extracted_2004_2012_2016_2020_2024.rds",
    "data/cleaned/anes_cdf_20260205_cleaned_2004_2012_2016_2020_2024.rds",
    "data/analysis_ready/anes_cdf_20260205_analysis_ready_nine_node_complete_case_2004_2012_2016_2020_2024.rds",
    "data/audit/anes_cdf_20260205_sample_flow_nine_node_complete_case_2004_2012_2016_2020_2024.csv",
    "data/audit/anes_cdf_20260205_analysis_ready_pairwise_respondent_overlap_2004_2012_2016_2020_2024.csv",
    "data/audit/anes_cdf_20260205_analysis_ready_pairwise_cell_audit_2004_2012_2016_2020_2024.csv",
    "data/audit/anes_cdf_20260205_nine_node_complete_case_polychoric_diagnostics_2004_2012_2016_2020_2024.rds",
    "data/audit/anes_cdf_20260205_nine_node_complete_case_polychoric_diagnostics_2004_2012_2016_2020_2024.csv"
  )
)

rds_files <- expected_output_files[
  grepl(
    "\\.rds$",
    expected_output_files,
    ignore.case = TRUE
  )
]

csv_files <- expected_output_files[
  grepl(
    "\\.csv$",
    expected_output_files,
    ignore.case = TRUE
  )
]
```

```{r preflight-helpers}
read_yaml_header <- function(path) {
  lines <- readLines(
    path,
    warn = FALSE
  )
  if (
    length(lines) < 3L ||
      trimws(lines[1L]) != "---"
  ) {
    stop(
      "Opening YAML delimiter is missing."
    )
  }
  closing_relative <- which(
    trimws(lines[-1L]) == "---"
  )
  if (length(closing_relative) < 1L) {
    stop(
      "Closing YAML delimiter is missing."
    )
  }
  closing_line <- closing_relative[1L] + 1L
  yaml_text <- paste(
    lines[2L:(closing_line - 1L)],
    collapse = "\n"
  )
  yaml::yaml.load(yaml_text)
}

safe_read_rds <- function(path) {
  tryCatch(
    list(
      value = readRDS(path),
      error = NA_character_
    ),
    error = function(condition) {
      list(
        value = NULL,
        error = conditionMessage(condition)
      )
    }
  )
}

safe_read_csv <- function(path) {
  tryCatch(
    list(
      value = utils::read.csv(
        path,
        stringsAsFactors = FALSE,
        check.names = FALSE
      ),
      error = NA_character_
    ),
    error = function(condition) {
      list(
        value = NULL,
        error = conditionMessage(condition)
      )
    }
  )
}

sha256_file <- function(path) {
  digest::digest(
    path,
    algo = "sha256",
    file = TRUE,
    serialize = FALSE
  )
}
```

```{r preflight-code-and-environment}
cleaning_lines <- readLines(
  cleaning_rmd,
  warn = FALSE
)

cleaning_root_portable <-
  any(
    grepl(
      "ANES[.]Rproj",
      cleaning_lines
    )
  ) &&
  !any(
    grepl(
      'normalizePath\\([[:space:]]*"[.][.]"',
      cleaning_lines
    )
  )

analysis_ready_yaml_error <- NA_character_
analysis_ready_yaml <- tryCatch(
  read_yaml_header(analysis_ready_rmd),
  error = function(condition) {
    analysis_ready_yaml_error <<- conditionMessage(condition)
    NULL
  }
)
analysis_ready_yaml_valid <-
  is.na(analysis_ready_yaml_error) &&
  is.list(analysis_ready_yaml) &&
  !is.null(analysis_ready_yaml$title) &&
  !is.null(analysis_ready_yaml$output)

lock_read_error <- NA_character_
renv_lock <- tryCatch(
  jsonlite::fromJSON(
    renv_lockfile,
    simplifyVector = FALSE
  ),
  error = function(condition) {
    lock_read_error <<- conditionMessage(condition)
    NULL
  }
)

locked_bootnet_version <- if (
  is.list(renv_lock) &&
    is.list(renv_lock$Packages$bootnet)
) {
  as.character(
    renv_lock$Packages$bootnet$Version
  )
} else {
  NA_character_
}

installed_bootnet_version <- if (
  requireNamespace(
    "bootnet",
    quietly = TRUE
  )
) {
  as.character(
    utils::packageVersion("bootnet")
  )
} else {
  NA_character_
}

bootnet_exact_version <-
  is.na(lock_read_error) &&
  identical(
    locked_bootnet_version,
    "1.9.1"
  ) &&
  identical(
    installed_bootnet_version,
    "1.9.1"
  )

code_environment_checks <- data.frame(
  Check = c(
    "03 project-root resolver is portable",
    "04 YAML header parses",
    "renv.lock parses as JSON",
    "bootnet is locked and installed at 1.9.1"
  ),
  Passed = c(
    cleaning_root_portable,
    analysis_ready_yaml_valid,
    is.na(lock_read_error),
    bootnet_exact_version
  ),
  Detail = c(
    NA_character_,
    analysis_ready_yaml_error,
    lock_read_error,
    paste0(
      "locked=",
      locked_bootnet_version,
      "; installed=",
      installed_bootnet_version
    )
  ),
  stringsAsFactors = FALSE
)

code_environment_checks
```

```{r preflight-saved-artifacts}
html_files_complete <- all(
  file.exists(html_files) &
    file.info(html_files)$size > 0
)

expected_outputs_complete <- all(
  file.exists(expected_output_files) &
    file.info(expected_output_files)$size > 0
)

rds_checks <- setNames(
  lapply(
    rds_files,
    safe_read_rds
  ),
  basename(rds_files)
)

csv_checks <- setNames(
  lapply(
    csv_files,
    safe_read_csv
  ),
  basename(csv_files)
)

rds_readable <- all(
  vapply(
    rds_checks,
    function(check) is.na(check$error),
    logical(1)
  )
)

csv_readable <- all(
  vapply(
    csv_checks,
    function(check) {
      is.na(check$error) &&
        is.data.frame(check$value) &&
        nrow(check$value) > 0L
    },
    logical(1)
  )
)

analysis_ready_name <- basename(
  expected_output_files[3L]
)
polychoric_name <- basename(
  expected_output_files[7L]
)

analysis_ready_data <- rds_checks[[analysis_ready_name]]$value
polychoric_diagnostics <- rds_checks[[polychoric_name]]$value

expected_years <- c(
  2004L,
  2012L,
  2016L,
  2020L,
  2024L
)
expected_counts <- c(
  `2004` = 789L,
  `2012` = 4353L,
  `2016` = 2709L,
  `2020` = 5390L,
  `2024` = 3525L
)
network_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223"
)

analysis_ready_structure_valid <-
  is.data.frame(analysis_ready_data) &&
  all(c("VCF0004", network_vars) %in%
        names(analysis_ready_data)) &&
  nrow(analysis_ready_data) == sum(expected_counts) &&
  all(
    stats::complete.cases(
      analysis_ready_data[, network_vars]
    )
  )

analysis_ready_counts_valid <- FALSE
if (analysis_ready_structure_valid) {
  observed_counts <- table(
    analysis_ready_data$VCF0004
  )
  analysis_ready_counts_valid <-
    identical(
      names(observed_counts),
      names(expected_counts)
    ) &&
    identical(
      as.integer(observed_counts),
      as.integer(expected_counts)
    )
}

analysis_ready_hashes_valid <- FALSE
polychoric_diagnostics_valid <- FALSE
if (is.list(polychoric_diagnostics)) {
  actual_analysis_ready_file_sha <- sha256_file(
    expected_output_files[3L]
  )
  actual_analysis_ready_object_sha <- digest::digest(
    analysis_ready_data,
    algo = "sha256",
    serialize = TRUE
  )
  analysis_ready_hashes_valid <- identical(
    tolower(actual_analysis_ready_file_sha),
    tolower(polychoric_diagnostics$Analysis_ready_file_SHA256)
  ) &&
    identical(
      tolower(actual_analysis_ready_object_sha),
      tolower(polychoric_diagnostics$Analysis_ready_object_SHA256)
    ) &&
    identical(
      tolower(actual_analysis_ready_object_sha),
      tolower(polychoric_diagnostics$Expected_analysis_ready_object_SHA256)
    )

  matrix_audit <- polychoric_diagnostics$Matrix_audit
  warning_audit <- polychoric_diagnostics$Warning_audit

  polychoric_diagnostics_valid <-
    identical(
      as.integer(polychoric_diagnostics$Election_years),
      expected_years
    ) &&
    identical(
      as.character(polychoric_diagnostics$Network_vars),
      network_vars
    ) &&
    is.data.frame(matrix_audit) &&
    nrow(matrix_audit) == 5L &&
    all(matrix_audit$Dimension_ok) &&
    all(matrix_audit$Node_order_ok) &&
    all(matrix_audit$All_values_finite) &&
    all(matrix_audit$Positive_definite) &&
    is.data.frame(warning_audit) &&
    nrow(warning_audit) == 5L &&
    all(warning_audit$Warning_n == 0L) &&
    length(polychoric_diagnostics$Matrices) == 5L
}

reproduction_report_passed <- FALSE
if (file.exists(reproduction_report)) {
  reproduction_lines <- readLines(
    reproduction_report,
    warn = FALSE
  )
  reproduction_report_passed <- any(
    trimws(reproduction_lines) ==
      "OVERALL_STATUS: PASS"
  )
}

artifact_checks <- data.frame(
  Check = c(
    "00-06 HTML files exist and are non-empty",
    "Expected 02-06 outputs exist and are non-empty",
    "Saved RDS files are readable",
    "Saved audit CSV files are readable and non-empty",
    "Analysis-ready object has the frozen nodes and complete cases",
    "Wave counts match 789, 4353, 2709, 5390 and 3525",
    "Analysis-ready file and object SHA-256 values match the diagnostic object",
    "All five saved polychoric matrices pass recorded diagnostics",
    "Blank-session 00-06 reproduction report records PASS"
  ),
  Passed = c(
    html_files_complete,
    expected_outputs_complete,
    rds_readable,
    csv_readable,
    analysis_ready_structure_valid,
    analysis_ready_counts_valid,
    analysis_ready_hashes_valid,
    polychoric_diagnostics_valid,
    reproduction_report_passed
  ),
  stringsAsFactors = FALSE
)

artifact_checks
```

```{r preflight-gate}
preflight_audit <- rbind(
  code_environment_checks[, c("Check", "Passed")],
  artifact_checks
)

preflight_passed <- all(
  preflight_audit$Passed
)

preflight_audit

data.frame(
  Preflight_passed = preflight_passed
)
```

## File-Hash Inventory

```{r file-hash-inventory}
files_for_hashing <- c(
  cleaning_rmd,
  analysis_ready_rmd,
  renv_lockfile,
  html_files,
  expected_output_files,
  reproduction_report
)

hash_inventory <- data.frame(
  File = basename(files_for_hashing),
  SHA256 = vapply(
    files_for_hashing,
    function(path) {
      if (file.exists(path)) {
        sha256_file(path)
      } else {
        NA_character_
      }
    },
    character(1)
  ),
  Size_bytes = ifelse(
    file.exists(files_for_hashing),
    file.info(files_for_hashing)$size,
    NA_real_
  ),
  stringsAsFactors = FALSE
)

hash_inventory
```

## Method Boundary

The local full texts by Epskamp and Fried, Epskamp, Borsboom and Fried, and Burger and colleagues support transparent ordinal/polychoric estimation, row-level nonparametric resampling, case-dropping stability assessment, and cautious interpretation. The preflight checks the D15 and D21 approval records and technical readiness separately.

## Execution Boundary and Excluded Analyses

- The preflight section does not read or modify an ANES raw file, estimate a network, or run a bootstrap.
- D15 and D21 passed before the formal five-wave execution. The current registry records D21 as `COMPLETE_AND_VALIDATED`.
- The standard HTML report defaults to reloading and revalidating the saved formal model objects; it does not rerun point estimation or bootstrap resampling.
- A new formal rerun requires D21 to be returned explicitly to an authorised execution state and requires the corresponding run control to be enabled.
- Case-dropping bootstrap, centrality, NCT, education-group analysis, sensitivity analysis, and substantive political interpretation remain outside Step 07.

## Conditions Applied to the Completed Formal Execution

1. D15 recorded the approved primary unweighted treatment and its restricted inferential scope; D21 recorded researcher approval before execution and now records completion and validation. Both rows were unique and readable from the configured registry.
2. Every technical preflight check passed before model execution.
3. The formal workflow comprised five point networks and five wave-specific 1,000-repetition nonparametric edge-weight bootstraps.

## Bounded 2004 Pilot Authorisation

The historical 2004 pilot uses the separate P07 authorisation record. D15 and D21 govern the formal five-wave analysis.

```{r pilot-authorisation}
pilot_authorisation_config <- file.path(
  project_root,
  "config",
  "step07_pilot_authorisation_path.txt"
)

pilot_authorisation_path <- if (
  file.exists(pilot_authorisation_config)
) {
  configured_lines <- trimws(
    readLines(
      pilot_authorisation_config,
      warn = FALSE
    )
  )

  configured_lines <- configured_lines[
    nzchar(configured_lines)
  ]

  if (length(configured_lines) == 1L) {
    configured_lines
  } else {
    NA_character_
  }
} else {
  NA_character_
}

pilot_authorisation_error <- NA_character_

pilot_authorisation <- tryCatch(
  {
    if (
      length(pilot_authorisation_path) != 1L ||
        is.na(pilot_authorisation_path) ||
        !file.exists(pilot_authorisation_path)
    ) {
      stop(
        "Configured pilot authorisation is unavailable."
      )
    }

    utils::read.csv(
      pilot_authorisation_path,
      stringsAsFactors = FALSE,
      check.names = FALSE
    )
  },
  error = function(condition) {
    pilot_authorisation_error <<-
      conditionMessage(condition)
    data.frame()
  }
)

required_pilot_columns <- c(
  "authorisation_id",
  "researcher_approval_status",
  "implementation_status",
  "election_year",
  "sample_scope",
  "node_count",
  "missing_data",
  "node_type",
  "correlation",
  "estimator",
  "gamma",
  "weight_treatment",
  "bootstrap_type",
  "bootstrap_counts",
  "allowed_statistics",
  "prohibited_analyses",
  "approved_at",
  "approval_source"
)

pilot_schema_valid <- all(
  required_pilot_columns %in%
    names(pilot_authorisation)
)

pilot_row <- if (pilot_schema_valid) {
  pilot_authorisation[
    pilot_authorisation$authorisation_id ==
      "P07_2004_BOOTSTRAP_PARAMETER_PILOT",
    ,
    drop = FALSE
  ]
} else {
  data.frame()
}

pilot_row_unique <- nrow(pilot_row) == 1L

pilot_prohibitions_required <- c(
  "FIVE_WAVE_FORMAL_GGM",
  "CENTRALITY",
  "NCT",
  "SENSITIVITY_ANALYSES",
  "SUBSTANTIVE_INTERPRETATION"
)

pilot_prohibitions_recorded <- if (pilot_row_unique) {
  recorded <- strsplit(
    as.character(
      pilot_row$prohibited_analyses
    ),
    "\\|"
  )[[1L]]

  all(
    pilot_prohibitions_required %in%
      recorded
  )
} else {
  FALSE
}

pilot_authorisation_passed <-
  is.na(pilot_authorisation_error) &&
  pilot_schema_valid &&
  pilot_row_unique &&
  normalise_status(
    pilot_row$researcher_approval_status
  ) == "APPROVED" &&
  normalise_status(
    pilot_row$implementation_status
  ) %in% c(
    "AUTHORISED",
    "AUTHORIZED"
  ) &&
  as.integer(pilot_row$election_year) == 2004L &&
  normalise_status(pilot_row$sample_scope) ==
    "OVERALL_COMPLETE_CASE" &&
  as.integer(pilot_row$node_count) == 9L &&
  normalise_status(pilot_row$missing_data) ==
    "COMPLETE_CASE" &&
  normalise_status(pilot_row$node_type) ==
    "ORDERED" &&
  normalise_status(pilot_row$correlation) ==
    "POLYCHORIC" &&
  normalise_status(pilot_row$estimator) ==
    "EBICGLASSO" &&
  isTRUE(
    all.equal(
      as.numeric(pilot_row$gamma),
      0.50,
      tolerance = 1e-12
    )
  ) &&
  normalise_status(pilot_row$weight_treatment) ==
    "UNWEIGHTED" &&
  normalise_status(pilot_row$bootstrap_type) ==
    "NONPARAMETRIC" &&
  identical(
    as.character(pilot_row$bootstrap_counts),
    "10|1000|2500"
  ) &&
  normalise_status(pilot_row$allowed_statistics) ==
    "EDGE" &&
  normalise_status(pilot_row$approval_source) ==
    "EXPLICIT_RESEARCHER_APPROVAL" &&
  pilot_prohibitions_recorded

pilot_authorisation_audit <- data.frame(
  Authorisation_path = pilot_authorisation_path,
  Read_error = pilot_authorisation_error,
  Schema_valid = pilot_schema_valid,
  Authorisation_row_unique = pilot_row_unique,
  Required_prohibitions_recorded =
    pilot_prohibitions_recorded,
  Pilot_authorisation_passed =
    pilot_authorisation_passed,
  stringsAsFactors = FALSE
)

pilot_authorisation_audit
```

## Session Information

```{r session-information}
sessionInfo()
```

## Hard Execution Boundary

All model-estimation code must appear after this chunk. Formal analysis requires the existing D15/D21 gate. The bounded 2004 pilot requires the separate P07 gate. If neither execution route passes together with the technical preflight, rendering ends here; interactive execution raises an error.

```{r fail-closed-execution-boundary}
formal_model_execution_allowed <-
  authorisation_passed &&
  preflight_passed

pilot_execution_allowed <-
  pilot_authorisation_passed &&
  preflight_passed

# Retain the historical object name only as a formal-analysis alias.
# Pilot code must never test this alias.
model_execution_allowed <-
  formal_model_execution_allowed

data.frame(
  Formal_authorisation_passed =
    authorisation_passed,
  Pilot_authorisation_passed =
    pilot_authorisation_passed,
  Technical_preflight_passed = preflight_passed,
  Formal_model_execution_allowed =
    formal_model_execution_allowed,
  Pilot_execution_allowed =
    pilot_execution_allowed
)

if (
  !formal_model_execution_allowed &&
    !pilot_execution_allowed
) {
  stop_message <- paste(
    "STEP 07 EXECUTION STOPPED:",
    "neither formal D15/D21 authorisation nor bounded P07 pilot",
    "authorisation passed together with all technical checks."
  )

  if (isTRUE(knitr::is_html_output())) {
    message(stop_message)
    knitr::knit_exit()
  } else {
    stop(stop_message)
  }
}
```

## Step-07 2004 Pilot Network and Bootstrap Parameter Assessment

This section is executed only after the bounded P07 pilot authorisation and
technical preflight gates have passed. It estimates the approved 2004 pilot
network and conducts staged nonparametric edge-weight bootstrap checks.
It does not run centrality, NCT, sensitivity analyses or substantive
political interpretation.

### Pilot Run Controls

Only one bootstrap run should be activated at a time. Start with the
10-repetition technical test. The 1,000- and 2,500-repetition runs must
remain disabled until the preceding stage has completed successfully.

```{r pilot-run-controls}
stopifnot(
  isTRUE(pilot_execution_allowed)
)

RUN_BOOTSTRAP_10 <- FALSE
RUN_BOOTSTRAP_1000 <- FALSE
RUN_BOOTSTRAP_2500 <- FALSE
RUN_PARAMETER_COMPARISON <- FALSE

active_bootstrap_runs <- sum(
  c(
    RUN_BOOTSTRAP_10,
    RUN_BOOTSTRAP_1000,
    RUN_BOOTSTRAP_2500
  )
)

stopifnot(
  active_bootstrap_runs <= 1L
)

data.frame(
  Run = c(
    "10-repetition technical test",
    "1,000-repetition candidate",
    "2,500-repetition candidate",
    "Compare saved 1,000 and 2,500 runs"
  ),
  Enabled = c(
    RUN_BOOTSTRAP_10,
    RUN_BOOTSTRAP_1000,
    RUN_BOOTSTRAP_2500,
    RUN_PARAMETER_COMPARISON
  ),
  stringsAsFactors = FALSE
)
```

### Confirm Required Packages

```{r pilot-package-check}
required_model_packages <- c(
  "bootnet",
  "qgraph",
  "lavaan",
  "ggplot2"
)

model_package_check <- data.frame(
  Package = required_model_packages,
  Installed = vapply(
    required_model_packages,
    requireNamespace,
    quietly = TRUE,
    FUN.VALUE = logical(1)
  ),
  Version = vapply(
    required_model_packages,
    function(package_name) {
      if (
        requireNamespace(
          package_name,
          quietly = TRUE
        )
      ) {
        as.character(
          utils::packageVersion(package_name)
        )
      } else {
        NA_character_
      }
    },
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

model_package_check

stopifnot(
  all(model_package_check$Installed),
  identical(
    model_package_check$Version[
      model_package_check$Package == "bootnet"
    ],
    "1.9.1"
  )
)
```

### Prepare Output Directories

These directories contain only derived pilot outputs. No raw ANES file
is modified.

```{r pilot-output-directories}
pilot_model_dir <- file.path(
  project_root,
  "outputs",
  "models",
  "pilot"
)

pilot_table_dir <- file.path(
  project_root,
  "outputs",
  "tables",
  "pilot"
)

pilot_figure_dir <- file.path(
  project_root,
  "outputs",
  "figures",
  "pilot"
)

pilot_log_dir <- file.path(
  project_root,
  "outputs",
  "logs"
)

dir.create(
  pilot_model_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

dir.create(
  pilot_table_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

dir.create(
  pilot_figure_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

dir.create(
  pilot_log_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

stopifnot(
  dir.exists(pilot_model_dir),
  dir.exists(pilot_table_dir),
  dir.exists(pilot_figure_dir),
  dir.exists(pilot_log_dir)
)
```

### Prepare the Ordered 2004 Pilot Data

```{r prepare-2004-pilot-data}
network_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223"
)

expected_categories <- list(
  VCF0806  = c(1, 2, 3, 4, 5, 6, 7),
  VCF0809  = c(1, 2, 3, 4, 5, 6, 7),
  VCF0838  = c(1, 2, 3, 4),
  VCF0839  = c(1, 2, 3, 4, 5, 6, 7),
  VCF0879a = c(1, 3, 5),
  VCF0888  = c(1, 2, 3),
  VCF0890  = c(1, 2, 3),
  VCF0894  = c(1, 2, 3),
  VCF9223  = c(1, 2, 3, 4)
)

analysis_ready_file <- file.path(
  project_root,
  "data",
  "analysis_ready",
  paste0(
    "anes_cdf_20260205_analysis_ready_nine_node_",
    "complete_case_2004_2012_2016_2020_2024.rds"
  )
)

stopifnot(
  file.exists(analysis_ready_file)
)

anes_analysis_ready <- readRDS(
  analysis_ready_file
)

pilot_data_2004 <- anes_analysis_ready[
  anes_analysis_ready$VCF0004 == 2004,
  network_vars,
  drop = FALSE
]

for (variable_name in network_vars) {
  pilot_data_2004[[variable_name]] <- ordered(
    as.numeric(
      pilot_data_2004[[variable_name]]
    ),
    levels = expected_categories[[variable_name]]
  )
}

pilot_input_audit <- data.frame(
  Check = c(
    "Input contains 789 respondents",
    "Input contains nine nodes",
    "Node order matches the frozen specification",
    "All nodes are ordered factors",
    "Input contains no missing values"
  ),
  Passed = c(
    nrow(pilot_data_2004) == 789L,
    ncol(pilot_data_2004) == 9L,
    identical(
      names(pilot_data_2004),
      network_vars
    ),
    all(
      vapply(
        pilot_data_2004,
        is.ordered,
        logical(1)
      )
    ),
    all(
      stats::complete.cases(
        pilot_data_2004
      )
    )
  ),
  stringsAsFactors = FALSE
)

pilot_input_audit

stopifnot(
  all(pilot_input_audit$Passed)
)
```

### Estimate the 2004 Pilot Network

The pilot uses the approved primary specification: complete cases,
polychoric correlations, EBICglasso and gamma = 0.50.

```{r estimate-2004-pilot-network}
set.seed(20260829)

network_2004 <- bootnet::estimateNetwork(
  data = pilot_data_2004,
  default = "EBICglasso",
  tuning = 0.50,
  corMethod = "cor_auto",
  missing = "listwise",
  sampleSize = "minimum",
  corArgs = list(
    detectOrdinal = TRUE,
    ordinalLevelMax = 7,
    forcePD = FALSE
  ),
  refit = FALSE,
  lambda.min.ratio = 0.01,
  nlambda = 100,
  threshold = FALSE,
  nonPositiveDefinite = "stop",
  verbose = TRUE
)

print(network_2004)

pilot_network_audit <- data.frame(
  Check = c(
    "Network contains 789 respondents",
    "Network contains nine nodes",
    "Network node order is correct",
    "All edge weights are finite",
    "Network is symmetric",
    "Network diagonal is zero"
  ),
  Passed = c(
    network_2004$nPerson == 789L,
    network_2004$nNode == 9L,
    identical(
      rownames(network_2004$graph),
      network_vars
    ) &&
      identical(
        colnames(network_2004$graph),
        network_vars
      ),
    all(
      is.finite(
        network_2004$graph
      )
    ),
    max(
      abs(
        network_2004$graph -
          t(network_2004$graph)
      )
    ) < 1e-12,
    max(
      abs(
        diag(network_2004$graph)
      )
    ) < 1e-12
  ),
  stringsAsFactors = FALSE
)

pilot_network_audit

stopifnot(
  all(pilot_network_audit$Passed)
)
```

### Verify the Pilot Network Against the Saved Polychoric Matrix

```{r verify-2004-pilot-network}
polychoric_bundle_file <- file.path(
  project_root,
  "data",
  "audit",
  paste0(
    "anes_cdf_20260205_nine_node_complete_case_",
    "polychoric_diagnostics_2004_2012_2016_2020_2024.rds"
  )
)

stopifnot(
  file.exists(polychoric_bundle_file)
)

polychoric_bundle <- readRDS(
  polychoric_bundle_file
)

stopifnot(
  "2004" %in%
    names(polychoric_bundle$Matrices)
)

reference_network_2004 <- qgraph::EBICglasso(
  S = polychoric_bundle$Matrices[["2004"]],
  n = 789,
  gamma = 0.50,
  nlambda = 100,
  lambda.min.ratio = 0.01,
  refit = FALSE,
  threshold = FALSE
)

maximum_network_difference <- max(
  abs(
    network_2004$graph -
      reference_network_2004
  )
)

network_reproduction_audit <- data.frame(
  Maximum_absolute_difference =
    maximum_network_difference,
  Matches_saved_polychoric_input =
    maximum_network_difference < 1e-8,
  stringsAsFactors = FALSE
)

network_reproduction_audit

stopifnot(
  maximum_network_difference < 1e-8
)
```

### Save and Plot the 2004 Pilot Network

```{r save-2004-pilot-network}
pilot_network_file <- file.path(
  pilot_model_dir,
  "2004_pilot_network_gamma050.rds"
)

saveRDS(
  network_2004,
  pilot_network_file
)

png(
  filename = file.path(
    pilot_figure_dir,
    "2004_pilot_network_gamma050.png"
  ),
  width = 2400,
  height = 2000,
  res = 300
)

plot(
  network_2004,
  layout = "spring",
  theme = "colorblind"
)

grDevices::dev.off()

stopifnot(
  file.exists(pilot_network_file)
)
```

### Define the Nonparametric Edge-Weight Bootstrap Function

This function stores only edge weights. It does not calculate
centrality. The warning log records warnings exposed by `bootnet`,
including retry warnings. Internal estimator warnings suppressed by
`bootnet` cannot be interpreted as having been comprehensively audited.

```{r define-edge-bootstrap-function}
run_edge_bootstrap <- function(
  network_object,
  n_boots,
  random_seed
) {
  recorded_warnings <- character(0)

  set.seed(random_seed)

  elapsed_time <- system.time({
    bootstrap_object <- withCallingHandlers(
      bootnet::bootnet(
        data = network_object,
        nBoots = n_boots,
        type = "nonparametric",
        statistics = "edge",
        computeCentrality = FALSE,
        nCores = 1,
        memorysaver = TRUE,
        maxErrors = 10,
        verbose = TRUE
      ),
      warning = function(warning_condition) {
        recorded_warnings <<- c(
          recorded_warnings,
          conditionMessage(warning_condition)
        )

        invokeRestart(
          "muffleWarning"
        )
      }
    )
  })

  list(
    object = bootstrap_object,
    nBoots = n_boots,
    seed = random_seed,
    elapsed_time = elapsed_time,
    bootnet_level_warnings =
      unique(recorded_warnings),
    R_version = R.version.string,
    bootnet_version = as.character(
      utils::packageVersion("bootnet")
    )
  )
}

write_bootstrap_log <- function(
  result_object,
  output_path
) {
  warning_text <- if (
    length(
      result_object$bootnet_level_warnings
    ) == 0L
  ) {
    "NO BOOTNET-LEVEL WARNINGS CAPTURED"
  } else {
    paste(
      result_object$bootnet_level_warnings,
      collapse = "\n"
    )
  }

  log_lines <- c(
    paste0(
      "Requested bootstrap repetitions: ",
      result_object$nBoots
    ),
    paste0(
      "Random seed: ",
      result_object$seed
    ),
    paste0(
      "bootnet version: ",
      result_object$bootnet_version
    ),
    paste0(
      "Elapsed seconds: ",
      unname(
        result_object$elapsed_time[["elapsed"]]
      )
    ),
    "",
    "Captured bootnet-level warnings:",
    warning_text
  )

  writeLines(
    log_lines,
    con = output_path
  )
}
```
### Validate Completed Edge Bootstrap Runs

The following function verifies that the requested number of bootstrap
networks was returned and that all stored edge estimates and network
matrices are structurally usable. Retry counts are reported separately
because no universal acceptable retry threshold is supplied by the
methodological literature.

```{r define-edge-bootstrap-validation}
count_retry_warnings <- function(warning_text) {
  if (length(warning_text) == 0L) {
    return(0L)
  }

  retry_pattern <-
    "^([0-9]+) bootstrap estimation\\(s\\) failed and were resampled\\.$"

  warning_parts <- regmatches(
    warning_text,
    regexec(
      retry_pattern,
      warning_text
    )
  )

  retry_values <- vapply(
    warning_parts,
    function(value) {
      if (length(value) >= 2L) {
        as.integer(value[2L])
      } else {
        0L
      }
    },
    integer(1)
  )

  sum(retry_values)
}

validate_edge_bootstrap <- function(
  result_object,
  expected_n_boots,
  expected_node_names,
  expected_n_person,
  reference_graph
) {
  boot_object <- result_object$object
  node_n <- length(expected_node_names)
  edge_n <- choose(node_n, 2L)

  graph_checks <- vapply(
    boot_object$boots,
    function(one_boot) {
      graph <- one_boot$graph

      is.matrix(graph) &&
        identical(
          dim(graph),
          c(node_n, node_n)
        ) &&
        identical(
          rownames(graph),
          expected_node_names
        ) &&
        identical(
          colnames(graph),
          expected_node_names
        ) &&
        all(is.finite(graph)) &&
        max(
          abs(
            graph - t(graph)
          )
        ) < 1e-12 &&
        max(
          abs(
            diag(graph)
          )
        ) < 1e-12 &&
        identical(
          as.integer(one_boot$nPerson),
          as.integer(expected_n_person)
        )
    },
    logical(1)
  )

  sample_graph_difference <- max(
    abs(
      boot_object$sample$graph -
        reference_graph
    )
  )

  validation_audit <- data.frame(
    Check = c(
      "Requested bootstrap count recorded",
      "Returned bootstrap network count",
      "Bootstrap edge-table row count",
      "Distinct bootstrap identifiers",
      "Only edge statistics stored",
      "All bootstrap edge values finite",
      "Original-sample edge table complete",
      "All returned bootstrap graphs valid",
      "Bootstrap original sample matches pilot graph"
    ),
    Observed = c(
      as.character(result_object$nBoots),
      as.character(length(boot_object$boots)),
      as.character(nrow(boot_object$bootTable)),
      as.character(
        length(
          unique(
            boot_object$bootTable$name
          )
        )
      ),
      paste(
        unique(
          boot_object$bootTable$type
        ),
        collapse = "|"
      ),
      as.character(
        sum(
          is.finite(
            boot_object$bootTable$value
          )
        )
      ),
      as.character(
        nrow(
          boot_object$sampleTable
        )
      ),
      paste0(
        sum(graph_checks),
        "/",
        length(graph_checks)
      ),
      format(
        sample_graph_difference,
        scientific = TRUE
      )
    ),
    Expected = c(
      as.character(expected_n_boots),
      as.character(expected_n_boots),
      as.character(
        expected_n_boots * edge_n
      ),
      as.character(expected_n_boots),
      "edge",
      as.character(
        expected_n_boots * edge_n
      ),
      as.character(edge_n),
      paste0(
        expected_n_boots,
        "/",
        expected_n_boots
      ),
      "< 1e-8"
    ),
    Passed = c(
      identical(
        as.integer(result_object$nBoots),
        as.integer(expected_n_boots)
      ),
      length(boot_object$boots) ==
        expected_n_boots,
      nrow(boot_object$bootTable) ==
        expected_n_boots * edge_n,
      length(
        unique(
          boot_object$bootTable$name
        )
      ) == expected_n_boots,
      identical(
        unique(
          boot_object$bootTable$type
        ),
        "edge"
      ),
      all(
        is.finite(
          boot_object$bootTable$value
        )
      ),
      nrow(boot_object$sampleTable) ==
        edge_n,
      all(graph_checks),
      sample_graph_difference < 1e-8
    ),
    stringsAsFactors = FALSE
  )

  list(
    audit = validation_audit,
    top_level_retry_count =
      count_retry_warnings(
        result_object$bootnet_level_warnings
      ),
    captured_warning_n =
      length(
        result_object$bootnet_level_warnings
      ),
    captured_warnings =
      result_object$bootnet_level_warnings
  )
}
```
### Run the 10-Repetition Technical Test

This run checks only whether the code completes. Its estimates must not
be reported or interpreted.

```{r run-edge-bootstrap-10}
if (RUN_BOOTSTRAP_10) {
  bootstrap_10 <- run_edge_bootstrap(
    network_object = network_2004,
    n_boots = 10,
    random_seed = 20260829
  )

  print(
    bootstrap_10$object
  )

  bootstrap_10$elapsed_time
  bootstrap_10$bootnet_level_warnings

  bootstrap_10_validation <- validate_edge_bootstrap(
  result_object = bootstrap_10,
  expected_n_boots = 10L,
  expected_node_names = network_vars,
  expected_n_person = 789L,
  reference_graph = network_2004$graph
)

bootstrap_10_validation$audit

bootstrap_10_diagnostics <- data.frame(
  Requested_bootstraps = 10L,
  Returned_bootstraps =
    length(
      bootstrap_10$object$boots
    ),
  Top_level_retry_count =
    bootstrap_10_validation$
      top_level_retry_count,
  Captured_warning_n =
    bootstrap_10_validation$
      captured_warning_n,
  Captured_warnings =
    if (
      bootstrap_10_validation$
        captured_warning_n == 0L
    ) {
      "NONE"
    } else {
      paste(
        bootstrap_10_validation$
          captured_warnings,
        collapse = " | "
      )
    },
  stringsAsFactors = FALSE
)

bootstrap_10_diagnostics

stopifnot(
  all(
    bootstrap_10_validation$
      audit$Passed
  )
)

write.csv(
  bootstrap_10_validation$audit,
  file.path(
    pilot_table_dir,
    "2004_edge_bootstrap_10_validation.csv"
  ),
  row.names = FALSE
)

write.csv(
  bootstrap_10_diagnostics,
  file.path(
    pilot_table_dir,
    "2004_edge_bootstrap_10_diagnostics.csv"
  ),
  row.names = FALSE
)

bootstrap_10_file <- file.path(
  pilot_model_dir,
  "2004_edge_bootstrap_10_technical_test.rds"
)

saveRDS(
  bootstrap_10,
  bootstrap_10_file
)

bootstrap_10_reloaded <- readRDS(
  bootstrap_10_file
)

stopifnot(
  length(
    bootstrap_10_reloaded$
      object$boots
  ) == 10L
)
write_bootstrap_log(
  result_object = bootstrap_10,
  output_path = file.path(
    pilot_log_dir,
    "07_2004_edge_bootstrap_10.log"
  )
)
}
```

### Run the 1,000-Repetition Candidate

Run this candidate only after the 10-repetition technical test has
completed successfully. This remains a non-inferential computational
pilot and must not be used for substantive edge interpretation.

```{r run-edge-bootstrap-1000}
if (RUN_BOOTSTRAP_1000) {

  # ------------------------------------------------
  # Confirm that the preceding 10-bootstrap gate passed
  # ------------------------------------------------

  bootstrap_10_validation_file <- file.path(
    pilot_table_dir,
    "2004_edge_bootstrap_10_validation.csv"
  )

  bootstrap_10_diagnostics_file <- file.path(
    pilot_table_dir,
    "2004_edge_bootstrap_10_diagnostics.csv"
  )

  stopifnot(
    file.exists(
      bootstrap_10_validation_file
    ),
    file.exists(
      bootstrap_10_diagnostics_file
    )
  )

  bootstrap_10_validation_saved <- read.csv(
    bootstrap_10_validation_file,
    stringsAsFactors = FALSE
  )

  bootstrap_10_diagnostics_saved <- read.csv(
    bootstrap_10_diagnostics_file,
    stringsAsFactors = FALSE
  )

  stopifnot(
    nrow(
      bootstrap_10_validation_saved
    ) == 9L,
    all(
      bootstrap_10_validation_saved$
        Passed
    ),
    bootstrap_10_diagnostics_saved$
      Requested_bootstraps == 10L,
    bootstrap_10_diagnostics_saved$
      Returned_bootstraps == 10L,
    bootstrap_10_diagnostics_saved$
      Top_level_retry_count == 0L,
    bootstrap_10_diagnostics_saved$
      Captured_warning_n == 0L
  )

  # ------------------------------------------------
  # Run the authorised 1,000-bootstrap candidate
  # ------------------------------------------------

  bootstrap_1000 <- run_edge_bootstrap(
    network_object = network_2004,
    n_boots = 1000L,
    random_seed = 20260829
  )

  print(
    bootstrap_1000$object
  )

  bootstrap_1000$elapsed_time
  bootstrap_1000$bootnet_level_warnings

  # ------------------------------------------------
  # Validate the returned bootstrap object
  # ------------------------------------------------

  bootstrap_1000_validation <-
    validate_edge_bootstrap(
      result_object = bootstrap_1000,
      expected_n_boots = 1000L,
      expected_node_names = network_vars,
      expected_n_person = 789L,
      reference_graph = network_2004$graph
    )

  bootstrap_1000_validation$audit

  bootstrap_1000_diagnostics <- data.frame(
    Requested_bootstraps = 1000L,
    Returned_bootstraps =
      length(
        bootstrap_1000$
          object$boots
      ),
    Elapsed_seconds =
      unname(
        bootstrap_1000$
          elapsed_time[["elapsed"]]
      ),
    Top_level_retry_count =
      bootstrap_1000_validation$
        top_level_retry_count,
    Captured_warning_n =
      bootstrap_1000_validation$
        captured_warning_n,
    Captured_warnings =
      if (
        bootstrap_1000_validation$
          captured_warning_n == 0L
      ) {
        "NONE"
      } else {
        paste(
          bootstrap_1000_validation$
            captured_warnings,
          collapse = " | "
        )
      },
    stringsAsFactors = FALSE
  )

  bootstrap_1000_diagnostics

  stopifnot(
    all(
      bootstrap_1000_validation$
        audit$Passed
    )
  )

  # ------------------------------------------------
  # Save and read back the complete bootstrap object
  # ------------------------------------------------

  bootstrap_1000_file <- file.path(
    pilot_model_dir,
    "2004_edge_bootstrap_1000.rds"
  )

  saveRDS(
    bootstrap_1000,
    bootstrap_1000_file
  )

  stopifnot(
    file.exists(
      bootstrap_1000_file
    )
  )

  bootstrap_1000_reloaded <- readRDS(
    bootstrap_1000_file
  )

  stopifnot(
    length(
      bootstrap_1000_reloaded$
        object$boots
    ) == 1000L,
    nrow(
      bootstrap_1000_reloaded$
        object$bootTable
    ) == 36000L,
    all(
      is.finite(
        bootstrap_1000_reloaded$
          object$bootTable$value
      )
    )
  )

  # ------------------------------------------------
  # Produce and save the edge-accuracy summary
  # ------------------------------------------------

  edge_summary_1000 <- as.data.frame(
    summary(
      bootstrap_1000$object,
      statistics = "edge"
    )
  )

  write.csv(
    edge_summary_1000,
    file.path(
      pilot_table_dir,
      "2004_edge_bootstrap_1000_summary.csv"
    ),
    row.names = FALSE
  )

  write.csv(
    bootstrap_1000_validation$audit,
    file.path(
      pilot_table_dir,
      "2004_edge_bootstrap_1000_validation.csv"
    ),
    row.names = FALSE
  )

  write.csv(
    bootstrap_1000_diagnostics,
    file.path(
      pilot_table_dir,
      "2004_edge_bootstrap_1000_diagnostics.csv"
    ),
    row.names = FALSE
  )

  # ------------------------------------------------
  # Produce and save the edge-accuracy figure
  # ------------------------------------------------

  edge_plot_1000 <- plot(
    bootstrap_1000$object,
    statistics = "edge",
    labels = FALSE,
    order = "sample"
  )

  ggplot2::ggsave(
    filename = file.path(
      pilot_figure_dir,
      "2004_edge_bootstrap_1000.png"
    ),
    plot = edge_plot_1000,
    width = 10,
    height = 8,
    dpi = 300
  )

  # ------------------------------------------------
  # Save the execution log
  # ------------------------------------------------

  write_bootstrap_log(
    result_object = bootstrap_1000,
    output_path = file.path(
      pilot_log_dir,
      "07_2004_edge_bootstrap_1000.log"
    )
  )

  # ------------------------------------------------
  # Confirm that every expected output was written
  # ------------------------------------------------

  bootstrap_1000_expected_outputs <- c(
    bootstrap_1000_file,
    file.path(
      pilot_table_dir,
      "2004_edge_bootstrap_1000_summary.csv"
    ),
    file.path(
      pilot_table_dir,
      "2004_edge_bootstrap_1000_validation.csv"
    ),
    file.path(
      pilot_table_dir,
      "2004_edge_bootstrap_1000_diagnostics.csv"
    ),
    file.path(
      pilot_figure_dir,
      "2004_edge_bootstrap_1000.png"
    ),
    file.path(
      pilot_log_dir,
      "07_2004_edge_bootstrap_1000.log"
    )
  )

  stopifnot(
    all(
      file.exists(
        bootstrap_1000_expected_outputs
      )
    ),
    all(
      file.info(
        bootstrap_1000_expected_outputs
      )$size > 0
    )
  )

  data.frame(
    Pilot_stage = "1,000-repetition candidate",
    Requested_bootstraps = 1000L,
    Returned_bootstraps =
      length(
        bootstrap_1000$
          object$boots
      ),
    Structural_validation_passed =
      all(
        bootstrap_1000_validation$
          audit$Passed
      ),
    Top_level_retry_count =
      bootstrap_1000_validation$
        top_level_retry_count,
    Captured_warning_n =
      bootstrap_1000_validation$
        captured_warning_n,
    Elapsed_seconds =
      unname(
        bootstrap_1000$
          elapsed_time[["elapsed"]]
      ),
    Output_files_complete =
      all(
        file.exists(
          bootstrap_1000_expected_outputs
        )
      ),
    stringsAsFactors = FALSE
  )
}
```
### Run the 2,500-Repetition Candidate

Run this candidate only after the 1,000-repetition candidate has
completed successfully and its saved validation, diagnostics and model
files have been checked. This remains a non-inferential computational
pilot and must not be used for substantive edge interpretation.

```{r run-edge-bootstrap-2500}
if (RUN_BOOTSTRAP_2500) {

  # ------------------------------------------------
  # Confirm that the preceding 1,000-bootstrap gate passed
  # ------------------------------------------------

  bootstrap_1000_validation_file <- file.path(
    pilot_table_dir,
    "2004_edge_bootstrap_1000_validation.csv"
  )

  bootstrap_1000_diagnostics_file <- file.path(
    pilot_table_dir,
    "2004_edge_bootstrap_1000_diagnostics.csv"
  )

  bootstrap_1000_summary_file <- file.path(
    pilot_table_dir,
    "2004_edge_bootstrap_1000_summary.csv"
  )

  bootstrap_1000_model_file <- file.path(
    pilot_model_dir,
    "2004_edge_bootstrap_1000.rds"
  )

  bootstrap_1000_figure_file <- file.path(
    pilot_figure_dir,
    "2004_edge_bootstrap_1000.png"
  )

  bootstrap_1000_log_file <- file.path(
    pilot_log_dir,
    "07_2004_edge_bootstrap_1000.log"
  )

  bootstrap_1000_required_files <- c(
    bootstrap_1000_validation_file,
    bootstrap_1000_diagnostics_file,
    bootstrap_1000_summary_file,
    bootstrap_1000_model_file,
    bootstrap_1000_figure_file,
    bootstrap_1000_log_file
  )

  stopifnot(
    all(
      file.exists(
        bootstrap_1000_required_files
      )
    ),
    all(
      file.info(
        bootstrap_1000_required_files
      )$size > 0
    )
  )

  bootstrap_1000_validation_saved <- read.csv(
    bootstrap_1000_validation_file,
    stringsAsFactors = FALSE
  )

  bootstrap_1000_diagnostics_saved <- read.csv(
    bootstrap_1000_diagnostics_file,
    stringsAsFactors = FALSE
  )

  bootstrap_1000_model_saved <- readRDS(
    bootstrap_1000_model_file
  )

  stopifnot(
    nrow(
      bootstrap_1000_validation_saved
    ) == 9L,
    all(
      as.logical(
        bootstrap_1000_validation_saved$Passed
      )
    ),
    nrow(
      bootstrap_1000_diagnostics_saved
    ) == 1L,
    bootstrap_1000_diagnostics_saved$
      Requested_bootstraps == 1000L,
    bootstrap_1000_diagnostics_saved$
      Returned_bootstraps == 1000L,
    bootstrap_1000_diagnostics_saved$
      Top_level_retry_count == 0L,
    bootstrap_1000_diagnostics_saved$
      Captured_warning_n == 0L,
    length(
      bootstrap_1000_model_saved$
        object$boots
    ) == 1000L,
    nrow(
      bootstrap_1000_model_saved$
        object$bootTable
    ) == 36000L,
    all(
      is.finite(
        bootstrap_1000_model_saved$
          object$bootTable$value
      )
    )
  )

  # ------------------------------------------------
  # Run the authorised 2,500-bootstrap candidate
  # ------------------------------------------------

  bootstrap_2500 <- run_edge_bootstrap(
    network_object = network_2004,
    n_boots = 2500L,
    random_seed = 20260829
  )

  print(
    bootstrap_2500$object
  )

  bootstrap_2500$elapsed_time
  bootstrap_2500$bootnet_level_warnings

  # ------------------------------------------------
  # Validate the returned bootstrap object
  # ------------------------------------------------

  bootstrap_2500_validation <-
    validate_edge_bootstrap(
      result_object = bootstrap_2500,
      expected_n_boots = 2500L,
      expected_node_names = network_vars,
      expected_n_person = 789L,
      reference_graph = network_2004$graph
    )

  bootstrap_2500_validation$audit

  bootstrap_2500_diagnostics <- data.frame(
    Requested_bootstraps = 2500L,
    Returned_bootstraps =
      length(
        bootstrap_2500$
          object$boots
      ),
    Elapsed_seconds =
      unname(
        bootstrap_2500$
          elapsed_time[["elapsed"]]
      ),
    Top_level_retry_count =
      bootstrap_2500_validation$
        top_level_retry_count,
    Captured_warning_n =
      bootstrap_2500_validation$
        captured_warning_n,
    Captured_warnings =
      if (
        bootstrap_2500_validation$
          captured_warning_n == 0L
      ) {
        "NONE"
      } else {
        paste(
          bootstrap_2500_validation$
            captured_warnings,
          collapse = " | "
        )
      },
    stringsAsFactors = FALSE
  )

  bootstrap_2500_diagnostics

  stopifnot(
    all(
      bootstrap_2500_validation$
        audit$Passed
    )
  )

  # ------------------------------------------------
  # Save and read back the complete bootstrap object
  # ------------------------------------------------

  bootstrap_2500_file <- file.path(
    pilot_model_dir,
    "2004_edge_bootstrap_2500.rds"
  )

  saveRDS(
    bootstrap_2500,
    bootstrap_2500_file
  )

  stopifnot(
    file.exists(
      bootstrap_2500_file
    )
  )

  bootstrap_2500_reloaded <- readRDS(
    bootstrap_2500_file
  )

  stopifnot(
    length(
      bootstrap_2500_reloaded$
        object$boots
    ) == 2500L,
    nrow(
      bootstrap_2500_reloaded$
        object$bootTable
    ) == 90000L,
    all(
      is.finite(
        bootstrap_2500_reloaded$
          object$bootTable$value
      )
    )
  )

  # ------------------------------------------------
  # Produce and save the edge-accuracy summary
  # ------------------------------------------------

  edge_summary_2500 <- as.data.frame(
    summary(
      bootstrap_2500$object,
      statistics = "edge"
    )
  )

  write.csv(
    edge_summary_2500,
    file.path(
      pilot_table_dir,
      "2004_edge_bootstrap_2500_summary.csv"
    ),
    row.names = FALSE
  )

  write.csv(
    bootstrap_2500_validation$audit,
    file.path(
      pilot_table_dir,
      "2004_edge_bootstrap_2500_validation.csv"
    ),
    row.names = FALSE
  )

  write.csv(
    bootstrap_2500_diagnostics,
    file.path(
      pilot_table_dir,
      "2004_edge_bootstrap_2500_diagnostics.csv"
    ),
    row.names = FALSE
  )

  # ------------------------------------------------
  # Produce and save the edge-accuracy figure
  # ------------------------------------------------

  edge_plot_2500 <- plot(
    bootstrap_2500$object,
    statistics = "edge",
    labels = FALSE,
    order = "sample"
  )

  ggplot2::ggsave(
    filename = file.path(
      pilot_figure_dir,
      "2004_edge_bootstrap_2500.png"
    ),
    plot = edge_plot_2500,
    width = 10,
    height = 8,
    dpi = 300
  )

  # ------------------------------------------------
  # Save the execution log
  # ------------------------------------------------

  write_bootstrap_log(
    result_object = bootstrap_2500,
    output_path = file.path(
      pilot_log_dir,
      "07_2004_edge_bootstrap_2500.log"
    )
  )

  # ------------------------------------------------
  # Confirm that every expected output was written
  # ------------------------------------------------

  bootstrap_2500_expected_outputs <- c(
    bootstrap_2500_file,
    file.path(
      pilot_table_dir,
      "2004_edge_bootstrap_2500_summary.csv"
    ),
    file.path(
      pilot_table_dir,
      "2004_edge_bootstrap_2500_validation.csv"
    ),
    file.path(
      pilot_table_dir,
      "2004_edge_bootstrap_2500_diagnostics.csv"
    ),
    file.path(
      pilot_figure_dir,
      "2004_edge_bootstrap_2500.png"
    ),
    file.path(
      pilot_log_dir,
      "07_2004_edge_bootstrap_2500.log"
    )
  )

  stopifnot(
    all(
      file.exists(
        bootstrap_2500_expected_outputs
      )
    ),
    all(
      file.info(
        bootstrap_2500_expected_outputs
      )$size > 0
    )
  )

  data.frame(
    Pilot_stage = "2,500-repetition candidate",
    Requested_bootstraps = 2500L,
    Returned_bootstraps =
      length(
        bootstrap_2500$
          object$boots
      ),
    Structural_validation_passed =
      all(
        bootstrap_2500_validation$
          audit$Passed
      ),
    Top_level_retry_count =
      bootstrap_2500_validation$
        top_level_retry_count,
    Captured_warning_n =
      bootstrap_2500_validation$
        captured_warning_n,
    Elapsed_seconds =
      unname(
        bootstrap_2500$
          elapsed_time[["elapsed"]]
      ),
    Output_files_complete =
      all(
        file.exists(
          bootstrap_2500_expected_outputs
        )
      ),
    stringsAsFactors = FALSE
  )
}
```
### Compare the Saved 1,000- and 2,500-Repetition Runs

Enable this section only after both candidate RDS files exist. This
section assembles diagnostics; it does not automatically choose a
preferred parameter according to substantive results.

```{r compare-bootstrap-parameters}
if (RUN_PARAMETER_COMPARISON) {
  bootstrap_1000_file <- file.path(
    pilot_model_dir,
    "2004_edge_bootstrap_1000.rds"
  )

  bootstrap_2500_file <- file.path(
    pilot_model_dir,
    "2004_edge_bootstrap_2500.rds"
  )

  stopifnot(
    file.exists(bootstrap_1000_file),
    file.exists(bootstrap_2500_file)
  )

  bootstrap_1000_saved <- readRDS(
    bootstrap_1000_file
  )

  bootstrap_2500_saved <- readRDS(
    bootstrap_2500_file
  )

  comparison_summary_1000 <- as.data.frame(
    summary(
      bootstrap_1000_saved$object,
      statistics = "edge"
    )
  )

  comparison_summary_2500 <- as.data.frame(
    summary(
      bootstrap_2500_saved$object,
      statistics = "edge"
    )
  )

  comparison_summary_1000$Bootstrap_n <- 1000L
  comparison_summary_2500$Bootstrap_n <- 2500L

  bootstrap_edge_summary_combined <- rbind(
    comparison_summary_1000,
    comparison_summary_2500
  )

  bootstrap_run_diagnostics <- data.frame(
    Bootstrap_n = c(
      1000L,
      2500L
    ),
    Seed = c(
      bootstrap_1000_saved$seed,
      bootstrap_2500_saved$seed
    ),
    Elapsed_seconds = c(
      unname(
        bootstrap_1000_saved$
          elapsed_time[["elapsed"]]
      ),
      unname(
        bootstrap_2500_saved$
          elapsed_time[["elapsed"]]
      )
    ),
    Bootnet_warning_n = c(
      length(
        bootstrap_1000_saved$
          bootnet_level_warnings
      ),
      length(
        bootstrap_2500_saved$
          bootnet_level_warnings
      )
    ),
    Bootnet_warnings = c(
      paste(
        bootstrap_1000_saved$
          bootnet_level_warnings,
        collapse = " | "
      ),
      paste(
        bootstrap_2500_saved$
          bootnet_level_warnings,
        collapse = " | "
      )
    ),
    stringsAsFactors = FALSE
  )

  write.csv(
    bootstrap_edge_summary_combined,
    file.path(
      pilot_table_dir,
      paste0(
        "2004_edge_bootstrap_",
        "1000_vs_2500_combined.csv"
      )
    ),
    row.names = FALSE
  )

  write.csv(
    bootstrap_run_diagnostics,
    file.path(
      pilot_table_dir,
      paste0(
        "2004_edge_bootstrap_",
        "parameter_run_diagnostics.csv"
      )
    ),
    row.names = FALSE
  )

  bootstrap_run_diagnostics
}
```

### Post-Pilot Session Information

```{r post-pilot-session-information}
capture.output(
  sessionInfo(),
  file = file.path(
    pilot_log_dir,
    "07_bootstrap_parameter_pilot_sessionInfo.txt"
  )
)

sessionInfo()
```

## Formal Five-Wave Primary Point Networks and Edge Accuracy

This section is governed by D21 and runs only when the formal D15/D21
gate and the technical preflight both pass. It estimates the five frozen
overall-sample point networks and the frozen 1,000-repetition
nonparametric edge-weight bootstrap. It does not calculate centrality,
case-dropping stability, NCT, education-group networks, sensitivities or
substantive political interpretations.

### Formal Run Controls and Output Directories

```{r formal-run-controls-and-directories}
stopifnot(
  isTRUE(formal_model_execution_allowed)
)

# Safe reporting defaults: reuse the completed, validated formal outputs.
# Change a RUN_* flag to TRUE only under a new, explicit rerun decision.
RUN_FORMAL_POINT_NETWORKS <- FALSE
RUN_FORMAL_EDGE_BOOTSTRAPS <- FALSE
REUSE_VALIDATED_FORMAL_OUTPUTS <- TRUE

stopifnot(
  RUN_FORMAL_POINT_NETWORKS ||
    REUSE_VALIDATED_FORMAL_OUTPUTS,
  RUN_FORMAL_EDGE_BOOTSTRAPS ||
    REUSE_VALIDATED_FORMAL_OUTPUTS
)

if (
  RUN_FORMAL_POINT_NETWORKS ||
    RUN_FORMAL_EDGE_BOOTSTRAPS
) {
  stopifnot(
    normalise_status(
      d21$implementation_status
    ) %in% c(
      "AUTHORISED",
      "AUTHORIZED"
    )
  )
}

formal_waves <- c(
  2004L,
  2012L,
  2016L,
  2020L,
  2024L
)

formal_expected_n <- c(
  `2004` = 789L,
  `2012` = 4353L,
  `2016` = 2709L,
  `2020` = 5390L,
  `2024` = 3525L
)

formal_bootstrap_seeds <- c(
  `2004` = 2026083001L,
  `2012` = 2026083002L,
  `2016` = 2026083003L,
  `2020` = 2026083004L,
  `2024` = 2026083005L
)

formal_model_dir <- file.path(
  project_root,
  "outputs",
  "models",
  "formal"
)

formal_table_dir <- file.path(
  project_root,
  "outputs",
  "tables",
  "formal"
)

formal_log_dir <- file.path(
  project_root,
  "outputs",
  "logs"
)

dir.create(
  formal_model_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

dir.create(
  formal_table_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

dir.create(
  formal_log_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

stopifnot(
  dir.exists(formal_model_dir),
  dir.exists(formal_table_dir),
  dir.exists(formal_log_dir),
  identical(
    names(formal_expected_n),
    as.character(formal_waves)
  ),
  identical(
    names(formal_bootstrap_seeds),
    as.character(formal_waves)
  )
)

data.frame(
  Wave = formal_waves,
  Expected_N = unname(formal_expected_n),
  Bootstrap_repetitions = 1000L,
  Bootstrap_seed = unname(formal_bootstrap_seeds),
  Point_network_enabled = RUN_FORMAL_POINT_NETWORKS,
  Edge_bootstrap_enabled = RUN_FORMAL_EDGE_BOOTSTRAPS,
  Reuse_validated_outputs =
    REUSE_VALIDATED_FORMAL_OUTPUTS,
  stringsAsFactors = FALSE
)
```

### Prepare and Validate the Five Ordered Inputs

```{r prepare-formal-five-wave-inputs}
stopifnot(
  exists("anes_analysis_ready"),
  exists("network_vars"),
  exists("expected_categories"),
  exists("polychoric_bundle")
)

prepare_formal_wave_data <- function(wave) {
  wave_data <- anes_analysis_ready[
    anes_analysis_ready$VCF0004 == wave,
    network_vars,
    drop = FALSE
  ]

  for (variable_name in network_vars) {
    wave_data[[variable_name]] <- ordered(
      as.numeric(wave_data[[variable_name]]),
      levels = expected_categories[[variable_name]]
    )
  }

  wave_data
}

formal_wave_data <- setNames(
  lapply(
    formal_waves,
    prepare_formal_wave_data
  ),
  as.character(formal_waves)
)

formal_input_validation <- do.call(
  rbind,
  lapply(
    as.character(formal_waves),
    function(wave_key) {
      wave_data <- formal_wave_data[[wave_key]]
      observed_categories <- lapply(
        wave_data,
        function(value) {
          as.numeric(
            levels(value)[
              levels(value) %in%
                as.character(unique(value))
            ]
          )
        }
      )

      data.frame(
        Wave = as.integer(wave_key),
        N = nrow(wave_data),
        Expected_N = unname(
          formal_expected_n[[wave_key]]
        ),
        Nine_nodes = ncol(wave_data) == 9L,
        Correct_node_order = identical(
          names(wave_data),
          network_vars
        ),
        All_ordered = all(
          vapply(
            wave_data,
            is.ordered,
            logical(1)
          )
        ),
        Complete_cases = all(
          stats::complete.cases(wave_data)
        ),
        Categories_within_specification = all(
          vapply(
            network_vars,
            function(variable_name) {
              all(
                observed_categories[[variable_name]] %in%
                  expected_categories[[variable_name]]
              )
            },
            logical(1)
          )
        ),
        stringsAsFactors = FALSE
      )
    }
  )
)

formal_input_validation$Passed <- with(
  formal_input_validation,
  N == Expected_N &
    Nine_nodes &
    Correct_node_order &
    All_ordered &
    Complete_cases &
    Categories_within_specification
)

write.csv(
  formal_input_validation,
  file.path(
    formal_table_dir,
    "07_formal_five_wave_input_validation.csv"
  ),
  row.names = FALSE
)

formal_input_validation

stopifnot(
  all(formal_input_validation$Passed)
)
```

### Estimate and Validate the Five Point Networks

```{r estimate-and-validate-formal-point-networks}
estimate_formal_network <- function(
  wave_key,
  wave_data
) {
  captured_warnings <- character(0)
  captured_messages <- character(0)

  set.seed(
    formal_bootstrap_seeds[[wave_key]]
  )

  elapsed <- system.time({
    network_object <- withCallingHandlers(
      bootnet::estimateNetwork(
        data = wave_data,
        default = "EBICglasso",
        tuning = 0.50,
        corMethod = "cor_auto",
        missing = "listwise",
        sampleSize = "minimum",
        corArgs = list(
          detectOrdinal = TRUE,
          ordinalLevelMax = 7,
          forcePD = FALSE
        ),
        refit = FALSE,
        lambda.min.ratio = 0.01,
        nlambda = 100,
        threshold = FALSE,
        nonPositiveDefinite = "stop",
        verbose = FALSE
      ),
      warning = function(condition) {
        captured_warnings <<- c(
          captured_warnings,
          conditionMessage(condition)
        )
        invokeRestart("muffleWarning")
      },
      message = function(condition) {
        captured_messages <<- c(
          captured_messages,
          conditionMessage(condition)
        )
        invokeRestart("muffleMessage")
      }
    )
  })

  list(
    object = network_object,
    warnings = unique(captured_warnings),
    messages = unique(captured_messages),
    elapsed = elapsed
  )
}

formal_network_results <- list()
formal_point_validation_list <- list()
formal_point_diagnostics_list <- list()
formal_edge_table_list <- list()

if (RUN_FORMAL_POINT_NETWORKS) {
  for (wave_key in as.character(formal_waves)) {
    wave_result <- estimate_formal_network(
      wave_key = wave_key,
      wave_data = formal_wave_data[[wave_key]]
    )

    wave_network <- wave_result$object
    expected_n <- unname(
      formal_expected_n[[wave_key]]
    )

    reference_graph <- qgraph::EBICglasso(
      S = polychoric_bundle$Matrices[[wave_key]],
      n = expected_n,
      gamma = 0.50,
      nlambda = 100,
      lambda.min.ratio = 0.01,
      refit = FALSE,
      threshold = FALSE
    )

    maximum_reference_difference <- max(
      abs(
        wave_network$graph -
          reference_graph
      )
    )

    validation <- data.frame(
      Wave = as.integer(wave_key),
      Check = c(
        "Expected respondent count",
        "Expected node count",
        "Frozen node order",
        "Finite graph",
        "Symmetric graph",
        "Zero diagonal",
        "Matches saved polychoric input"
      ),
      Observed = c(
        as.character(wave_network$nPerson),
        as.character(wave_network$nNode),
        paste(rownames(wave_network$graph), collapse = "|"),
        as.character(all(is.finite(wave_network$graph))),
        format(
          max(
            abs(
              wave_network$graph -
                t(wave_network$graph)
            )
          ),
          scientific = TRUE
        ),
        format(
          max(abs(diag(wave_network$graph))),
          scientific = TRUE
        ),
        format(
          maximum_reference_difference,
          scientific = TRUE
        )
      ),
      Expected = c(
        as.character(expected_n),
        "9",
        paste(network_vars, collapse = "|"),
        "TRUE",
        "< 1e-12",
        "< 1e-12",
        "< 1e-8"
      ),
      Passed = c(
        identical(
          as.integer(wave_network$nPerson),
          as.integer(expected_n)
        ),
        identical(
          as.integer(wave_network$nNode),
          9L
        ),
        identical(
          rownames(wave_network$graph),
          network_vars
        ) && identical(
          colnames(wave_network$graph),
          network_vars
        ),
        all(is.finite(wave_network$graph)),
        max(
          abs(
            wave_network$graph -
              t(wave_network$graph)
          )
        ) < 1e-12,
        max(abs(diag(wave_network$graph))) < 1e-12,
        maximum_reference_difference < 1e-8
      ),
      stringsAsFactors = FALSE
    )

    formal_point_validation_list[[wave_key]] <- validation

    graph_upper <- upper.tri(
      wave_network$graph,
      diag = FALSE
    )

    formal_edge_table_list[[wave_key]] <- data.frame(
      Wave = as.integer(wave_key),
      Node_1 = rownames(wave_network$graph)[
        row(wave_network$graph)[graph_upper]
      ],
      Node_2 = colnames(wave_network$graph)[
        col(wave_network$graph)[graph_upper]
      ],
      Edge_weight = wave_network$graph[graph_upper],
      Retained_under_EBICglasso =
        wave_network$graph[graph_upper] != 0,
      stringsAsFactors = FALSE
    )

    formal_point_diagnostics_list[[wave_key]] <- data.frame(
      Wave = as.integer(wave_key),
      N = expected_n,
      Nodes = nrow(wave_network$graph),
      Possible_edges = choose(nrow(wave_network$graph), 2L),
      Retained_edges = sum(
        wave_network$graph[graph_upper] != 0
      ),
      Global_strength_descriptive = sum(
        abs(wave_network$graph[graph_upper])
      ),
      Maximum_reference_difference =
        maximum_reference_difference,
      Estimation_elapsed_seconds = unname(
        wave_result$elapsed[["elapsed"]]
      ),
      Captured_warning_n = length(wave_result$warnings),
      Captured_warnings = paste(
        wave_result$warnings,
        collapse = " | "
      ),
      Captured_message_n = length(wave_result$messages),
      Captured_messages = paste(
        wave_result$messages,
        collapse = " | "
      ),
      stringsAsFactors = FALSE
    )

    formal_network_results[[wave_key]] <- wave_result

    saveRDS(
      wave_result,
      file.path(
        formal_model_dir,
        paste0(
          "07_formal_network_",
          wave_key,
          "_gamma050.rds"
        )
      )
    )

    write.csv(
      validation,
      file.path(
        formal_table_dir,
        paste0(
          "07_formal_network_",
          wave_key,
          "_validation.csv"
        )
      ),
      row.names = FALSE
    )

    if (!all(validation$Passed)) {
      stop(
        paste0(
          "Formal point-network validation failed for ",
          wave_key,
          "."
        )
      )
    }
  }
} else if (REUSE_VALIDATED_FORMAL_OUTPUTS) {
  saved_formal_network_file <- file.path(
    formal_model_dir,
    "07_formal_five_wave_networks_gamma050.rds"
  )

  saved_point_validation_file <- file.path(
    formal_table_dir,
    "07_formal_five_wave_point_network_validation.csv"
  )

  saved_point_diagnostics_file <- file.path(
    formal_table_dir,
    "07_formal_five_wave_point_network_diagnostics.csv"
  )

  saved_edge_table_file <- file.path(
    formal_table_dir,
    "07_formal_five_wave_edge_weights.csv"
  )

  saved_point_files <- c(
    saved_formal_network_file,
    saved_point_validation_file,
    saved_point_diagnostics_file,
    saved_edge_table_file
  )

  stopifnot(
    all(file.exists(saved_point_files)),
    all(file.info(saved_point_files)$size > 0)
  )

  formal_network_results <- readRDS(
    saved_formal_network_file
  )

  saved_point_validation <- read.csv(
    saved_point_validation_file,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  saved_point_diagnostics <- read.csv(
    saved_point_diagnostics_file,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  saved_edge_table <- read.csv(
    saved_edge_table_file,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  stopifnot(
    identical(
      names(formal_network_results),
      as.character(formal_waves)
    ),
    nrow(saved_point_validation) == 35L,
    all(saved_point_validation$Passed),
    nrow(saved_point_diagnostics) == 5L,
    nrow(saved_edge_table) == 180L,
    identical(
      sort(unique(saved_point_validation$Wave)),
      formal_waves
    )
  )

  for (wave_key in as.character(formal_waves)) {
    saved_result <- formal_network_results[[wave_key]]
    saved_graph <- saved_result$object$graph
    expected_n <- unname(
      formal_expected_n[[wave_key]]
    )

    reference_graph <- qgraph::EBICglasso(
      S = polychoric_bundle$Matrices[[wave_key]],
      n = expected_n,
      gamma = 0.50,
      nlambda = 100,
      lambda.min.ratio = 0.01,
      refit = FALSE,
      threshold = FALSE
    )

    stopifnot(
      identical(
        as.integer(saved_result$object$nPerson),
        as.integer(expected_n)
      ),
      identical(
        rownames(saved_graph),
        network_vars
      ),
      identical(
        colnames(saved_graph),
        network_vars
      ),
      all(is.finite(saved_graph)),
      max(abs(saved_graph - t(saved_graph))) < 1e-12,
      max(abs(diag(saved_graph))) < 1e-12,
      max(abs(saved_graph - reference_graph)) < 1e-8
    )
  }

  formal_point_validation_list <- split(
    saved_point_validation,
    as.character(saved_point_validation$Wave)
  )

  formal_point_diagnostics_list <- split(
    saved_point_diagnostics,
    as.character(saved_point_diagnostics$Wave)
  )

  formal_edge_table_list <- split(
    saved_edge_table,
    as.character(saved_edge_table$Wave)
  )
}

formal_point_validation <- do.call(
  rbind,
  formal_point_validation_list
)

formal_point_diagnostics <- do.call(
  rbind,
  formal_point_diagnostics_list
)

formal_edge_table <- do.call(
  rbind,
  formal_edge_table_list
)

if (RUN_FORMAL_POINT_NETWORKS) {
  write.csv(
    formal_point_validation,
    file.path(
      formal_table_dir,
      "07_formal_five_wave_point_network_validation.csv"
    ),
    row.names = FALSE
  )

  write.csv(
    formal_point_diagnostics,
    file.path(
      formal_table_dir,
      "07_formal_five_wave_point_network_diagnostics.csv"
    ),
    row.names = FALSE
  )

  write.csv(
    formal_edge_table,
    file.path(
      formal_table_dir,
      "07_formal_five_wave_edge_weights.csv"
    ),
    row.names = FALSE
  )

  saveRDS(
    formal_network_results,
    file.path(
      formal_model_dir,
      "07_formal_five_wave_networks_gamma050.rds"
    )
  )
}

formal_point_diagnostics
```

### Run and Validate 1,000 Edge-Weight Bootstraps per Wave

```{r run-and-validate-formal-edge-bootstraps}
stopifnot(
  exists("formal_network_results"),
  length(formal_network_results) == 5L,
  all(formal_point_validation$Passed)
)

formal_bootstrap_results <- list()
formal_bootstrap_validation_list <- list()
formal_bootstrap_diagnostics_list <- list()
formal_bootstrap_summary_list <- list()

if (RUN_FORMAL_EDGE_BOOTSTRAPS) {
  for (wave_key in as.character(formal_waves)) {
    network_object <- formal_network_results[[
      wave_key
    ]]$object

    bootstrap_result <- run_edge_bootstrap(
      network_object = network_object,
      n_boots = 1000L,
      random_seed = unname(
        formal_bootstrap_seeds[[wave_key]]
      )
    )

    validation_result <- validate_edge_bootstrap(
      result_object = bootstrap_result,
      expected_n_boots = 1000L,
      expected_node_names = network_vars,
      expected_n_person = unname(
        formal_expected_n[[wave_key]]
      ),
      reference_graph = network_object$graph
    )

    validation_table <- validation_result$audit
    validation_table$Wave <- as.integer(wave_key)
    validation_table <- validation_table[
      ,
      c(
        "Wave",
        "Check",
        "Observed",
        "Expected",
        "Passed"
      )
    ]

    formal_bootstrap_validation_list[[wave_key]] <-
      validation_table

    formal_bootstrap_diagnostics_list[[wave_key]] <-
      data.frame(
        Wave = as.integer(wave_key),
        Requested_bootstraps = 1000L,
        Returned_bootstraps = length(
          bootstrap_result$object$boots
        ),
        Seed = bootstrap_result$seed,
        Elapsed_seconds = unname(
          bootstrap_result$elapsed_time[["elapsed"]]
        ),
        Top_level_retry_count =
          validation_result$top_level_retry_count,
        Captured_warning_n =
          validation_result$captured_warning_n,
        Captured_warnings = paste(
          validation_result$captured_warnings,
          collapse = " | "
        ),
        All_validation_checks_passed = all(
          validation_table$Passed
        ),
        stringsAsFactors = FALSE
      )

    bootstrap_summary <- as.data.frame(
      summary(
        bootstrap_result$object,
        statistics = "edge"
      )
    )
    bootstrap_summary$Wave <- as.integer(wave_key)
    formal_bootstrap_summary_list[[wave_key]] <-
      bootstrap_summary

    saveRDS(
      bootstrap_result,
      file.path(
        formal_model_dir,
        paste0(
          "07_formal_edge_bootstrap_",
          wave_key,
          "_1000.rds"
        )
      )
    )

    write.csv(
      validation_table,
      file.path(
        formal_table_dir,
        paste0(
          "07_formal_edge_bootstrap_",
          wave_key,
          "_validation.csv"
        )
      ),
      row.names = FALSE
    )

    write.csv(
      bootstrap_summary,
      file.path(
        formal_table_dir,
        paste0(
          "07_formal_edge_bootstrap_",
          wave_key,
          "_summary.csv"
        )
      ),
      row.names = FALSE
    )

    write_bootstrap_log(
      result_object = bootstrap_result,
      output_path = file.path(
        formal_log_dir,
        paste0(
          "07_formal_edge_bootstrap_",
          wave_key,
          "_1000.log"
        )
      )
    )

    formal_bootstrap_results[[wave_key]] <-
      bootstrap_result

    if (!all(validation_table$Passed)) {
      stop(
        paste0(
          "Formal edge-bootstrap validation failed for ",
          wave_key,
          ". Inspect the saved validation and log files."
        )
      )
    }
  }
} else if (REUSE_VALIDATED_FORMAL_OUTPUTS) {
  for (wave_key in as.character(formal_waves)) {
    saved_bootstrap_file <- file.path(
      formal_model_dir,
      paste0(
        "07_formal_edge_bootstrap_",
        wave_key,
        "_1000.rds"
      )
    )

    saved_bootstrap_summary_file <- file.path(
      formal_table_dir,
      paste0(
        "07_formal_edge_bootstrap_",
        wave_key,
        "_summary.csv"
      )
    )

    stopifnot(
      file.exists(saved_bootstrap_file),
      file.info(saved_bootstrap_file)$size > 0,
      file.exists(saved_bootstrap_summary_file),
      file.info(saved_bootstrap_summary_file)$size > 0
    )

    bootstrap_result <- readRDS(
      saved_bootstrap_file
    )

    network_object <- formal_network_results[[
      wave_key
    ]]$object

    validation_result <- validate_edge_bootstrap(
      result_object = bootstrap_result,
      expected_n_boots = 1000L,
      expected_node_names = network_vars,
      expected_n_person = unname(
        formal_expected_n[[wave_key]]
      ),
      reference_graph = network_object$graph
    )

    validation_table <- validation_result$audit
    validation_table$Wave <- as.integer(wave_key)
    validation_table <- validation_table[
      ,
      c(
        "Wave",
        "Check",
        "Observed",
        "Expected",
        "Passed"
      )
    ]

    bootstrap_summary <- read.csv(
      saved_bootstrap_summary_file,
      stringsAsFactors = FALSE,
      check.names = FALSE
    )

    stopifnot(
      nrow(bootstrap_summary) == 36L,
      all(validation_table$Passed)
    )

    formal_bootstrap_results[[wave_key]] <-
      bootstrap_result

    formal_bootstrap_validation_list[[wave_key]] <-
      validation_table

    formal_bootstrap_diagnostics_list[[wave_key]] <-
      data.frame(
        Wave = as.integer(wave_key),
        Requested_bootstraps = 1000L,
        Returned_bootstraps = length(
          bootstrap_result$object$boots
        ),
        Seed = bootstrap_result$seed,
        Elapsed_seconds = unname(
          bootstrap_result$elapsed_time[["elapsed"]]
        ),
        Top_level_retry_count =
          validation_result$top_level_retry_count,
        Captured_warning_n =
          validation_result$captured_warning_n,
        Captured_warnings = paste(
          validation_result$captured_warnings,
          collapse = " | "
        ),
        All_validation_checks_passed = all(
          validation_table$Passed
        ),
        stringsAsFactors = FALSE
      )

    bootstrap_summary$Wave <- as.integer(wave_key)
    formal_bootstrap_summary_list[[wave_key]] <-
      bootstrap_summary
  }
}

formal_bootstrap_validation <- do.call(
  rbind,
  formal_bootstrap_validation_list
)

formal_bootstrap_diagnostics <- do.call(
  rbind,
  formal_bootstrap_diagnostics_list
)

formal_bootstrap_edge_summary <- do.call(
  rbind,
  formal_bootstrap_summary_list
)

if (RUN_FORMAL_EDGE_BOOTSTRAPS) {
  write.csv(
    formal_bootstrap_validation,
    file.path(
      formal_table_dir,
      "07_formal_five_wave_edge_bootstrap_validation.csv"
    ),
    row.names = FALSE
  )

  write.csv(
    formal_bootstrap_diagnostics,
    file.path(
      formal_table_dir,
      "07_formal_five_wave_edge_bootstrap_diagnostics.csv"
    ),
    row.names = FALSE
  )

  write.csv(
    formal_bootstrap_edge_summary,
    file.path(
      formal_table_dir,
      "07_formal_five_wave_edge_bootstrap_summary.csv"
    ),
    row.names = FALSE
  )
}

formal_bootstrap_diagnostics

stopifnot(
  all(formal_bootstrap_validation$Passed),
  all(
    formal_bootstrap_diagnostics$
      All_validation_checks_passed
  )
)
```

### Formal Step-07 Completion Record

```{r formal-step07-completion-record}
formal_completion_file <- file.path(
  formal_table_dir,
  "07_formal_step07_completion_status.csv"
)

if (RUN_FORMAL_EDGE_BOOTSTRAPS) {
  formal_step07_completion <- data.frame(
    Component = c(
      "Five ordered complete-case inputs",
      "Five primary point networks",
      "Five 1,000-repetition edge bootstraps",
      "Centrality",
      "Case-dropping bootstrap",
      "NCT",
      "Education-group networks",
      "Sensitivity analyses",
      "Political interpretation"
    ),
    Status = c(
      "COMPLETE_AND_VALIDATED",
      "COMPLETE_AND_VALIDATED",
      "COMPLETE_AND_VALIDATED",
      "NOT_AUTHORISED_NOT_RUN",
      "NOT_AUTHORISED_NOT_RUN",
      "NOT_AUTHORISED_NOT_RUN",
      "NOT_AUTHORISED_NOT_RUN",
      "NOT_AUTHORISED_NOT_RUN",
      "NOT_AUTHORISED_NOT_RUN"
    ),
    stringsAsFactors = FALSE
  )

  write.csv(
    formal_step07_completion,
    formal_completion_file,
    row.names = FALSE
  )

  capture.output(
    sessionInfo(),
    file = file.path(
      formal_log_dir,
      "07_formal_five_wave_sessionInfo.txt"
    )
  )
} else {
  stopifnot(
    REUSE_VALIDATED_FORMAL_OUTPUTS,
    file.exists(formal_completion_file),
    file.info(formal_completion_file)$size > 0
  )

  formal_step07_completion <- read.csv(
    formal_completion_file,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  stopifnot(
    nrow(formal_step07_completion) == 9L,
    identical(
      formal_step07_completion$Status[1:3],
      rep("COMPLETE_AND_VALIDATED", 3L)
    )
  )
}

formal_step07_completion
```

