---
---
title: "07A Edge-Weight Accuracy Review"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    df_print: paged
---

## 1. Purpose and analysis boundary

This notebook performs a read-only accuracy review of the five formal
edge-weight bootstrap results produced in Step 07.

It does not estimate networks, rerun bootstrap resampling, calculate
centrality, conduct case-dropping bootstrap, perform NCT, analyse
education groups, or interpret political change.

Bootstrap intervals are treated as indicators of estimation accuracy.
Whether an interval includes zero is not treated as a formal significance
test.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = TRUE,
  message = TRUE
)

options(stringsAsFactors = FALSE)
```

## 2. Locate the project

```{r project-root}
current_dir <- normalizePath(
  ".",
  winslash = "/",
  mustWork = TRUE
)

if (
  file.exists(
    file.path(
      current_dir,
      "ANES.Rproj"
    )
  )
) {
  project_root <- current_dir
} else if (
  basename(current_dir) == "notebooks" &&
    file.exists(
      file.path(
        dirname(current_dir),
        "ANES.Rproj"
      )
    )
) {
  project_root <- dirname(current_dir)
} else {
  stop(
    "Project root could not be identified. Open the ANES_dissertation RStudio project before knitting."
  )
}

knitr::opts_knit$set(
  root.dir = project_root
)

project_root
```

## 3. Define input and output locations

```{r paths}
formal_table_dir <- file.path(
  project_root,
  "outputs",
  "tables",
  "formal"
)

formal_model_dir <- file.path(
  project_root,
  "outputs",
  "models",
  "formal"
)

formal_figure_dir <- file.path(
  project_root,
  "outputs",
  "figures",
  "formal"
)

log_dir <- file.path(
  project_root,
  "outputs",
  "logs"
)

dir.create(
  formal_figure_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

dir.create(
  log_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

point_edge_file <- file.path(
  formal_table_dir,
  "07_formal_five_wave_edge_weights.csv"
)

bootstrap_summary_file <- file.path(
  formal_table_dir,
  "07_formal_five_wave_edge_bootstrap_summary.csv"
)

bootstrap_diagnostics_file <- file.path(
  formal_table_dir,
  "07_formal_five_wave_edge_bootstrap_diagnostics.csv"
)

bootstrap_validation_file <- file.path(
  formal_table_dir,
  "07_formal_five_wave_edge_bootstrap_validation.csv"
)

point_diagnostics_file <- file.path(
  formal_table_dir,
  "07_formal_five_wave_point_network_diagnostics.csv"
)

model_files <- c(
  file.path(
    formal_model_dir,
    "07_formal_five_wave_networks_gamma050.rds"
  ),
  file.path(
    formal_model_dir,
    paste0(
      "07_formal_edge_bootstrap_",
      c(2004, 2012, 2016, 2020, 2024),
      "_1000.rds"
    )
  )
)

table_input_files <- c(
  point_edge_file,
  bootstrap_summary_file,
  bootstrap_diagnostics_file,
  bootstrap_validation_file,
  point_diagnostics_file
)

all_source_files <- c(
  table_input_files,
  model_files
)

missing_source_files <- all_source_files[
  !file.exists(all_source_files)
]

if (length(missing_source_files) > 0L) {
  stop(
    paste(
      "Required Step-07 source files are missing:",
      paste(missing_source_files, collapse = "; ")
    )
  )
}
```

## 4. Record source-file provenance

```{r input-inventory}
model_md5_before <- unname(
  tools::md5sum(model_files)
)

input_inventory <- data.frame(
  File = basename(all_source_files),
  Path = normalizePath(
    all_source_files,
    winslash = "/"
  ),
  Size_bytes = file.info(all_source_files)$size,
  Last_modified = format(
    file.info(all_source_files)$mtime,
    "%Y-%m-%d %H:%M:%S"
  ),
  MD5 = unname(
    tools::md5sum(all_source_files)
  ),
  stringsAsFactors = FALSE
)

input_inventory_file <- file.path(
  formal_table_dir,
  "07A_input_inventory.csv"
)

write.csv(
  input_inventory,
  input_inventory_file,
  row.names = FALSE,
  na = ""
)

input_inventory
```

## 5. Read the validated Step-07 outputs

```{r read-inputs}
point_edges <- read.csv(
  point_edge_file,
  check.names = FALSE
)

bootstrap_summary <- read.csv(
  bootstrap_summary_file,
  check.names = FALSE
)

bootstrap_diagnostics <- read.csv(
  bootstrap_diagnostics_file,
  check.names = FALSE
)

bootstrap_validation <- read.csv(
  bootstrap_validation_file,
  check.names = FALSE
)

point_diagnostics <- read.csv(
  point_diagnostics_file,
  check.names = FALSE
)

assert_columns <- function(
  data,
  required_columns,
  object_name
) {
  missing_columns <- setdiff(
    required_columns,
    names(data)
  )

  if (length(missing_columns) > 0L) {
    stop(
      paste(
        object_name,
        "is missing required columns:",
        paste(missing_columns, collapse = ", ")
      )
    )
  }
}

assert_columns(
  point_edges,
  c(
    "Wave",
    "Node_1",
    "Node_2",
    "Edge_weight",
    "Retained_under_EBICglasso"
  ),
  "Point-edge table"
)

assert_columns(
  bootstrap_summary,
  c(
    "Wave",
    "node1",
    "node2",
    "sample",
    "mean",
    "sd",
    "CIlower",
    "CIupper",
    "q2.5",
    "q97.5",
    "prop0"
  ),
  "Bootstrap-summary table"
)

assert_columns(
  bootstrap_diagnostics,
  c(
    "Wave",
    "Requested_bootstraps",
    "Returned_bootstraps",
    "Captured_warning_n",
    "All_validation_checks_passed"
  ),
  "Bootstrap-diagnostics table"
)

assert_columns(
  bootstrap_validation,
  c(
    "Wave",
    "Check",
    "Passed"
  ),
  "Bootstrap-validation table"
)
```

## 6. Join point estimates and bootstrap results

```{r construct-review-table}
expected_waves <- c(
  2004L,
  2012L,
  2016L,
  2020L,
  2024L
)

point_edges$Wave <- as.integer(
  point_edges$Wave
)

bootstrap_summary$Wave <- as.integer(
  bootstrap_summary$Wave
)

bootstrap_diagnostics$Wave <- as.integer(
  bootstrap_diagnostics$Wave
)

bootstrap_validation$Wave <- as.integer(
  bootstrap_validation$Wave
)

point_edges$Edge_weight <- as.numeric(
  point_edges$Edge_weight
)

point_edges$Retained_under_EBICglasso <-
  toupper(
    as.character(
      point_edges$Retained_under_EBICglasso
    )
  ) == "TRUE"

bootstrap_numeric_columns <- c(
  "sample",
  "mean",
  "sd",
  "CIlower",
  "CIupper",
  "q2.5",
  "q97.5",
  "prop0"
)

for (column_name in bootstrap_numeric_columns) {
  bootstrap_summary[[column_name]] <-
    as.numeric(
      bootstrap_summary[[column_name]]
    )
}

point_edges$Edge_ID <- paste(
  pmin(
    point_edges$Node_1,
    point_edges$Node_2
  ),
  pmax(
    point_edges$Node_1,
    point_edges$Node_2
  ),
  sep = "--"
)

bootstrap_summary$Edge_ID <- paste(
  pmin(
    bootstrap_summary$node1,
    bootstrap_summary$node2
  ),
  pmax(
    bootstrap_summary$node1,
    bootstrap_summary$node2
  ),
  sep = "--"
)

review_joined <- merge(
  point_edges,
  bootstrap_summary,
  by = c(
    "Wave",
    "Edge_ID"
  ),
  all = FALSE,
  sort = FALSE
)

if (nrow(review_joined) == 0L) {
  stop(
    "The point-edge and bootstrap tables could not be joined."
  )
}

edge_accuracy <- data.frame(
  Wave = review_joined$Wave,
  Edge_ID = review_joined$Edge_ID,
  Node_1 = review_joined$Node_1,
  Node_2 = review_joined$Node_2,
  Point_edge = review_joined$Edge_weight,
  Retained_under_EBICglasso =
    review_joined$Retained_under_EBICglasso,
  Bootstrap_mean = review_joined$mean,
  Bootstrap_sd = review_joined$sd,
  Reported_CI_lower = review_joined$CIlower,
  Reported_CI_upper = review_joined$CIupper,
  Percentile_2_5 = review_joined[["q2.5"]],
  Percentile_97_5 = review_joined[["q97.5"]],
  Bootstrap_zero_proportion =
    review_joined$prop0,
  stringsAsFactors = FALSE
)

edge_accuracy$Reported_CI_width <-
  edge_accuracy$Reported_CI_upper -
  edge_accuracy$Reported_CI_lower

edge_accuracy$Percentile_interval_width <-
  edge_accuracy$Percentile_97_5 -
  edge_accuracy$Percentile_2_5

edge_accuracy$Percentile_interval_includes_zero <-
  edge_accuracy$Percentile_2_5 <= 0 &
  edge_accuracy$Percentile_97_5 >= 0

edge_accuracy$Bootstrap_mean_minus_point <-
  edge_accuracy$Bootstrap_mean -
  edge_accuracy$Point_edge

edge_accuracy$Retained_edge_sign_reproduced <-
  ifelse(
    edge_accuracy$Retained_under_EBICglasso,
    sign(edge_accuracy$Point_edge) ==
      sign(edge_accuracy$Bootstrap_mean),
    NA
  )

edge_accuracy <- edge_accuracy[
  order(
    edge_accuracy$Wave,
    edge_accuracy$Node_1,
    edge_accuracy$Node_2
  ),
]

row.names(edge_accuracy) <- NULL
```

## 7. Validation gate

```{r validation-gate}
numeric_review_columns <- c(
  "Point_edge",
  "Bootstrap_mean",
  "Bootstrap_sd",
  "Reported_CI_lower",
  "Reported_CI_upper",
  "Percentile_2_5",
  "Percentile_97_5",
  "Bootstrap_zero_proportion"
)

all_numeric_values_finite <- all(
  vapply(
    edge_accuracy[numeric_review_columns],
    function(x) {
      all(is.finite(x))
    },
    logical(1)
  )
)

duplicate_key_n <- sum(
  duplicated(
    paste(
      edge_accuracy$Wave,
      edge_accuracy$Edge_ID,
      sep = "::"
    )
  )
)

maximum_point_difference <- max(
  abs(
    review_joined$Edge_weight -
    review_joined$sample
  )
)

bootstrap_validation_passed <- all(
  toupper(
    as.character(
      bootstrap_validation$Passed
    )
  ) == "TRUE"
)

bootstrap_diagnostics_passed <- all(
  bootstrap_diagnostics$Requested_bootstraps == 1000L &
  bootstrap_diagnostics$Returned_bootstraps == 1000L &
  bootstrap_diagnostics$Captured_warning_n == 0L &
  toupper(
    as.character(
      bootstrap_diagnostics$
        All_validation_checks_passed
    )
  ) == "TRUE"
)

model_md5_after <- unname(
  tools::md5sum(model_files)
)

model_hashes_unchanged <- identical(
  model_md5_before,
  model_md5_after
)

validation <- data.frame(
  Check = c(
    "All required source files exist",
    "Point-edge table has 180 rows",
    "Bootstrap summary has 180 rows",
    "Joined review table has 180 rows",
    "Expected five waves are present",
    "Wave-edge keys are unique",
    "Point estimates match bootstrap sample values",
    "All reviewed numeric values are finite",
    "Reported intervals have valid ordering",
    "Percentile intervals have valid ordering",
    "Bootstrap zero proportions are between zero and one",
    "Original bootstrap validations all passed",
    "Bootstrap diagnostics record 1000 returned runs per wave",
    "Formal model-object hashes remained unchanged"
  ),
  Observed = c(
    length(missing_source_files),
    nrow(point_edges),
    nrow(bootstrap_summary),
    nrow(edge_accuracy),
    paste(
      sort(unique(edge_accuracy$Wave)),
      collapse = ", "
    ),
    duplicate_key_n,
    format(
      maximum_point_difference,
      scientific = TRUE
    ),
    all_numeric_values_finite,
    all(
      edge_accuracy$Reported_CI_lower <=
        edge_accuracy$Reported_CI_upper
    ),
    all(
      edge_accuracy$Percentile_2_5 <=
        edge_accuracy$Percentile_97_5
    ),
    all(
      edge_accuracy$Bootstrap_zero_proportion >= 0 &
      edge_accuracy$Bootstrap_zero_proportion <= 1
    ),
    bootstrap_validation_passed,
    bootstrap_diagnostics_passed,
    model_hashes_unchanged
  ),
  Expected = c(
    "0",
    "180",
    "180",
    "180",
    "2004, 2012, 2016, 2020, 2024",
    "0",
    "<= 1e-12",
    "TRUE",
    "TRUE",
    "TRUE",
    "TRUE",
    "TRUE",
    "TRUE",
    "TRUE"
  ),
  Passed = c(
    length(missing_source_files) == 0L,
    nrow(point_edges) == 180L,
    nrow(bootstrap_summary) == 180L,
    nrow(edge_accuracy) == 180L,
    identical(
      sort(unique(edge_accuracy$Wave)),
      expected_waves
    ),
    duplicate_key_n == 0L,
    maximum_point_difference <= 1e-12,
    all_numeric_values_finite,
    all(
      edge_accuracy$Reported_CI_lower <=
        edge_accuracy$Reported_CI_upper
    ),
    all(
      edge_accuracy$Percentile_2_5 <=
        edge_accuracy$Percentile_97_5
    ),
    all(
      edge_accuracy$Bootstrap_zero_proportion >= 0 &
      edge_accuracy$Bootstrap_zero_proportion <= 1
    ),
    bootstrap_validation_passed,
    bootstrap_diagnostics_passed,
    model_hashes_unchanged
  ),
  stringsAsFactors = FALSE
)

validation_file <- file.path(
  formal_table_dir,
  "07A_edge_accuracy_validation.csv"
)

write.csv(
  validation,
  validation_file,
  row.names = FALSE,
  na = ""
)

validation

if (!all(validation$Passed)) {
  stop(
    paste(
      "07A validation failed.",
      "Inspect 07A_edge_accuracy_validation.csv."
    )
  )
}
```

## 8. Produce the edge-accuracy tables

```{r produce-tables}
edge_accuracy_file <- file.path(
  formal_table_dir,
  "07A_edge_accuracy.csv"
)

write.csv(
  edge_accuracy,
  edge_accuracy_file,
  row.names = FALSE,
  na = ""
)

wave_summary <- do.call(
  rbind,
  lapply(
    split(
      edge_accuracy,
      edge_accuracy$Wave
    ),
    function(wave_data) {
      data.frame(
        Wave = unique(wave_data$Wave),
        Edges = nrow(wave_data),
        Retained_edges = sum(
          wave_data$Retained_under_EBICglasso
        ),
        All_edges_percentile_interval_excluding_zero =
          sum(
            !wave_data$
              Percentile_interval_includes_zero
          ),
        Retained_edges_percentile_interval_excluding_zero =
          sum(
            wave_data$Retained_under_EBICglasso &
            !wave_data$
              Percentile_interval_includes_zero
          ),
        Median_percentile_interval_width =
          median(
            wave_data$
              Percentile_interval_width
          ),
        Maximum_percentile_interval_width =
          max(
            wave_data$
              Percentile_interval_width
          ),
        Median_bootstrap_zero_proportion =
          median(
            wave_data$
              Bootstrap_zero_proportion
          ),
        Maximum_absolute_mean_point_difference =
          max(
            abs(
              wave_data$
                Bootstrap_mean_minus_point
            )
          ),
        Retained_edges_with_reproduced_sign =
          sum(
            wave_data$
              Retained_edge_sign_reproduced,
            na.rm = TRUE
          ),
        stringsAsFactors = FALSE
      )
    }
  )
)

row.names(wave_summary) <- NULL

wave_summary_file <- file.path(
  formal_table_dir,
  "07A_edge_accuracy_by_wave.csv"
)

write.csv(
  wave_summary,
  wave_summary_file,
  row.names = FALSE,
  na = ""
)

widest_intervals <- do.call(
  rbind,
  lapply(
    split(
      edge_accuracy,
      edge_accuracy$Wave
    ),
    function(wave_data) {
      ordered_data <- wave_data[
        order(
          -wave_data$
            Percentile_interval_width
        ),
      ]

      head(
        ordered_data,
        10L
      )
    }
  )
)

row.names(widest_intervals) <- NULL

widest_intervals_file <- file.path(
  formal_table_dir,
  "07A_edge_accuracy_widest_intervals.csv"
)

write.csv(
  widest_intervals,
  widest_intervals_file,
  row.names = FALSE,
  na = ""
)

conditions <- merge(
  point_diagnostics,
  bootstrap_diagnostics,
  by = "Wave",
  all = TRUE,
  suffixes = c(
    "_point_network",
    "_edge_bootstrap"
  )
)

conditions_file <- file.path(
  formal_table_dir,
  "07A_edge_accuracy_conditions.csv"
)

write.csv(
  conditions,
  conditions_file,
  row.names = FALSE,
  na = ""
)

wave_summary
```

The counts above are descriptive accuracy diagnostics. They are not
formal tests that an edge exists or differs across election years.

## 9. Identify the widest intervals

```{r widest-intervals}
widest_intervals[
  ,
  c(
    "Wave",
    "Edge_ID",
    "Point_edge",
    "Percentile_2_5",
    "Percentile_97_5",
    "Percentile_interval_width",
    "Bootstrap_zero_proportion"
  )
]
```

## 10. Produce one accuracy figure per wave

```{r create-figures}
figure_files <- character(
  length(expected_waves)
)

for (wave_index in seq_along(expected_waves)) {
  current_wave <- expected_waves[wave_index]

  wave_data <- edge_accuracy[
    edge_accuracy$Wave == current_wave,
  ]

  wave_data <- wave_data[
    order(wave_data$Point_edge),
  ]

  y_position <- seq_len(
    nrow(wave_data)
  )

  point_colours <- ifelse(
    wave_data$Retained_under_EBICglasso,
    "#0072B2",
    "#999999"
  )

  x_limits <- range(
    c(
      wave_data$Percentile_2_5,
      wave_data$Percentile_97_5
    ),
    finite = TRUE
  )

  x_padding <- diff(x_limits) * 0.05

  if (!is.finite(x_padding) || x_padding == 0) {
    x_padding <- 0.05
  }

  figure_file <- file.path(
    formal_figure_dir,
    paste0(
      "07A_edge_accuracy_",
      current_wave,
      ".png"
    )
  )

  png(
    filename = figure_file,
    width = 2000,
    height = 1700,
    res = 180
  )

  par(
    mar = c(
      5,
      12,
      4,
      2
    ) + 0.1
  )

  plot(
    wave_data$Point_edge,
    y_position,
    xlim = c(
      x_limits[1] - x_padding,
      x_limits[2] + x_padding
    ),
    ylim = c(
      0.5,
      nrow(wave_data) + 0.5
    ),
    yaxt = "n",
    pch = 19,
    col = point_colours,
    xlab = "Regularised partial-correlation edge weight",
    ylab = "",
    main = paste(
      current_wave,
      "edge-weight accuracy"
    )
  )

  segments(
    x0 = wave_data$Percentile_2_5,
    y0 = y_position,
    x1 = wave_data$Percentile_97_5,
    y1 = y_position,
    col = point_colours,
    lwd = 1.5
  )

  points(
    wave_data$Point_edge,
    y_position,
    pch = 19,
    col = point_colours
  )

  axis(
    side = 2,
    at = y_position,
    labels = wave_data$Edge_ID,
    las = 2,
    cex.axis = 0.65
  )

  abline(
    v = 0,
    lty = 2,
    col = "#444444"
  )

  legend(
    "bottomright",
    legend = c(
      "Retained under EBICglasso",
      "Not retained under EBICglasso"
    ),
    col = c(
      "#0072B2",
      "#999999"
    ),
    pch = 19,
    bty = "n",
    cex = 0.8
  )

  dev.off()

  figure_files[wave_index] <- figure_file
}

figure_files
```

## 11. Display the figures
```{r display-figures, echo=FALSE, out.width="95%"}
figure_files_for_html <- normalizePath(
  figure_files,
  winslash = "/",
  mustWork = TRUE
)

if (!all(file.exists(figure_files_for_html))) {
  stop(
    "One or more 07A figure files are missing."
  )
}

knitr::include_graphics(
  figure_files_for_html,
  rel_path = FALSE
)
```

## 12. Save the review log and software record

```{r log-and-session-info}
log_file <- file.path(
  log_dir,
  "07A_edge_accuracy_review.log"
)
```

session_file <- file.path(
  log_dir,
  "07A_edge_accuracy_sessionInfo.txt"
)

log_lines <- c(
  "STEP 07A EDGE-WEIGHT ACCURACY REVIEW",
  paste(
    "Completed:",
    format(
      Sys.time(),
      "%Y-%m-%d %H:%M:%S"
    )
  ),
  paste(
    "Waves:",
    paste(
      expected_waves,
      collapse = ", "
    )
  ),
  paste(
    "Reviewed edge rows:",
    nrow(edge_accuracy)
  ),
  paste(
    "Validation checks passed:",
    sum(validation$Passed),
    "of",
    nrow(validation)
  ),
  paste(
    "Formal model hashes unchanged:",
    model_hashes_unchanged
  ),
  "Networks re-estimated: NO",
  "Bootstrap resampling rerun: NO",
  "Centrality calculated: NO",
  "NCT performed: NO",
  "Political interpretation performed: NO"
)

writeLines(
  log_lines,
  log_file
)

capture.output(
  sessionInfo(),
  file = session_file
)

cat(
  paste(
    log_lines,
    collapse = "\n"
  )
)
```

## 13. Completion status

Step 07A is complete only when every validation check is `TRUE`, all five
figures exist, and the formal model-object hashes remain unchanged.

The accuracy intervals must not be interpreted as proof of the presence or
absence of an edge, as causal evidence, or as evidence of between-wave
differences.