---
title: "07B Expected Influence and Case-Dropping Stability"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

## Purpose and analysis boundary

This notebook implements D23 only. It calculates one-step Expected
Influence as the focal exploratory signed centrality measure and Strength
as an auxiliary reference for the five validated overall-sample networks.
It then performs a staged case-dropping bootstrap stability assessment.

The notebook does not run education-group networks, Bridge Expected
Influence, closeness, betweenness, NCT, cross-wave centrality significance
tests, sensitivity analyses or substantive political interpretation. It
does not modify the completed Step-07 point networks or edge bootstraps.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  error = FALSE,
  warning = TRUE,
  message = TRUE
)
```

## Locate the project

```{r project-root}
current_dir <- normalizePath(
  ".",
  winslash = "/"
)

if (
  file.exists(
    file.path(
      current_dir,
      "ANES.Rproj"
    )
  )
) {
  project_root <- current_dir
} else if (
  basename(current_dir) == "notebooks" &&
    file.exists(
      file.path(
        dirname(current_dir),
        "ANES.Rproj"
      )
    )
) {
  project_root <- dirname(current_dir)
} else {
  stop(
    "Project root could not be identified. Open the R project or knit from the notebooks folder."
  )
}

knitr::opts_knit$set(
  root.dir = project_root
)

project_root
```

## D23 authorisation gate

The current canonical decision registry is read at render time. No local
copy of the D23 status is treated as authority.

```{r d23-authorisation-gate}
normalise_status <- function(value) {
  toupper(
    gsub(
      "[^A-Z0-9]+",
      "_",
      trimws(value)
    )
  )
}

registry_config <- file.path(
  project_root,
  "config",
  "decision_registry_path.txt"
)

stopifnot(
  file.exists(registry_config)
)

registry_lines <- trimws(
  readLines(
    registry_config,
    warn = FALSE
  )
)
registry_lines <- registry_lines[nzchar(registry_lines)]

stopifnot(
  length(registry_lines) == 1L
)

registry_path <- registry_lines[[1L]]

stopifnot(
  file.exists(registry_path)
)

decision_registry <- read.csv(
  registry_path,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

required_registry_columns <- c(
  "decision_id",
  "decision_status",
  "implementation_status",
  "researcher_approval_status"
)

stopifnot(
  all(
    required_registry_columns %in%
      names(decision_registry)
  )
)

d23 <- decision_registry[
  decision_registry$decision_id == "D23",
  ,
  drop = FALSE
]

stopifnot(
  nrow(d23) == 1L
)

d23_authorised <-
  normalise_status(d23$decision_status) == "FROZEN" &&
  normalise_status(
    d23$researcher_approval_status
  ) == "APPROVED" &&
  normalise_status(
    d23$implementation_status
  ) %in% c(
    "AUTHORISED_TECHNICAL_PILOT_NOT_STARTED",
    "TECHNICAL_PILOT_COMPLETE_FORMAL_NOT_STARTED",
    "FORMAL_EXECUTION_AUTHORISED",
    "COMPLETE_AND_VALIDATED"
  )

d23_gate <- data.frame(
  Decision_ID = d23$decision_id,
  Decision_status = d23$decision_status,
  Implementation_status = d23$implementation_status,
  Researcher_approval = d23$researcher_approval_status,
  Authorised = d23_authorised,
  stringsAsFactors = FALSE
)

d23_gate

stopifnot(
  isTRUE(d23_authorised)
)
```

## Run controls

Only the authorised 2004 ten-repetition technical test is enabled in the
initial file. Do not enable the formal five-wave run until the saved pilot
validation and warning files have been reviewed.

```{r run-controls}
RUN_2004_TECHNICAL_TEST <- FALSE
RUN_FORMAL_FIVE_WAVE_CASE_DROP <- FALSE
REUSE_VALIDATED_FORMAL_OUTPUTS <- TRUE

active_execution_modes <- sum(
  c(
    RUN_2004_TECHNICAL_TEST,
    RUN_FORMAL_FIVE_WAVE_CASE_DROP,
    REUSE_VALIDATED_FORMAL_OUTPUTS
  )
)

stopifnot(
  active_execution_modes <= 1L
)

data.frame(
  Mode = c(
    "2004 ten-repetition technical test",
    "Five-wave 1,000-repetition formal case-dropping run",
    "Reuse previously validated 07B outputs"
  ),
  Enabled = c(
    RUN_2004_TECHNICAL_TEST,
    RUN_FORMAL_FIVE_WAVE_CASE_DROP,
    REUSE_VALIDATED_FORMAL_OUTPUTS
  ),
  stringsAsFactors = FALSE
)
```

## Packages and frozen parameters

```{r package-and-parameter-check}
required_packages <- c(
  "bootnet",
  "digest",
  "ggplot2"
)

package_check <- data.frame(
  Package = required_packages,
  Installed = vapply(
    required_packages,
    requireNamespace,
    quietly = TRUE,
    FUN.VALUE = logical(1)
  ),
  Version = vapply(
    required_packages,
    function(package_name) {
      if (
        requireNamespace(
          package_name,
          quietly = TRUE
        )
      ) {
        as.character(
          utils::packageVersion(package_name)
        )
      } else {
        NA_character_
      }
    },
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

package_check

stopifnot(
  all(package_check$Installed),
  identical(
    package_check$Version[
      package_check$Package == "bootnet"
    ],
    "1.9.1"
  )
)

formal_waves <- c(
  2004L,
  2012L,
  2016L,
  2020L,
  2024L
)

formal_expected_n <- c(
  `2004` = 789L,
  `2012` = 4353L,
  `2016` = 2709L,
  `2020` = 5390L,
  `2024` = 3525L
)

network_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223"
)

expected_categories <- list(
  VCF0806 = c(1, 2, 3, 4, 5, 6, 7),
  VCF0809 = c(1, 2, 3, 4, 5, 6, 7),
  VCF0838 = c(1, 2, 3, 4),
  VCF0839 = c(1, 2, 3, 4, 5, 6, 7),
  VCF0879a = c(1, 3, 5),
  VCF0888 = c(1, 2, 3),
  VCF0890 = c(1, 2, 3),
  VCF0894 = c(1, 2, 3),
  VCF9223 = c(1, 2, 3, 4)
)

case_drop_parameters <- data.frame(
  Statistic = "one-step Expected Influence; auxiliary Strength",
  Type = "case",
  Formal_repetitions_per_wave = 1000L,
  Technical_repetitions_2004 = 10L,
  Case_min = 0.05,
  Case_max = 0.75,
  Case_levels = 10L,
  CS_correlation = 0.70,
  Cores = 1L,
  Max_errors_per_bootstrap = 10L,
  stringsAsFactors = FALSE
)

case_drop_parameters
```

## Input and output locations

```{r locations}
analysis_ready_file <- file.path(
  project_root,
  "data",
  "analysis_ready",
  paste0(
    "anes_cdf_20260205_analysis_ready_nine_node_",
    "complete_case_2004_2012_2016_2020_2024.rds"
  )
)

formal_network_file <- file.path(
  project_root,
  "outputs",
  "models",
  "formal",
  "07_formal_five_wave_networks_gamma050.rds"
)

formal_point_validation_file <- file.path(
  project_root,
  "outputs",
  "tables",
  "formal",
  "07_formal_five_wave_point_network_validation.csv"
)

formal_edge_table_file <- file.path(
  project_root,
  "outputs",
  "tables",
  "formal",
  "07_formal_five_wave_edge_weights.csv"
)

model_dir <- file.path(
  project_root,
  "outputs",
  "models",
  "formal"
)

table_dir <- file.path(
  project_root,
  "outputs",
  "tables",
  "formal"
)

pilot_table_dir <- file.path(
  project_root,
  "outputs",
  "tables",
  "pilot"
)

figure_dir <- file.path(
  project_root,
  "outputs",
  "figures",
  "formal"
)

log_dir <- file.path(
  project_root,
  "outputs",
  "logs"
)

for (
  output_directory in c(
    model_dir,
    table_dir,
    pilot_table_dir,
    figure_dir,
    log_dir
  )
) {
  dir.create(
    output_directory,
    recursive = TRUE,
    showWarnings = FALSE
  )
}

required_input_files <- c(
  analysis_ready_file,
  formal_network_file,
  formal_point_validation_file,
  formal_edge_table_file,
  registry_path
)

stopifnot(
  all(file.exists(required_input_files)),
  all(file.info(required_input_files)$size > 0)
)
```

## Source provenance and read-only input gate

```{r source-provenance}
sha256_file <- function(path) {
  digest::digest(
    path,
    algo = "sha256",
    file = TRUE,
    serialize = FALSE
  )
}

source_inventory <- data.frame(
  File = basename(required_input_files),
  Path = normalizePath(
    required_input_files,
    winslash = "/",
    mustWork = TRUE
  ),
  SHA256 = vapply(
    required_input_files,
    sha256_file,
    FUN.VALUE = character(1)
  ),
  Size_bytes = unname(
    file.info(required_input_files)$size
  ),
  stringsAsFactors = FALSE
)

write.csv(
  source_inventory,
  file.path(
    table_dir,
    "07B_input_inventory.csv"
  ),
  row.names = FALSE
)

source_inventory
```

## Load and validate the frozen five-wave inputs

```{r load-and-validate-inputs}
anes_analysis_ready <- readRDS(
  analysis_ready_file
)

formal_network_results <- readRDS(
  formal_network_file
)

formal_point_validation <- read.csv(
  formal_point_validation_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

formal_edge_table <- read.csv(
  formal_edge_table_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

stopifnot(
  is.data.frame(anes_analysis_ready),
  all(c("VCF0004", network_vars) %in% names(anes_analysis_ready)),
  identical(
    names(formal_network_results),
    as.character(formal_waves)
  ),
  nrow(formal_point_validation) == 35L,
  all(formal_point_validation$Passed),
  nrow(formal_edge_table) == 180L
)

prepare_wave_data <- function(wave) {
  wave_data <- anes_analysis_ready[
    anes_analysis_ready$VCF0004 == wave,
    network_vars,
    drop = FALSE
  ]

  for (variable_name in network_vars) {
    wave_data[[variable_name]] <- ordered(
      as.numeric(
        wave_data[[variable_name]]
      ),
      levels = expected_categories[[variable_name]]
    )
  }

  wave_data
}

formal_wave_data <- setNames(
  lapply(
    formal_waves,
    prepare_wave_data
  ),
  as.character(formal_waves)
)

input_validation_list <- lapply(
  as.character(formal_waves),
  function(wave_key) {
    wave_data <- formal_wave_data[[wave_key]]
    saved_result <- formal_network_results[[wave_key]]
    network_object <- saved_result$object
    graph <- network_object$graph

    data.frame(
      Wave = as.integer(wave_key),
      N = nrow(wave_data),
      Expected_N = unname(
        formal_expected_n[[wave_key]]
      ),
      Nine_nodes = ncol(wave_data) == 9L,
      Frozen_node_order = identical(
        names(wave_data),
        network_vars
      ),
      All_ordered = all(
        vapply(
          wave_data,
          is.ordered,
          logical(1)
        )
      ),
      Complete_cases = all(
        stats::complete.cases(wave_data)
      ),
      Saved_network_N_matches = identical(
        as.integer(network_object$nPerson),
        as.integer(
          formal_expected_n[[wave_key]]
        )
      ),
      Graph_finite = all(is.finite(graph)),
      Graph_symmetric = max(
        abs(graph - t(graph))
      ) < 1e-12,
      Graph_zero_diagonal = max(
        abs(diag(graph))
      ) < 1e-12,
      Graph_node_order = identical(
        rownames(graph),
        network_vars
      ) && identical(
        colnames(graph),
        network_vars
      ),
      stringsAsFactors = FALSE
    )
  }
)

input_validation <- do.call(
  rbind,
  input_validation_list
)

input_validation$Passed <- with(
  input_validation,
  N == Expected_N &
    Nine_nodes &
    Frozen_node_order &
    All_ordered &
    Complete_cases &
    Saved_network_N_matches &
    Graph_finite &
    Graph_symmetric &
    Graph_zero_diagonal &
    Graph_node_order
)

write.csv(
  input_validation,
  file.path(
    table_dir,
    "07B_input_validation.csv"
  ),
  row.names = FALSE
)

input_validation

stopifnot(
  all(input_validation$Passed)
)
```

## Calculate raw exploratory centrality values

One-step Expected Influence is calculated as the signed sum of each
node's incident edge weights. Strength is the sum of the corresponding
absolute edge weights. These are descriptive values conditional on the
frozen node set, coding, sample and estimator.

```{r raw-centrality-values}
raw_centrality_list <- lapply(
  as.character(formal_waves),
  function(wave_key) {
    graph <- formal_network_results[[
      wave_key
    ]]$object$graph

    data.frame(
      Wave = as.integer(wave_key),
      Node = rownames(graph),
      Expected_Influence = rowSums(graph),
      Strength = rowSums(abs(graph)),
      Positive_retained_edges = rowSums(graph > 0),
      Negative_retained_edges = rowSums(graph < 0),
      Zero_edges = rowSums(graph == 0) - 1L,
      Interpretive_status = "STABILITY_NOT_YET_ESTABLISHED",
      stringsAsFactors = FALSE
    )
  }
)

raw_centrality_scores <- do.call(
  rbind,
  raw_centrality_list
)

stopifnot(
  nrow(raw_centrality_scores) == 45L,
  all(is.finite(raw_centrality_scores$Expected_Influence)),
  all(is.finite(raw_centrality_scores$Strength)),
  all(raw_centrality_scores$Strength >= 0)
)

write.csv(
  raw_centrality_scores,
  file.path(
    table_dir,
    "07B_raw_centrality_scores.csv"
  ),
  row.names = FALSE
)

raw_centrality_scores
```

## Case-dropping functions

```{r case-drop-functions}
count_retry_warnings <- function(warning_text) {
  if (length(warning_text) == 0L) {
    return(0L)
  }

  retry_pattern <-
    "^([0-9]+) bootstrap estimation\\(s\\) failed and were resampled\\.$"

  warning_parts <- regmatches(
    warning_text,
    regexec(
      retry_pattern,
      warning_text
    )
  )

  retry_values <- vapply(
    warning_parts,
    function(value) {
      if (length(value) >= 2L) {
        as.integer(value[[2L]])
      } else {
        0L
      }
    },
    integer(1)
  )

  sum(retry_values)
}

run_case_drop <- function(
  network_object,
  n_boots,
  random_seed
) {
  recorded_warnings <- character(0)
  recorded_messages <- character(0)

  set.seed(random_seed)

  elapsed_time <- system.time({
    case_drop_object <- withCallingHandlers(
      bootnet::bootnet(
        data = network_object,
        nBoots = n_boots,
        type = "case",
        statistics = c(
          "expectedInfluence",
          "strength"
        ),
        computeCentrality = TRUE,
        caseMin = 0.05,
        caseMax = 0.75,
        caseN = 10,
        nCores = 1,
        weighted = TRUE,
        signed = TRUE,
        memorysaver = TRUE,
        maxErrors = 10,
        verbose = TRUE
      ),
      warning = function(condition) {
        recorded_warnings <<- c(
          recorded_warnings,
          conditionMessage(condition)
        )
        invokeRestart("muffleWarning")
      },
      message = function(condition) {
        recorded_messages <<- c(
          recorded_messages,
          conditionMessage(condition)
        )
        invokeRestart("muffleMessage")
      }
    )
  })

  list(
    object = case_drop_object,
    nBoots = as.integer(n_boots),
    seed = as.integer(random_seed),
    elapsed_time = elapsed_time,
    bootnet_level_warnings = unique(recorded_warnings),
    bootnet_level_messages = unique(recorded_messages),
    bootnet_version = as.character(
      utils::packageVersion("bootnet")
    ),
    R_version = R.version.string
  )
}

validate_case_drop <- function(
  result_object,
  expected_n_boots,
  expected_node_names,
  original_n,
  reference_graph
) {
  boot_object <- result_object$object
  node_n <- length(expected_node_names)
  statistic_names <- c(
    "expectedInfluence",
    "strength"
  )

  graph_checks <- vapply(
    boot_object$boots,
    function(one_boot) {
      graph <- one_boot$graph

      is.matrix(graph) &&
        identical(
          dim(graph),
          c(node_n, node_n)
        ) &&
        identical(
          rownames(graph),
          expected_node_names
        ) &&
        identical(
          colnames(graph),
          expected_node_names
        ) &&
        all(is.finite(graph)) &&
        max(abs(graph - t(graph))) < 1e-12 &&
        max(abs(diag(graph))) < 1e-12
    },
    logical(1)
  )

  retained_n <- vapply(
    boot_object$boots,
    function(one_boot) {
      as.integer(one_boot$nPerson)
    },
    integer(1)
  )

  minimum_expected_retained <- floor(
    original_n * (1 - 0.75)
  ) - 1L
  maximum_expected_retained <- ceiling(
    original_n * (1 - 0.05)
  ) + 1L

  sample_graph_difference <- max(
    abs(
      boot_object$sample$graph -
        reference_graph
    )
  )

  observed_types <- sort(
    unique(
      as.character(
        boot_object$bootTable$type
      )
    )
  )

  expected_boot_rows <-
    expected_n_boots *
    node_n *
    length(statistic_names)

  expected_sample_rows <-
    node_n *
    length(statistic_names)

  retry_count <- count_retry_warnings(
    result_object$bootnet_level_warnings
  )

  validation <- data.frame(
    Check = c(
      "Requested bootstrap count recorded",
      "Returned bootstrap network count",
      "Distinct bootstrap identifiers",
      "Expected centrality statistics stored",
      "Bootstrap centrality row count",
      "Original-sample centrality row count",
      "All bootstrap centrality values finite",
      "All returned bootstrap graphs valid",
      "Retained sample sizes within frozen range",
      "Original-sample graph matches Step-07 graph",
      "No top-level estimation warnings captured",
      "No failed-and-resampled estimates reported"
    ),
    Observed = c(
      as.character(result_object$nBoots),
      as.character(length(boot_object$boots)),
      as.character(
        length(
          unique(
            boot_object$bootTable$name
          )
        )
      ),
      paste(observed_types, collapse = "|"),
      as.character(nrow(boot_object$bootTable)),
      as.character(nrow(boot_object$sampleTable)),
      as.character(
        all(
          is.finite(
            boot_object$bootTable$value
          )
        )
      ),
      paste0(
        sum(graph_checks),
        "/",
        length(graph_checks)
      ),
      paste0(
        min(retained_n),
        "-",
        max(retained_n)
      ),
      format(
        sample_graph_difference,
        scientific = TRUE
      ),
      as.character(
        length(
          result_object$bootnet_level_warnings
        )
      ),
      as.character(retry_count)
    ),
    Expected = c(
      as.character(expected_n_boots),
      as.character(expected_n_boots),
      as.character(expected_n_boots),
      paste(sort(statistic_names), collapse = "|"),
      as.character(expected_boot_rows),
      as.character(expected_sample_rows),
      "TRUE",
      paste0(
        expected_n_boots,
        "/",
        expected_n_boots
      ),
      paste0(
        minimum_expected_retained,
        " to ",
        maximum_expected_retained
      ),
      "< 1e-8",
      "0",
      "0"
    ),
    Passed = c(
      identical(
        as.integer(result_object$nBoots),
        as.integer(expected_n_boots)
      ),
      length(boot_object$boots) == expected_n_boots,
      length(
        unique(
          boot_object$bootTable$name
        )
      ) == expected_n_boots,
      identical(
        observed_types,
        sort(statistic_names)
      ),
      nrow(boot_object$bootTable) == expected_boot_rows,
      nrow(boot_object$sampleTable) == expected_sample_rows,
      all(is.finite(boot_object$bootTable$value)),
      all(graph_checks),
      all(
        retained_n >= minimum_expected_retained &
          retained_n <= maximum_expected_retained
      ),
      sample_graph_difference < 1e-8,
      length(result_object$bootnet_level_warnings) == 0L,
      retry_count == 0L
    ),
    stringsAsFactors = FALSE
  )

  diagnostics <- data.frame(
    Requested_bootstraps = expected_n_boots,
    Returned_bootstraps = length(boot_object$boots),
    Original_N = original_n,
    Minimum_retained_N = min(retained_n),
    Maximum_retained_N = max(retained_n),
    Unique_retained_sample_sizes = length(
      unique(retained_n)
    ),
    Case_min = 0.05,
    Case_max = 0.75,
    Case_levels = 10L,
    Retry_count = retry_count,
    Captured_warning_n = length(
      result_object$bootnet_level_warnings
    ),
    Captured_warnings = paste(
      result_object$bootnet_level_warnings,
      collapse = " | "
    ),
    Captured_message_n = length(
      result_object$bootnet_level_messages
    ),
    Captured_messages = paste(
      result_object$bootnet_level_messages,
      collapse = " | "
    ),
    Elapsed_seconds = unname(
      result_object$elapsed_time[["elapsed"]]
    ),
    All_validation_checks_passed = all(
      validation$Passed
    ),
    stringsAsFactors = FALSE
  )

  list(
    validation = validation,
    diagnostics = diagnostics
  )
}

write_case_drop_log <- function(
  result_object,
  output_path
) {
  warning_text <- if (
    length(
      result_object$bootnet_level_warnings
    ) == 0L
  ) {
    "NO BOOTNET-LEVEL WARNINGS CAPTURED"
  } else {
    paste(
      result_object$bootnet_level_warnings,
      collapse = "\n"
    )
  }

  message_text <- if (
    length(
      result_object$bootnet_level_messages
    ) == 0L
  ) {
    "NO BOOTNET-LEVEL MESSAGES CAPTURED"
  } else {
    paste(
      result_object$bootnet_level_messages,
      collapse = "\n"
    )
  }

  writeLines(
    c(
      paste0(
        "Requested case-dropping repetitions: ",
        result_object$nBoots
      ),
      paste0("Random seed: ", result_object$seed),
      paste0(
        "bootnet version: ",
        result_object$bootnet_version
      ),
      paste0("R version: ", result_object$R_version),
      paste0(
        "Elapsed seconds: ",
        unname(
          result_object$elapsed_time[["elapsed"]]
        )
      ),
      "",
      "Captured bootnet-level warnings:",
      warning_text,
      "",
      "Captured bootnet-level messages:",
      message_text
    ),
    con = output_path
  )
}

classify_cs <- function(cs_value) {
  if (!is.finite(cs_value)) {
    return("NOT_ESTABLISHED")
  }

  if (cs_value < 0.25) {
    return("DO_NOT_INTERPRET_RANKING")
  }

  if (cs_value < 0.50) {
    return("RESTRICTED_EXPLORATORY_ONLY")
  }

  "CAUTIOUS_EXPLORATORY_INTERPRETATION"
}
```

## Run the authorised 2004 ten-repetition technical test

This run checks code execution and output structure only. Its CS coefficient
must not be calculated or interpreted, and it must not be used as evidence
of centrality stability.

```{r run-2004-technical-test}
pilot_validation_file <- file.path(
  pilot_table_dir,
  "07B_case_drop_2004_10_validation.csv"
)

pilot_diagnostics_file <- file.path(
  pilot_table_dir,
  "07B_case_drop_2004_10_diagnostics.csv"
)

if (RUN_2004_TECHNICAL_TEST) {
  pilot_seed <- 2026083020L
  pilot_wave_key <- "2004"
  pilot_network <- formal_network_results[[
    pilot_wave_key
  ]]$object

  pilot_result <- run_case_drop(
    network_object = pilot_network,
    n_boots = 10L,
    random_seed = pilot_seed
  )

  pilot_check <- validate_case_drop(
    result_object = pilot_result,
    expected_n_boots = 10L,
    expected_node_names = network_vars,
    original_n = unname(
      formal_expected_n[[pilot_wave_key]]
    ),
    reference_graph = pilot_network$graph
  )

  pilot_validation <- pilot_check$validation
  pilot_validation$Wave <- 2004L
  pilot_validation$Run_type <- "TECHNICAL_TEST_ONLY"

  pilot_diagnostics <- pilot_check$diagnostics
  pilot_diagnostics$Wave <- 2004L
  pilot_diagnostics$Seed <- pilot_seed
  pilot_diagnostics$Interpretation <-
    "NO_STABILITY_INFERENCE_FROM_TEN_REPETITIONS"

  saveRDS(
    pilot_result,
    file.path(
      model_dir,
      "07B_case_drop_2004_10_technical.rds"
    )
  )

  write.csv(
    pilot_validation,
    pilot_validation_file,
    row.names = FALSE
  )

  write.csv(
    pilot_diagnostics,
    pilot_diagnostics_file,
    row.names = FALSE
  )

  write_case_drop_log(
    pilot_result,
    file.path(
      log_dir,
      "07B_case_drop_2004_10_technical.log"
    )
  )

  pilot_validation
  pilot_diagnostics

  if (!all(pilot_validation$Passed)) {
    stop(
      paste(
        "The 2004 ten-repetition technical test completed with one or more failed checks.",
        "Inspect the saved validation, diagnostics and log files before any formal run."
      )
    )
  }
}
```

## Formal five-wave case-dropping gate

```{r formal-case-drop-gate}
pilot_gate_passed <- FALSE

if (
  file.exists(pilot_validation_file) &&
    file.info(pilot_validation_file)$size > 0
) {
  saved_pilot_validation <- read.csv(
    pilot_validation_file,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  pilot_gate_passed <-
    nrow(saved_pilot_validation) == 12L &&
    all(saved_pilot_validation$Passed) &&
    all(saved_pilot_validation$Wave == 2004L) &&
    all(
      saved_pilot_validation$Run_type ==
        "TECHNICAL_TEST_ONLY"
    )
}

formal_case_drop_allowed <-
  isTRUE(d23_authorised) &&
  isTRUE(pilot_gate_passed) &&
  isTRUE(RUN_FORMAL_FIVE_WAVE_CASE_DROP)

formal_case_drop_gate <- data.frame(
  D23_authorised = d23_authorised,
  Pilot_validation_passed = pilot_gate_passed,
  Formal_run_switch = RUN_FORMAL_FIVE_WAVE_CASE_DROP,
  Formal_execution_allowed = formal_case_drop_allowed,
  stringsAsFactors = FALSE
)

formal_case_drop_gate

if (
  RUN_FORMAL_FIVE_WAVE_CASE_DROP &&
    !formal_case_drop_allowed
) {
  stop(
    paste(
      "Formal five-wave case-dropping is blocked.",
      "D23 and the saved 2004 technical validation must both pass."
    )
  )
}
```

## Run the five formal 1,000-repetition case-dropping analyses

The formal run is disabled initially. When later enabled, it runs one wave
at a time, writes each wave's files before advancing, and stops after any
failed validation. It does not silently change the estimator or case-drop
range.

```{r run-formal-five-wave-case-drop}
formal_case_drop_results <- list()
formal_validation_list <- list()
formal_diagnostics_list <- list()
formal_cs_list <- list()

formal_case_drop_seeds <- c(
  `2004` = 2026083011L,
  `2012` = 2026083012L,
  `2016` = 2026083013L,
  `2020` = 2026083014L,
  `2024` = 2026083015L
)

if (formal_case_drop_allowed) {
  for (wave_key in as.character(formal_waves)) {
    network_object <- formal_network_results[[
      wave_key
    ]]$object

    wave_result <- run_case_drop(
      network_object = network_object,
      n_boots = 1000L,
      random_seed = unname(
        formal_case_drop_seeds[[wave_key]]
      )
    )

    wave_check <- validate_case_drop(
      result_object = wave_result,
      expected_n_boots = 1000L,
      expected_node_names = network_vars,
      original_n = unname(
        formal_expected_n[[wave_key]]
      ),
      reference_graph = network_object$graph
    )

    wave_validation <- wave_check$validation
    wave_validation$Wave <- as.integer(wave_key)
    wave_validation$Run_type <- "FORMAL_1000_CASE_DROP"

    wave_diagnostics <- wave_check$diagnostics
    wave_diagnostics$Wave <- as.integer(wave_key)
    wave_diagnostics$Seed <- unname(
      formal_case_drop_seeds[[wave_key]]
    )

    cs_error <- NA_character_
    cs_warnings <- character(0)
    cs_messages <- character(0)
    cs_values <- tryCatch(
      withCallingHandlers(
        bootnet::corStability(
          wave_result$object,
          cor = 0.70,
          statistics = c(
            "expectedInfluence",
            "strength"
          ),
          verbose = TRUE
        ),
        warning = function(condition) {
          cs_warnings <<- c(
            cs_warnings,
            conditionMessage(condition)
          )
          invokeRestart("muffleWarning")
        },
        message = function(condition) {
          cs_messages <<- c(
            cs_messages,
            conditionMessage(condition)
          )
          invokeRestart("muffleMessage")
        }
      ),
      error = function(condition) {
        cs_error <<- conditionMessage(condition)
        c(
          expectedInfluence = NA_real_,
          strength = NA_real_
        )
      }
    )

    wave_cs <- data.frame(
      Wave = as.integer(wave_key),
      Statistic = names(cs_values),
      CS_coefficient = as.numeric(cs_values),
      Correlation_threshold = 0.70,
      Interpretation_status = vapply(
        as.numeric(cs_values),
        classify_cs,
        FUN.VALUE = character(1)
      ),
      CS_error = ifelse(
        is.na(cs_error),
        "",
        cs_error
      ),
      stringsAsFactors = FALSE
    )

    model_file <- file.path(
      model_dir,
      paste0(
        "07B_case_drop_",
        wave_key,
        "_1000.rds"
      )
    )

    validation_file <- file.path(
      table_dir,
      paste0(
        "07B_case_drop_",
        wave_key,
        "_validation.csv"
      )
    )

    diagnostics_file <- file.path(
      table_dir,
      paste0(
        "07B_case_drop_",
        wave_key,
        "_diagnostics.csv"
      )
    )

    cs_file <- file.path(
      table_dir,
      paste0(
        "07B_case_drop_",
        wave_key,
        "_CS.csv"
      )
    )

    figure_file <- file.path(
      figure_dir,
      paste0(
        "07B_centrality_stability_",
        wave_key,
        ".png"
      )
    )

    saveRDS(
      wave_result,
      model_file
    )

    write.csv(
      wave_validation,
      validation_file,
      row.names = FALSE
    )

    write.csv(
      wave_diagnostics,
      diagnostics_file,
      row.names = FALSE
    )

    write.csv(
      wave_cs,
      cs_file,
      row.names = FALSE
    )

    write_case_drop_log(
      wave_result,
      file.path(
        log_dir,
        paste0(
          "07B_case_drop_",
          wave_key,
          "_1000.log"
        )
      )
    )

    plot_error <- NA_character_
    stability_plot <- tryCatch(
      plot(
        wave_result$object,
        statistics = c(
          "expectedInfluence",
          "strength"
        )
      ),
      error = function(condition) {
        plot_error <<- conditionMessage(condition)
        NULL
      }
    )

    if (!is.null(stability_plot)) {
      ggplot2::ggsave(
        filename = figure_file,
        plot = stability_plot,
        width = 9,
        height = 6,
        units = "in",
        dpi = 300
      )
    }

    wave_diagnostics$CS_calculated <- all(
      is.finite(wave_cs$CS_coefficient)
    )
    wave_diagnostics$CS_warning_n <- length(
      unique(cs_warnings)
    )
    wave_diagnostics$CS_warnings <- paste(
      unique(cs_warnings),
      collapse = " | "
    )
    wave_diagnostics$CS_message_n <- length(
      unique(cs_messages)
    )
    wave_diagnostics$CS_messages <- paste(
      unique(cs_messages),
      collapse = " | "
    )
    wave_diagnostics$Stability_plot_created <-
      file.exists(figure_file) &&
      file.info(figure_file)$size > 0
    wave_diagnostics$Plot_error <- ifelse(
      is.na(plot_error),
      "",
      plot_error
    )

    write.csv(
      wave_diagnostics,
      diagnostics_file,
      row.names = FALSE
    )

    formal_case_drop_results[[wave_key]] <- wave_result
    formal_validation_list[[wave_key]] <- wave_validation
    formal_diagnostics_list[[wave_key]] <- wave_diagnostics
    formal_cs_list[[wave_key]] <- wave_cs

    if (
      !all(wave_validation$Passed) ||
        !all(is.finite(wave_cs$CS_coefficient)) ||
        !wave_diagnostics$Stability_plot_created
    ) {
      stop(
        paste0(
          "Formal case-dropping validation failed for ",
          wave_key,
          ". Inspect its saved files before continuing."
        )
      )
    }
  }

  formal_validation <- do.call(
    rbind,
    formal_validation_list
  )

  formal_diagnostics <- do.call(
    rbind,
    formal_diagnostics_list
  )

  formal_cs <- do.call(
    rbind,
    formal_cs_list
  )

  saveRDS(
    formal_case_drop_results,
    file.path(
      model_dir,
      "07B_case_drop_five_wave_1000.rds"
    )
  )

  write.csv(
    formal_validation,
    file.path(
      table_dir,
      "07B_case_drop_validation.csv"
    ),
    row.names = FALSE
  )

  write.csv(
    formal_diagnostics,
    file.path(
      table_dir,
      "07B_case_drop_diagnostics.csv"
    ),
    row.names = FALSE
  )

  write.csv(
    formal_cs,
    file.path(
      table_dir,
      "07B_CS_coefficients.csv"
    ),
    row.names = FALSE
  )
}
```

## Reuse and validate saved formal outputs

This section permits a later HTML render without rerunning the 5,000
formal bootstrap samples. It does not accept incomplete output sets.

```{r reuse-saved-formal-outputs}
if (REUSE_VALIDATED_FORMAL_OUTPUTS) {
  combined_formal_files <- c(
    file.path(
      model_dir,
      "07B_case_drop_five_wave_1000.rds"
    ),
    file.path(
      table_dir,
      "07B_case_drop_validation.csv"
    ),
    file.path(
      table_dir,
      "07B_case_drop_diagnostics.csv"
    ),
    file.path(
      table_dir,
      "07B_CS_coefficients.csv"
    )
  )

  expected_figures <- file.path(
    figure_dir,
    paste0(
      "07B_centrality_stability_",
      formal_waves,
      ".png"
    )
  )

  stopifnot(
    all(file.exists(combined_formal_files)),
    all(file.info(combined_formal_files)$size > 0),
    all(file.exists(expected_figures)),
    all(file.info(expected_figures)$size > 0)
  )

  formal_validation <- read.csv(
    combined_formal_files[[2L]],
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  formal_diagnostics <- read.csv(
    combined_formal_files[[3L]],
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  formal_cs <- read.csv(
    combined_formal_files[[4L]],
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  stopifnot(
    nrow(formal_validation) == 60L,
    all(formal_validation$Passed),
    nrow(formal_diagnostics) == 5L,
    all(
      formal_diagnostics$All_validation_checks_passed
    ),
    all(formal_diagnostics$CS_calculated),
    all(
      formal_diagnostics$Stability_plot_created
    ),
    nrow(formal_cs) == 10L,
    all(is.finite(formal_cs$CS_coefficient))
  )

  formal_cs
}
```

## Completion status and session information

```{r completion-and-session-info}
completion_status <- if (
  RUN_FORMAL_FIVE_WAVE_CASE_DROP
) {
  "FORMAL_CASE_DROPPING_EXECUTED_SUBJECT_TO_SAVED_VALIDATION"
} else if (RUN_2004_TECHNICAL_TEST) {
  "2004_TEN_REPETITION_TECHNICAL_TEST_EXECUTED_NO_STABILITY_INFERENCE"
} else if (REUSE_VALIDATED_FORMAL_OUTPUTS) {
  "VALIDATED_FORMAL_OUTPUTS_REUSED_NO_MODEL_RERUN"
} else {
  "PREFLIGHT_AND_RAW_CENTRALITY_ONLY"
}

completion_table <- data.frame(
  Component = c(
    "D23 authorisation",
    "Five-wave raw Expected Influence",
    "Five-wave auxiliary raw Strength",
    "2004 ten-repetition technical test",
    "Five-wave 1,000-repetition case-dropping",
    "Education-group centrality",
    "Bridge Expected Influence",
    "Closeness and betweenness",
    "NCT or cross-wave centrality tests",
    "Political interpretation"
  ),
  Status = c(
    "AUTHORISED",
    "CALCULATED_STABILITY_NOT_IMPLIED",
    "CALCULATED_STABILITY_NOT_IMPLIED",
    ifelse(
      RUN_2004_TECHNICAL_TEST,
      "EXECUTED_TECHNICAL_ONLY",
      "NOT_RUN_THIS_RENDER"
    ),
    ifelse(
      RUN_FORMAL_FIVE_WAVE_CASE_DROP,
      "EXECUTED_SUBJECT_TO_VALIDATION",
      ifelse(
        REUSE_VALIDATED_FORMAL_OUTPUTS,
        "REUSED_VALIDATED_OUTPUTS",
        "NOT_RUN_THIS_RENDER"
      )
    ),
    "NOT_AUTHORISED_NOT_RUN",
    "NOT_AUTHORISED_NOT_RUN",
    "NOT_AUTHORISED_NOT_RUN",
    "NOT_AUTHORISED_NOT_RUN",
    "NOT_AUTHORISED_NOT_RUN"
  ),
  stringsAsFactors = FALSE
)

write.csv(
  completion_table,
  file.path(
    table_dir,
    "07B_completion_status.csv"
  ),
  row.names = FALSE
)

writeLines(
  c(
    paste0("07B status: ", completion_status),
    paste0("Rendered at: ", Sys.time()),
    "",
    "Interpretation boundary:",
    paste(
      "Expected Influence is net signed conditional connectivity under",
      "the pre-specified coding. It is not causal influence, political",
      "importance, public salience or an intervention target."
    )
  ),
  con = file.path(
    log_dir,
    "07B_centrality_stability_status.log"
  )
)

capture.output(
  sessionInfo(),
  file = file.path(
    log_dir,
    "07B_centrality_stability_sessionInfo.txt"
  )
)

completion_table
```

## Interpretation rule registered before results

The case-dropping CS coefficient is interpreted at correlation 0.70 as
follows: below 0.25, do not interpret node rankings; from 0.25 to below
0.50, permit restricted exploratory description only; at or above 0.50,
permit cautious exploratory interpretation. These thresholds do not turn
centrality into a causal or confirmatory result. A low CS coefficient does
not remove the corresponding wave from the primary network analysis.
