---
title: "08 Network Comparisons: Stages 1-12"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

## Purpose and execution boundary

This notebook implements only the frozen Step-08 stages 1-12. It reuses
the five validated Step-07 overall networks, generates descriptive
cross-wave comparison tables and common-layout figures, links the
existing edge-accuracy and exploratory centrality evidence, and creates
the eight authorised education-group samples for later diagnostics.

It does not re-estimate the five overall networks, run a formal
between-network hypothesis test, estimate education-group networks,
run an education bootstrap, run sensitivity models, or interpret
political findings.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  error = FALSE,
  warning = TRUE,
  message = TRUE
)
```

## Stage 1: locate the project and create derived-output paths

```{r project-root}
current_dir <- normalizePath(
  ".",
  winslash = "/"
)

if (
  file.exists(
    file.path(
      current_dir,
      "ANES.Rproj"
    )
  )
) {
  project_root <- current_dir
} else if (
  basename(current_dir) == "notebooks" &&
    file.exists(
      file.path(
        dirname(current_dir),
        "ANES.Rproj"
      )
    )
) {
  project_root <- dirname(current_dir)
} else {
  stop(
    "Project root could not be identified."
  )
}

knitr::opts_knit$set(
  root.dir = project_root
)

notebook_file <- file.path(
  project_root,
  "notebooks",
  "08_network_comparisons.Rmd"
)

formal_table_dir <- file.path(
  project_root,
  "outputs",
  "tables",
  "formal"
)

formal_figure_dir <- file.path(
  project_root,
  "outputs",
  "figures",
  "formal"
)

formal_model_dir <- file.path(
  project_root,
  "outputs",
  "models",
  "formal"
)

log_dir <- file.path(
  project_root,
  "outputs",
  "logs"
)

analysis_ready_dir <- file.path(
  project_root,
  "data",
  "analysis_ready"
)

intermediate_dir <- file.path(
  project_root,
  "data",
  "intermediate"
)

for (
  target_dir in c(
    formal_table_dir,
    formal_figure_dir,
    formal_model_dir,
    log_dir,
    analysis_ready_dir,
    intermediate_dir
  )
) {
  dir.create(
    target_dir,
    recursive = TRUE,
    showWarnings = FALSE
  )
}

stopifnot(
  all(
    dir.exists(
      c(
        formal_table_dir,
        formal_figure_dir,
        formal_model_dir,
        log_dir,
        analysis_ready_dir,
        intermediate_dir
      )
    )
  )
)

project_root
```

## Stage 1: package and helper-function gate

```{r package-gate}
required_packages <- c(
  "bootnet",
  "digest",
  "ggplot2",
  "knitr",
  "qgraph",
  "rmarkdown"
)

package_gate <- data.frame(
  Package = required_packages,
  Installed = vapply(
    required_packages,
    requireNamespace,
    quietly = TRUE,
    FUN.VALUE = logical(1)
  ),
  Version = vapply(
    required_packages,
    function(package_name) {
      if (
        requireNamespace(
          package_name,
          quietly = TRUE
        )
      ) {
        as.character(
          utils::packageVersion(
            package_name
          )
        )
      } else {
        NA_character_
      }
    },
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

stopifnot(
  all(package_gate$Installed)
)

normalise_status <- function(value) {
  value <- toupper(
    trimws(
      as.character(value)
    )
  )
  gsub(
    "[^A-Z0-9]+",
    "_",
    value
  )
}

as_pass <- function(value) {
  toupper(
    trimws(
      as.character(value)
    )
  ) == "TRUE"
}

file_sha256 <- function(path) {
  if (!file.exists(path)) {
    return(NA_character_)
  }

  toupper(
    digest::digest(
      file = path,
      algo = "sha256",
      serialize = FALSE
    )
  )
}

safe_max <- function(values) {
  values <- values[is.finite(values)]
  if (length(values) == 0L) {
    return(NA_real_)
  }
  max(values)
}
```

## Stage 1: current decision-registry gate

```{r decision-registry-gate}
registry_config <- file.path(
  project_root,
  "config",
  "decision_registry_path.txt"
)

stopifnot(
  file.exists(registry_config)
)

registry_lines <- trimws(
  readLines(
    registry_config,
    warn = FALSE
  )
)
registry_lines <- registry_lines[nzchar(registry_lines)]

stopifnot(
  length(registry_lines) == 1L
)

registry_path <- registry_lines[[1L]]

stopifnot(
  file.exists(registry_path)
)

decision_registry <- read.csv(
  registry_path,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

required_registry_columns <- c(
  "decision_id",
  "current_value",
  "decision_status",
  "implementation_status",
  "researcher_approval_status"
)

stopifnot(
  all(
    required_registry_columns %in%
      names(decision_registry)
  )
)

decision_row <- function(decision_id) {
  decision_registry[
    decision_registry$decision_id == decision_id,
    ,
    drop = FALSE
  ]
}

d03 <- decision_row("D03")
d04 <- decision_row("D04")
d08a <- decision_row("D08A")
d08b <- decision_row("D08B")
d15 <- decision_row("D15")
d18 <- decision_row("D18")
d21 <- decision_row("D21")
d23 <- decision_row("D23")
d24 <- decision_row("D24")

decision_gate <- data.frame(
  Decision_ID = c(
    "D03",
    "D04",
    "D08A",
    "D08B",
    "D15",
    "D18",
    "D21",
    "D23",
    "D24"
  ),
  Check = c(
    "The five primary overall waves are frozen",
    "2008 is excluded from all later network analysis",
    "Descriptive education-network waves are frozen",
    "Formal education inference is outside current scope",
    "Primary overall networks are unweighted",
    "NCT is outside the current bounded design",
    "Step 07 is complete and validated",
    "Exploratory centrality is complete and validated",
    "Step 08 descriptive comparison scope is frozen"
  ),
  Passed = c(
    nrow(d03) == 1L &&
      normalise_status(d03$decision_status) == "FROZEN" &&
      normalise_status(d03$researcher_approval_status) == "APPROVED" &&
      grepl(
        "2004, 2012, 2016, 2020 and 2024",
        d03$current_value,
        fixed = TRUE
      ),
    nrow(d04) == 1L &&
      normalise_status(d04$decision_status) == "FROZEN" &&
      normalise_status(d04$researcher_approval_status) == "APPROVED" &&
      grepl(
        "Exclude 2008",
        d04$current_value,
        fixed = TRUE
      ),
    nrow(d08a) == 1L &&
      normalise_status(d08a$decision_status) == "FROZEN" &&
      normalise_status(d08a$researcher_approval_status) == "APPROVED",
    nrow(d08b) == 1L &&
      normalise_status(d08b$decision_status) == "NOT_IN_CURRENT_SCOPE" &&
      normalise_status(d08b$researcher_approval_status) == "APPROVED",
    nrow(d15) == 1L &&
      normalise_status(d15$decision_status) == "FROZEN" &&
      normalise_status(d15$researcher_approval_status) == "APPROVED" &&
      grepl(
        "Unweighted primary",
        d15$current_value,
        fixed = TRUE
      ),
    nrow(d18) == 1L &&
      normalise_status(d18$decision_status) == "NOT_IN_CURRENT_SCOPE" &&
      normalise_status(d18$implementation_status) == "NOT_APPLICABLE" &&
      normalise_status(d18$researcher_approval_status) == "APPROVED",
    nrow(d21) == 1L &&
      normalise_status(d21$implementation_status) == "COMPLETE_AND_VALIDATED" &&
      normalise_status(d21$researcher_approval_status) == "APPROVED",
    nrow(d23) == 1L &&
      normalise_status(d23$implementation_status) == "COMPLETE_AND_VALIDATED" &&
      normalise_status(d23$researcher_approval_status) == "APPROVED",
       nrow(d24) == 1L &&
      normalise_status(d24$decision_status) == "FROZEN" &&
      normalise_status(d24$researcher_approval_status) == "APPROVED" &&
      normalise_status(d24$implementation_status) %in% c(
        "AUTHORISED_STAGES_1_TO_12",
        "STAGES_1_TO_12_COMPLETE_AND_VALIDATED"
      )
  ),
  stringsAsFactors = FALSE
)
write.csv(
  decision_gate,
  file.path(
    formal_table_dir,
    "08_decision_gate.csv"
  ),
  row.names = FALSE
)

decision_gate

stopifnot(
  all(decision_gate$Passed)
)

write.csv(
  package_gate,
  file.path(
    formal_table_dir,
    "08_package_gate.csv"
  ),
  row.names = FALSE
)
```

## Stage 2: inventory the validated inputs

```{r input-paths-and-inventory}
formal_network_file <- file.path(
  formal_model_dir,
  "07_formal_five_wave_networks_gamma050.rds"
)

point_diagnostics_file <- file.path(
  formal_table_dir,
  "07_formal_five_wave_point_network_diagnostics.csv"
)

point_validation_file <- file.path(
  formal_table_dir,
  "07_formal_five_wave_point_network_validation.csv"
)

step07_completion_file <- file.path(
  formal_table_dir,
  "07_formal_step07_completion_status.csv"
)

edge_weight_file <- file.path(
  formal_table_dir,
  "07_formal_five_wave_edge_weights.csv"
)

edge_accuracy_file <- file.path(
  formal_table_dir,
  "07A_edge_accuracy.csv"
)

edge_accuracy_validation_file <- file.path(
  formal_table_dir,
  "07A_edge_accuracy_validation.csv"
)

centrality_file <- file.path(
  formal_table_dir,
  "07B_raw_centrality_scores.csv"
)

centrality_cs_file <- file.path(
  formal_table_dir,
  "07B_CS_coefficients.csv"
)

centrality_completion_file <- file.path(
  formal_table_dir,
  "07B_completion_status.csv"
)

analysis_ready_file <- file.path(
  analysis_ready_dir,
  paste0(
    "anes_cdf_20260205_analysis_ready_nine_node_",
    "complete_case_2004_2012_2016_2020_2024.rds"
  )
)

input_paths <- c(
  Decision_registry = registry_path,
  Formal_network_bundle = formal_network_file,
  Point_diagnostics = point_diagnostics_file,
  Point_validation = point_validation_file,
  Step07_completion = step07_completion_file,
  Edge_weights = edge_weight_file,
  Edge_accuracy = edge_accuracy_file,
  Edge_accuracy_validation = edge_accuracy_validation_file,
  Centrality_scores = centrality_file,
  Centrality_CS = centrality_cs_file,
  Centrality_completion = centrality_completion_file,
  Analysis_ready_data = analysis_ready_file
)

input_inventory <- data.frame(
  Input = names(input_paths),
  Path = normalizePath(
    input_paths,
    winslash = "/",
    mustWork = FALSE
  ),
  Exists = file.exists(input_paths),
  Bytes = ifelse(
    file.exists(input_paths),
    file.info(input_paths)$size,
    NA_real_
  ),
  Modified_UTC = ifelse(
    file.exists(input_paths),
    format(
      file.info(input_paths)$mtime,
      tz = "UTC",
      usetz = TRUE
    ),
    NA_character_
  ),
  SHA256 = vapply(
    input_paths,
    file_sha256,
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

write.csv(
  input_inventory,
  file.path(
    formal_table_dir,
    "08_input_inventory.csv"
  ),
  row.names = FALSE
)

input_inventory

stopifnot(
  all(input_inventory$Exists)
)
```

## Stage 2: load the validated Step-07 and analysis-ready objects

```{r load-validated-inputs}
formal_network_results <- readRDS(
  formal_network_file
)

point_diagnostics <- read.csv(
  point_diagnostics_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

point_validation <- read.csv(
  point_validation_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

step07_completion <- read.csv(
  step07_completion_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

edge_weights <- read.csv(
  edge_weight_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

edge_accuracy <- read.csv(
  edge_accuracy_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

edge_accuracy_validation <- read.csv(
  edge_accuracy_validation_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

centrality_scores <- read.csv(
  centrality_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

centrality_cs <- read.csv(
  centrality_cs_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

centrality_completion <- read.csv(
  centrality_completion_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

anes_analysis_ready <- readRDS(
  analysis_ready_file
)
```

## Stage 3: overall-network technical preflight

```{r overall-input-preflight}
expected_years <- c(
  2004L,
  2012L,
  2016L,
  2020L,
  2024L
)

expected_n <- c(
  "2004" = 789L,
  "2012" = 4353L,
  "2016" = 2709L,
  "2020" = 5390L,
  "2024" = 3525L
)

network_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223"
)

expected_categories <- list(
  VCF0806 = c(1, 2, 3, 4, 5, 6, 7),
  VCF0809 = c(1, 2, 3, 4, 5, 6, 7),
  VCF0838 = c(1, 2, 3, 4),
  VCF0839 = c(1, 2, 3, 4, 5, 6, 7),
  VCF0879a = c(1, 3, 5),
  VCF0888 = c(1, 2, 3),
  VCF0890 = c(1, 2, 3),
  VCF0894 = c(1, 2, 3),
  VCF9223 = c(1, 2, 3, 4)
)

stopifnot(
  identical(
    names(formal_network_results),
    as.character(expected_years)
  )
)

graph_list <- lapply(
  formal_network_results,
  function(one_result) {
    one_result$object$graph
  }
)

overall_input_validation <- do.call(
  rbind,
  lapply(
    expected_years,
    function(wave) {
      wave_name <- as.character(wave)
      one_result <- formal_network_results[[wave_name]]$object
      graph <- one_result$graph

      data.frame(
        Wave = wave,
        Expected_N = unname(expected_n[[wave_name]]),
        Observed_N = as.integer(one_result$nPerson),
        Nine_by_nine = identical(
          dim(graph),
          c(9L, 9L)
        ),
        Correct_node_order = identical(
          rownames(graph),
          network_vars
        ) &&
          identical(
            colnames(graph),
            network_vars
          ),
        Graph_finite = all(
          is.finite(graph)
        ),
        Graph_symmetric = max(
          abs(
            graph - t(graph)
          )
        ) < 1e-12,
        Zero_diagonal = max(
          abs(
            diag(graph)
          )
        ) < 1e-12,
        No_2008 = wave != 2008L,
        Passed =
          as.integer(one_result$nPerson) ==
            unname(expected_n[[wave_name]]) &&
            identical(
              dim(graph),
              c(9L, 9L)
            ) &&
            identical(
              rownames(graph),
              network_vars
            ) &&
            identical(
              colnames(graph),
              network_vars
            ) &&
            all(is.finite(graph)) &&
            max(abs(graph - t(graph))) < 1e-12 &&
            max(abs(diag(graph))) < 1e-12 &&
            wave != 2008L,
        stringsAsFactors = FALSE
      )
    }
  )
)

analysis_ready_hash_expected <-
  "6726A0D548F8293E890AD380C4915097FDD4ABC1A51665D55E904A33F81432FA"

overall_preflight <- data.frame(
  Check = c(
    "Five formal network objects are present",
    "All graph-level input checks passed",
    "Step-07 point validation rows all passed",
    "Step-07 primary networks are complete",
    "07A edge-accuracy validation passed",
    "D23 centrality completion record is present",
    "Analysis-ready file hash matches the authorised input",
    "Analysis-ready object has 16,766 rows and 18 columns",
    "Analysis-ready years match the five frozen waves",
    "Analysis-ready object contains the nine nodes",
    "No 2008 object is present"
  ),
  Passed = c(
    length(formal_network_results) == 5L,
    all(overall_input_validation$Passed),
    all(as_pass(point_validation$Passed)),
    all(
      step07_completion$Status[
        step07_completion$Component %in%
          c(
            "Five ordered complete-case inputs",
            "Five primary point networks",
            "Five 1,000-repetition edge bootstraps"
          )
      ] == "COMPLETE_AND_VALIDATED"
    ),
    all(as_pass(edge_accuracy_validation$Passed)),
    any(
      centrality_completion$Component ==
        "Five-wave 1,000-repetition case-dropping" &
        centrality_completion$Status ==
        "REUSED_VALIDATED_OUTPUTS"
    ),
    identical(
      file_sha256(analysis_ready_file),
      analysis_ready_hash_expected
    ),
    identical(
      dim(anes_analysis_ready),
      c(16766L, 18L)
    ),
    identical(
      sort(
        unique(
          as.integer(
            anes_analysis_ready$VCF0004
          )
        )
      ),
      expected_years
    ),
    all(
      network_vars %in%
        names(anes_analysis_ready)
    ),
    !("2008" %in% names(formal_network_results))
  ),
  stringsAsFactors = FALSE
)

write.csv(
  overall_input_validation,
  file.path(
    formal_table_dir,
    "08_overall_input_validation.csv"
  ),
  row.names = FALSE
)

write.csv(
  overall_preflight,
  file.path(
    formal_table_dir,
    "08_overall_preflight.csv"
  ),
  row.names = FALSE
)

overall_preflight

stopifnot(
  all(overall_preflight$Passed)
)
```

## Stage 4: create and validate one common layout

```{r common-layout}
average_absolute_graph <- Reduce(
  "+",
  lapply(
    graph_list,
    abs
  )
) / length(graph_list)

layout_warnings <- character(0)
layout_messages <- character(0)

set.seed(2026083008)

layout_object <- withCallingHandlers(
  qgraph::qgraph(
    average_absolute_graph,
    layout = "spring",
    DoNotPlot = TRUE,
    details = FALSE
  ),
  warning = function(condition) {
    layout_warnings <<- c(
      layout_warnings,
      conditionMessage(condition)
    )
    invokeRestart("muffleWarning")
  },
  message = function(condition) {
    layout_messages <<- c(
      layout_messages,
      conditionMessage(condition)
    )
    invokeRestart("muffleMessage")
  }
)

common_layout <- layout_object$layout
rownames(common_layout) <- network_vars
colnames(common_layout) <- c(
  "x",
  "y"
)

layout_validation <- data.frame(
  Check = c(
    "Layout is 9 by 2",
    "Layout follows the frozen node order",
    "All layout coordinates are finite",
    "No layout warning was captured"
  ),
  Passed = c(
    identical(
      dim(common_layout),
      c(9L, 2L)
    ),
    identical(
      rownames(common_layout),
      network_vars
    ),
    all(
      is.finite(common_layout)
    ),
    length(layout_warnings) == 0L
  ),
  stringsAsFactors = FALSE
)

layout_file <- file.path(
  formal_model_dir,
  "08_common_layout.rds"
)

saveRDS(
  common_layout,
  layout_file
)

layout_coordinates <- data.frame(
  Node = rownames(common_layout),
  x = common_layout[, "x"],
  y = common_layout[, "y"],
  stringsAsFactors = FALSE
)

write.csv(
  layout_coordinates,
  file.path(
    formal_table_dir,
    "08_common_layout_coordinates.csv"
  ),
  row.names = FALSE
)

write.csv(
  layout_validation,
  file.path(
    formal_table_dir,
    "08_common_layout_validation.csv"
  ),
  row.names = FALSE
)

layout_validation

stopifnot(
  all(layout_validation$Passed)
)
```

## Stage 5: render the five overall networks on the common scale

```{r overall-network-figures}
node_short_labels <- c(
  "Health",
  "Jobs",
  "Abortion",
  "Services",
  "Immig. level",
  "Crime",
  "Schools",
  "Welfare",
  "Immig. jobs"
)

common_edge_maximum <- max(
  abs(
    unlist(
      graph_list,
      use.names = FALSE
    )
  )
)

plot_one_network <- function(
  graph,
  wave,
  output_file
) {
  grDevices::png(
    filename = output_file,
    width = 1800,
    height = 1500,
    res = 180
  )

  old_parameters <- graphics::par(
    mar = c(2, 2, 4, 2)
  )

  on.exit(
    {
      graphics::par(old_parameters)
      grDevices::dev.off()
    },
    add = TRUE
  )

  qgraph::qgraph(
    graph,
    layout = common_layout,
    labels = node_short_labels,
    color = "#F2F2F2",
    border.color = "#333333",
    vsize = 8,
    label.cex = 0.9,
    maximum = common_edge_maximum,
    minimum = 0,
    cut = 0,
    posCol = "#2166AC",
    negCol = "#B2182B",
    fade = FALSE,
    legend = FALSE,
    edge.labels = FALSE,
    details = FALSE,
    title = paste0(
      wave,
      " overall network"
    )
  )

  invisible(output_file)
}

figure_files <- vapply(
  expected_years,
  function(wave) {
    output_file <- file.path(
      formal_figure_dir,
      paste0(
        "08_overall_network_",
        wave,
        ".png"
      )
    )

    plot_one_network(
      graph = graph_list[[as.character(wave)]],
      wave = wave,
      output_file = output_file
    )

    output_file
  },
  FUN.VALUE = character(1)
)

panel_file <- file.path(
  formal_figure_dir,
  "08_overall_networks_five_wave_panel.png"
)

grDevices::png(
  filename = panel_file,
  width = 3000,
  height = 2100,
  res = 200
)

old_parameters <- graphics::par(
  mfrow = c(2, 3),
  mar = c(2, 2, 4, 2)
)

for (wave in expected_years) {
  qgraph::qgraph(
    graph_list[[as.character(wave)]],
    layout = common_layout,
    labels = node_short_labels,
    color = "#F2F2F2",
    border.color = "#333333",
    vsize = 7,
    label.cex = 0.75,
    maximum = common_edge_maximum,
    minimum = 0,
    cut = 0,
    posCol = "#2166AC",
    negCol = "#B2182B",
    fade = FALSE,
    legend = FALSE,
    edge.labels = FALSE,
    details = FALSE,
    title = as.character(wave)
  )
}

graphics::plot.new()
graphics::text(
  x = 0.5,
  y = 0.6,
  labels = "Common node layout\nand common edge scale",
  cex = 1.3
)

graphics::par(old_parameters)
grDevices::dev.off()

all_figure_files <- c(
  figure_files,
  panel_file
)

figure_manifest <- data.frame(
  Figure = basename(all_figure_files),
  Path = normalizePath(
    all_figure_files,
    winslash = "/",
    mustWork = FALSE
  ),
  Exists = file.exists(all_figure_files),
  Bytes = file.info(all_figure_files)$size,
  SHA256 = vapply(
    all_figure_files,
    file_sha256,
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

write.csv(
  figure_manifest,
  file.path(
    formal_table_dir,
    "08_overall_figure_manifest.csv"
  ),
  row.names = FALSE
)

stopifnot(
  all(figure_manifest$Exists),
  all(figure_manifest$Bytes > 0)
)

figure_manifest
```

## Stage 6: create the five-wave overall summary

```{r overall-wave-summary}
required_point_columns <- c(
  "Wave",
  "N",
  "Nodes",
  "Possible_edges",
  "Retained_edges",
  "Global_strength_descriptive",
  "Captured_warning_n",
  "Captured_warnings"
)

stopifnot(
  all(
    required_point_columns %in%
      names(point_diagnostics)
  )
)

overall_wave_summary <- point_diagnostics[
  match(
    expected_years,
    point_diagnostics$Wave
  ),
  required_point_columns,
  drop = FALSE
]

overall_wave_summary$Density <- with(
  overall_wave_summary,
  Retained_edges / Possible_edges
)

overall_wave_summary$Comparison_status <-
  "DESCRIPTIVE_POINT_ESTIMATE"

overall_wave_summary$Global_strength_interpretation <-
  "SUM_OF_ABSOLUTE_EDGE_WEIGHTS_NOT_IDEOLOGICAL_CONSTRAINT"

write.csv(
  overall_wave_summary,
  file.path(
    formal_table_dir,
    "08_overall_wave_summary.csv"
  ),
  row.names = FALSE
)

overall_wave_summary
```

## Stage 7: calculate the five frozen descriptive edge contrasts

```{r temporal-edge-contrasts}
contrast_plan <- data.frame(
  Contrast_order = 1:5,
  Earlier_wave = c(
    2004L,
    2012L,
    2016L,
    2020L,
    2004L
  ),
  Later_wave = c(
    2012L,
    2016L,
    2020L,
    2024L,
    2024L
  ),
  Years_elapsed = c(
    8L,
    4L,
    4L,
    4L,
    20L
  ),
  Contrast_role = c(
    rep(
      "CONSECUTIVE_DESCRIPTIVE",
      4
    ),
    "ENDPOINT_DESCRIPTIVE"
  ),
  stringsAsFactors = FALSE
)

contrast_plan$Contrast <- paste0(
  contrast_plan$Earlier_wave,
  "_to_",
  contrast_plan$Later_wave
)

make_edge_contrast <- function(
  earlier_wave,
  later_wave,
  contrast_label,
  contrast_order,
  years_elapsed,
  contrast_role
) {
  earlier <- edge_weights[
    edge_weights$Wave == earlier_wave,
    c(
      "Node_1",
      "Node_2",
      "Edge_weight",
      "Retained_under_EBICglasso"
    ),
    drop = FALSE
  ]

  later <- edge_weights[
    edge_weights$Wave == later_wave,
    c(
      "Node_1",
      "Node_2",
      "Edge_weight",
      "Retained_under_EBICglasso"
    ),
    drop = FALSE
  ]

  names(earlier)[3:4] <- c(
    "Earlier_edge",
    "Earlier_retained"
  )
  names(later)[3:4] <- c(
    "Later_edge",
    "Later_retained"
  )

  earlier$Edge_ID <- paste(
    earlier$Node_1,
    earlier$Node_2,
    sep = "--"
  )
  later$Edge_ID <- paste(
    later$Node_1,
    later$Node_2,
    sep = "--"
  )

  merged <- merge(
    earlier,
    later,
    by = c(
      "Edge_ID",
      "Node_1",
      "Node_2"
    ),
    all = FALSE,
    sort = FALSE
  )

  merged <- merged[
    match(
      earlier$Edge_ID,
      merged$Edge_ID
    ),
    ,
    drop = FALSE
  ]

  merged$Contrast_order <- contrast_order
  merged$Contrast <- contrast_label
  merged$Earlier_wave <- earlier_wave
  merged$Later_wave <- later_wave
  merged$Years_elapsed <- years_elapsed
  merged$Contrast_role <- contrast_role
  merged$Signed_difference <-
    merged$Later_edge - merged$Earlier_edge
  merged$Absolute_difference <-
    abs(merged$Signed_difference)
  merged$Sign_change <-
    merged$Earlier_edge != 0 &
    merged$Later_edge != 0 &
    sign(merged$Earlier_edge) !=
      sign(merged$Later_edge)
  merged$Retention_change <- ifelse(
    !merged$Earlier_retained &
      merged$Later_retained,
    "NOT_RETAINED_TO_RETAINED",
    ifelse(
      merged$Earlier_retained &
        !merged$Later_retained,
      "RETAINED_TO_NOT_RETAINED",
      ifelse(
        merged$Earlier_retained &
          merged$Later_retained,
        "RETAINED_IN_BOTH",
        "NOT_RETAINED_IN_EITHER"
      )
    )
  )

  ranking_order <- order(
    -merged$Absolute_difference,
    merged$Edge_ID
  )
  merged$Absolute_difference_rank <- NA_integer_
  merged$Absolute_difference_rank[ranking_order] <-
    seq_along(ranking_order)
  merged$Inference_status <-
    "DESCRIPTIVE_NO_BETWEEN_WAVE_SIGNIFICANCE_TEST"

  merged[
    ,
    c(
      "Contrast_order",
      "Contrast",
      "Earlier_wave",
      "Later_wave",
      "Years_elapsed",
      "Contrast_role",
      "Edge_ID",
      "Node_1",
      "Node_2",
      "Earlier_edge",
      "Later_edge",
      "Signed_difference",
      "Absolute_difference",
      "Absolute_difference_rank",
      "Earlier_retained",
      "Later_retained",
      "Retention_change",
      "Sign_change",
      "Inference_status"
    ),
    drop = FALSE
  ]
}

edge_contrast_list <- lapply(
  seq_len(
    nrow(contrast_plan)
  ),
  function(row_index) {
    make_edge_contrast(
      earlier_wave =
        contrast_plan$Earlier_wave[[row_index]],
      later_wave =
        contrast_plan$Later_wave[[row_index]],
      contrast_label =
        contrast_plan$Contrast[[row_index]],
      contrast_order =
        contrast_plan$Contrast_order[[row_index]],
      years_elapsed =
        contrast_plan$Years_elapsed[[row_index]],
      contrast_role =
        contrast_plan$Contrast_role[[row_index]]
    )
  }
)

overall_edge_differences <- do.call(
  rbind,
  edge_contrast_list
)

overall_edge_differences <- overall_edge_differences[
  order(
    overall_edge_differences$Contrast_order,
    overall_edge_differences$Absolute_difference_rank
  ),
  ,
  drop = FALSE
]

overall_edge_differences_top5 <-
  overall_edge_differences[
    overall_edge_differences$Absolute_difference_rank <= 5L,
    ,
    drop = FALSE
  ]

write.csv(
  contrast_plan,
  file.path(
    formal_table_dir,
    "08_overall_contrast_plan.csv"
  ),
  row.names = FALSE
)

write.csv(
  overall_edge_differences,
  file.path(
    formal_table_dir,
    "08_overall_edge_differences.csv"
  ),
  row.names = FALSE
)

write.csv(
  overall_edge_differences_top5,
  file.path(
    formal_table_dir,
    "08_overall_edge_differences_top5.csv"
  ),
  row.names = FALSE
)

stopifnot(
  nrow(overall_edge_differences) == 180L,
  all(
    table(
      overall_edge_differences$Contrast
    ) == 36L
  ),
  nrow(overall_edge_differences_top5) == 25L,
  all(
    is.finite(
      overall_edge_differences$Signed_difference
    )
  )
)

overall_edge_differences_top5
```

## Stage 8: calculate descriptive global-strength contrasts

```{r global-strength-contrasts}
global_strength_lookup <- setNames(
  overall_wave_summary$Global_strength_descriptive,
  overall_wave_summary$Wave
)

overall_global_strength_contrasts <- contrast_plan

overall_global_strength_contrasts$Earlier_global_strength <-
  unname(
    global_strength_lookup[
      as.character(
        overall_global_strength_contrasts$Earlier_wave
      )
    ]
  )

overall_global_strength_contrasts$Later_global_strength <-
  unname(
    global_strength_lookup[
      as.character(
        overall_global_strength_contrasts$Later_wave
      )
    ]
  )

overall_global_strength_contrasts$Signed_difference <-
  with(
    overall_global_strength_contrasts,
    Later_global_strength -
      Earlier_global_strength
  )

overall_global_strength_contrasts$Absolute_difference <-
  abs(
    overall_global_strength_contrasts$Signed_difference
  )

overall_global_strength_contrasts$Inference_status <-
  "DESCRIPTIVE_NO_GLOBAL_STRENGTH_TEST"

overall_global_strength_contrasts$Construct_boundary <-
  "GLOBAL_STRENGTH_IS_NOT_EQUIVALENT_TO_IDEOLOGICAL_CONSTRAINT"

write.csv(
  overall_global_strength_contrasts,
  file.path(
    formal_table_dir,
    "08_overall_global_strength_contrasts.csv"
  ),
  row.names = FALSE
)

stopifnot(
  nrow(
    overall_global_strength_contrasts
  ) == 5L,
  all(
    is.finite(
      overall_global_strength_contrasts$Signed_difference
    )
  )
)

overall_global_strength_contrasts
```

## Stage 9: link within-wave edge-accuracy evidence

```{r edge-accuracy-link}
required_accuracy_columns <- c(
  "Wave",
  "Edge_ID",
  "Node_1",
  "Node_2",
  "Point_edge",
  "Retained_under_EBICglasso",
  "Bootstrap_mean",
  "Bootstrap_sd",
  "Percentile_2_5",
  "Percentile_97_5",
  "Bootstrap_zero_proportion",
  "Percentile_interval_width"
)

stopifnot(
  all(
    required_accuracy_columns %in%
      names(edge_accuracy)
  ),
  nrow(edge_accuracy) == 180L
)

overall_edges_with_accuracy <- edge_accuracy

overall_edges_with_accuracy$Interpretation_boundary <-
  paste0(
    "WITHIN_WAVE_ACCURACY_ONLY;",
    "NOT_A_BETWEEN_WAVE_DIFFERENCE_TEST"
  )

write.csv(
  overall_edges_with_accuracy,
  file.path(
    formal_table_dir,
    "08_overall_edges_with_accuracy.csv"
  ),
  row.names = FALSE
)

stopifnot(
  nrow(overall_edges_with_accuracy) == 180L,
  all(
    is.finite(
      overall_edges_with_accuracy$Point_edge
    )
  ),
  all(
    is.finite(
      overall_edges_with_accuracy$Bootstrap_mean
    )
  )
)
```

## Stage 10: prepare descriptive exploratory centrality outputs

```{r centrality-descriptive-tables}
expected_influence_cs <- centrality_cs[
  centrality_cs$Statistic ==
    "expectedInfluence",
  c(
    "Wave",
    "CS_coefficient",
    "Correlation_threshold",
    "Interpretation_status",
    "CS_error"
  ),
  drop = FALSE
]

names(expected_influence_cs)[2:5] <- c(
  "CS_coefficient",
  "CS_correlation_threshold",
  "CS_interpretation_status",
  "CS_error"
)

strength_cs <- centrality_cs[
  centrality_cs$Statistic ==
    "strength",
  c(
    "Wave",
    "CS_coefficient",
    "Correlation_threshold",
    "Interpretation_status",
    "CS_error"
  ),
  drop = FALSE
]

names(strength_cs)[2:5] <- c(
  "CS_coefficient",
  "CS_correlation_threshold",
  "CS_interpretation_status",
  "CS_error"
)

overall_expected_influence <- merge(
  centrality_scores[
    ,
    c(
      "Wave",
      "Node",
      "Expected_Influence",
      "Positive_retained_edges",
      "Negative_retained_edges",
      "Zero_edges"
    ),
    drop = FALSE
  ],
  expected_influence_cs,
  by = "Wave",
  all.x = TRUE,
  sort = FALSE
)

overall_strength <- merge(
  centrality_scores[
    ,
    c(
      "Wave",
      "Node",
      "Strength",
      "Positive_retained_edges",
      "Negative_retained_edges",
      "Zero_edges"
    ),
    drop = FALSE
  ],
  strength_cs,
  by = "Wave",
  all.x = TRUE,
  sort = FALSE
)

overall_expected_influence$Interpretation_boundary <-
  paste0(
    "EXPLORATORY_SIGNED_CONNECTIVITY;",
    "NO_CAUSAL_OR_CROSS_WAVE_SIGNIFICANCE_CLAIM"
  )

overall_strength$Interpretation_boundary <-
  paste0(
    "AUXILIARY_ABSOLUTE_CONNECTIVITY;",
    "NO_CAUSAL_OR_CROSS_WAVE_SIGNIFICANCE_CLAIM"
  )

overall_expected_influence <- overall_expected_influence[
  order(
    overall_expected_influence$Wave,
    match(
      overall_expected_influence$Node,
      network_vars
    )
  ),
  ,
  drop = FALSE
]

overall_strength <- overall_strength[
  order(
    overall_strength$Wave,
    match(
      overall_strength$Node,
      network_vars
    )
  ),
  ,
  drop = FALSE
]

write.csv(
  overall_expected_influence,
  file.path(
    formal_table_dir,
    "08_overall_expected_influence.csv"
  ),
  row.names = FALSE
)

write.csv(
  overall_strength,
  file.path(
    formal_table_dir,
    "08_overall_strength.csv"
  ),
  row.names = FALSE
)

centrality_stability <- centrality_cs
centrality_stability$Interpretation_boundary <- paste0(
  "WITHIN_WAVE_CASE_DROPPING_STABILITY;",
  "NO_CROSS_WAVE_CENTRALITY_TEST"
)

write.csv(
  centrality_stability,
  file.path(
    formal_table_dir,
    "08_overall_centrality_stability.csv"
  ),
  row.names = FALSE
)

stopifnot(
  nrow(overall_expected_influence) == 45L,
  nrow(overall_strength) == 45L,
  all(
    is.finite(
      overall_expected_influence$Expected_Influence
    )
  ),
  all(
    is.finite(
      overall_strength$Strength
    )
  ),
  all(
    is.finite(
      overall_expected_influence$CS_coefficient
    )
  ),
  all(
    is.finite(
      overall_strength$CS_coefficient
    )
  )
)
```

```{r centrality-descriptive-figures}
expected_influence_plot <- ggplot2::ggplot(
  overall_expected_influence,
  ggplot2::aes(
    x = Wave,
    y = Expected_Influence,
    group = Node
  )
) +
  ggplot2::geom_hline(
    yintercept = 0,
    colour = "grey75",
    linewidth = 0.35
  ) +
  ggplot2::geom_line(
    colour = "#2166AC",
    linewidth = 0.6
  ) +
  ggplot2::geom_point(
    colour = "#2166AC",
    size = 1.6
  ) +
  ggplot2::facet_wrap(
    ~ Node,
    ncol = 3,
    scales = "free_y"
  ) +
  ggplot2::scale_x_continuous(
    breaks = expected_years
  ) +
  ggplot2::labs(
    title = "Exploratory one-step Expected Influence",
    subtitle = "Descriptive trajectories; no cross-wave significance test",
    x = "Election year",
    y = "Expected Influence"
  ) +
  ggplot2::theme_bw(
    base_size = 10
  ) +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(
      angle = 45,
      hjust = 1
    )
  )

strength_plot <- ggplot2::ggplot(
  overall_strength,
  ggplot2::aes(
    x = Wave,
    y = Strength,
    group = Node
  )
) +
  ggplot2::geom_line(
    colour = "#4D4D4D",
    linewidth = 0.6
  ) +
  ggplot2::geom_point(
    colour = "#4D4D4D",
    size = 1.6
  ) +
  ggplot2::facet_wrap(
    ~ Node,
    ncol = 3,
    scales = "free_y"
  ) +
  ggplot2::scale_x_continuous(
    breaks = expected_years
  ) +
  ggplot2::labs(
    title = "Auxiliary Strength",
    subtitle = "Descriptive trajectories; no cross-wave significance test",
    x = "Election year",
    y = "Strength"
  ) +
  ggplot2::theme_bw(
    base_size = 10
  ) +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(
      angle = 45,
      hjust = 1
    )
  )

expected_influence_figure <- file.path(
  formal_figure_dir,
  "08_expected_influence_over_time.png"
)

strength_figure <- file.path(
  formal_figure_dir,
  "08_strength_over_time.png"
)

ggplot2::ggsave(
  filename = expected_influence_figure,
  plot = expected_influence_plot,
  width = 12,
  height = 9,
  units = "in",
  dpi = 180
)

ggplot2::ggsave(
  filename = strength_figure,
  plot = strength_plot,
  width = 12,
  height = 9,
  units = "in",
  dpi = 180
)

stopifnot(
  file.exists(expected_influence_figure),
  file.exists(strength_figure),
  file.info(expected_influence_figure)$size > 0,
  file.info(strength_figure)$size > 0
)
```

## Stage 11: validate the overall comparison module

```{r overall-comparison-validation}
source_lines <- readLines(
  notebook_file,
  warn = FALSE
)

forbidden_call_patterns <- c(
  paste0(
    "NetworkComparisonTest::",
    "NCT("
  ),
  paste0(
    "bootnet::",
    "estimateNetwork("
  ),
  paste0(
    "qgraph::",
    "EBICglasso("
  )
)

forbidden_call_scan <- data.frame(
  Pattern = forbidden_call_patterns,
  Found = vapply(
    forbidden_call_patterns,
    function(pattern) {
      any(
        grepl(
          pattern,
          source_lines,
          fixed = TRUE
        )
      )
    },
    FUN.VALUE = logical(1)
  ),
  stringsAsFactors = FALSE
)

write.csv(
  forbidden_call_scan,
  file.path(
    formal_table_dir,
    "08_forbidden_call_scan.csv"
  ),
  row.names = FALSE
)

comparison_year_values <- unique(
  c(
    overall_wave_summary$Wave,
    overall_edge_differences$Earlier_wave,
    overall_edge_differences$Later_wave,
    overall_global_strength_contrasts$Earlier_wave,
    overall_global_strength_contrasts$Later_wave,
    overall_expected_influence$Wave,
    overall_strength$Wave
  )
)

overall_output_files <- c(
  file.path(
    formal_table_dir,
    "08_overall_wave_summary.csv"
  ),
  file.path(
    formal_table_dir,
    "08_overall_edge_differences.csv"
  ),
  file.path(
    formal_table_dir,
    "08_overall_edge_differences_top5.csv"
  ),
  file.path(
    formal_table_dir,
    "08_overall_global_strength_contrasts.csv"
  ),
  file.path(
    formal_table_dir,
    "08_overall_edges_with_accuracy.csv"
  ),
  file.path(
    formal_table_dir,
    "08_overall_expected_influence.csv"
  ),
  file.path(
    formal_table_dir,
    "08_overall_strength.csv"
  ),
  file.path(
    formal_table_dir,
    "08_overall_centrality_stability.csv"
  ),
  all_figure_files,
  expected_influence_figure,
  strength_figure,
  layout_file
)

overall_validation <- data.frame(
  Check = c(
    "Decision gate passed",
    "Overall input preflight passed",
    "Common layout validation passed",
    "Five wave-summary rows were produced",
    "Five contrasts contain 36 edges each",
    "Twenty-five predeclared top-five rows were produced",
    "Five global-strength contrasts were produced",
    "Edge-accuracy table contains 180 rows",
    "Expected Influence table contains 45 rows",
    "Strength table contains 45 rows",
    "All overall years belong to the frozen five-wave set",
    "2008 is absent from every overall comparison table",
    "No prohibited model-estimation or NCT call was found",
    "All required overall files exist and are non-empty"
  ),
  Passed = c(
    all(decision_gate$Passed),
    all(overall_preflight$Passed),
    all(layout_validation$Passed),
    nrow(overall_wave_summary) == 5L,
    nrow(overall_edge_differences) == 180L &&
      all(
        table(
          overall_edge_differences$Contrast
        ) == 36L
      ),
    nrow(overall_edge_differences_top5) == 25L,
    nrow(overall_global_strength_contrasts) == 5L,
    nrow(overall_edges_with_accuracy) == 180L,
    nrow(overall_expected_influence) == 45L,
    nrow(overall_strength) == 45L,
    all(
      comparison_year_values %in%
        expected_years
    ),
    !(2008L %in% comparison_year_values),
    !any(forbidden_call_scan$Found),
    all(file.exists(overall_output_files)) &&
      all(file.info(overall_output_files)$size > 0)
  ),
  stringsAsFactors = FALSE
)

write.csv(
  overall_validation,
  file.path(
    formal_table_dir,
    "08_overall_validation.csv"
  ),
  row.names = FALSE
)

overall_validation

stopifnot(
  all(overall_validation$Passed)
)
```

## Stage 12: rebuild and validate the eight education-group samples

```{r education-source-gate}
anes_analysis_ready_stage12 <- readRDS(
  analysis_ready_file
)

education_years <- c(
  2012L,
  2016L,
  2020L,
  2024L
)

stage12_source_gate <- data.frame(
  Check = c(
    "Analysis-ready SHA-256 matches the authorised file",
    "Analysis-ready object is 16,766 by 18",
    "All required columns are present",
    "The nine nodes are complete",
    "Nonmissing VCF0110 values are only codes 1-4",
    "Target education years exclude 2004 and 2008"
  ),
  Passed = c(
    identical(
      file_sha256(analysis_ready_file),
      analysis_ready_hash_expected
    ),
    identical(
      dim(anes_analysis_ready_stage12),
      c(16766L, 18L)
    ),
    all(
      c(
        "VCF0004",
        "VCF0006",
        "VCF0006a",
        "VCF0110",
        network_vars
      ) %in%
        names(anes_analysis_ready_stage12)
    ),
    all(
      complete.cases(
        anes_analysis_ready_stage12[
          network_vars
        ]
      )
    ),
    all(
      na.omit(
        unique(
          anes_analysis_ready_stage12$VCF0110
        )
      ) %in%
        1:4
    ),
    !any(
      c(
        2004L,
        2008L
      ) %in%
        education_years
    )
  ),
  stringsAsFactors = FALSE
)

write.csv(
  stage12_source_gate,
  file.path(
    formal_table_dir,
    "08_education_source_gate.csv"
  ),
  row.names = FALSE
)

stopifnot(
  all(stage12_source_gate$Passed)
)
```

```{r build-education-group-samples}
education_target <- as.data.frame(
  anes_analysis_ready_stage12[
    anes_analysis_ready_stage12$VCF0004 %in%
      education_years,
    ,
    drop = FALSE
  ]
)

education_target$Education_group <- ifelse(
  education_target$VCF0110 %in%
    1:3,
  "No college degree",
  ifelse(
    education_target$VCF0110 == 4,
    "College or advanced degree",
    NA_character_
  )
)

expected_education_groups <- data.frame(
  Wave = rep(
    education_years,
    each = 2L
  ),
  Education_group = rep(
    c(
      "No college degree",
      "College or advanced degree"
    ),
    times = length(education_years)
  ),
  Group_key = rep(
    c(
      "no_college_degree",
      "college_or_advanced"
    ),
    times = length(education_years)
  ),
  Expected_N = c(
    2773L,
    1531L,
    1478L,
    1216L,
    2624L,
    2691L,
    1729L,
    1781L
  ),
  stringsAsFactors = FALSE
)

education_samples <- vector(
  mode = "list",
  length = nrow(expected_education_groups)
)

names(education_samples) <- paste0(
  expected_education_groups$Wave,
  "__",
  expected_education_groups$Group_key
)

for (
  row_index in seq_len(
    nrow(expected_education_groups)
  )
) {
  wave <- expected_education_groups$Wave[[row_index]]
  group_label <-
    expected_education_groups$Education_group[[row_index]]

  selected_rows <-
    education_target$VCF0004 == wave &
    education_target$Education_group ==
      group_label &
    !is.na(
      education_target$Education_group
    )

  full_data <- education_target[
    selected_rows,
    ,
    drop = FALSE
  ]

  ordinal_nodes <- as.data.frame(
    full_data[
      network_vars
    ]
  )

  for (node_name in network_vars) {
    ordinal_nodes[[node_name]] <- ordered(
      as.numeric(
        ordinal_nodes[[node_name]]
      ),
      levels =
        expected_categories[[node_name]]
    )
  }

  education_samples[[row_index]] <- list(
    Wave = wave,
    Education_group = group_label,
    Education_codes = if (
      group_label ==
        "No college degree"
    ) {
      1:3
    } else {
      4
    },
    N = nrow(full_data),
    Full_data = full_data,
    Ordinal_nodes = ordinal_nodes
  )
}

education_group_bundle <- list(
  Source_file = normalizePath(
    analysis_ready_file,
    winslash = "/"
  ),
  Source_SHA256 = file_sha256(
    analysis_ready_file
  ),
  Education_years = education_years,
  Node_order = network_vars,
  Expected_categories =
    expected_categories,
  Group_definition = c(
    "No college degree" =
      "VCF0110 codes 1-3",
    "College or advanced degree" =
      "VCF0110 code 4"
  ),
  Samples = education_samples
)
```

```{r education-sample-flow-and-validation}
education_sample_flow <- do.call(
  rbind,
  lapply(
    education_years,
    function(wave) {
      wave_rows <-
        education_target$VCF0004 == wave

      data.frame(
        Wave = wave,
        Overall_nine_node_complete_N =
          sum(wave_rows),
        Education_valid_N = sum(
          wave_rows &
            !is.na(
              education_target$Education_group
            )
        ),
        Education_missing_N = sum(
          wave_rows &
            is.na(
              education_target$Education_group
            )
        ),
        No_college_degree_N = sum(
          wave_rows &
            education_target$Education_group ==
              "No college degree",
          na.rm = TRUE
        ),
        College_or_advanced_N = sum(
          wave_rows &
            education_target$Education_group ==
              "College or advanced degree",
          na.rm = TRUE
        ),
        stringsAsFactors = FALSE
      )
    }
  )
)

expected_sample_flow <- data.frame(
  Wave = education_years,
  Overall_nine_node_complete_N = c(
    4353L,
    2709L,
    5390L,
    3525L
  ),
  Education_valid_N = c(
    4304L,
    2694L,
    5315L,
    3510L
  ),
  Education_missing_N = c(
    49L,
    15L,
    75L,
    15L
  ),
  No_college_degree_N = c(
    2773L,
    1478L,
    2624L,
    1729L
  ),
  College_or_advanced_N = c(
    1531L,
    1216L,
    2691L,
    1781L
  )
)

education_group_validation <- do.call(
  rbind,
  lapply(
    seq_along(
      education_samples
    ),
    function(sample_index) {
      one_sample <-
        education_samples[[sample_index]]
      expected_row <-
        expected_education_groups[
          sample_index,
          ,
          drop = FALSE
        ]
      full_data <- one_sample$Full_data
      ordinal_nodes <-
        one_sample$Ordinal_nodes

      expected_codes <- if (
        one_sample$Education_group ==
          "No college degree"
      ) {
        1:3
      } else {
        4
      }

      level_checks <- vapply(
        network_vars,
        function(node_name) {
          identical(
            as.numeric(
              levels(
                ordinal_nodes[[node_name]]
              )
            ),
            as.numeric(
              expected_categories[[node_name]]
            )
          )
        },
        FUN.VALUE = logical(1)
      )

      respondent_key <- paste(
        full_data$VCF0004,
        full_data$VCF0006,
        sep = "::"
      )

      data.frame(
        Wave = one_sample$Wave,
        Education_group =
          one_sample$Education_group,
        Expected_N =
          expected_row$Expected_N,
        Observed_N = nrow(full_data),
        Correct_N =
          nrow(full_data) ==
            expected_row$Expected_N,
        Nine_nodes_present = identical(
          names(ordinal_nodes),
          network_vars
        ),
        Nine_nodes_complete = all(
          complete.cases(
            ordinal_nodes
          )
        ),
        All_nodes_ordered = all(
          vapply(
            ordinal_nodes,
            is.ordered,
            FUN.VALUE = logical(1)
          )
        ),
        Frozen_levels_retained =
          all(level_checks),
        Education_rule_valid = all(
          full_data$VCF0110 %in%
            expected_codes
        ),
        Within_group_ID_unique =
          !anyDuplicated(respondent_key),
        Passed =
          nrow(full_data) ==
            expected_row$Expected_N &&
            identical(
              names(ordinal_nodes),
              network_vars
            ) &&
            all(
              complete.cases(
                ordinal_nodes
              )
            ) &&
            all(
              vapply(
                ordinal_nodes,
                is.ordered,
                FUN.VALUE = logical(1)
              )
            ) &&
            all(level_checks) &&
            all(
              full_data$VCF0110 %in%
                expected_codes
            ) &&
            !anyDuplicated(respondent_key),
        stringsAsFactors = FALSE
      )
    }
  )
)

education_wave_partition_validation <- data.frame(
  Wave = education_years,
  Expected_flow_matches = vapply(
    seq_along(education_years),
    function(row_index) {
      all(
        as.numeric(
          education_sample_flow[
            row_index,
            names(expected_sample_flow),
            drop = FALSE
          ]
        ) ==
          as.numeric(
            expected_sample_flow[
              row_index,
              ,
              drop = FALSE
            ]
          )
      )
    },
    FUN.VALUE = logical(1)
  ),
  Groups_sum_to_education_valid = with(
    education_sample_flow,
    No_college_degree_N +
      College_or_advanced_N ==
      Education_valid_N
  ),
  Overall_partitions_correctly = with(
    education_sample_flow,
    Education_valid_N +
      Education_missing_N ==
      Overall_nine_node_complete_N
  ),
  No_2004_or_2008 =
    !(education_years %in%
        c(
          2004L,
          2008L
        )),
  stringsAsFactors = FALSE
)

education_wave_partition_validation$Passed <-
  with(
    education_wave_partition_validation,
    Expected_flow_matches &
      Groups_sum_to_education_valid &
      Overall_partitions_correctly &
      No_2004_or_2008
  )

write.csv(
  education_sample_flow,
  file.path(
    formal_table_dir,
    "08_education_sample_flow.csv"
  ),
  row.names = FALSE
)

write.csv(
  education_group_validation,
  file.path(
    formal_table_dir,
    "08_education_group_validation.csv"
  ),
  row.names = FALSE
)

write.csv(
  education_wave_partition_validation,
  file.path(
    formal_table_dir,
    "08_education_wave_partition_validation.csv"
  ),
  row.names = FALSE
)

stopifnot(
  all(education_group_validation$Passed),
  all(
    education_wave_partition_validation$Passed
  ),
  sum(
    education_sample_flow$Education_valid_N
  ) == 15823L
)
```

```{r save-and-reload-education-samples}
education_bundle_file <- file.path(
  intermediate_dir,
  "08_education_group_samples_2012_2016_2020_2024.rds"
)

saveRDS(
  education_group_bundle,
  education_bundle_file,
  version = 3,
  compress = "gzip"
)

education_bundle_reloaded <- readRDS(
  education_bundle_file
)

education_bundle_validation <- data.frame(
  Check = c(
    "Education bundle exists",
    "Education bundle is non-empty",
    "Reloaded bundle is identical",
    "Bundle contains eight samples",
    "All eight sample names are unique",
    "No education network object is stored"
  ),
  Passed = c(
    file.exists(education_bundle_file),
    file.info(education_bundle_file)$size > 0,
    identical(
      education_group_bundle,
      education_bundle_reloaded
    ),
    length(
      education_bundle_reloaded$Samples
    ) == 8L,
    anyDuplicated(
      names(
        education_bundle_reloaded$Samples
      )
    ) == 0L,
    all(
      !vapply(
        education_bundle_reloaded$Samples,
        function(one_sample) {
          any(
            c(
              "graph",
              "network",
              "bootstrap"
            ) %in%
              names(one_sample)
          )
        },
        FUN.VALUE = logical(1)
      )
    )
  ),
  stringsAsFactors = FALSE
)

education_input_inventory <- data.frame(
  Role = c(
    "Authorised analysis-ready input",
    "Derived eight-group Stage-12 bundle"
  ),
  Path = normalizePath(
    c(
      analysis_ready_file,
      education_bundle_file
    ),
    winslash = "/"
  ),
  SHA256 = c(
    file_sha256(analysis_ready_file),
    file_sha256(education_bundle_file)
  ),
  Bytes = file.info(
    c(
      analysis_ready_file,
      education_bundle_file
    )
  )$size,
  stringsAsFactors = FALSE
)

write.csv(
  education_bundle_validation,
  file.path(
    formal_table_dir,
    "08_education_bundle_validation.csv"
  ),
  row.names = FALSE
)

write.csv(
  education_input_inventory,
  file.path(
    formal_table_dir,
    "08_education_input_inventory.csv"
  ),
  row.names = FALSE
)

stopifnot(
  all(
    education_bundle_validation$Passed
  )
)

education_input_inventory
```

## Stage 12 completion record and reproducibility log

```{r stage1-12-completion-record}
stage1_12_completion <- data.frame(
  Component = c(
    "Decision and package gates",
    "Validated Step-07 inputs loaded",
    "Common layout created",
    "Five common-scale overall figures",
    "Five-wave overall summary",
    "Five descriptive edge contrasts",
    "Five descriptive global-strength contrasts",
    "Within-wave edge accuracy linked",
    "Exploratory Expected Influence and Strength prepared",
    "Overall comparison validation",
    "Eight education-group samples created",
    "Education sample partition and bundle validation",
    "Overall network re-estimation",
    "Education network estimation",
    "Education bootstrap",
    "NCT",
    "Political interpretation"
  ),
  Status = c(
    "COMPLETE_AND_VALIDATED",
    "COMPLETE_AND_VALIDATED",
    "COMPLETE_AND_VALIDATED",
    "COMPLETE_AND_VALIDATED",
    "COMPLETE_AND_VALIDATED",
    "COMPLETE_AND_VALIDATED",
    "COMPLETE_AND_VALIDATED",
    "COMPLETE_AND_VALIDATED",
    "COMPLETE_AND_VALIDATED",
    "COMPLETE_AND_VALIDATED",
    "COMPLETE_AND_VALIDATED",
    "COMPLETE_AND_VALIDATED",
    "NOT_RUN_REUSED_VALIDATED_STEP07_OBJECTS",
    "NOT_RUN_STAGE13_SCOPE",
    "NOT_RUN_STAGE15_SCOPE",
    "NOT_IN_CURRENT_SCOPE_NOT_RUN",
    "NOT_AUTHORISED_NOT_RUN"
  ),
  Required_for_stage1_12 = c(
    rep(
      TRUE,
      12
    ),
    rep(
      FALSE,
      5
    )
  ),
  Passed = c(
    all(decision_gate$Passed) &&
      all(package_gate$Installed),
    all(overall_preflight$Passed),
    all(layout_validation$Passed),
    all(figure_manifest$Exists),
    nrow(overall_wave_summary) == 5L,
    nrow(overall_edge_differences) == 180L,
    nrow(
      overall_global_strength_contrasts
    ) == 5L,
    nrow(overall_edges_with_accuracy) == 180L,
    nrow(overall_expected_influence) == 45L &&
      nrow(overall_strength) == 45L,
    all(overall_validation$Passed),
    length(education_samples) == 8L,
    all(education_group_validation$Passed) &&
      all(
        education_wave_partition_validation$Passed
      ) &&
      all(
        education_bundle_validation$Passed
      ),
    !any(
      forbidden_call_scan$Pattern %in%
        forbidden_call_patterns[2:3] &
        forbidden_call_scan$Found
    ),
    TRUE,
    TRUE,
    !any(
      forbidden_call_scan$Pattern ==
        forbidden_call_patterns[[1L]] &
        forbidden_call_scan$Found
    ),
    TRUE
  ),
  stringsAsFactors = FALSE
)

stage1_12_complete <- all(
  stage1_12_completion$Passed[
    stage1_12_completion$Required_for_stage1_12
  ]
)

write.csv(
  stage1_12_completion,
  file.path(
    formal_table_dir,
    "08_stage1_12_completion_status.csv"
  ),
  row.names = FALSE
)

stopifnot(
  isTRUE(stage1_12_complete)
)

stage1_12_completion
```

```{r output-inventory-log-and-session}
stage1_12_output_paths <- c(
  list.files(
    formal_table_dir,
    pattern = "^08_",
    full.names = TRUE
  ),
  list.files(
    formal_figure_dir,
    pattern = "^08_",
    full.names = TRUE
  ),
  layout_file,
  education_bundle_file
)

stage1_12_output_paths <- sort(
  unique(
    stage1_12_output_paths[
      file.exists(stage1_12_output_paths)
    ]
  )
)

output_inventory <- data.frame(
  File = basename(
    stage1_12_output_paths
  ),
  Path = normalizePath(
    stage1_12_output_paths,
    winslash = "/"
  ),
  Bytes = file.info(
    stage1_12_output_paths
  )$size,
  SHA256 = vapply(
    stage1_12_output_paths,
    file_sha256,
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

output_inventory_file <- file.path(
  formal_table_dir,
  "08_stage1_12_output_inventory.csv"
)

write.csv(
  output_inventory,
  output_inventory_file,
  row.names = FALSE
)

session_file <- file.path(
  log_dir,
  "08_stage1_12_sessionInfo.txt"
)

writeLines(
  capture.output(
    sessionInfo()
  ),
  con = session_file
)

log_file <- file.path(
  log_dir,
  "08_stage1_12_execution.log"
)

writeLines(
  c(
    "STEP 08 STAGES 1-12 EXECUTION LOG",
    paste(
      "Project root:",
      normalizePath(
        project_root,
        winslash = "/"
      )
    ),
    paste(
      "Decision gate passed:",
      all(decision_gate$Passed)
    ),
    paste(
      "Overall validation passed:",
      all(overall_validation$Passed)
    ),
    paste(
      "Education groups created:",
      length(education_samples)
    ),
    paste(
      "Education group N total:",
      sum(
        education_sample_flow$Education_valid_N
      )
    ),
    paste(
      "Stage 1-12 complete:",
      stage1_12_complete
    ),
    "Overall network re-estimation: NOT RUN",
    "Education network estimation: NOT RUN",
    "Education bootstrap: NOT RUN",
    "NCT: NOT RUN",
    "Political interpretation: NOT RUN",
    paste(
      "Existing dense-network warnings carried forward:",
      sum(
        overall_wave_summary$Captured_warning_n
      )
    ),
    paste(
      "Layout warnings captured:",
      length(layout_warnings)
    ),
    paste(
      "Output files inventoried:",
      nrow(output_inventory)
    )
  ),
  con = log_file
)

stopifnot(
  file.exists(output_inventory_file),
  file.exists(session_file),
  file.exists(log_file),
  file.info(session_file)$size > 0,
  file.info(log_file)$size > 0
)
```

## Execution result

Stages 1-12 are complete when the saved completion table records all
required components as passed. The education-group bundle is an input
for later diagnostics only. It contains no education network or
bootstrap result.

No ANES raw file was read or modified. No protected source file,
literature PDF, Step-07 model object or Step-07 bootstrap object was
modified.
