---
title: "08A Education-Group Polychoric Input Diagnostics"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

# Execution boundary

This notebook implements authorised decision D25, Stage 13 only. It
audits the eight education-group inputs and estimates their unweighted
polychoric correlation matrices. It does not estimate a GGM, apply
EBICglasso, run a bootstrap, calculate centrality, run NCT, execute a
sensitivity analysis, compare education groups, or interpret political
results. Sparse cells are recorded but are not used with an unvalidated
automatic exclusion threshold. Matrices are never smoothed or repaired.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  error = FALSE,
  warning = TRUE,
  message = TRUE
)
```

# Stage 13.1: project and package gate

```{r project-root-and-packages}
current_dir <- normalizePath(
  ".",
  winslash = "/"
)

if (
  file.exists(
    file.path(
      current_dir,
      "ANES.Rproj"
    )
  )
) {
  project_root <- current_dir
} else if (
  basename(current_dir) == "notebooks" &&
    file.exists(
      file.path(
        dirname(current_dir),
        "ANES.Rproj"
      )
    )
) {
  project_root <- dirname(current_dir)
} else {
  stop(
    "Project root could not be identified."
  )
}

knitr::opts_knit$set(
  root.dir = project_root
)

notebook_file <- file.path(
  project_root,
  "notebooks",
  "08A_education_polychoric_input_diagnostics.Rmd"
)

formal_table_dir <- file.path(
  project_root,
  "outputs",
  "tables",
  "formal"
)

formal_model_dir <- file.path(
  project_root,
  "outputs",
  "models",
  "formal"
)

log_dir <- file.path(
  project_root,
  "outputs",
  "logs"
)

intermediate_dir <- file.path(
  project_root,
  "data",
  "intermediate"
)

for (
  target_dir in c(
    formal_table_dir,
    formal_model_dir,
    log_dir,
    intermediate_dir
  )
) {
  dir.create(
    target_dir,
    recursive = TRUE,
    showWarnings = FALSE
  )
}

required_packages <- c(
  "digest",
  "knitr",
  "lavaan",
  "rmarkdown"
)

package_gate <- data.frame(
  Package = required_packages,
  Installed = vapply(
    required_packages,
    requireNamespace,
    quietly = TRUE,
    FUN.VALUE = logical(1)
  ),
  Version = vapply(
    required_packages,
    function(package_name) {
      if (
        requireNamespace(
          package_name,
          quietly = TRUE
        )
      ) {
        as.character(
          utils::packageVersion(
            package_name
          )
        )
      } else {
        NA_character_
      }
    },
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

stopifnot(
  all(package_gate$Installed)
)

normalise_status <- function(value) {
  value <- toupper(
    trimws(
      as.character(value)
    )
  )
  gsub(
    "[^A-Z0-9]+",
    "_",
    value
  )
}

as_pass <- function(value) {
  toupper(
    trimws(
      as.character(value)
    )
  ) == "TRUE"
}

file_sha256 <- function(path) {
  if (!file.exists(path)) {
    return(NA_character_)
  }

  toupper(
    digest::digest(
      file = path,
      algo = "sha256",
      serialize = FALSE
    )
  )
}

ordered_values <- function(x) {
  as.numeric(
    as.character(x)
  )
}

project_root
package_gate
```

# Stage 13.2: canonical decision gate

```{r decision-gate}
registry_config <- file.path(
  project_root,
  "config",
  "decision_registry_path.txt"
)

stopifnot(
  file.exists(registry_config)
)

registry_lines <- trimws(
  readLines(
    registry_config,
    warn = FALSE
  )
)
registry_lines <- registry_lines[nzchar(registry_lines)]

stopifnot(
  length(registry_lines) == 1L,
  file.exists(registry_lines[[1L]])
)

registry_path <- registry_lines[[1L]]
decision_registry <- read.csv(
  registry_path,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

decision_row <- function(decision_id) {
  decision_registry[
    decision_registry$decision_id == decision_id,
    ,
    drop = FALSE
  ]
}

g01 <- decision_row("G01")
d04 <- decision_row("D04")
d06 <- decision_row("D06")
d07 <- decision_row("D07")
d08a <- decision_row("D08A")
d08b <- decision_row("D08B")
d09a <- decision_row("D09A")
d11 <- decision_row("D11")
d13 <- decision_row("D13")
d15 <- decision_row("D15")
d18 <- decision_row("D18")
d24 <- decision_row("D24")
d25 <- decision_row("D25")

decision_gate <- data.frame(
  Decision_ID = c(
    "G01",
    "D04",
    "D06",
    "D07",
    "D08A",
    "D08B",
    "D09A",
    "D11",
    "D13",
    "D15",
    "D18",
    "D24",
    "D25"
  ),
  Check = c(
    "Canonical registry is active for execution",
    "2008 is excluded",
    "Nine-node set is frozen",
    "Education grouping is frozen",
    "Descriptive education module is approved",
    "Formal education inference is outside scope",
    "Node coding direction is verified",
    "Complete-case primary rule is verified",
    "Ordinal polychoric method family is frozen",
    "Primary education inputs are unweighted",
    "NCT is outside current scope",
    "Stages 1-12 are complete and validated",
    "Stage 13 is authorised or already complete and may be reproduced"
  ),
  Passed = c(
    nrow(g01) == 1L &&
      normalise_status(g01$implementation_status) ==
        "ACTIVE_FOR_EXECUTION_ONLY" &&
      normalise_status(g01$researcher_approval_status) ==
        "APPROVED",
    nrow(d04) == 1L &&
      normalise_status(d04$decision_status) == "FROZEN" &&
      normalise_status(d04$implementation_status) == "VERIFIED" &&
      grepl("Exclude 2008", d04$current_value, fixed = TRUE),
    nrow(d06) == 1L &&
      normalise_status(d06$decision_status) == "FROZEN" &&
      normalise_status(d06$implementation_status) == "VERIFIED",
    nrow(d07) == 1L &&
      normalise_status(d07$decision_status) == "FROZEN" &&
      normalise_status(d07$implementation_status) == "VERIFIED",
    nrow(d08a) == 1L &&
      normalise_status(d08a$decision_status) == "FROZEN" &&
      normalise_status(d08a$researcher_approval_status) == "APPROVED",
    nrow(d08b) == 1L &&
      normalise_status(d08b$decision_status) == "NOT_IN_CURRENT_SCOPE" &&
      normalise_status(d08b$implementation_status) == "NOT_APPLICABLE",
    nrow(d09a) == 1L &&
      normalise_status(d09a$decision_status) == "FROZEN" &&
      normalise_status(d09a$implementation_status) == "VERIFIED",
    nrow(d11) == 1L &&
      normalise_status(d11$decision_status) == "FROZEN" &&
      normalise_status(d11$implementation_status) %in%
        c("VERIFIED", "COMPLETE_WITH_WARNINGS") &&
      grepl("Complete-case primary", d11$current_value, fixed = TRUE),
    nrow(d13) == 1L &&
      normalise_status(d13$decision_status) == "FROZEN" &&
      normalise_status(d13$researcher_approval_status) == "APPROVED" &&
      grepl("Polychoric", d13$current_value, fixed = TRUE),
    nrow(d15) == 1L &&
      normalise_status(d15$decision_status) == "FROZEN" &&
      normalise_status(d15$researcher_approval_status) == "APPROVED" &&
      grepl("Unweighted primary", d15$current_value, fixed = TRUE),
    nrow(d18) == 1L &&
      normalise_status(d18$decision_status) == "NOT_IN_CURRENT_SCOPE" &&
      normalise_status(d18$implementation_status) == "NOT_APPLICABLE",
    nrow(d24) == 1L &&
      normalise_status(d24$implementation_status) ==
        "STAGES_1_TO_12_COMPLETE_AND_VALIDATED",
    nrow(d25) == 1L &&
      normalise_status(d25$decision_status) == "FROZEN" &&
      normalise_status(d25$implementation_status) %in%
        c(
          "AUTHORISED_STAGE_13_ONLY",
          "STAGE_13_COMPLETE_AND_VALIDATED"
        ) &&
      normalise_status(d25$researcher_approval_status) == "APPROVED"
  ),
  stringsAsFactors = FALSE
)

stopifnot(
  all(decision_gate$Passed)
)

write.csv(
  package_gate,
  file.path(
    formal_table_dir,
    "08A_package_gate.csv"
  ),
  row.names = FALSE
)

write.csv(
  decision_gate,
  file.path(
    formal_table_dir,
    "08A_decision_gate.csv"
  ),
  row.names = FALSE
)

decision_gate
```

# Stage 13.3: authorised input gate

```{r input-inventory-and-load}
education_bundle_file <- file.path(
  intermediate_dir,
  "08_education_group_samples_2012_2016_2020_2024.rds"
)

education_group_validation_file <- file.path(
  formal_table_dir,
  "08_education_group_validation.csv"
)

education_partition_validation_file <- file.path(
  formal_table_dir,
  "08_education_wave_partition_validation.csv"
)

education_bundle_validation_file <- file.path(
  formal_table_dir,
  "08_education_bundle_validation.csv"
)

stage1_12_completion_file <- file.path(
  formal_table_dir,
  "08_stage1_12_completion_status.csv"
)

input_paths <- c(
  Decision_registry = registry_path,
  Education_bundle = education_bundle_file,
  Education_group_validation = education_group_validation_file,
  Education_partition_validation = education_partition_validation_file,
  Education_bundle_validation = education_bundle_validation_file,
  Stage1_12_completion = stage1_12_completion_file
)

input_inventory <- data.frame(
  Input = names(input_paths),
  Path = normalizePath(
    input_paths,
    winslash = "/",
    mustWork = FALSE
  ),
  Exists = file.exists(input_paths),
  Bytes = ifelse(
    file.exists(input_paths),
    file.info(input_paths)$size,
    NA_real_
  ),
  SHA256 = vapply(
    input_paths,
    file_sha256,
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

stopifnot(
  all(input_inventory$Exists),
  all(input_inventory$Bytes > 0)
)

education_bundle <- readRDS(
  education_bundle_file
)

education_group_validation <- read.csv(
  education_group_validation_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

education_partition_validation <- read.csv(
  education_partition_validation_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

education_bundle_validation <- read.csv(
  education_bundle_validation_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

stage1_12_completion <- read.csv(
  stage1_12_completion_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

write.csv(
  input_inventory,
  file.path(
    formal_table_dir,
    "08A_input_inventory.csv"
  ),
  row.names = FALSE
)

input_inventory
```

```{r validate-stage12-bundle}
expected_bundle_sha256 <-
  "48D8320E2F26E609A30FFBBA54601B6625896BEBF2565F24345396E0FE50E4AF"

expected_source_sha256 <-
  "6726A0D548F8293E890AD380C4915097FDD4ABC1A51665D55E904A33F81432FA"

network_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223"
)

expected_categories <- list(
  VCF0806 = 1:7,
  VCF0809 = 1:7,
  VCF0838 = 1:4,
  VCF0839 = 1:7,
  VCF0879a = c(1, 3, 5),
  VCF0888 = 1:3,
  VCF0890 = 1:3,
  VCF0894 = 1:3,
  VCF9223 = 1:4
)

expected_groups <- data.frame(
  Group_key = c(
    "2012__no_college_degree",
    "2012__college_or_advanced",
    "2016__no_college_degree",
    "2016__college_or_advanced",
    "2020__no_college_degree",
    "2020__college_or_advanced",
    "2024__no_college_degree",
    "2024__college_or_advanced"
  ),
  Wave = c(
    2012L,
    2012L,
    2016L,
    2016L,
    2020L,
    2020L,
    2024L,
    2024L
  ),
  Education_group = rep(
    c(
      "No college degree",
      "College or advanced degree"
    ),
    times = 4L
  ),
  Expected_N = c(
    2773L,
    1531L,
    1478L,
    1216L,
    2624L,
    2691L,
    1729L,
    1781L
  ),
  stringsAsFactors = FALSE
)

samples <- education_bundle$Samples

overall_input_validation <- data.frame(
  Check = c(
    "Stage-12 bundle SHA-256 matches",
    "Embedded analysis-ready source SHA-256 matches",
    "Stage-12 validation tables all passed",
    "Stage-12 required components all passed",
    "Four education waves are exact",
    "Eight sample keys are exact",
    "Node order is exact",
    "Expected category specification is exact",
    "No 2004 or 2008 sample is present"
  ),
  Passed = c(
    identical(
      file_sha256(education_bundle_file),
      expected_bundle_sha256
    ),
    identical(
      education_bundle$Source_SHA256,
      expected_source_sha256
    ),
    all(as_pass(education_group_validation$Passed)) &&
      all(as_pass(education_partition_validation$Passed)) &&
      all(as_pass(education_bundle_validation$Passed)),
    all(
      as_pass(
        stage1_12_completion$Passed[
          as_pass(
            stage1_12_completion$Required_for_stage1_12
          )
        ]
      )
    ),
    identical(
      as.integer(education_bundle$Education_years),
      c(2012L, 2016L, 2020L, 2024L)
    ),
    identical(
      names(samples),
      expected_groups$Group_key
    ),
    identical(
      education_bundle$Node_order,
      network_vars
    ),
    isTRUE(
      all.equal(
        education_bundle$Expected_categories,
        expected_categories,
        check.attributes = FALSE
      )
    ),
    !any(
      vapply(
        samples,
        function(one_sample) {
          one_sample$Wave %in% c(2004L, 2008L)
        },
        FUN.VALUE = logical(1)
      )
    )
  ),
  stringsAsFactors = FALSE
)

sample_input_validation <- do.call(
  rbind,
  lapply(
    seq_len(nrow(expected_groups)),
    function(row_index) {
      group_key <- expected_groups$Group_key[[row_index]]
      expected_row <- expected_groups[row_index, , drop = FALSE]
      one_sample <- samples[[group_key]]
      ordinal_nodes <- one_sample$Ordinal_nodes
      full_data <- one_sample$Full_data

      ordered_and_levels <- vapply(
        network_vars,
        function(node_name) {
          is.ordered(ordinal_nodes[[node_name]]) &&
            identical(
              as.numeric(
                levels(
                  ordinal_nodes[[node_name]]
                )
              ),
              as.numeric(
                expected_categories[[node_name]]
              )
            )
        },
        FUN.VALUE = logical(1)
      )

      values_match <- vapply(
        network_vars,
        function(node_name) {
          identical(
            ordered_values(
              ordinal_nodes[[node_name]]
            ),
            as.numeric(
              full_data[[node_name]]
            )
          )
        },
        FUN.VALUE = logical(1)
      )

      respondent_key <- paste(
        full_data$VCF0004,
        full_data$VCF0006,
        sep = "::"
      )

      passed <-
        one_sample$Wave == expected_row$Wave &&
        one_sample$Education_group ==
          expected_row$Education_group &&
        one_sample$N == expected_row$Expected_N &&
        nrow(full_data) == expected_row$Expected_N &&
        nrow(ordinal_nodes) == expected_row$Expected_N &&
        identical(names(ordinal_nodes), network_vars) &&
        all(complete.cases(ordinal_nodes)) &&
        all(ordered_and_levels) &&
        all(values_match) &&
        !anyDuplicated(respondent_key)

      data.frame(
        Group_key = group_key,
        Wave = one_sample$Wave,
        Education_group = one_sample$Education_group,
        Expected_N = expected_row$Expected_N,
        Observed_N = nrow(ordinal_nodes),
        Correct_node_order = identical(
          names(ordinal_nodes),
          network_vars
        ),
        Nine_nodes_complete = all(
          complete.cases(ordinal_nodes)
        ),
        Ordered_levels_exact = all(ordered_and_levels),
        Ordinal_values_match_full_data = all(values_match),
        Respondent_ID_unique = !anyDuplicated(respondent_key),
        Passed = passed,
        stringsAsFactors = FALSE
      )
    }
  )
)

same_wave_group_overlap <- do.call(
  rbind,
  lapply(
    c(2012L, 2016L, 2020L, 2024L),
    function(wave) {
      no_college_key <- paste0(
        wave,
        "__no_college_degree"
      )
      college_key <- paste0(
        wave,
        "__college_or_advanced"
      )

      respondent_key <- function(one_sample) {
        paste(
          one_sample$Full_data$VCF0004,
          one_sample$Full_data$VCF0006,
          sep = "::"
        )
      }

      overlap_n <- length(
        intersect(
          respondent_key(samples[[no_college_key]]),
          respondent_key(samples[[college_key]])
        )
      )

      data.frame(
        Wave = wave,
        Overlap_N = overlap_n,
        Passed = overlap_n == 0L,
        stringsAsFactors = FALSE
      )
    }
  )
)

write.csv(
  overall_input_validation,
  file.path(
    formal_table_dir,
    "08A_overall_input_validation.csv"
  ),
  row.names = FALSE
)

write.csv(
  sample_input_validation,
  file.path(
    formal_table_dir,
    "08A_sample_input_validation.csv"
  ),
  row.names = FALSE
)

write.csv(
  same_wave_group_overlap,
  file.path(
    formal_table_dir,
    "08A_same_wave_group_overlap.csv"
  ),
  row.names = FALSE
)

stopifnot(
  all(overall_input_validation$Passed),
  all(sample_input_validation$Passed),
  all(same_wave_group_overlap$Passed)
)

overall_input_validation
sample_input_validation
same_wave_group_overlap
```

# Stage 13.4: marginal-category diagnostics

```{r category-frequency-diagnostics}
category_frequency <- do.call(
  rbind,
  lapply(
    names(samples),
    function(group_key) {
      one_sample <- samples[[group_key]]

      do.call(
        rbind,
        lapply(
          network_vars,
          function(node_name) {
            values <- ordered_values(
              one_sample$Ordinal_nodes[[node_name]]
            )
            counts <- table(
              factor(
                values,
                levels = expected_categories[[node_name]]
              )
            )

            data.frame(
              Group_key = group_key,
              Wave = one_sample$Wave,
              Education_group = one_sample$Education_group,
              Node = node_name,
              Category = as.numeric(names(counts)),
              Count = as.integer(counts),
              Percent = 100 * as.integer(counts) /
                one_sample$N,
              Expected_category = TRUE,
              Observed_positive = as.integer(counts) > 0L,
              stringsAsFactors = FALSE
            )
          }
        )
      )
    }
  )
)

node_category_diagnostics <- do.call(
  rbind,
  lapply(
    split(
      category_frequency,
      list(
        category_frequency$Group_key,
        category_frequency$Node
      ),
      drop = TRUE
    ),
    function(one_node) {
      data.frame(
        Group_key = one_node$Group_key[[1L]],
        Wave = one_node$Wave[[1L]],
        Education_group = one_node$Education_group[[1L]],
        Node = one_node$Node[[1L]],
        Sample_N = sum(one_node$Count),
        Expected_category_n = nrow(one_node),
        Observed_category_n = sum(one_node$Count > 0L),
        Zero_category_n = sum(one_node$Count == 0L),
        Minimum_category_count = min(one_node$Count),
        Minimum_category_percent = min(one_node$Percent),
        Category_gate_passed = all(one_node$Count > 0L),
        stringsAsFactors = FALSE
      )
    }
  )
)

rownames(node_category_diagnostics) <- NULL
node_category_diagnostics <- node_category_diagnostics[
  order(
    match(
      node_category_diagnostics$Group_key,
      expected_groups$Group_key
    ),
    match(
      node_category_diagnostics$Node,
      network_vars
    )
  ),
  ,
  drop = FALSE
]

write.csv(
  category_frequency,
  file.path(
    formal_table_dir,
    "08A_category_frequencies.csv"
  ),
  row.names = FALSE
)

write.csv(
  node_category_diagnostics,
  file.path(
    formal_table_dir,
    "08A_node_category_diagnostics.csv"
  ),
  row.names = FALSE
)

stopifnot(
  nrow(category_frequency) == 328L,
  nrow(node_category_diagnostics) == 72L,
  all(
    node_category_diagnostics$Sample_N ==
      expected_groups$Expected_N[
        match(
          node_category_diagnostics$Group_key,
          expected_groups$Group_key
        )
      ]
  )
)

node_category_diagnostics
```

# Stage 13.5: all 36 pairwise contingency tables per group

```{r pairwise-cell-diagnostics}
node_pairs <- combn(
  network_vars,
  2,
  simplify = FALSE
)

pairwise_cell_diagnostics <- do.call(
  rbind,
  lapply(
    names(samples),
    function(group_key) {
      one_sample <- samples[[group_key]]

      do.call(
        rbind,
        lapply(
          node_pairs,
          function(node_pair) {
            cell_counts <- table(
              factor(
                ordered_values(
                  one_sample$Ordinal_nodes[[node_pair[[1L]]]]
                ),
                levels = expected_categories[[node_pair[[1L]]]]
              ),
              factor(
                ordered_values(
                  one_sample$Ordinal_nodes[[node_pair[[2L]]]]
                ),
                levels = expected_categories[[node_pair[[2L]]]]
              )
            )

            data.frame(
              Group_key = group_key,
              Wave = one_sample$Wave,
              Education_group = one_sample$Education_group,
              Node_1 = node_pair[[1L]],
              Node_2 = node_pair[[2L]],
              Pair_N = sum(cell_counts),
              Total_cells = length(cell_counts),
              Cells_0 = sum(cell_counts == 0L),
              Cells_1_4 = sum(
                cell_counts >= 1L &
                  cell_counts <= 4L
              ),
              Cells_5_9 = sum(
                cell_counts >= 5L &
                  cell_counts <= 9L
              ),
              Cells_10_plus = sum(
                cell_counts >= 10L
              ),
              Minimum_cell = min(cell_counts),
              Pair_with_zero = as.integer(
                any(cell_counts == 0L)
              ),
              Pair_with_1_4 = as.integer(
                any(
                  cell_counts >= 1L &
                    cell_counts <= 4L
                )
              ),
              Pair_with_5_9 = as.integer(
                any(
                  cell_counts >= 5L &
                    cell_counts <= 9L
                )
              ),
              stringsAsFactors = FALSE
            )
          }
        )
      )
    }
  )
)

pairwise_cell_summary <- do.call(
  rbind,
  lapply(
    split(
      pairwise_cell_diagnostics,
      pairwise_cell_diagnostics$Group_key
    ),
    function(one_group) {
      data.frame(
        Group_key = one_group$Group_key[[1L]],
        Wave = one_group$Wave[[1L]],
        Education_group = one_group$Education_group[[1L]],
        Sample_N = unique(one_group$Pair_N),
        Total_pairs = nrow(one_group),
        Total_cells = sum(one_group$Total_cells),
        Cells_0 = sum(one_group$Cells_0),
        Cells_1_4 = sum(one_group$Cells_1_4),
        Cells_5_9 = sum(one_group$Cells_5_9),
        Cells_10_plus = sum(one_group$Cells_10_plus),
        Pairs_with_zero = sum(one_group$Pair_with_zero),
        Pairs_with_1_4 = sum(one_group$Pair_with_1_4),
        Pairs_with_5_9 = sum(one_group$Pair_with_5_9),
        Minimum_cell = min(one_group$Minimum_cell),
        Sparse_cell_status = if (
          any(
            one_group$Cells_0 > 0L |
              one_group$Cells_1_4 > 0L
          )
        ) {
          "PASS_WITH_SPARSE_CELL_WARNING"
        } else {
          "PASS_NO_SPARSE_CELL_FLAG"
        },
        stringsAsFactors = FALSE
      )
    }
  )
)

rownames(pairwise_cell_summary) <- NULL
pairwise_cell_summary <- pairwise_cell_summary[
  match(
    expected_groups$Group_key,
    pairwise_cell_summary$Group_key
  ),
  ,
  drop = FALSE
]

pairwise_structure_passed <-
  nrow(pairwise_cell_diagnostics) == 288L &&
  all(
    table(pairwise_cell_diagnostics$Group_key) == 36L
  ) &&
  all(
    pairwise_cell_diagnostics$Pair_N ==
      expected_groups$Expected_N[
        match(
          pairwise_cell_diagnostics$Group_key,
          expected_groups$Group_key
        )
      ]
  ) &&
  all(
    pairwise_cell_diagnostics$Total_cells ==
      pairwise_cell_diagnostics$Cells_0 +
      pairwise_cell_diagnostics$Cells_1_4 +
      pairwise_cell_diagnostics$Cells_5_9 +
      pairwise_cell_diagnostics$Cells_10_plus
  ) &&
  all(pairwise_cell_summary$Total_cells == 733L)

write.csv(
  pairwise_cell_diagnostics,
  file.path(
    formal_table_dir,
    "08A_pairwise_cell_diagnostics.csv"
  ),
  row.names = FALSE
)

write.csv(
  pairwise_cell_summary,
  file.path(
    formal_table_dir,
    "08A_pairwise_cell_summary.csv"
  ),
  row.names = FALSE
)

stopifnot(
  pairwise_structure_passed
)

pairwise_cell_summary
```

# Stage 13.6: warning-audited polychoric estimation

```{r estimate-polychoric-matrices}
estimate_polychoric <- function(
  ordinal_nodes,
  node_order
) {
  captured_warnings <- character(0)
  captured_messages <- character(0)
  captured_error <- character(0)

  matrix_result <- tryCatch(
    withCallingHandlers(
      lavaan::lavCor(
        object = ordinal_nodes,
        ordered = node_order,
        missing = "listwise",
        estimator = "two.step",
        meanstructure = FALSE,
        cor_smooth = FALSE,
        output = "cor"
      ),
      warning = function(condition) {
        captured_warnings <<- c(
          captured_warnings,
          conditionMessage(condition)
        )
        invokeRestart("muffleWarning")
      },
      message = function(condition) {
        captured_messages <<- c(
          captured_messages,
          conditionMessage(condition)
        )
        invokeRestart("muffleMessage")
      }
    ),
    error = function(condition) {
      captured_error <<- conditionMessage(condition)
      NULL
    }
  )

  if (!is.null(matrix_result)) {
    matrix_result <- unclass(matrix_result)
  }

  list(
    Matrix = matrix_result,
    Warnings = unique(captured_warnings),
    Messages = unique(captured_messages),
    Error = unique(captured_error)
  )
}

polychoric_results <- lapply(
  samples,
  function(one_sample) {
    estimate_polychoric(
      ordinal_nodes = one_sample$Ordinal_nodes,
      node_order = network_vars
    )
  }
)

names(polychoric_results) <- names(samples)

condition_records <- do.call(
  rbind,
  lapply(
    names(polychoric_results),
    function(group_key) {
      one_result <- polychoric_results[[group_key]]
      one_sample <- samples[[group_key]]
      rows <- list()

      if (length(one_result$Warnings) > 0L) {
        rows[[length(rows) + 1L]] <- data.frame(
          Condition_type = "WARNING",
          Condition_text = one_result$Warnings,
          stringsAsFactors = FALSE
        )
      }

      if (length(one_result$Messages) > 0L) {
        rows[[length(rows) + 1L]] <- data.frame(
          Condition_type = "MESSAGE",
          Condition_text = one_result$Messages,
          stringsAsFactors = FALSE
        )
      }

      if (length(one_result$Error) > 0L) {
        rows[[length(rows) + 1L]] <- data.frame(
          Condition_type = "ERROR",
          Condition_text = one_result$Error,
          stringsAsFactors = FALSE
        )
      }

      if (length(rows) == 0L) {
        rows[[1L]] <- data.frame(
          Condition_type = "NONE",
          Condition_text = "",
          stringsAsFactors = FALSE
        )
      }

      one_table <- do.call(rbind, rows)
      data.frame(
        Group_key = group_key,
        Wave = one_sample$Wave,
        Education_group = one_sample$Education_group,
        one_table,
        stringsAsFactors = FALSE
      )
    }
  )
)

write.csv(
  condition_records,
  file.path(
    formal_table_dir,
    "08A_polychoric_conditions.csv"
  ),
  row.names = FALSE
)

condition_records
```

# Stage 13.7: matrix diagnostics and Stage-14 readiness

```{r validate-polychoric-matrices}
matrix_diagnostics <- do.call(
  rbind,
  lapply(
    names(polychoric_results),
    function(group_key) {
      one_result <- polychoric_results[[group_key]]
      one_sample <- samples[[group_key]]
      matrix_value <- one_result$Matrix

      estimation_succeeded <-
        is.matrix(matrix_value) &&
        length(one_result$Error) == 0L

      if (estimation_succeeded) {
        dimension_ok <- identical(
          dim(matrix_value),
          c(9L, 9L)
        )
        node_order_ok <-
          identical(
            rownames(matrix_value),
            network_vars
          ) &&
          identical(
            colnames(matrix_value),
            network_vars
          )
        all_values_finite <- all(
          is.finite(matrix_value)
        )
        maximum_asymmetry <- max(
          abs(
            matrix_value -
              t(matrix_value)
          )
        )
        maximum_diagonal_deviation <- max(
          abs(
            diag(matrix_value) - 1
          )
        )
        off_diagonal <- matrix_value[
          lower.tri(matrix_value)
        ]
        minimum_off_diagonal <- min(off_diagonal)
        maximum_off_diagonal <- max(off_diagonal)
        correlations_strictly_within_bounds <-
          minimum_off_diagonal > -1 &&
          maximum_off_diagonal < 1
        eigenvalues <- tryCatch(
          eigen(
            matrix_value,
            symmetric = TRUE,
            only.values = TRUE
          )$values,
          error = function(condition) {
            numeric(0)
          }
        )
        minimum_eigenvalue <- if (
          length(eigenvalues) == 9L
        ) {
          min(eigenvalues)
        } else {
          NA_real_
        }
        maximum_eigenvalue <- if (
          length(eigenvalues) == 9L
        ) {
          max(eigenvalues)
        } else {
          NA_real_
        }
        positive_definite <-
          length(eigenvalues) == 9L &&
          all(is.finite(eigenvalues)) &&
          all(eigenvalues > 0)
        condition_number <- if (positive_definite) {
          maximum_eigenvalue /
            minimum_eigenvalue
        } else {
          NA_real_
        }
      } else {
        dimension_ok <- FALSE
        node_order_ok <- FALSE
        all_values_finite <- FALSE
        maximum_asymmetry <- NA_real_
        maximum_diagonal_deviation <- NA_real_
        minimum_off_diagonal <- NA_real_
        maximum_off_diagonal <- NA_real_
        correlations_strictly_within_bounds <- FALSE
        minimum_eigenvalue <- NA_real_
        maximum_eigenvalue <- NA_real_
        condition_number <- NA_real_
        positive_definite <- FALSE
      }

      hard_matrix_gate_passed <-
        estimation_succeeded &&
        dimension_ok &&
        node_order_ok &&
        all_values_finite &&
        is.finite(maximum_asymmetry) &&
        maximum_asymmetry < 1e-12 &&
        is.finite(maximum_diagonal_deviation) &&
        maximum_diagonal_deviation < 1e-12 &&
        correlations_strictly_within_bounds &&
        positive_definite

      data.frame(
        Group_key = group_key,
        Wave = one_sample$Wave,
        Education_group = one_sample$Education_group,
        Sample_N = one_sample$N,
        Correlation_type = "Polychoric",
        Function = "lavaan::lavCor",
        Estimator = "two.step",
        Missing = "listwise",
        Cor_smooth = FALSE,
        Warning_n = length(one_result$Warnings),
        Message_n = length(one_result$Messages),
        Error_n = length(one_result$Error),
        Estimation_succeeded = estimation_succeeded,
        Dimension_ok = dimension_ok,
        Node_order_ok = node_order_ok,
        All_values_finite = all_values_finite,
        Maximum_asymmetry = maximum_asymmetry,
        Maximum_diagonal_deviation =
          maximum_diagonal_deviation,
        Minimum_off_diagonal = minimum_off_diagonal,
        Maximum_off_diagonal = maximum_off_diagonal,
        Correlations_strictly_within_bounds =
          correlations_strictly_within_bounds,
        Minimum_eigenvalue = minimum_eigenvalue,
        Maximum_eigenvalue = maximum_eigenvalue,
        Condition_number = condition_number,
        Positive_definite = positive_definite,
        Warning_free = length(one_result$Warnings) == 0L,
        Hard_matrix_gate_passed =
          hard_matrix_gate_passed,
        stringsAsFactors = FALSE
      )
    }
  )
)

matrix_diagnostics <- matrix_diagnostics[
  match(
    expected_groups$Group_key,
    matrix_diagnostics$Group_key
  ),
  ,
  drop = FALSE
]

matrix_values <- do.call(
  rbind,
  lapply(
    names(polychoric_results),
    function(group_key) {
      matrix_value <- polychoric_results[[group_key]]$Matrix
      one_sample <- samples[[group_key]]

      if (!is.matrix(matrix_value)) {
        return(NULL)
      }

      index <- expand.grid(
        Row_index = seq_along(network_vars),
        Column_index = seq_along(network_vars)
      )

      data.frame(
        Group_key = group_key,
        Wave = one_sample$Wave,
        Education_group = one_sample$Education_group,
        Row_node = network_vars[index$Row_index],
        Column_node = network_vars[index$Column_index],
        Correlation = matrix_value[
          cbind(
            index$Row_index,
            index$Column_index
          )
        ],
        stringsAsFactors = FALSE
      )
    }
  )
)

eigenvalue_diagnostics <- do.call(
  rbind,
  lapply(
    names(polychoric_results),
    function(group_key) {
      matrix_value <- polychoric_results[[group_key]]$Matrix
      one_sample <- samples[[group_key]]

      if (!is.matrix(matrix_value)) {
        return(NULL)
      }

      eigenvalues <- eigen(
        matrix_value,
        symmetric = TRUE,
        only.values = TRUE
      )$values

      data.frame(
        Group_key = group_key,
        Wave = one_sample$Wave,
        Education_group = one_sample$Education_group,
        Eigenvalue_index = seq_along(eigenvalues),
        Eigenvalue = eigenvalues,
        stringsAsFactors = FALSE
      )
    }
  )
)

group_category_gate <- aggregate(
  Category_gate_passed ~ Group_key,
  data = node_category_diagnostics,
  FUN = all
)

stage14_readiness <- merge(
  matrix_diagnostics,
  group_category_gate,
  by = "Group_key",
  all.x = TRUE,
  sort = FALSE
)

stage14_readiness <- merge(
  stage14_readiness,
  pairwise_cell_summary[
    ,
    c(
      "Group_key",
      "Cells_0",
      "Cells_1_4",
      "Pairs_with_zero",
      "Pairs_with_1_4",
      "Sparse_cell_status"
    )
  ],
  by = "Group_key",
  all.x = TRUE,
  sort = FALSE
)

stage14_readiness <- stage14_readiness[
  match(
    expected_groups$Group_key,
    stage14_readiness$Group_key
  ),
  ,
  drop = FALSE
]

stage14_readiness$Input_gate_passed <-
  sample_input_validation$Passed[
    match(
      stage14_readiness$Group_key,
      sample_input_validation$Group_key
    )
  ]

stage14_readiness$Ready_for_stage14 <- with(
  stage14_readiness,
  Input_gate_passed &
    Category_gate_passed &
    Warning_free &
    Error_n == 0L &
    Hard_matrix_gate_passed
)

stage14_readiness$Readiness_status <- ifelse(
  stage14_readiness$Ready_for_stage14,
  ifelse(
    stage14_readiness$Cells_0 > 0L |
      stage14_readiness$Cells_1_4 > 0L,
    "READY_WITH_SPARSE_CELL_WARNING",
    "READY"
  ),
  ifelse(
    stage14_readiness$Warning_n > 0L &&
      stage14_readiness$Error_n == 0L &&
      stage14_readiness$Hard_matrix_gate_passed,
    "REVIEW_REQUIRED_ESTIMATION_WARNING",
    "NOT_READY_INPUT_OR_MATRIX_FAILURE"
  )
)

write.csv(
  matrix_diagnostics,
  file.path(
    formal_table_dir,
    "08A_polychoric_matrix_diagnostics.csv"
  ),
  row.names = FALSE
)

write.csv(
  matrix_values,
  file.path(
    formal_table_dir,
    "08A_polychoric_matrix_values.csv"
  ),
  row.names = FALSE
)

write.csv(
  eigenvalue_diagnostics,
  file.path(
    formal_table_dir,
    "08A_polychoric_eigenvalues.csv"
  ),
  row.names = FALSE
)

write.csv(
  stage14_readiness,
  file.path(
    formal_table_dir,
    "08A_stage14_readiness.csv"
  ),
  row.names = FALSE
)

matrix_diagnostics
stage14_readiness
```

# Stage 13.8: save and round-trip the diagnostic bundle

```{r save-diagnostic-bundle}
polychoric_matrices <- lapply(
  polychoric_results,
  function(one_result) {
    one_result$Matrix
  }
)

diagnostic_bundle <- list(
  Source_education_bundle = normalizePath(
    education_bundle_file,
    winslash = "/"
  ),
  Source_education_bundle_SHA256 =
    file_sha256(education_bundle_file),
  Source_analysis_ready_SHA256 =
    education_bundle$Source_SHA256,
  Decision_ID = "D25",
  Group_specification = expected_groups,
  Node_order = network_vars,
  Expected_categories = expected_categories,
  Correlation_specification = list(
    Function = "lavaan::lavCor",
    Ordered = network_vars,
    Missing = "listwise",
    Estimator = "two.step",
    Meanstructure = FALSE,
    Cor_smooth = FALSE,
    Output = "cor",
    Weighted = FALSE
  ),
  Package_versions = package_gate,
  Category_frequencies = category_frequency,
  Node_category_diagnostics =
    node_category_diagnostics,
  Pairwise_cell_diagnostics =
    pairwise_cell_diagnostics,
  Pairwise_cell_summary = pairwise_cell_summary,
  Conditions = condition_records,
  Matrix_diagnostics = matrix_diagnostics,
  Stage14_readiness = stage14_readiness,
  Matrices = polychoric_matrices,
  R_version = R.version.string
)

diagnostic_bundle_file <- file.path(
  formal_model_dir,
  "08A_education_polychoric_input_diagnostics.rds"
)

saveRDS(
  diagnostic_bundle,
  diagnostic_bundle_file,
  version = 3,
  compress = "gzip"
)

diagnostic_bundle_reloaded <- readRDS(
  diagnostic_bundle_file
)

roundtrip_validation <- data.frame(
  Check = c(
    "Diagnostic bundle exists",
    "Diagnostic bundle is non-empty",
    "Reloaded diagnostic bundle is identical",
    "Eight matrix slots are preserved",
    "Source education-bundle SHA-256 is preserved",
    "No graph, network or bootstrap object is stored"
  ),
  Passed = c(
    file.exists(diagnostic_bundle_file),
    file.info(diagnostic_bundle_file)$size > 0,
    identical(
      diagnostic_bundle,
      diagnostic_bundle_reloaded
    ),
    length(
      diagnostic_bundle_reloaded$Matrices
    ) == 8L,
    identical(
      diagnostic_bundle_reloaded$
        Source_education_bundle_SHA256,
      expected_bundle_sha256
    ),
    !any(
      c(
        "graph",
        "network",
        "bootstrap"
      ) %in%
        names(diagnostic_bundle_reloaded)
    )
  ),
  stringsAsFactors = FALSE
)

write.csv(
  roundtrip_validation,
  file.path(
    formal_table_dir,
    "08A_roundtrip_validation.csv"
  ),
  row.names = FALSE
)

stopifnot(
  all(roundtrip_validation$Passed)
)

roundtrip_validation
```

# Stage 13.9: prohibited-call and completion audit

```{r prohibited-call-scan}
source_lines <- readLines(
  notebook_file,
  warn = FALSE
)

prohibited_call_patterns <- c(
  paste0(
    "bootnet::",
    "estimateNetwork("
  ),
  paste0(
    "qgraph::",
    "EBICglasso("
  ),
  paste0(
    "bootnet::",
    "bootnet("
  ),
  paste0(
    "NetworkComparisonTest::",
    "NCT("
  )
)

prohibited_call_scan <- data.frame(
  Pattern = prohibited_call_patterns,
  Found = vapply(
    prohibited_call_patterns,
    function(pattern) {
      any(
        grepl(
          pattern,
          source_lines,
          fixed = TRUE
        )
      )
    },
    FUN.VALUE = logical(1)
  ),
  stringsAsFactors = FALSE
)

write.csv(
  prohibited_call_scan,
  file.path(
    formal_table_dir,
    "08A_prohibited_call_scan.csv"
  ),
  row.names = FALSE
)

stopifnot(
  !any(prohibited_call_scan$Found)
)

prohibited_call_scan
```

```{r stage13-status-and-validation}
all_groups_ready_for_stage14 <- all(
  stage14_readiness$Ready_for_stage14
)

stage13_validation <- data.frame(
  Check = c(
    "Package gate passed",
    "Decision gate passed",
    "Overall Stage-12 input gate passed",
    "Eight sample-input gates passed",
    "Same-wave education groups do not overlap",
    "328 category-frequency rows were created",
    "72 node-category rows were created",
    "288 pairwise-cell rows were created",
    "Each group has 36 pairs and 733 expected cells",
    "Eight polychoric attempts were recorded",
    "Condition capture covers all eight groups",
    "Matrix diagnostic table contains eight groups",
    "Round-trip validation passed",
    "No prohibited call was found"
  ),
  Passed = c(
    all(package_gate$Installed),
    all(decision_gate$Passed),
    all(overall_input_validation$Passed),
    all(sample_input_validation$Passed),
    all(same_wave_group_overlap$Passed),
    nrow(category_frequency) == 328L,
    nrow(node_category_diagnostics) == 72L,
    nrow(pairwise_cell_diagnostics) == 288L,
    pairwise_structure_passed,
    length(polychoric_results) == 8L,
    all(
      expected_groups$Group_key %in%
        condition_records$Group_key
    ),
    nrow(matrix_diagnostics) == 8L,
    all(roundtrip_validation$Passed),
    !any(prohibited_call_scan$Found)
  ),
  stringsAsFactors = FALSE
)

stage13_status <- data.frame(
  Stage = "Stage 13 education polychoric input diagnostics",
  Execution_status = if (
    all(stage13_validation$Passed)
  ) {
    if (all_groups_ready_for_stage14) {
      "STAGE_13_COMPLETE_AND_VALIDATED"
    } else {
      "STAGE_13_COMPLETE_WITH_STAGE14_BLOCKERS"
    }
  } else {
    "STAGE_13_IMPLEMENTATION_FAILURE"
  },
  Groups_processed = nrow(stage14_readiness),
  Groups_ready_for_stage14 = sum(
    stage14_readiness$Ready_for_stage14
  ),
  Groups_blocked_from_stage14 = sum(
    !stage14_readiness$Ready_for_stage14
  ),
  Groups_with_zero_cells = sum(
    stage14_readiness$Cells_0 > 0L
  ),
  Groups_with_cells_1_4 = sum(
    stage14_readiness$Cells_1_4 > 0L
  ),
  Education_GGM_run = FALSE,
  Bootstrap_run = FALSE,
  NCT_run = FALSE,
  Sensitivity_run = FALSE,
  Political_interpretation_run = FALSE,
  stringsAsFactors = FALSE
)

stage13_completion <- data.frame(
  Component = c(
    "Governance and package gates",
    "Stage-12 source validation",
    "Marginal category diagnostics",
    "Pairwise-cell diagnostics",
    "Polychoric attempts and condition capture",
    "Matrix and eigenvalue diagnostics",
    "Diagnostic bundle round trip",
    "Prohibited-call scan",
    "Education network estimation",
    "Education edge bootstrap",
    "NCT",
    "Political interpretation"
  ),
  Status = c(
    rep("COMPLETE_AND_VALIDATED", 8L),
    "NOT_RUN_STAGE14_SCOPE",
    "NOT_RUN_LATER_SCOPE",
    "NOT_IN_CURRENT_SCOPE_NOT_RUN",
    "NOT_AUTHORISED_NOT_RUN"
  ),
  Required_for_stage13 = c(
    rep(TRUE, 8L),
    rep(FALSE, 4L)
  ),
  Passed = c(
    all(package_gate$Installed) &&
      all(decision_gate$Passed),
    all(overall_input_validation$Passed) &&
      all(sample_input_validation$Passed),
    nrow(category_frequency) == 328L &&
      nrow(node_category_diagnostics) == 72L,
    pairwise_structure_passed,
    length(polychoric_results) == 8L &&
      nrow(matrix_diagnostics) == 8L,
    nrow(matrix_diagnostics) == 8L,
    all(roundtrip_validation$Passed),
    !any(prohibited_call_scan$Found),
    TRUE,
    TRUE,
    TRUE,
    TRUE
  ),
  stringsAsFactors = FALSE
)

write.csv(
  stage13_validation,
  file.path(
    formal_table_dir,
    "08A_stage13_validation.csv"
  ),
  row.names = FALSE
)

write.csv(
  stage13_status,
  file.path(
    formal_table_dir,
    "08A_stage13_status.csv"
  ),
  row.names = FALSE
)

write.csv(
  stage13_completion,
  file.path(
    formal_table_dir,
    "08A_stage13_completion_status.csv"
  ),
  row.names = FALSE
)

stage13_validation
stage13_status
stage13_completion

stopifnot(
  all(stage13_validation$Passed),
  all(
    stage13_completion$Passed[
      stage13_completion$Required_for_stage13
    ]
  )
)
```

# Stage 13.10: output inventory, log and session information

```{r output-inventory-log-and-session}
stage13_output_paths <- c(
  list.files(
    formal_table_dir,
    pattern = "^08A_",
    full.names = TRUE
  ),
  diagnostic_bundle_file
)

stage13_output_paths <- sort(
  unique(
    stage13_output_paths[
      file.exists(stage13_output_paths)
    ]
  )
)

output_inventory <- data.frame(
  File = basename(stage13_output_paths),
  Path = normalizePath(
    stage13_output_paths,
    winslash = "/"
  ),
  Bytes = file.info(stage13_output_paths)$size,
  SHA256 = vapply(
    stage13_output_paths,
    file_sha256,
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

output_inventory_file <- file.path(
  formal_table_dir,
  "08A_stage13_output_inventory.csv"
)

write.csv(
  output_inventory,
  output_inventory_file,
  row.names = FALSE
)

session_file <- file.path(
  log_dir,
  "08A_stage13_sessionInfo.txt"
)

writeLines(
  capture.output(
    sessionInfo()
  ),
  con = session_file
)

log_file <- file.path(
  log_dir,
  "08A_stage13_execution.log"
)

writeLines(
  c(
    "STEP 08A STAGE 13 EXECUTION LOG",
    paste(
      "Stage-12 bundle SHA-256:",
      file_sha256(education_bundle_file)
    ),
    paste(
      "Groups processed:",
      nrow(stage14_readiness)
    ),
    paste(
      "Groups ready for Stage 14:",
      sum(stage14_readiness$Ready_for_stage14)
    ),
    paste(
      "Groups blocked from Stage 14:",
      sum(!stage14_readiness$Ready_for_stage14)
    ),
    paste(
      "Polychoric warnings:",
      sum(matrix_diagnostics$Warning_n)
    ),
    paste(
      "Polychoric errors:",
      sum(matrix_diagnostics$Error_n)
    ),
    paste(
      "Positive-definite matrices:",
      sum(matrix_diagnostics$Positive_definite)
    ),
    paste(
      "Groups with zero pairwise cells:",
      sum(stage14_readiness$Cells_0 > 0L)
    ),
    paste(
      "Stage-13 status:",
      stage13_status$Execution_status
    ),
    "Education network estimation: NOT RUN",
    "Education edge bootstrap: NOT RUN",
    "NCT: NOT RUN",
    "Sensitivity analysis: NOT RUN",
    "Political interpretation: NOT RUN"
  ),
  con = log_file
)

stopifnot(
  file.exists(output_inventory_file),
  file.exists(session_file),
  file.exists(log_file),
  file.info(output_inventory_file)$size > 0,
  file.info(session_file)$size > 0,
  file.info(log_file)$size > 0
)

output_inventory
```

# Stage-13 boundary

The Stage-13 diagnostic result determines whether each education-group
input is technically eligible for a separately authorised Stage 14.
No Stage-14 model is executed by this notebook.
