---
title: "08B Education Network Estimation and Edge-Weight Accuracy"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: show
params:
  execute_models: false
---

# Scope and execution boundary

This notebook implements the bounded Stage 14 and Stage 15 education-network
workflow authorised by decision D26. It estimates eight descriptive,
unweighted education-group point networks and runs one 1,000-repetition
nonparametric edge-weight bootstrap for each network. It does not run NCT,
centrality, case-dropping bootstrap, sensitivity analysis, weighted network
estimation, formal education-group inference, or political interpretation.

The Stage-12 sample bundle and Stage-13 diagnostic bundle are read-only inputs.
Sparse cells remain documented; the notebook does not merge categories, remove
nodes, smooth or repair matrices, or change the frozen model specification.

Ordinary rendering reuses the already validated Stage-14 and Stage-15 outputs.
Model estimation and bootstrap execution occur only when
`params$execute_models = TRUE` is supplied explicitly. A successful ordinary
render therefore documents the saved results and may refresh audit tables; it
is not a new model or bootstrap run.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = TRUE,
  message = TRUE,
  error = FALSE
)
```

# 1. Project, packages and helpers

```{r project-root-packages-helpers}
current_dir <- normalizePath(".", winslash = "/")

if (file.exists(file.path(current_dir, "ANES.Rproj"))) {
  project_root <- current_dir
} else if (
  basename(current_dir) == "notebooks" &&
    file.exists(file.path(dirname(current_dir), "ANES.Rproj"))
) {
  project_root <- dirname(current_dir)
} else {
  stop("Project root could not be identified.")
}

knitr::opts_knit$set(root.dir = project_root)

notebook_file <- file.path(
  project_root,
  "notebooks",
  "08B_education_network_estimation_and_edge_accuracy.Rmd"
)

intermediate_dir <- file.path(project_root, "data", "intermediate")
formal_model_dir <- file.path(project_root, "outputs", "models", "formal")
formal_table_dir <- file.path(project_root, "outputs", "tables", "formal")
pilot_model_dir <- file.path(project_root, "outputs", "models", "pilot")
pilot_table_dir <- file.path(project_root, "outputs", "tables", "pilot")
log_dir <- file.path(project_root, "outputs", "logs")

for (target_dir in c(
  formal_model_dir,
  formal_table_dir,
  pilot_model_dir,
  pilot_table_dir,
  log_dir
)) {
  dir.create(target_dir, recursive = TRUE, showWarnings = FALSE)
}

required_packages <- c(
  "bootnet",
  "digest",
  "jsonlite",
  "knitr",
  "lavaan",
  "qgraph",
  "rmarkdown",
  "renv"
)

package_gate <- data.frame(
  Package = required_packages,
  Installed = vapply(
    required_packages,
    requireNamespace,
    quietly = TRUE,
    FUN.VALUE = logical(1)
  ),
  Version = vapply(
    required_packages,
    function(package_name) {
      if (requireNamespace(package_name, quietly = TRUE)) {
        as.character(utils::packageVersion(package_name))
      } else {
        NA_character_
      }
    },
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

normalise_version <- function(value) {
  gsub("-", ".", as.character(value), fixed = TRUE)
}

normalise_status <- function(value) {
  value <- toupper(trimws(as.character(value)))
  gsub("[^A-Z0-9]+", "_", value)
}

file_sha256 <- function(path) {
  if (!file.exists(path)) {
    return(NA_character_)
  }
  toupper(
    digest::digest(
      file = path,
      algo = "sha256",
      serialize = FALSE
    )
  )
}

ordered_values <- function(value) {
  as.numeric(as.character(value))
}

known_dense_warning <- function(warnings) {
  if (length(warnings) == 0L) {
    return(TRUE)
  }
  all(
    grepl(
      "A dense regularized network was selected",
      warnings,
      fixed = TRUE
    )
  )
}

retry_warning_pattern <-
  "^([0-9]+) bootstrap estimation\\(s\\) failed and were resampled\\.$"

known_bootstrap_warning <- function(warnings) {
  if (length(warnings) == 0L) {
    return(TRUE)
  }
  all(grepl(retry_warning_pattern, warnings))
}

count_retry_warnings <- function(warnings) {
  if (length(warnings) == 0L) {
    return(0L)
  }

  warning_parts <- regmatches(
    warnings,
    regexec(retry_warning_pattern, warnings)
  )

  retry_values <- vapply(
    warning_parts,
    function(value) {
      if (length(value) >= 2L) {
        as.integer(value[[2L]])
      } else {
        0L
      }
    },
    FUN.VALUE = integer(1)
  )

  sum(retry_values)
}

stopifnot(all(package_gate$Installed))

installed_version_gate <- data.frame(
  Package = c("bootnet", "qgraph", "lavaan"),
  Expected = c("1.9.1", "1.10.1", "0.7.2"),
  Observed = c(
    as.character(utils::packageVersion("bootnet")),
    as.character(utils::packageVersion("qgraph")),
    as.character(utils::packageVersion("lavaan"))
  ),
  stringsAsFactors = FALSE
)
installed_version_gate$Passed <-
  normalise_version(installed_version_gate$Observed) ==
    normalise_version(installed_version_gate$Expected)

renv_lock_file <- file.path(project_root, "renv.lock")
stopifnot(file.exists(renv_lock_file))
renv_lock <- renv::lockfile_read(renv_lock_file)

locked_version <- function(package_name) {
  package_record <- renv_lock$Packages[[package_name]]
  if (is.null(package_record) || is.null(package_record$Version)) {
    return(NA_character_)
  }
  as.character(package_record$Version)
}

lock_version_gate <- data.frame(
  Package = c("bootnet", "qgraph", "lavaan"),
  Expected = c("1.9.1", "1.10.1", "0.7.2"),
  Observed = vapply(
    c("bootnet", "qgraph", "lavaan"),
    locked_version,
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)
lock_version_gate$Passed <-
  normalise_version(lock_version_gate$Observed) ==
    normalise_version(lock_version_gate$Expected)

stopifnot(
  all(installed_version_gate$Passed),
  all(lock_version_gate$Passed)
)

write.csv(
  package_gate,
  file.path(formal_table_dir, "08B_package_gate.csv"),
  row.names = FALSE
)
write.csv(
  installed_version_gate,
  file.path(formal_table_dir, "08B_installed_version_gate.csv"),
  row.names = FALSE
)
write.csv(
  lock_version_gate,
  file.path(formal_table_dir, "08B_lock_version_gate.csv"),
  row.names = FALSE
)

package_gate
installed_version_gate
lock_version_gate
```

# 2. Canonical governance gate

```{r governance-gate}
registry_config <- file.path(
  project_root,
  "config",
  "decision_registry_path.txt"
)

stopifnot(file.exists(registry_config))

registry_lines <- trimws(readLines(registry_config, warn = FALSE))
registry_lines <- registry_lines[nzchar(registry_lines)]

stopifnot(
  length(registry_lines) == 1L,
  file.exists(registry_lines[[1L]])
)

registry_path <- registry_lines[[1L]]
decision_registry <- read.csv(
  registry_path,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

decision_row <- function(decision_id) {
  decision_registry[
    decision_registry$decision_id == decision_id,
    ,
    drop = FALSE
  ]
}

d08a <- decision_row("D08A")
d08b <- decision_row("D08B")
d13 <- decision_row("D13")
d14 <- decision_row("D14")
d15 <- decision_row("D15")
d18 <- decision_row("D18")
d22 <- decision_row("D22")
d25 <- decision_row("D25")
d26 <- decision_row("D26")

d26_allowed_status <- c(
  "AUTHORIZED",
  "AUTHORISED",
  "AUTHORIZED_STAGE_14_15_BOUNDED",
  "AUTHORISED_STAGE_14_15_BOUNDED",
  if (!isTRUE(params$execute_models)) {
    "COMPLETE_AND_VALIDATED"
  }
)

decision_gate <- data.frame(
  Decision_ID = c(
    "D08A",
    "D08B",
    "D13",
    "D14",
    "D15",
    "D18",
    "D22",
    "D25",
    "D26"
  ),
  Check = c(
    "Bounded descriptive education module is frozen and approved",
    "Formal education inference remains outside current scope",
    "Ordinal polychoric EBICglasso estimator is frozen",
    "Primary gamma is 0.50",
    "Primary education networks are unweighted",
    "NCT remains outside current scope",
    "One thousand nonparametric edge bootstraps are frozen",
    "Stage 13 is complete and validated",
    if (isTRUE(params$execute_models)) {
      "Stage 14 and 15 bounded execution is explicitly authorised"
    } else {
      "Stage 14 and 15 outputs are complete and may be reused"
    }
  ),
  Passed = c(
    nrow(d08a) == 1L &&
      normalise_status(d08a$decision_status) == "FROZEN" &&
      normalise_status(d08a$researcher_approval_status) == "APPROVED",
    nrow(d08b) == 1L &&
      normalise_status(d08b$decision_status) == "NOT_IN_CURRENT_SCOPE" &&
      normalise_status(d08b$implementation_status) == "NOT_APPLICABLE",
    nrow(d13) == 1L &&
      normalise_status(d13$decision_status) == "FROZEN" &&
      grepl("Polychoric", d13$current_value, fixed = TRUE),
    nrow(d14) == 1L &&
      normalise_status(d14$decision_status) == "FROZEN" &&
      grepl("0.50 primary", d14$current_value, fixed = TRUE),
    nrow(d15) == 1L &&
      normalise_status(d15$decision_status) == "FROZEN" &&
      grepl("Unweighted primary", d15$current_value, fixed = TRUE),
    nrow(d18) == 1L &&
      normalise_status(d18$decision_status) == "NOT_IN_CURRENT_SCOPE" &&
      normalise_status(d18$implementation_status) == "NOT_APPLICABLE",
    nrow(d22) == 1L &&
      normalise_status(d22$decision_status) == "FROZEN" &&
      grepl("1,000 nonparametric edge-weight bootstrap", d22$current_value,
        fixed = TRUE
      ),
    nrow(d25) == 1L &&
      normalise_status(d25$implementation_status) ==
        "STAGE_13_COMPLETE_AND_VALIDATED",
    nrow(d26) == 1L &&
      normalise_status(d26$researcher_approval_status) == "APPROVED" &&
      normalise_status(d26$implementation_status) %in%
        d26_allowed_status
  ),
  stringsAsFactors = FALSE
)

write.csv(
  decision_gate,
  file.path(formal_table_dir, "08B_decision_gate.csv"),
  row.names = FALSE
)

decision_gate
stopifnot(all(decision_gate$Passed))
```

# 3. Read-only source and hash gate

```{r source-and-hash-gate}
education_bundle_file <- file.path(
  intermediate_dir,
  "08_education_group_samples_2012_2016_2020_2024.rds"
)

stage13_bundle_file <- file.path(
  formal_model_dir,
  "08A_education_polychoric_input_diagnostics.rds"
)

stage13_readiness_file <- file.path(
  formal_table_dir,
  "08A_stage14_readiness.csv"
)

expected_education_bundle_sha256 <-
  "48D8320E2F26E609A30FFBBA54601B6625896BEBF2565F24345396E0FE50E4AF"

expected_stage13_bundle_sha256 <-
  "A7B18CE2872E35D181C6736908F0CF785EBF39C283D79FDAFF87A8E733479BCC"

source_gate <- data.frame(
  Input = c(
    "Canonical decision registry",
    "Stage-12 education sample bundle",
    "Stage-13 polychoric diagnostic bundle",
    "Stage-13 readiness table"
  ),
  Path = normalizePath(
    c(
      registry_path,
      education_bundle_file,
      stage13_bundle_file,
      stage13_readiness_file
    ),
    winslash = "/",
    mustWork = FALSE
  ),
  Exists = file.exists(
    c(
      registry_path,
      education_bundle_file,
      stage13_bundle_file,
      stage13_readiness_file
    )
  ),
  SHA256 = vapply(
    c(
      registry_path,
      education_bundle_file,
      stage13_bundle_file,
      stage13_readiness_file
    ),
    file_sha256,
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

source_gate$Expected_SHA256 <- c(
  NA_character_,
  expected_education_bundle_sha256,
  expected_stage13_bundle_sha256,
  NA_character_
)

source_gate$Passed <-
  source_gate$Exists &
    (
      is.na(source_gate$Expected_SHA256) |
        source_gate$SHA256 == source_gate$Expected_SHA256
    )

write.csv(
  source_gate,
  file.path(formal_table_dir, "08B_source_gate.csv"),
  row.names = FALSE
)

source_gate
stopifnot(all(source_gate$Passed))

education_bundle <- readRDS(education_bundle_file)
stage13_bundle <- readRDS(stage13_bundle_file)
stage13_readiness <- read.csv(
  stage13_readiness_file,
  stringsAsFactors = FALSE,
  check.names = FALSE
)
```

# 4. Fixed groups, nodes and ordered complete-case inputs

```{r fixed-input-validation}
network_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223"
)

expected_categories <- list(
  VCF0806 = 1:7,
  VCF0809 = 1:7,
  VCF0838 = 1:4,
  VCF0839 = 1:7,
  VCF0879a = c(1, 3, 5),
  VCF0888 = 1:3,
  VCF0890 = 1:3,
  VCF0894 = 1:3,
  VCF9223 = 1:4
)

expected_groups <- data.frame(
  Group_key = c(
    "2012__no_college_degree",
    "2012__college_or_advanced",
    "2016__no_college_degree",
    "2016__college_or_advanced",
    "2020__no_college_degree",
    "2020__college_or_advanced",
    "2024__no_college_degree",
    "2024__college_or_advanced"
  ),
  Wave = c(2012L, 2012L, 2016L, 2016L, 2020L, 2020L, 2024L, 2024L),
  Education_group = rep(
    c("No college degree", "College or advanced degree"),
    times = 4L
  ),
  Expected_N = c(2773L, 1531L, 1478L, 1216L, 2624L, 2691L, 1729L, 1781L),
  stringsAsFactors = FALSE
)

samples <- education_bundle$Samples
stage13_matrices <- stage13_bundle$Matrices

bundle_gate <- data.frame(
  Check = c(
    "Stage-12 group keys are exact",
    "Stage-13 matrix keys are exact",
    "Stage-13 readiness keys are exact",
    "Stage-12 node order is exact",
    "Stage-13 node order is exact",
    "Stage-12 category specification is exact",
    "Stage-13 category specification is exact",
    "Stage-13 records the Stage-12 bundle hash",
    "Every group is ready for Stage 14"
  ),
  Passed = c(
    identical(names(samples), expected_groups$Group_key),
    identical(names(stage13_matrices), expected_groups$Group_key),
    identical(stage13_readiness$Group_key, expected_groups$Group_key),
    identical(education_bundle$Node_order, network_vars),
    identical(stage13_bundle$Node_order, network_vars),
    isTRUE(all.equal(
      education_bundle$Expected_categories,
      expected_categories,
      check.attributes = FALSE
    )),
    isTRUE(all.equal(
      stage13_bundle$Expected_categories,
      expected_categories,
      check.attributes = FALSE
    )),
    identical(
      stage13_bundle$Source_education_bundle_SHA256,
      expected_education_bundle_sha256
    ),
    nrow(stage13_readiness) == 8L &&
      all(stage13_readiness$Ready_for_stage14)
  ),
  stringsAsFactors = FALSE
)

input_validation <- do.call(
  rbind,
  lapply(
    seq_len(nrow(expected_groups)),
    function(row_index) {
      specification <- expected_groups[row_index, , drop = FALSE]
      group_key <- specification$Group_key[[1L]]
      one_sample <- samples[[group_key]]
      ordinal_nodes <- one_sample$Ordinal_nodes
      matrix_value <- stage13_matrices[[group_key]]

      ordered_levels_exact <- vapply(
        network_vars,
        function(node_name) {
          is.ordered(ordinal_nodes[[node_name]]) &&
            identical(
              as.numeric(levels(ordinal_nodes[[node_name]])),
              as.numeric(expected_categories[[node_name]])
            )
        },
        FUN.VALUE = logical(1)
      )

      matrix_eigenvalues <- eigen(
        matrix_value,
        symmetric = TRUE,
        only.values = TRUE
      )$values

      data.frame(
        Group_key = group_key,
        Wave = specification$Wave,
        Education_group = specification$Education_group,
        Expected_N = specification$Expected_N,
        Observed_N = nrow(ordinal_nodes),
        Correct_node_order = identical(names(ordinal_nodes), network_vars),
        All_ordered_levels_exact = all(ordered_levels_exact),
        Complete_cases = all(stats::complete.cases(ordinal_nodes)),
        Matrix_9_by_9 = identical(dim(matrix_value), c(9L, 9L)),
        Matrix_node_order_exact =
          identical(rownames(matrix_value), network_vars) &&
            identical(colnames(matrix_value), network_vars),
        Matrix_finite = all(is.finite(matrix_value)),
        Matrix_symmetric = max(abs(matrix_value - t(matrix_value))) < 1e-12,
        Matrix_unit_diagonal = max(abs(diag(matrix_value) - 1)) < 1e-12,
        Matrix_positive_definite = all(matrix_eigenvalues > 0),
        stringsAsFactors = FALSE
      )
    }
  )
)

input_validation$Passed <- with(
  input_validation,
  Expected_N == Observed_N &
    Correct_node_order &
    All_ordered_levels_exact &
    Complete_cases &
    Matrix_9_by_9 &
    Matrix_node_order_exact &
    Matrix_finite &
    Matrix_symmetric &
    Matrix_unit_diagonal &
    Matrix_positive_definite
)

write.csv(
  bundle_gate,
  file.path(formal_table_dir, "08B_bundle_gate.csv"),
  row.names = FALSE
)
write.csv(
  input_validation,
  file.path(formal_table_dir, "08B_input_validation.csv"),
  row.names = FALSE
)

bundle_gate
input_validation

stopifnot(
  all(bundle_gate$Passed),
  all(input_validation$Passed)
)
```

# 5. Frozen controls and random seeds

```{r frozen-run-controls}
RUN_STAGE14_POINT_NETWORKS <- isTRUE(params$execute_models)
RUN_STAGE15_TECHNICAL_PILOT <- isTRUE(params$execute_models)
RUN_STAGE15_EDGE_BOOTSTRAPS <- isTRUE(params$execute_models)
REUSE_VALIDATED_FORMAL_OUTPUTS <- !isTRUE(params$execute_models)

technical_pilot_group <- "2016__college_or_advanced"
technical_pilot_seed <- 2026083010L

point_seeds <- setNames(
  2026083021L:2026083028L,
  expected_groups$Group_key
)

bootstrap_seeds <- setNames(
  2026083011L:2026083018L,
  expected_groups$Group_key
)

run_control_table <- data.frame(
  Group_key = expected_groups$Group_key,
  Wave = expected_groups$Wave,
  Education_group = expected_groups$Education_group,
  N = expected_groups$Expected_N,
  Point_seed = unname(point_seeds),
  Bootstrap_seed = unname(bootstrap_seeds),
  Gamma = 0.50,
  Bootstrap_repetitions = 1000L,
  Weighted = FALSE,
  Execution_mode = if (
    REUSE_VALIDATED_FORMAL_OUTPUTS
  ) {
    "REUSE_VALIDATED_OUTPUTS"
  } else {
    "RECOMPUTE_MODELS_AND_BOOTSTRAPS"
  },
  stringsAsFactors = FALSE
)

write.csv(
  run_control_table,
  file.path(formal_table_dir, "08B_run_controls.csv"),
  row.names = FALSE
)

stopifnot(
  xor(
    all(c(
      RUN_STAGE14_POINT_NETWORKS,
      RUN_STAGE15_TECHNICAL_PILOT,
      RUN_STAGE15_EDGE_BOOTSTRAPS
    )),
    REUSE_VALIDATED_FORMAL_OUTPUTS
  ),
  technical_pilot_group %in% expected_groups$Group_key,
  identical(names(point_seeds), expected_groups$Group_key),
  identical(names(bootstrap_seeds), expected_groups$Group_key),
  !anyDuplicated(point_seeds),
  !anyDuplicated(bootstrap_seeds)
)

run_control_table
```

# 6. Stage 14 point-network functions

```{r define-point-network-functions}
estimate_point_network <- function(ordinal_nodes, random_seed) {
  captured_warnings <- character(0)
  captured_messages <- character(0)
  captured_error <- character(0)
  network_object <- NULL

  set.seed(random_seed)

  elapsed <- system.time({
    network_object <- tryCatch(
      withCallingHandlers(
        bootnet::estimateNetwork(
          data = ordinal_nodes,
          default = "EBICglasso",
          tuning = 0.50,
          corMethod = "cor_auto",
          missing = "listwise",
          sampleSize = "minimum",
          corArgs = list(
            detectOrdinal = TRUE,
            ordinalLevelMax = 7,
            forcePD = FALSE
          ),
          refit = FALSE,
          lambda.min.ratio = 0.01,
          nlambda = 100,
          threshold = FALSE,
          nonPositiveDefinite = "stop",
          verbose = FALSE
        ),
        warning = function(condition) {
          captured_warnings <<- c(
            captured_warnings,
            conditionMessage(condition)
          )
          invokeRestart("muffleWarning")
        },
        message = function(condition) {
          captured_messages <<- c(
            captured_messages,
            conditionMessage(condition)
          )
          invokeRestart("muffleMessage")
        }
      ),
      error = function(condition) {
        captured_error <<- conditionMessage(condition)
        NULL
      }
    )
  })

  list(
    object = network_object,
    warnings = unique(captured_warnings),
    messages = unique(captured_messages),
    error = unique(captured_error),
    elapsed = elapsed,
    seed = random_seed
  )
}

estimate_reference_graph <- function(matrix_value, sample_n) {
  captured_warnings <- character(0)
  captured_error <- character(0)
  reference_graph <- NULL

  reference_graph <- tryCatch(
    withCallingHandlers(
      qgraph::EBICglasso(
        S = matrix_value,
        n = sample_n,
        gamma = 0.50,
        nlambda = 100,
        lambda.min.ratio = 0.01,
        refit = FALSE,
        threshold = FALSE
      ),
      warning = function(condition) {
        captured_warnings <<- c(
          captured_warnings,
          conditionMessage(condition)
        )
        invokeRestart("muffleWarning")
      }
    ),
    error = function(condition) {
      captured_error <<- conditionMessage(condition)
      NULL
    }
  )

  list(
    graph = reference_graph,
    warnings = unique(captured_warnings),
    error = unique(captured_error)
  )
}
```

# 7. Estimate and validate all eight point networks

```{r stage14-estimate-point-networks}
point_results <- list()
point_validation_list <- list()
point_diagnostics_list <- list()
edge_weight_list <- list()

if (RUN_STAGE14_POINT_NETWORKS) {
  for (group_key in expected_groups$Group_key) {
    specification <- expected_groups[
      expected_groups$Group_key == group_key,
      ,
      drop = FALSE
    ]
    ordinal_nodes <- samples[[group_key]]$Ordinal_nodes
    expected_n <- specification$Expected_N[[1L]]

    point_result <- estimate_point_network(
      ordinal_nodes = ordinal_nodes,
      random_seed = unname(point_seeds[[group_key]])
    )

    reference_result <- estimate_reference_graph(
      matrix_value = stage13_matrices[[group_key]],
      sample_n = expected_n
    )

    point_result$reference <- reference_result

    point_file <- file.path(
      formal_model_dir,
      paste0("08B_education_network_", group_key, "_gamma050.rds")
    )
    saveRDS(point_result, point_file)

    if (is.null(point_result$object) || is.null(reference_result$graph)) {
      failure_table <- data.frame(
        Group_key = group_key,
        Point_error = paste(point_result$error, collapse = " | "),
        Reference_error = paste(reference_result$error, collapse = " | "),
        stringsAsFactors = FALSE
      )
      write.csv(
        failure_table,
        file.path(
          formal_table_dir,
          paste0("08B_education_network_", group_key, "_failure.csv")
        ),
        row.names = FALSE
      )
      writeLines(
        c(
          paste0("Group key: ", group_key),
          paste0("Point seed: ", point_result$seed),
          paste0(
            "Elapsed seconds: ",
            unname(point_result$elapsed[["elapsed"]])
          ),
          paste0(
            "Point error: ",
            paste(point_result$error, collapse = " | ")
          ),
          paste0(
            "Reference error: ",
            paste(reference_result$error, collapse = " | ")
          )
        ),
        file.path(
          log_dir,
          paste0("08B_education_network_", group_key, ".log")
        )
      )
      stop(paste0("Point-network estimation failed for ", group_key, "."))
    }

    graph <- point_result$object$graph
    reference_graph <- reference_result$graph
    maximum_reference_difference <- max(abs(graph - reference_graph))

    point_warnings_known <-
      known_dense_warning(point_result$warnings) &&
        known_dense_warning(reference_result$warnings)

    validation <- data.frame(
      Group_key = group_key,
      Check = c(
        "Expected respondent count",
        "Expected node count",
        "Frozen node order",
        "Finite graph",
        "Symmetric graph",
        "Zero diagonal",
        "Matches saved Stage-13 polychoric input",
        "No unclassified point-estimator warning",
        "No point-estimator error"
      ),
      Observed = c(
        as.character(point_result$object$nPerson),
        as.character(point_result$object$nNode),
        paste(rownames(graph), collapse = "|"),
        as.character(all(is.finite(graph))),
        format(max(abs(graph - t(graph))), scientific = TRUE),
        format(max(abs(diag(graph))), scientific = TRUE),
        format(maximum_reference_difference, scientific = TRUE),
        as.character(point_warnings_known),
        paste(
          c(point_result$error, reference_result$error),
          collapse = " | "
        )
      ),
      Expected = c(
        as.character(expected_n),
        "9",
        paste(network_vars, collapse = "|"),
        "TRUE",
        "< 1e-12",
        "< 1e-12",
        "< 1e-8",
        "TRUE",
        ""
      ),
      Passed = c(
        identical(as.integer(point_result$object$nPerson), as.integer(expected_n)),
        identical(as.integer(point_result$object$nNode), 9L),
        identical(rownames(graph), network_vars) &&
          identical(colnames(graph), network_vars),
        all(is.finite(graph)),
        max(abs(graph - t(graph))) < 1e-12,
        max(abs(diag(graph))) < 1e-12,
        maximum_reference_difference < 1e-8,
        point_warnings_known,
        length(point_result$error) == 0L &&
          length(reference_result$error) == 0L
      ),
      stringsAsFactors = FALSE
    )

    upper_index <- upper.tri(graph, diag = FALSE)
    edge_weights <- data.frame(
      Group_key = group_key,
      Wave = specification$Wave[[1L]],
      Education_group = specification$Education_group[[1L]],
      N = expected_n,
      Node_1 = rownames(graph)[row(graph)[upper_index]],
      Node_2 = colnames(graph)[col(graph)[upper_index]],
      Edge_weight = graph[upper_index],
      Retained_under_EBICglasso = graph[upper_index] != 0,
      stringsAsFactors = FALSE
    )

    diagnostics <- data.frame(
      Group_key = group_key,
      Wave = specification$Wave[[1L]],
      Education_group = specification$Education_group[[1L]],
      N = expected_n,
      Nodes = 9L,
      Possible_edges = 36L,
      Retained_edges = sum(graph[upper_index] != 0),
      Global_strength_descriptive = sum(abs(graph[upper_index])),
      Maximum_reference_difference = maximum_reference_difference,
      Estimation_elapsed_seconds = unname(point_result$elapsed[["elapsed"]]),
      Point_warning_n = length(point_result$warnings),
      Point_warnings = paste(point_result$warnings, collapse = " | "),
      Reference_warning_n = length(reference_result$warnings),
      Reference_warnings = paste(reference_result$warnings, collapse = " | "),
      Message_n = length(point_result$messages),
      Messages = paste(point_result$messages, collapse = " | "),
      All_validation_checks_passed = all(validation$Passed),
      stringsAsFactors = FALSE
    )

    write.csv(
      validation,
      file.path(
        formal_table_dir,
        paste0("08B_education_network_", group_key, "_validation.csv")
      ),
      row.names = FALSE
    )
    write.csv(
      diagnostics,
      file.path(
        formal_table_dir,
        paste0("08B_education_network_", group_key, "_diagnostics.csv")
      ),
      row.names = FALSE
    )
    write.csv(
      edge_weights,
      file.path(
        formal_table_dir,
        paste0("08B_education_network_", group_key, "_edge_weights.csv")
      ),
      row.names = FALSE
    )
    writeLines(
      c(
        paste0("Group key: ", group_key),
        paste0("Point seed: ", point_result$seed),
        paste0(
          "Elapsed seconds: ",
          unname(point_result$elapsed[["elapsed"]])
        ),
        paste0(
          "Maximum reference-graph difference: ",
          maximum_reference_difference
        ),
        paste0(
          "Point-estimator warnings: ",
          if (length(point_result$warnings) == 0L) {
            "NONE"
          } else {
            paste(point_result$warnings, collapse = " | ")
          }
        ),
        paste0(
          "Reference-estimator warnings: ",
          if (length(reference_result$warnings) == 0L) {
            "NONE"
          } else {
            paste(reference_result$warnings, collapse = " | ")
          }
        ),
        paste0("All validation checks passed: ", all(validation$Passed))
      ),
      file.path(
        log_dir,
        paste0("08B_education_network_", group_key, ".log")
      )
    )

    point_results[[group_key]] <- point_result
    point_validation_list[[group_key]] <- validation
    point_diagnostics_list[[group_key]] <- diagnostics
    edge_weight_list[[group_key]] <- edge_weights

    if (!all(validation$Passed)) {
      stop(paste0(
        "Point-network validation failed for ",
        group_key,
        ". Inspect its saved validation and diagnostics files."
      ))
    }
  }
} else if (REUSE_VALIDATED_FORMAL_OUTPUTS) {
  point_model_file <- file.path(
    formal_model_dir,
    "08B_education_eight_networks_gamma050.rds"
  )
  point_validation_file <- file.path(
    formal_table_dir,
    "08B_education_point_network_validation.csv"
  )
  point_diagnostics_file <- file.path(
    formal_table_dir,
    "08B_education_point_network_diagnostics.csv"
  )
  edge_weights_file <- file.path(
    formal_table_dir,
    "08B_education_edge_weights.csv"
  )

  stopifnot(all(file.exists(c(
    point_model_file,
    point_validation_file,
    point_diagnostics_file,
    edge_weights_file
  ))))

  point_results <- readRDS(point_model_file)
  point_validation_list <- split(
    read.csv(point_validation_file, stringsAsFactors = FALSE),
    read.csv(point_validation_file, stringsAsFactors = FALSE)$Group_key
  )
  point_diagnostics_list <- split(
    read.csv(point_diagnostics_file, stringsAsFactors = FALSE),
    read.csv(point_diagnostics_file, stringsAsFactors = FALSE)$Group_key
  )
  edge_weight_list <- split(
    read.csv(edge_weights_file, stringsAsFactors = FALSE),
    read.csv(edge_weights_file, stringsAsFactors = FALSE)$Group_key
  )
}

point_validation <- do.call(rbind, point_validation_list)
point_diagnostics <- do.call(rbind, point_diagnostics_list)
edge_weights <- do.call(rbind, edge_weight_list)

if (RUN_STAGE14_POINT_NETWORKS) {
  write.csv(
    point_validation,
    file.path(formal_table_dir, "08B_education_point_network_validation.csv"),
    row.names = FALSE
  )
  write.csv(
    point_diagnostics,
    file.path(formal_table_dir, "08B_education_point_network_diagnostics.csv"),
    row.names = FALSE
  )
  write.csv(
    edge_weights,
    file.path(formal_table_dir, "08B_education_edge_weights.csv"),
    row.names = FALSE
  )
  saveRDS(
    point_results,
    file.path(formal_model_dir, "08B_education_eight_networks_gamma050.rds"),
    version = 3,
    compress = "gzip"
  )
}

stopifnot(
  length(point_results) == 8L,
  nrow(point_validation) == 72L,
  all(point_validation$Passed),
  nrow(point_diagnostics) == 8L,
  nrow(edge_weights) == 288L
)

point_diagnostics
```

# 8. Stage 15 edge-bootstrap functions

```{r define-edge-bootstrap-functions}
run_edge_bootstrap <- function(network_object, n_boots, random_seed) {
  captured_warnings <- character(0)
  captured_error <- character(0)
  bootstrap_object <- NULL

  set.seed(random_seed)

  elapsed_time <- system.time({
    bootstrap_object <- tryCatch(
      withCallingHandlers(
        bootnet::bootnet(
          data = network_object,
          nBoots = n_boots,
          type = "nonparametric",
          statistics = "edge",
          computeCentrality = FALSE,
          nCores = 1,
          memorysaver = TRUE,
          maxErrors = 10,
          verbose = TRUE
        ),
        warning = function(condition) {
          captured_warnings <<- c(
            captured_warnings,
            conditionMessage(condition)
          )
          invokeRestart("muffleWarning")
        }
      ),
      error = function(condition) {
        captured_error <<- conditionMessage(condition)
        NULL
      }
    )
  })

  list(
    object = bootstrap_object,
    nBoots = n_boots,
    seed = random_seed,
    elapsed_time = elapsed_time,
    bootnet_level_warnings = unique(captured_warnings),
    error = unique(captured_error),
    R_version = R.version.string,
    bootnet_version = as.character(utils::packageVersion("bootnet"))
  )
}

validate_edge_bootstrap <- function(
  result_object,
  expected_n_boots,
  expected_node_names,
  expected_n_person,
  reference_graph
) {
  node_n <- length(expected_node_names)
  edge_n <- choose(node_n, 2L)
  boot_object <- result_object$object

  if (is.null(boot_object)) {
    return(data.frame(
      Check = c(
        "Bootstrap object exists",
        "No bootstrap execution error"
      ),
      Observed = c("FALSE", paste(result_object$error, collapse = " | ")),
      Expected = c("TRUE", ""),
      Passed = c(FALSE, FALSE),
      stringsAsFactors = FALSE
    ))
  }

  graph_checks <- vapply(
    boot_object$boots,
    function(one_boot) {
      graph <- one_boot$graph
      is.matrix(graph) &&
        identical(dim(graph), c(node_n, node_n)) &&
        identical(rownames(graph), expected_node_names) &&
        identical(colnames(graph), expected_node_names) &&
        all(is.finite(graph)) &&
        max(abs(graph - t(graph))) < 1e-12 &&
        max(abs(diag(graph))) < 1e-12 &&
        identical(
          as.integer(one_boot$nPerson),
          as.integer(expected_n_person)
        )
    },
    FUN.VALUE = logical(1)
  )

  sample_graph_difference <- max(
    abs(boot_object$sample$graph - reference_graph)
  )

  warning_classified <- known_bootstrap_warning(
    result_object$bootnet_level_warnings
  )

  data.frame(
    Check = c(
      "Bootstrap object exists",
      "Requested bootstrap count recorded",
      "Returned bootstrap network count",
      "Bootstrap edge-table row count",
      "Distinct bootstrap identifiers",
      "Only edge statistics stored",
      "All bootstrap edge values finite",
      "Original-sample edge table complete",
      "All returned bootstrap graphs valid",
      "Bootstrap original sample matches point graph",
      "No unclassified bootstrap warning",
      "No bootstrap execution error"
    ),
    Observed = c(
      "TRUE",
      as.character(result_object$nBoots),
      as.character(length(boot_object$boots)),
      as.character(nrow(boot_object$bootTable)),
      as.character(length(unique(boot_object$bootTable$name))),
      paste(unique(boot_object$bootTable$type), collapse = "|"),
      as.character(sum(is.finite(boot_object$bootTable$value))),
      as.character(nrow(boot_object$sampleTable)),
      paste0(sum(graph_checks), "/", length(graph_checks)),
      format(sample_graph_difference, scientific = TRUE),
      as.character(warning_classified),
      paste(result_object$error, collapse = " | ")
    ),
    Expected = c(
      "TRUE",
      as.character(expected_n_boots),
      as.character(expected_n_boots),
      as.character(expected_n_boots * edge_n),
      as.character(expected_n_boots),
      "edge",
      as.character(expected_n_boots * edge_n),
      as.character(edge_n),
      paste0(expected_n_boots, "/", expected_n_boots),
      "< 1e-8",
      "TRUE",
      ""
    ),
    Passed = c(
      TRUE,
      identical(as.integer(result_object$nBoots), as.integer(expected_n_boots)),
      length(boot_object$boots) == expected_n_boots,
      nrow(boot_object$bootTable) == expected_n_boots * edge_n,
      length(unique(boot_object$bootTable$name)) == expected_n_boots,
      identical(unique(boot_object$bootTable$type), "edge"),
      all(is.finite(boot_object$bootTable$value)),
      nrow(boot_object$sampleTable) == edge_n,
      all(graph_checks),
      sample_graph_difference < 1e-8,
      warning_classified,
      length(result_object$error) == 0L
    ),
    stringsAsFactors = FALSE
  )
}

bootstrap_diagnostics <- function(
  group_key,
  wave,
  education_group,
  result_object,
  validation
) {
  data.frame(
    Group_key = group_key,
    Wave = wave,
    Education_group = education_group,
    Requested_bootstraps = result_object$nBoots,
    Returned_bootstraps = if (is.null(result_object$object)) {
      0L
    } else {
      length(result_object$object$boots)
    },
    Seed = result_object$seed,
    Elapsed_seconds = unname(result_object$elapsed_time[["elapsed"]]),
    Top_level_retry_count = count_retry_warnings(
      result_object$bootnet_level_warnings
    ),
    Captured_warning_n = length(result_object$bootnet_level_warnings),
    Captured_warnings = paste(
      result_object$bootnet_level_warnings,
      collapse = " | "
    ),
    Error_n = length(result_object$error),
    Error = paste(result_object$error, collapse = " | "),
    All_validation_checks_passed = all(validation$Passed),
    stringsAsFactors = FALSE
  )
}

write_bootstrap_log <- function(result_object, output_path) {
  warning_text <- if (length(result_object$bootnet_level_warnings) == 0L) {
    "NO BOOTNET-LEVEL WARNINGS CAPTURED"
  } else {
    paste(result_object$bootnet_level_warnings, collapse = "\n")
  }

  error_text <- if (length(result_object$error) == 0L) {
    "NO BOOTSTRAP EXECUTION ERROR"
  } else {
    paste(result_object$error, collapse = "\n")
  }

  writeLines(
    c(
      paste0("Requested bootstrap repetitions: ", result_object$nBoots),
      paste0("Random seed: ", result_object$seed),
      paste0("bootnet version: ", result_object$bootnet_version),
      paste0(
        "Elapsed seconds: ",
        unname(result_object$elapsed_time[["elapsed"]])
      ),
      "",
      "Captured bootnet-level warnings:",
      warning_text,
      "",
      "Execution error:",
      error_text
    ),
    con = output_path
  )
}
```

# 9. Ten-repetition technical pilot

The technical pilot uses the prespecified smallest and most sparse retained
sample, 2016 college or advanced degree. It tests code execution only. Its
edge estimates must not be reported or used to change the frozen parameters.

```{r stage15-technical-pilot}
if (RUN_STAGE15_TECHNICAL_PILOT) {
  pilot_network <- point_results[[technical_pilot_group]]$object
  pilot_specification <- expected_groups[
    expected_groups$Group_key == technical_pilot_group,
    ,
    drop = FALSE
  ]

  pilot_result <- run_edge_bootstrap(
    network_object = pilot_network,
    n_boots = 10L,
    random_seed = technical_pilot_seed
  )

  pilot_validation <- validate_edge_bootstrap(
    result_object = pilot_result,
    expected_n_boots = 10L,
    expected_node_names = network_vars,
    expected_n_person = pilot_specification$Expected_N[[1L]],
    reference_graph = pilot_network$graph
  )

  pilot_validation$Group_key <- technical_pilot_group
  pilot_validation <- pilot_validation[
    ,
    c("Group_key", "Check", "Observed", "Expected", "Passed")
  ]

  pilot_diagnostics <- bootstrap_diagnostics(
    group_key = technical_pilot_group,
    wave = pilot_specification$Wave[[1L]],
    education_group = pilot_specification$Education_group[[1L]],
    result_object = pilot_result,
    validation = pilot_validation
  )

  saveRDS(
    pilot_result,
    file.path(
      pilot_model_dir,
      "08B_education_2016_college_edge_bootstrap_10_technical.rds"
    )
  )
  write.csv(
    pilot_validation,
    file.path(
      pilot_table_dir,
      "08B_education_2016_college_edge_bootstrap_10_validation.csv"
    ),
    row.names = FALSE
  )
  write.csv(
    pilot_diagnostics,
    file.path(
      pilot_table_dir,
      "08B_education_2016_college_edge_bootstrap_10_diagnostics.csv"
    ),
    row.names = FALSE
  )
  write_bootstrap_log(
    pilot_result,
    file.path(log_dir, "08B_education_2016_college_bootstrap_10.log")
  )

  stopifnot(all(pilot_validation$Passed))
} else if (REUSE_VALIDATED_FORMAL_OUTPUTS) {
  pilot_validation <- read.csv(
    file.path(
      pilot_table_dir,
      "08B_education_2016_college_edge_bootstrap_10_validation.csv"
    ),
    stringsAsFactors = FALSE
  )
  pilot_diagnostics <- read.csv(
    file.path(
      pilot_table_dir,
      "08B_education_2016_college_edge_bootstrap_10_diagnostics.csv"
    ),
    stringsAsFactors = FALSE
  )

  stopifnot(all(as.logical(pilot_validation$Passed)))
}

pilot_diagnostics
```

# 10. Eight 1,000-repetition edge-weight bootstraps

```{r stage15-run-edge-bootstraps}
bootstrap_results <- list()
bootstrap_validation_list <- list()
bootstrap_diagnostics_list <- list()
bootstrap_summary_list <- list()

if (RUN_STAGE15_EDGE_BOOTSTRAPS) {
  for (group_key in expected_groups$Group_key) {
    specification <- expected_groups[
      expected_groups$Group_key == group_key,
      ,
      drop = FALSE
    ]
    point_network <- point_results[[group_key]]$object

    result <- run_edge_bootstrap(
      network_object = point_network,
      n_boots = 1000L,
      random_seed = unname(bootstrap_seeds[[group_key]])
    )

    validation <- validate_edge_bootstrap(
      result_object = result,
      expected_n_boots = 1000L,
      expected_node_names = network_vars,
      expected_n_person = specification$Expected_N[[1L]],
      reference_graph = point_network$graph
    )
    validation$Group_key <- group_key
    validation <- validation[
      ,
      c("Group_key", "Check", "Observed", "Expected", "Passed")
    ]

    diagnostics <- bootstrap_diagnostics(
      group_key = group_key,
      wave = specification$Wave[[1L]],
      education_group = specification$Education_group[[1L]],
      result_object = result,
      validation = validation
    )

    model_path <- file.path(
      formal_model_dir,
      paste0("08B_education_edge_bootstrap_", group_key, "_1000.rds")
    )
    validation_path <- file.path(
      formal_table_dir,
      paste0("08B_education_edge_bootstrap_", group_key, "_validation.csv")
    )
    diagnostics_path <- file.path(
      formal_table_dir,
      paste0("08B_education_edge_bootstrap_", group_key, "_diagnostics.csv")
    )
    summary_path <- file.path(
      formal_table_dir,
      paste0("08B_education_edge_bootstrap_", group_key, "_summary.csv")
    )
    log_path <- file.path(
      log_dir,
      paste0("08B_education_edge_bootstrap_", group_key, "_1000.log")
    )

    saveRDS(result, model_path)
    write.csv(validation, validation_path, row.names = FALSE)
    write.csv(diagnostics, diagnostics_path, row.names = FALSE)
    write_bootstrap_log(result, log_path)

    if (!all(validation$Passed)) {
      stop(paste0(
        "Edge-bootstrap validation failed for ",
        group_key,
        ". Inspect its saved RDS, validation, diagnostics and log files."
      ))
    }

    summary_table <- as.data.frame(
      summary(result$object, statistics = "edge")
    )
    summary_table$Group_key <- group_key
    summary_table$Wave <- specification$Wave[[1L]]
    summary_table$Education_group <-
      specification$Education_group[[1L]]

    stopifnot(nrow(summary_table) == 36L)
    write.csv(summary_table, summary_path, row.names = FALSE)

    bootstrap_results[[group_key]] <- result
    bootstrap_validation_list[[group_key]] <- validation
    bootstrap_diagnostics_list[[group_key]] <- diagnostics
    bootstrap_summary_list[[group_key]] <- summary_table
  }
} else if (REUSE_VALIDATED_FORMAL_OUTPUTS) {
  bootstrap_model_file <- file.path(
    formal_model_dir,
    "08B_education_eight_edge_bootstraps_1000.rds"
  )
  bootstrap_validation_file <- file.path(
    formal_table_dir,
    "08B_education_edge_bootstrap_validation.csv"
  )
  bootstrap_diagnostics_file <- file.path(
    formal_table_dir,
    "08B_education_edge_bootstrap_diagnostics.csv"
  )
  bootstrap_summary_file <- file.path(
    formal_table_dir,
    "08B_education_edge_bootstrap_summary.csv"
  )

  stopifnot(all(file.exists(c(
    bootstrap_model_file,
    bootstrap_validation_file,
    bootstrap_diagnostics_file,
    bootstrap_summary_file
  ))))

  bootstrap_results <- readRDS(bootstrap_model_file)
  bootstrap_validation_saved <- read.csv(
    bootstrap_validation_file,
    stringsAsFactors = FALSE
  )
  bootstrap_diagnostics_saved <- read.csv(
    bootstrap_diagnostics_file,
    stringsAsFactors = FALSE
  )
  bootstrap_summary_saved <- read.csv(
    bootstrap_summary_file,
    stringsAsFactors = FALSE
  )
  bootstrap_validation_list <- split(
    bootstrap_validation_saved,
    bootstrap_validation_saved$Group_key
  )
  bootstrap_diagnostics_list <- split(
    bootstrap_diagnostics_saved,
    bootstrap_diagnostics_saved$Group_key
  )
  bootstrap_summary_list <- split(
    bootstrap_summary_saved,
    bootstrap_summary_saved$Group_key
  )
}

bootstrap_validation <- do.call(rbind, bootstrap_validation_list)
bootstrap_diagnostics_table <- do.call(rbind, bootstrap_diagnostics_list)
bootstrap_summary <- do.call(rbind, bootstrap_summary_list)

if (RUN_STAGE15_EDGE_BOOTSTRAPS) {
  write.csv(
    bootstrap_validation,
    file.path(
      formal_table_dir,
      "08B_education_edge_bootstrap_validation.csv"
    ),
    row.names = FALSE
  )
  write.csv(
    bootstrap_diagnostics_table,
    file.path(
      formal_table_dir,
      "08B_education_edge_bootstrap_diagnostics.csv"
    ),
    row.names = FALSE
  )
  write.csv(
    bootstrap_summary,
    file.path(
      formal_table_dir,
      "08B_education_edge_bootstrap_summary.csv"
    ),
    row.names = FALSE
  )
  saveRDS(
    bootstrap_results,
    file.path(
      formal_model_dir,
      "08B_education_eight_edge_bootstraps_1000.rds"
    ),
    version = 3,
    compress = "gzip"
  )
}

stopifnot(
  length(bootstrap_results) == 8L,
  nrow(bootstrap_validation) == 96L,
  all(bootstrap_validation$Passed),
  nrow(bootstrap_diagnostics_table) == 8L,
  nrow(bootstrap_summary) == 288L,
  all(bootstrap_diagnostics_table$Returned_bootstraps == 1000L)
)

bootstrap_diagnostics_table
```

# 11. Prohibited-call audit

```{r prohibited-call-audit}
source_lines <- readLines(notebook_file, warn = FALSE)

prohibited_call_patterns <- c(
  paste0("NetworkComparisonTest::", "NCT("),
  paste0("bootnet::", "corStability("),
  paste0("networktools::", "bridge("),
  paste0("qgraph::", "centrality_auto("),
  paste0("qgraph::", "centralityPlot(")
)

prohibited_call_scan <- data.frame(
  Pattern = prohibited_call_patterns,
  Found = vapply(
    prohibited_call_patterns,
    function(pattern) {
      any(grepl(pattern, source_lines, fixed = TRUE))
    },
    FUN.VALUE = logical(1)
  ),
  stringsAsFactors = FALSE
)

write.csv(
  prohibited_call_scan,
  file.path(formal_table_dir, "08B_prohibited_call_scan.csv"),
  row.names = FALSE
)

prohibited_call_scan
stopifnot(!any(prohibited_call_scan$Found))
```

# 12. Completion status, session information and output hashes

```{r completion-session-hash-inventory}
completion_status <- data.frame(
  Component = c(
    "Governance and package gates",
    "Stage-12 and Stage-13 source hashes",
    "Eight ordered complete-case inputs",
    "Eight gamma-0.50 point networks",
    "Ten-repetition technical pilot",
    "Eight 1,000-repetition edge bootstraps",
    "Point-network structural validation",
    "Bootstrap structural validation",
    "NCT",
    "Centrality",
    "Sensitivity analysis",
    "Political interpretation"
  ),
  Status = c(
    rep("COMPLETE_AND_VALIDATED", 8L),
    "NOT_IN_CURRENT_SCOPE_NOT_RUN",
    "NOT_AUTHORISED_NOT_RUN",
    "NOT_AUTHORISED_NOT_RUN",
    "NOT_AUTHORISED_NOT_RUN"
  ),
  Required_for_D08A = c(rep(TRUE, 8L), rep(FALSE, 4L)),
  Passed = c(
    all(decision_gate$Passed) &&
      all(installed_version_gate$Passed) &&
      all(lock_version_gate$Passed),
    all(source_gate$Passed) && all(bundle_gate$Passed),
    all(input_validation$Passed),
    length(point_results) == 8L,
    all(pilot_validation$Passed),
    length(bootstrap_results) == 8L,
    all(point_validation$Passed),
    all(bootstrap_validation$Passed),
    TRUE,
    TRUE,
    TRUE,
    TRUE
  ),
  stringsAsFactors = FALSE
)

write.csv(
  completion_status,
  file.path(formal_table_dir, "08B_education_completion_status.csv"),
  row.names = FALSE
)

stopifnot(
  all(
    completion_status$Passed[
      completion_status$Required_for_D08A
    ]
  )
)

session_file <- file.path(log_dir, "08B_education_sessionInfo.txt")
writeLines(capture.output(sessionInfo()), session_file)

execution_log_file <- file.path(log_dir, "08B_education_execution.log")
writeLines(
  c(
    paste0(
      "Execution mode: ",
      if (REUSE_VALIDATED_FORMAL_OUTPUTS) {
        "REUSE_VALIDATED_OUTPUTS"
      } else {
        "RECOMPUTE_MODELS_AND_BOOTSTRAPS"
      }
    ),
    paste0("Registry: ", normalizePath(registry_path, winslash = "/")),
    paste0("Registry SHA-256: ", file_sha256(registry_path)),
    paste0("Stage-12 bundle SHA-256: ", file_sha256(education_bundle_file)),
    paste0("Stage-13 bundle SHA-256: ", file_sha256(stage13_bundle_file)),
    paste0("Point networks returned: ", length(point_results), "/8"),
    paste0(
      "Formal bootstrap networks returned: ",
      sum(bootstrap_diagnostics_table$Returned_bootstraps),
      "/8000"
    ),
    paste0(
      "Top-level retry count: ",
      sum(bootstrap_diagnostics_table$Top_level_retry_count)
    ),
    paste0(
      "All required completion checks passed: ",
      all(
        completion_status$Passed[
          completion_status$Required_for_D08A
        ]
      )
    ),
    "NCT: NOT RUN",
    "Centrality: NOT RUN",
    "Sensitivity analysis: NOT RUN",
    "Political interpretation: NOT RUN"
  ),
  execution_log_file
)

output_candidates <- c(
  list.files(
    formal_model_dir,
    pattern = "^08B_",
    full.names = TRUE
  ),
  list.files(
    formal_table_dir,
    pattern = "^08B_",
    full.names = TRUE
  ),
  list.files(
    pilot_model_dir,
    pattern = "^08B_",
    full.names = TRUE
  ),
  list.files(
    pilot_table_dir,
    pattern = "^08B_",
    full.names = TRUE
  ),
  list.files(
    log_dir,
    pattern = "^08B_",
    full.names = TRUE
  )
)
output_candidates <- sort(unique(output_candidates))

output_inventory_path <- file.path(
  formal_table_dir,
  "08B_output_hash_inventory.csv"
)
output_candidates <- setdiff(
  normalizePath(output_candidates, winslash = "/", mustWork = FALSE),
  normalizePath(output_inventory_path, winslash = "/", mustWork = FALSE)
)

output_inventory <- data.frame(
  File = basename(output_candidates),
  Path = normalizePath(
    output_candidates,
    winslash = "/",
    mustWork = FALSE
  ),
  Bytes = file.info(output_candidates)$size,
  SHA256 = vapply(
    output_candidates,
    file_sha256,
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

write.csv(
  output_inventory,
  output_inventory_path,
  row.names = FALSE
)

stopifnot(
  nrow(output_inventory) > 0L,
  all(output_inventory$Bytes > 0),
  all(nzchar(output_inventory$SHA256))
)

completion_status
output_inventory
```

# Completion status

If the notebook reaches this section, all eight descriptive education-group
point networks and all eight 1,000-repetition nonparametric edge-weight
bootstraps have passed the prespecified structural checks. The outputs support
descriptive reporting; this notebook performs no formal between-group test.
