---
title: "08C Education Result Readiness"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: show
---

# Scope and interpretation boundary

This notebook organises the completed D08A education-network outputs for
result-readiness review. It reads the eight validated point networks and their
existing 1,000-repetition nonparametric edge-bootstrap summaries. It does not
estimate or re-estimate a network, run a bootstrap, calculate centrality, run
NCT, test education-group differences, or make political interpretations.

The primary accuracy interval used below is the 95% bootstrap percentile
interval (`q2.5` to `q97.5`). Bootstrap intervals are used to describe
estimation accuracy through their width. Whether an interval includes zero is
retained as audit information only and is not used as a significance test, a
second edge-selection rule, or a reporting gate. An edge retained under the
frozen EBICglasso procedure may be described cautiously within its single
network when its point estimate, interval and interval width are reported
together. Wider intervals require greater interpretive caution. A regularised
zero means only that the edge was not retained under the selected procedure.

The five widest intervals per group are listed as review priorities. Their rank
is descriptive and is not a universal instability threshold.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = TRUE,
  message = TRUE,
  error = FALSE
)
```

# 1. Project paths, packages and helpers

```{r project-paths-packages-helpers}
current_dir <- normalizePath(".", winslash = "/")

if (file.exists(file.path(current_dir, "ANES.Rproj"))) {
  project_root <- current_dir
} else if (
  basename(current_dir) == "notebooks" &&
    file.exists(file.path(dirname(current_dir), "ANES.Rproj"))
) {
  project_root <- dirname(current_dir)
} else {
  stop("Project root could not be identified.")
}

knitr::opts_knit$set(root.dir = project_root)

notebook_file <- file.path(
  project_root,
  "notebooks",
  "08C_education_result_readiness.Rmd"
)

formal_model_dir <- file.path(project_root, "outputs", "models", "formal")
formal_table_dir <- file.path(project_root, "outputs", "tables", "formal")
formal_figure_dir <- file.path(project_root, "outputs", "figures", "formal")
log_dir <- file.path(project_root, "outputs", "logs")

for (target_dir in c(formal_table_dir, formal_figure_dir, log_dir)) {
  dir.create(target_dir, recursive = TRUE, showWarnings = FALSE)
}

required_packages <- c("digest", "knitr", "rmarkdown")

package_gate <- data.frame(
  Package = required_packages,
  Installed = vapply(
    required_packages,
    requireNamespace,
    quietly = TRUE,
    FUN.VALUE = logical(1)
  ),
  Version = vapply(
    required_packages,
    function(package_name) {
      if (requireNamespace(package_name, quietly = TRUE)) {
        as.character(utils::packageVersion(package_name))
      } else {
        NA_character_
      }
    },
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

stopifnot(all(package_gate$Installed))

normalise_status <- function(value) {
  value <- toupper(trimws(as.character(value)))
  gsub("[^A-Z0-9]+", "_", value)
}

file_sha256 <- function(path) {
  if (!file.exists(path)) {
    return(NA_character_)
  }

  toupper(
    digest::digest(
      file = path,
      algo = "sha256",
      serialize = FALSE
    )
  )
}

safe_message <- function(value) {
  value <- as.character(value)
  value[is.na(value) | !nzchar(trimws(value))] <- "NONE"
  value
}

write.csv(
  package_gate,
  file.path(formal_table_dir, "08C_package_gate.csv"),
  row.names = FALSE
)

package_gate
```

# 2. Governance gate

```{r governance-gate}
registry_config <- file.path(
  project_root,
  "config",
  "decision_registry_path.txt"
)

stopifnot(file.exists(registry_config))

registry_lines <- trimws(readLines(registry_config, warn = FALSE))
registry_lines <- registry_lines[nzchar(registry_lines)]

stopifnot(
  length(registry_lines) == 1L,
  file.exists(registry_lines[[1L]])
)

registry_path <- registry_lines[[1L]]
decision_registry <- read.csv(
  registry_path,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

decision_row <- function(decision_id) {
  decision_registry[
    decision_registry$decision_id == decision_id,
    ,
    drop = FALSE
  ]
}

d08a <- decision_row("D08A")
d08b <- decision_row("D08B")
d22 <- decision_row("D22")
d26 <- decision_row("D26")

decision_gate <- data.frame(
  Decision_ID = c("D08A", "D08B", "D22", "D26"),
  Check = c(
    "Descriptive education-network module is frozen and complete",
    "Formal education-group inference remains outside current scope",
    "One thousand nonparametric edge bootstraps are frozen and complete",
    "Eight education networks and edge bootstraps are complete and validated"
  ),
  Passed = c(
    nrow(d08a) == 1L &&
      normalise_status(d08a$decision_status) == "FROZEN" &&
      normalise_status(d08a$implementation_status) ==
        "COMPLETE_AND_VALIDATED",
    nrow(d08b) == 1L &&
      normalise_status(d08b$decision_status) ==
        "NOT_IN_CURRENT_SCOPE" &&
      normalise_status(d08b$implementation_status) == "NOT_APPLICABLE",
    nrow(d22) == 1L &&
      normalise_status(d22$decision_status) == "FROZEN" &&
      normalise_status(d22$implementation_status) ==
        "COMPLETE_AND_VALIDATED",
    nrow(d26) == 1L &&
      normalise_status(d26$decision_status) == "FROZEN" &&
      normalise_status(d26$implementation_status) ==
        "COMPLETE_AND_VALIDATED"
  ),
  stringsAsFactors = FALSE
)

write.csv(
  decision_gate,
  file.path(formal_table_dir, "08C_decision_gate.csv"),
  row.names = FALSE
)

decision_gate
stopifnot(all(decision_gate$Passed))
```

# 3. Read-only source and hash gate

```{r source-and-hash-gate}
source_paths <- c(
  point_models = file.path(
    formal_model_dir,
    "08B_education_eight_networks_gamma050.rds"
  ),
  bootstrap_models = file.path(
    formal_model_dir,
    "08B_education_eight_edge_bootstraps_1000.rds"
  ),
  point_diagnostics = file.path(
    formal_table_dir,
    "08B_education_point_network_diagnostics.csv"
  ),
  edge_weights = file.path(
    formal_table_dir,
    "08B_education_edge_weights.csv"
  ),
  bootstrap_summary = file.path(
    formal_table_dir,
    "08B_education_edge_bootstrap_summary.csv"
  ),
  bootstrap_diagnostics = file.path(
    formal_table_dir,
    "08B_education_edge_bootstrap_diagnostics.csv"
  ),
  completion_status = file.path(
    formal_table_dir,
    "08B_education_completion_status.csv"
  ),
  stage14_readiness = file.path(
    formal_table_dir,
    "08A_stage14_readiness.csv"
  )
)

expected_source_hashes <- c(
  point_models =
    "C0516E5248C907D832BF17DF8896D4F96EB7A7E12A1F4FE0EBE5C6817E05D9E9",
  bootstrap_models =
    "A1F4C09D39F1339FF5CE95A4B55A08BF8BD2781F91C0F72062BB2861C160BFED",
  point_diagnostics =
    "396DF68A713A7CF70D09A13A1420782FEEAF9C8D85E0C192516A95C67AAC9386",
  edge_weights =
    "094D1F9916196A898BB5322E68DD44FF7558513A9EB8A8016FEA49535AFD1BB7",
  bootstrap_summary =
    "3BE4F9A3F302C2B2682A922C68BDDD2F2DF41CFBDE36BB99B62EC66612959F2A",
  bootstrap_diagnostics =
    "B14718A9D4999518A2750FD13DBECD7B719F24D857C892741575B678E4CF789C",
  completion_status =
    "6A52971C7504A1AF043D1D84074215BDAEA981CFDAD91F1A544E13FD339CD2BE",
  stage14_readiness =
    "0506FC088C0439F6030785F9019899E5A38AD8AA0E1E2D09345A40CF1A460BF2"
)

source_gate <- data.frame(
  Input = names(source_paths),
  Path = normalizePath(source_paths, winslash = "/", mustWork = FALSE),
  Exists = file.exists(source_paths),
  Expected_SHA256 = unname(expected_source_hashes[names(source_paths)]),
  Observed_SHA256 = vapply(
    source_paths,
    file_sha256,
    FUN.VALUE = character(1)
  ),
  stringsAsFactors = FALSE
)

source_gate$Passed <-
  source_gate$Exists &
    source_gate$Observed_SHA256 == source_gate$Expected_SHA256

write.csv(
  source_gate,
  file.path(formal_table_dir, "08C_source_hash_gate.csv"),
  row.names = FALSE
)

source_gate
stopifnot(all(source_gate$Passed))
```

# 4. Read the completed D08A results

```{r read-existing-results}
point_models <- readRDS(source_paths[["point_models"]])
bootstrap_models <- readRDS(source_paths[["bootstrap_models"]])

point_diagnostics <- read.csv(
  source_paths[["point_diagnostics"]],
  stringsAsFactors = FALSE,
  check.names = FALSE
)

edge_weights <- read.csv(
  source_paths[["edge_weights"]],
  stringsAsFactors = FALSE,
  check.names = FALSE
)

bootstrap_summary <- read.csv(
  source_paths[["bootstrap_summary"]],
  stringsAsFactors = FALSE,
  check.names = FALSE
)

bootstrap_diagnostics <- read.csv(
  source_paths[["bootstrap_diagnostics"]],
  stringsAsFactors = FALSE,
  check.names = FALSE
)

completion_status <- read.csv(
  source_paths[["completion_status"]],
  stringsAsFactors = FALSE,
  check.names = FALSE
)

stage14_readiness <- read.csv(
  source_paths[["stage14_readiness"]],
  stringsAsFactors = FALSE,
  check.names = FALSE
)

network_vars <- c(
  "VCF0806",
  "VCF0809",
  "VCF0838",
  "VCF0839",
  "VCF0879a",
  "VCF0888",
  "VCF0890",
  "VCF0894",
  "VCF9223"
)

node_labels <- c(
  VCF0806 = "Health",
  VCF0809 = "Jobs",
  VCF0838 = "Abortion",
  VCF0839 = "Services",
  VCF0879a = "Immig. level",
  VCF0888 = "Crime spend",
  VCF0890 = "School spend",
  VCF0894 = "Welfare",
  VCF9223 = "Immig. jobs"
)

expected_groups <- data.frame(
  Group_key = c(
    "2012__no_college_degree",
    "2012__college_or_advanced",
    "2016__no_college_degree",
    "2016__college_or_advanced",
    "2020__no_college_degree",
    "2020__college_or_advanced",
    "2024__no_college_degree",
    "2024__college_or_advanced"
  ),
  Wave = c(2012L, 2012L, 2016L, 2016L, 2020L, 2020L, 2024L, 2024L),
  Education_group = rep(
    c("No college degree", "College or advanced degree"),
    times = 4L
  ),
  Expected_N = c(2773L, 1531L, 1478L, 1216L, 2624L, 2691L, 1729L, 1781L),
  stringsAsFactors = FALSE
)
```

# 5. Structural validation and prohibited-call audit

```{r structural-validation}
required_bootstrap_columns <- c(
  "type", "id", "node1", "node2", "sample", "mean", "sd",
  "CIlower", "CIupper", "q2.5", "q97.5", "prop0",
  "Group_key", "Wave", "Education_group"
)

point_graphs <- lapply(
  point_models,
  function(value) value$object$graph
)

point_graph_valid <- vapply(
  point_graphs,
  function(graph) {
    is.matrix(graph) &&
      identical(dim(graph), c(9L, 9L)) &&
      identical(rownames(graph), network_vars) &&
      identical(colnames(graph), network_vars) &&
      all(is.finite(graph)) &&
      max(abs(graph - t(graph))) < 1e-12 &&
      max(abs(diag(graph))) < 1e-12
  },
  FUN.VALUE = logical(1)
)

bootstrap_counts <- vapply(
  bootstrap_models,
  function(value) length(value$object$boots),
  FUN.VALUE = integer(1)
)

bootstrap_graph_valid <- vapply(
  bootstrap_models,
  function(value) {
    all(vapply(
      value$object$boots,
      function(boot_value) {
        graph <- boot_value$graph
        is.matrix(graph) &&
          identical(dim(graph), c(9L, 9L)) &&
          all(is.finite(graph)) &&
          max(abs(graph - t(graph))) < 1e-12 &&
          max(abs(diag(graph))) < 1e-12
      },
      FUN.VALUE = logical(1)
    ))
  },
  FUN.VALUE = logical(1)
)

structural_validation <- data.frame(
  Check = c(
    "Eight point-model group keys are exact and ordered",
    "Eight bootstrap-model group keys are exact and ordered",
    "Point diagnostics contain eight ordered groups",
    "Edge-weight table contains 8 x 36 rows",
    "Bootstrap summary contains 8 x 36 rows",
    "Bootstrap diagnostics contain eight ordered groups",
    "Stage-14 readiness contains eight ordered groups",
    "Bootstrap summary contains required columns",
    "Every point graph is a valid frozen 9 x 9 graph",
    "Every stored bootstrap object contains 1,000 graphs",
    "Every stored bootstrap graph is structurally valid",
    "Every bootstrap diagnostic reports 1,000 returned networks",
    "Every bootstrap diagnostic passed validation",
    "Every Stage-14 input was ready with sparse-cell status recorded",
    "All required 08B completion rows passed"
  ),
  Passed = c(
    identical(names(point_models), expected_groups$Group_key),
    identical(names(bootstrap_models), expected_groups$Group_key),
    identical(point_diagnostics$Group_key, expected_groups$Group_key),
    nrow(edge_weights) == 288L,
    nrow(bootstrap_summary) == 288L,
    identical(bootstrap_diagnostics$Group_key, expected_groups$Group_key),
    identical(stage14_readiness$Group_key, expected_groups$Group_key),
    all(required_bootstrap_columns %in% names(bootstrap_summary)),
    all(point_graph_valid),
    identical(unname(bootstrap_counts), rep(1000L, 8L)),
    all(bootstrap_graph_valid),
    all(bootstrap_diagnostics$Returned_bootstraps == 1000L),
    all(bootstrap_diagnostics$All_validation_checks_passed),
    all(stage14_readiness$Ready_for_stage14) &&
      all(grepl("SPARSE_CELL_WARNING", stage14_readiness$Sparse_cell_status)),
    all(
      completion_status$Passed[
        completion_status$Required_for_D08A
      ]
    )
  ),
  stringsAsFactors = FALSE
)

write.csv(
  structural_validation,
  file.path(formal_table_dir, "08C_input_validation.csv"),
  row.names = FALSE
)

structural_validation
stopifnot(all(structural_validation$Passed))
```

```{r prohibited-call-audit}
code_lines <- readLines(notebook_file, warn = FALSE)

prohibited_patterns <- c(
  paste0("bootnet::", "estimateNetwork("),
  paste0("qgraph::", "EBICglasso("),
  paste0("bootnet::", "bootnet("),
  paste0("NetworkComparisonTest::", "NCT("),
  paste0("bootnet::", "corStability("),
  paste0("qgraph::", "centrality_auto("),
  paste0("stats::", "t.test("),
  paste0("stats::", "wilcox.test(")
)

prohibited_call_scan <- data.frame(
  Pattern = prohibited_patterns,
  Found = vapply(
    prohibited_patterns,
    function(pattern) any(grepl(pattern, code_lines, fixed = TRUE)),
    FUN.VALUE = logical(1)
  ),
  stringsAsFactors = FALSE
)

write.csv(
  prohibited_call_scan,
  file.path(formal_table_dir, "08C_prohibited_call_scan.csv"),
  row.names = FALSE
)

prohibited_call_scan
stopifnot(!any(prohibited_call_scan$Found))
```

# 6. Build the edge-accuracy review table

```{r build-edge-accuracy-review}
canonicalise_edges <- function(data, node1_name, node2_name) {
  node1 <- as.character(data[[node1_name]])
  node2 <- as.character(data[[node2_name]])
  index1 <- match(node1, network_vars)
  index2 <- match(node2, network_vars)

  if (anyNA(index1) || anyNA(index2) || any(index1 == index2)) {
    stop("An edge contains an unexpected node name or a self-loop.")
  }

  first_node <- ifelse(index1 < index2, node1, node2)
  second_node <- ifelse(index1 < index2, node2, node1)

  data$Node_1 <- first_node
  data$Node_2 <- second_node
  data$Edge_ID <- paste(first_node, second_node, sep = "--")
  data
}

point_edges <- canonicalise_edges(edge_weights, "Node_1", "Node_2")
bootstrap_edges <- canonicalise_edges(bootstrap_summary, "node1", "node2")

review_joined <- merge(
  point_edges,
  bootstrap_edges,
  by = c(
    "Group_key",
    "Wave",
    "Education_group",
    "Node_1",
    "Node_2",
    "Edge_ID"
  ),
  all = FALSE,
  sort = FALSE,
  suffixes = c("_point", "_bootstrap")
)

edge_accuracy_review <- data.frame(
  Group_key = review_joined$Group_key,
  Wave = review_joined$Wave,
  Education_group = review_joined$Education_group,
  N = review_joined$N,
  Edge_ID = review_joined$Edge_ID,
  Node_1 = review_joined$Node_1,
  Node_2 = review_joined$Node_2,
  Edge_label = paste(
    unname(node_labels[review_joined$Node_1]),
    unname(node_labels[review_joined$Node_2]),
    sep = " -- "
  ),
  Point_edge = review_joined$Edge_weight,
  Retained_under_EBICglasso =
    review_joined$Retained_under_EBICglasso,
  Bootstrap_mean = review_joined$mean,
  Bootstrap_sd = review_joined$sd,
  Reported_CI_lower_sample_plus_or_minus_2SD = review_joined$CIlower,
  Reported_CI_upper_sample_plus_or_minus_2SD = review_joined$CIupper,
  Percentile_CI_2_5 = review_joined[["q2.5"]],
  Percentile_CI_97_5 = review_joined[["q97.5"]],
  Bootstrap_zero_proportion = review_joined$prop0,
  stringsAsFactors = FALSE
)

edge_accuracy_review$Reported_CI_width <-
  edge_accuracy_review$Reported_CI_upper_sample_plus_or_minus_2SD -
    edge_accuracy_review$Reported_CI_lower_sample_plus_or_minus_2SD

edge_accuracy_review$Percentile_CI_width <-
  edge_accuracy_review$Percentile_CI_97_5 -
    edge_accuracy_review$Percentile_CI_2_5

edge_accuracy_review$Percentile_CI_includes_zero <-
  edge_accuracy_review$Percentile_CI_2_5 <= 0 &
    edge_accuracy_review$Percentile_CI_97_5 >= 0

edge_accuracy_review$Point_sign <- ifelse(
  edge_accuracy_review$Point_edge > 0,
  "POSITIVE",
  ifelse(edge_accuracy_review$Point_edge < 0, "NEGATIVE", "ZERO")
)

edge_accuracy_review$Interpretation_status <- ifelse(
  edge_accuracy_review$Retained_under_EBICglasso,
  "RETAINED_UNDER_EBICGLASSO_REPORT_ESTIMATE_AND_INTERVAL",
  "NOT_RETAINED_UNDER_REGULARISATION_DO_NOT_CALL_ABSENT"
)

edge_accuracy_review$Descriptive_use <- ifelse(
  edge_accuracy_review$Retained_under_EBICglasso,
  paste(
    "May describe direction and approximate magnitude within this",
    "single network while reporting the bootstrap interval and its width;",
    "interval inclusion of zero is audit information, not a second",
    "edge-selection or significance rule; not evidence of a",
    "between-group difference."
  ),
  paste(
    "State only that the edge was not retained under the frozen",
    "regularisation procedure; do not claim absence."
  )
)
edge_accuracy_review$Between_group_inference <-
  "NOT_PERFORMED_AND_NOT_PERMITTED_FROM_THIS_TABLE"

group_order <- setNames(
  seq_len(nrow(expected_groups)),
  expected_groups$Group_key
)

node_order <- setNames(seq_along(network_vars), network_vars)

edge_accuracy_review <- edge_accuracy_review[
  order(
    unname(group_order[edge_accuracy_review$Group_key]),
    unname(node_order[edge_accuracy_review$Node_1]),
    unname(node_order[edge_accuracy_review$Node_2])
  ),
]

row.names(edge_accuracy_review) <- NULL

edge_review_validation <- data.frame(
  Check = c(
    "Joined table contains 8 x 36 unique edges",
    "Group keys are exact",
    "Every point estimate matches the bootstrap sample estimate",
    "All accuracy quantities are finite",
    "Every percentile interval is ordered",
    "Every reported plus-or-minus-2SD interval is ordered",
    "Every bootstrap zero proportion lies in [0, 1]",
    "Interval-zero audit flag agrees with the recorded endpoints",
    "Interpretation status depends only on regularised retention"
  ),
  Passed = c(
    nrow(edge_accuracy_review) == 288L &&
      !any(
        duplicated(
          paste(
            edge_accuracy_review$Group_key,
            edge_accuracy_review$Edge_ID,
            sep = "::"
          )
        )
      ),
    setequal(
      edge_accuracy_review$Group_key,
      expected_groups$Group_key
    ),
    max(
      abs(
        review_joined$Edge_weight -
          review_joined$sample
      )
    ) < 1e-12,
    all(
      vapply(
        edge_accuracy_review[
          c(
            "Point_edge",
            "Bootstrap_mean",
            "Bootstrap_sd",
            "Reported_CI_lower_sample_plus_or_minus_2SD",
            "Reported_CI_upper_sample_plus_or_minus_2SD",
            "Percentile_CI_2_5",
            "Percentile_CI_97_5",
            "Bootstrap_zero_proportion"
          )
        ],
        function(value) all(is.finite(value)),
        FUN.VALUE = logical(1)
      )
    ),
    all(
      edge_accuracy_review$Percentile_CI_2_5 <=
        edge_accuracy_review$Percentile_CI_97_5
    ),
    all(
      edge_accuracy_review$
        Reported_CI_lower_sample_plus_or_minus_2SD <=
        edge_accuracy_review$
          Reported_CI_upper_sample_plus_or_minus_2SD
    ),
    all(
      edge_accuracy_review$Bootstrap_zero_proportion >= 0 &
        edge_accuracy_review$Bootstrap_zero_proportion <= 1
    ),
    all(
      edge_accuracy_review$Percentile_CI_includes_zero ==
        (
          edge_accuracy_review$Percentile_CI_2_5 <= 0 &
            edge_accuracy_review$Percentile_CI_97_5 >= 0
        )
    ),
    all(
      ifelse(
        edge_accuracy_review$Retained_under_EBICglasso,
        edge_accuracy_review$Interpretation_status ==
          paste0(
            "RETAINED_UNDER_EBICGLASSO_",
            "REPORT_ESTIMATE_AND_INTERVAL"
          ),
        edge_accuracy_review$Interpretation_status ==
          paste0(
            "NOT_RETAINED_UNDER_REGULARISATION_",
            "DO_NOT_CALL_ABSENT"
          )
      )
    )
  ),
  stringsAsFactors = FALSE
)

write.csv(
  edge_review_validation,
  file.path(formal_table_dir, "08C_edge_review_validation.csv"),
  row.names = FALSE
)

stopifnot(all(edge_review_validation$Passed))

edge_accuracy_review_file <- file.path(
  formal_table_dir,
  "08C_education_edge_accuracy_review.csv"
)

write.csv(
  edge_accuracy_review,
  edge_accuracy_review_file,
  row.names = FALSE,
  na = ""
)

edge_review_validation
```

# 7. Widest intervals and group readiness

```{r widest-intervals-and-readiness}
widest_intervals <- do.call(
  rbind,
  lapply(
    expected_groups$Group_key,
    function(group_key) {
      group_data <- edge_accuracy_review[
        edge_accuracy_review$Group_key == group_key,
        ,
        drop = FALSE
      ]

      group_data <- group_data[
        order(-group_data$Percentile_CI_width, group_data$Edge_ID),
        ,
        drop = FALSE
      ]

      group_data$Width_rank_within_group <- seq_len(nrow(group_data))
      head(group_data, 5L)
    }
  )
)

row.names(widest_intervals) <- NULL

widest_intervals <- widest_intervals[
  ,
  c(
    "Group_key",
    "Wave",
    "Education_group",
    "N",
    "Width_rank_within_group",
    "Edge_ID",
    "Edge_label",
    "Point_edge",
    "Retained_under_EBICglasso",
    "Percentile_CI_2_5",
    "Percentile_CI_97_5",
    "Percentile_CI_width",
    "Percentile_CI_includes_zero",
    "Bootstrap_zero_proportion",
    "Interpretation_status"
  )
]

widest_intervals_file <- file.path(
  formal_table_dir,
  "08C_education_widest_intervals.csv"
)

write.csv(
  widest_intervals,
  widest_intervals_file,
  row.names = FALSE,
  na = ""
)

result_readiness <- do.call(
  rbind,
  lapply(
    expected_groups$Group_key,
    function(group_key) {
      edge_data <- edge_accuracy_review[
        edge_accuracy_review$Group_key == group_key,
        ,
        drop = FALSE
      ]

      point_row <- point_diagnostics[
        point_diagnostics$Group_key == group_key,
        ,
        drop = FALSE
      ]

      bootstrap_row <- bootstrap_diagnostics[
        bootstrap_diagnostics$Group_key == group_key,
        ,
        drop = FALSE
      ]

      readiness_row <- stage14_readiness[
        stage14_readiness$Group_key == group_key,
        ,
        drop = FALSE
      ]

                  retained <-
        edge_data$Retained_under_EBICglasso

      retained_interval_excludes_zero_audit <-
        retained &
          !edge_data$Percentile_CI_includes_zero

      retained_interval_includes_zero_audit <-
        retained &
          edge_data$Percentile_CI_includes_zero

      unretained <- !retained

      widest_index <- which.max(edge_data$Percentile_CI_width)

      all_required_valid <-
        point_row$All_validation_checks_passed &&
          bootstrap_row$All_validation_checks_passed &&
          bootstrap_row$Returned_bootstraps == 1000L &&
          readiness_row$Ready_for_stage14

      data.frame(
        Group_key = group_key,
        Wave = point_row$Wave,
        Education_group = point_row$Education_group,
        N = point_row$N,
        Nodes = point_row$Nodes,
        Possible_edges = point_row$Possible_edges,
        Retained_edges = point_row$Retained_edges,
        Not_retained_edges = sum(unretained),
        Bootstrap_repetitions = bootstrap_row$Returned_bootstraps,
                All_edges_percentile_CI_excludes_zero_audit_only =
          sum(
            !edge_data$Percentile_CI_includes_zero
          ),
        All_edges_percentile_CI_includes_zero_audit_only =
          sum(
            edge_data$Percentile_CI_includes_zero
          ),
        Retained_edges_available_for_cautious_description =
          sum(retained),
        Retained_edges_percentile_CI_excludes_zero_audit_only =
          sum(
            retained_interval_excludes_zero_audit
          ),
        Retained_edges_percentile_CI_includes_zero_audit_only =
          sum(
            retained_interval_includes_zero_audit
          ),
        Median_percentile_CI_width = median(edge_data$Percentile_CI_width),
        Maximum_percentile_CI_width = max(edge_data$Percentile_CI_width),
        Widest_interval_edge = edge_data$Edge_ID[[widest_index]],
        Widest_interval_label = edge_data$Edge_label[[widest_index]],
        Point_warning_n = point_row$Point_warning_n,
        Point_warnings = safe_message(point_row$Point_warnings),
        Bootstrap_warning_n = bootstrap_row$Captured_warning_n,
        Bootstrap_warnings = safe_message(bootstrap_row$Captured_warnings),
        Sparse_cell_status = readiness_row$Sparse_cell_status,
        Cells_0 = readiness_row$Cells_0,
        Cells_1_4 = readiness_row$Cells_1_4,
        Pairs_with_zero = readiness_row$Pairs_with_zero,
        Pairs_with_1_4 = readiness_row$Pairs_with_1_4,
        Input_and_bootstrap_checks_passed = all_required_valid,
        Result_readiness = ifelse(
          all_required_valid,
          "READY_FOR_CAUTIOUS_WITHIN_NETWORK_DESCRIPTION_WITH_LIMITATIONS",
          "REVIEW_REQUIRED"
        ),
        Permitted_scope = paste(
          "Within-network descriptive edge patterns and precision only;",
          "no education-group significance claim."
        ),
        stringsAsFactors = FALSE
      )
    }
  )
)

row.names(result_readiness) <- NULL

result_readiness_file <- file.path(
  formal_table_dir,
  "08C_education_result_readiness.csv"
)

write.csv(
  result_readiness,
  result_readiness_file,
  row.names = FALSE,
  na = ""
)

stopifnot(
  nrow(widest_intervals) == 40L,
  nrow(result_readiness) == 8L,
  all(
    result_readiness$
      Input_and_bootstrap_checks_passed
  ),
  all(
    result_readiness$Retained_edges ==
      result_readiness$
        Retained_edges_available_for_cautious_description
  ),
  all(
    result_readiness$Retained_edges ==
      result_readiness$
        Retained_edges_percentile_CI_excludes_zero_audit_only +
      result_readiness$
        Retained_edges_percentile_CI_includes_zero_audit_only
  ),
  all(
    result_readiness$Possible_edges ==
      result_readiness$Retained_edges +
        result_readiness$Not_retained_edges
  )
)

result_readiness[
  ,
  c(
    "Group_key",
    "N",
    "Retained_edges",
    "Retained_edges_available_for_cautious_description",
    "Retained_edges_percentile_CI_includes_zero_audit_only",
    "Not_retained_edges",
    "Median_percentile_CI_width",
    "Maximum_percentile_CI_width",
    "Sparse_cell_status",
    "Result_readiness"
  )
]
```

# 8. Fixed visual specification

```{r fixed-visual-specification}
layout_angle <- pi / 2 - (seq_along(network_vars) - 1) * 2 * pi /
  length(network_vars)

fixed_layout <- cbind(cos(layout_angle), sin(layout_angle))
rownames(fixed_layout) <- network_vars

all_point_edges <- unlist(
  lapply(point_graphs, function(graph) graph[upper.tri(graph)]),
  use.names = FALSE
)

observed_maximum_absolute_edge <- max(abs(all_point_edges), na.rm = TRUE)
network_edge_scale_maximum <- 0.50

stopifnot(observed_maximum_absolute_edge <= network_edge_scale_maximum)

global_maximum_absolute_interval <- max(
  abs(c(
    edge_accuracy_review$Percentile_CI_2_5,
    edge_accuracy_review$Percentile_CI_97_5,
    edge_accuracy_review$Point_edge
  )),
  na.rm = TRUE
)

accuracy_axis_limit <- ceiling(global_maximum_absolute_interval * 20) / 20

positive_colour <- "#0072B2"
negative_colour <- "#D55E00"
unretained_colour <- "#888888"
node_fill_colour <- "#F7F7F7"
node_border_colour <- "#222222"

edge_template_index <- which(
  upper.tri(matrix(0, nrow = 9L, ncol = 9L)),
  arr.ind = TRUE
)

edge_template <- data.frame(
  Node_1 = network_vars[edge_template_index[, "row"]],
  Node_2 = network_vars[edge_template_index[, "col"]],
  stringsAsFactors = FALSE
)

edge_template$Edge_ID <- paste(
  edge_template$Node_1,
  edge_template$Node_2,
  sep = "--"
)

edge_template$Edge_label <- paste(
  unname(node_labels[edge_template$Node_1]),
  unname(node_labels[edge_template$Node_2]),
  sep = " -- "
)

visual_specification <- data.frame(
  Element = c(
    "Node order",
    "Layout",
    "Positive edge",
    "Negative edge",
    "Network edge-width scale",
    "Accuracy x-axis",
    "Accuracy interval",
    "Figure export"
  ),
  Specification = c(
    paste(network_vars, collapse = " | "),
    "Fixed nine-node circular layout for every network",
    paste0(positive_colour, "; solid line"),
    paste0(negative_colour, "; dashed line"),
    paste0(
      "Fixed common maximum absolute edge = ",
      format(network_edge_scale_maximum, digits = 6),
      "; observed maximum = ",
      format(observed_maximum_absolute_edge, digits = 6)
    ),
    paste0(
      "Common symmetric limits = +/-",
      format(accuracy_axis_limit, digits = 6)
    ),
    "95% bootstrap percentile interval (q2.5 to q97.5)",
    "PNG, opaque white background, fixed pixel dimensions and resolution"
  ),
  stringsAsFactors = FALSE
)

write.csv(
  visual_specification,
  file.path(formal_table_dir, "08C_visual_specification.csv"),
  row.names = FALSE
)

visual_specification
```

# 9. Generate eight fixed-layout network figures

```{r generate-network-figures}
network_figure_files <- character(nrow(expected_groups))

for (group_index in seq_len(nrow(expected_groups))) {
  group_key <- expected_groups$Group_key[[group_index]]
  group_row <- result_readiness[
    result_readiness$Group_key == group_key,
    ,
    drop = FALSE
  ]
  graph <- point_graphs[[group_key]]

  figure_file <- file.path(
    formal_figure_dir,
    paste0("08C_education_network_", group_key, ".png")
  )

  grDevices::png(
    filename = figure_file,
    width = 2400,
    height = 2400,
    res = 300,
    bg = "white"
  )

  par(mar = c(6.5, 2.5, 5.0, 2.5) + 0.1, xpd = FALSE)

  plot(
    fixed_layout,
    type = "n",
    xlim = c(-1.35, 1.35),
    ylim = c(-1.30, 1.28),
    xaxs = "i",
    yaxs = "i",
    axes = FALSE,
    xlab = "",
    ylab = "",
    asp = 1,
    main = paste(group_row$Wave, group_row$Education_group, sep = " - ")
  )

  mtext(
    paste0(
      "N = ",
      format(group_row$N, big.mark = ","),
      "; retained edges = ",
      group_row$Retained_edges,
      "/36"
    ),
    side = 3,
    line = 0.2,
    cex = 0.83,
    col = "#333333"
  )

  retained_index <- which(
    upper.tri(graph) & abs(graph) > .Machine$double.eps,
    arr.ind = TRUE
  )

  if (nrow(retained_index) > 0L) {
    for (edge_index in seq_len(nrow(retained_index))) {
      row_index <- retained_index[edge_index, "row"]
      column_index <- retained_index[edge_index, "col"]
      edge_value <- graph[row_index, column_index]
      line_width <- max(
        0.55,
        7.5 * abs(edge_value) / network_edge_scale_maximum
      )

      segments(
        x0 = fixed_layout[row_index, 1],
        y0 = fixed_layout[row_index, 2],
        x1 = fixed_layout[column_index, 1],
        y1 = fixed_layout[column_index, 2],
        col = ifelse(edge_value > 0, positive_colour, negative_colour),
        lty = ifelse(edge_value > 0, 1, 2),
        lwd = line_width
      )
    }
  }

  points(
    fixed_layout[, 1],
    fixed_layout[, 2],
    pch = 21,
    cex = 5.1,
    bg = node_fill_colour,
    col = node_border_colour,
    lwd = 1.4
  )

  text(
    fixed_layout[, 1],
    fixed_layout[, 2],
    labels = unname(node_labels[network_vars]),
    cex = 0.72,
    col = "#111111"
  )

  legend(
    "bottom",
    inset = -0.17,
    horiz = TRUE,
    legend = c("Positive conditional association", "Negative conditional association"),
    col = c(positive_colour, negative_colour),
    lty = c(1, 2),
    lwd = c(3, 3),
    bty = "n",
    cex = 0.83,
    xpd = NA
  )

  mtext(
    paste0(
      "Common layout and edge-width scale across all eight panels; ",
      "regularised zero edges are not drawn."
    ),
    side = 1,
    line = 5.0,
    cex = 0.72,
    col = "#333333"
  )

  grDevices::dev.off()

  network_figure_files[[group_index]] <- figure_file
}

stopifnot(
  length(network_figure_files) == 8L,
  all(file.exists(network_figure_files)),
  all(file.info(network_figure_files)$size > 0)
)

network_figure_files
```

# 10. Generate eight edge-accuracy figures

```{r generate-accuracy-figures}
accuracy_figure_files <- character(nrow(expected_groups))

for (group_index in seq_len(nrow(expected_groups))) {
  group_key <- expected_groups$Group_key[[group_index]]
  group_row <- result_readiness[
    result_readiness$Group_key == group_key,
    ,
    drop = FALSE
  ]
  group_data <- edge_accuracy_review[
    edge_accuracy_review$Group_key == group_key,
    ,
    drop = FALSE
  ]

  group_data <- group_data[
    match(edge_template$Edge_ID, group_data$Edge_ID),
    ,
    drop = FALSE
  ]

  stopifnot(identical(group_data$Edge_ID, edge_template$Edge_ID))

  retained_positive <-
    group_data$Retained_under_EBICglasso &
      group_data$Point_edge > 0

  retained_negative <-
    group_data$Retained_under_EBICglasso &
      group_data$Point_edge < 0

  point_colours <- rep(
    unretained_colour,
    nrow(group_data)
  )

  point_symbols <- rep(
    1L,
    nrow(group_data)
  )

  interval_line_types <- rep(
    3L,
    nrow(group_data)
  )

  point_colours[retained_positive] <-
    positive_colour

  point_symbols[retained_positive] <-
    16L

  interval_line_types[retained_positive] <-
    1L

  point_colours[retained_negative] <-
    negative_colour

  point_symbols[retained_negative] <-
    15L

  interval_line_types[retained_negative] <-
    2L

  y_position <- rev(seq_len(nrow(group_data)))

  figure_file <- file.path(
    formal_figure_dir,
    paste0("08C_education_edge_accuracy_", group_key, ".png")
  )

  grDevices::png(
    filename = figure_file,
    width = 2700,
    height = 3000,
    res = 300,
    bg = "white"
  )

  par(mar = c(12.0, 14.5, 6.0, 2.5) + 0.1, xpd = FALSE)

  plot(
    group_data$Point_edge,
    y_position,
    type = "n",
    xlim = c(-accuracy_axis_limit, accuracy_axis_limit),
    ylim = c(0.5, nrow(group_data) + 0.5),
    yaxt = "n",
    xaxs = "i",
    xlab = "Regularised partial-correlation edge weight",
    ylab = "",
    main = paste0(
      group_row$Wave,
      " - ",
      group_row$Education_group,
      "\nWithin-network edge accuracy"
    ),
    cex.main = 1.05
  )

  abline(v = 0, lty = 3, col = "#222222", lwd = 1)

  segments(
    x0 = group_data$Percentile_CI_2_5,
    y0 = y_position,
    x1 = group_data$Percentile_CI_97_5,
    y1 = y_position,
    col = point_colours,
    lty = interval_line_types,
    lwd = 1.7
  )

  points(
    group_data$Point_edge,
    y_position,
    pch = point_symbols,
    col = point_colours,
    bg = point_colours,
    cex = 0.9
  )

  axis(
    side = 2,
    at = y_position,
    labels = edge_template$Edge_label,
    las = 2,
    cex.axis = 0.62
  )

  legend(
    "bottom",
    inset = -0.19,
    ncol = 3,
    legend = c(
      "Retained positive edge",
      "Retained negative edge",
      "Not retained under regularisation"
    ),
    col = c(
      positive_colour,
      negative_colour,
      unretained_colour
    ),
    pch = c(16, 15, 1),
    lty = c(1, 2, 3),
    bty = "n",
    cex = 0.72,
    xpd = NA
  )

  mtext(
    "Points are original-sample edges; bars are 95% bootstrap percentile intervals.",
    side = 1,
    line = 9.3,
    cex = 0.68,
    col = "#333333"
  )

  mtext(
    paste(
      "Interval inclusion of zero is audit information only;",
      "it is not a reporting gate or an education-group test."
    ),
    side = 1,
    line = 10.3,
    cex = 0.68,
    col = "#333333"
  )

  grDevices::dev.off()

  accuracy_figure_files[[group_index]] <- figure_file
}

stopifnot(
  length(accuracy_figure_files) == 8L,
  all(file.exists(accuracy_figure_files)),
  all(file.info(accuracy_figure_files)$size > 0)
)

accuracy_figure_files
```

# 11. Figure manifest and final validation

```{r figure-manifest-and-final-validation}
figure_manifest <- rbind(
  data.frame(
    Figure_type = "FIXED_LAYOUT_NETWORK",
    Group_key = expected_groups$Group_key,
    Wave = expected_groups$Wave,
    Education_group = expected_groups$Education_group,
    Path = normalizePath(
      network_figure_files,
      winslash = "/",
      mustWork = TRUE
    ),
    Width_px = 2400L,
    Height_px = 2400L,
    Resolution_ppi = 300L,
    Common_scale = paste0(
      "Maximum absolute edge = ",
      format(network_edge_scale_maximum, digits = 8),
      "; observed maximum = ",
      format(observed_maximum_absolute_edge, digits = 8)
    ),
    stringsAsFactors = FALSE
  ),
  data.frame(
    Figure_type = "EDGE_ACCURACY",
    Group_key = expected_groups$Group_key,
    Wave = expected_groups$Wave,
    Education_group = expected_groups$Education_group,
    Path = normalizePath(
      accuracy_figure_files,
      winslash = "/",
      mustWork = TRUE
    ),
    Width_px = 2700L,
    Height_px = 3000L,
    Resolution_ppi = 300L,
    Common_scale = paste0(
      "Symmetric x-axis +/-",
      format(accuracy_axis_limit, digits = 8)
    ),
    stringsAsFactors = FALSE
  )
)

figure_manifest$Bytes <- file.info(figure_manifest$Path)$size
figure_manifest$SHA256 <- vapply(
  figure_manifest$Path,
  file_sha256,
  FUN.VALUE = character(1)
)

figure_manifest_file <- file.path(
  formal_table_dir,
  "08C_education_figure_manifest.csv"
)

write.csv(
  figure_manifest,
  figure_manifest_file,
  row.names = FALSE
)

final_validation <- data.frame(
  Check = c(
    "Source hash gate passed",
    "Governance gate passed",
    "Structural input validation passed",
    "Prohibited-call audit passed",
    "Edge-review validation passed",
    "Edge-accuracy review contains 288 rows",
    "Widest-interval table contains five rows per group",
    "Result-readiness table contains eight groups",
    "Eight network figures exist and are non-empty",
    "Eight accuracy figures exist and are non-empty",
    "All figures have recorded hashes",
    "No formal between-group inference field was populated"
  ),
  Passed = c(
    all(source_gate$Passed),
    all(decision_gate$Passed),
    all(structural_validation$Passed),
    !any(prohibited_call_scan$Found),
    all(edge_review_validation$Passed),
    nrow(edge_accuracy_review) == 288L,
    nrow(widest_intervals) == 40L &&
      all(table(widest_intervals$Group_key) == 5L),
    nrow(result_readiness) == 8L,
    length(network_figure_files) == 8L &&
      all(file.info(network_figure_files)$size > 0),
    length(accuracy_figure_files) == 8L &&
      all(file.info(accuracy_figure_files)$size > 0),
    all(nzchar(figure_manifest$SHA256)),
    all(
      edge_accuracy_review$Between_group_inference ==
        "NOT_PERFORMED_AND_NOT_PERMITTED_FROM_THIS_TABLE"
    )
  ),
  stringsAsFactors = FALSE
)

write.csv(
  final_validation,
  file.path(formal_table_dir, "08C_final_validation.csv"),
  row.names = FALSE
)

final_validation
stopifnot(all(final_validation$Passed))
```

# 12. Readiness summary

```{r readiness-summary-table, echo=FALSE}
knitr::kable(
  result_readiness[
    ,
    c(
      "Wave",
      "Education_group",
      "N",
      "Retained_edges",
      "Retained_edges_available_for_cautious_description",
      "Retained_edges_percentile_CI_includes_zero_audit_only",
      "Not_retained_edges",
      "Not_retained_edges",
      "Median_percentile_CI_width",
      "Maximum_percentile_CI_width",
      "Result_readiness"
    )
  ],
  digits = 3,
  caption = paste(
    "Education-network result readiness.",
    "The interval-width summaries describe within-network estimation accuracy.
Counts based on whether an interval includes zero are retained for audit only
and are not edge-selection, significance or reporting criteria. None of these
quantities provides evidence that the two education groups differ."
  )
)
```

# 13. Widest intervals requiring the greatest caution

```{r widest-intervals-table, echo=FALSE}
knitr::kable(
  widest_intervals[
    ,
    c(
      "Wave",
      "Education_group",
      "Width_rank_within_group",
      "Edge_label",
      "Point_edge",
      "Percentile_CI_2_5",
      "Percentile_CI_97_5",
      "Percentile_CI_width",
      "Interpretation_status"
    )
  ],
  digits = 3,
  caption = paste(
    "Five widest 95% bootstrap percentile intervals per group.",
    "Width rank is an audit aid, not a universal cutoff."
  )
)
```

# 14. Fixed-layout education network figures

Every panel uses the same node order, circular coordinates, positive and
negative edge encodings, and global edge-width scale. Visual differences are
therefore not created by separate layout optimisation. The figures remain
descriptive; they do not establish education-group differences.

```{r display-network-figures, echo=FALSE, out.width="49%"}
network_figure_files_for_html <- normalizePath(
  network_figure_files,
  winslash = "/",
  mustWork = TRUE
)

knitr::include_graphics(
  network_figure_files_for_html,
  rel_path = FALSE
)
```

# 15. Education edge-accuracy figures

Every panel uses the same edge order and symmetric horizontal scale. The bars
are 95% bootstrap percentile intervals. Visual categories distinguish retained
positive edges, retained negative edges and edges not retained under the frozen
regularisation procedure. Interval width describes estimation accuracy.
Whether an interval includes zero is audit information only and is not a
reporting gate, significance test or education-group comparison.

```{r display-accuracy-figures, echo=FALSE, out.width="49%"}
accuracy_figure_files_for_html <- normalizePath(
  accuracy_figure_files,
  winslash = "/",
  mustWork = TRUE
)

knitr::include_graphics(
  accuracy_figure_files_for_html,
  rel_path = FALSE
)
```

# 16. Log, session information and output inventory

```{r log-session-and-output-inventory}
session_file <- file.path(
  log_dir,
  "08C_education_result_readiness_sessionInfo.txt"
)

writeLines(capture.output(sessionInfo()), session_file)

log_file <- file.path(
  log_dir,
  "08C_education_result_readiness.log"
)

writeLines(
  c(
    "STEP 08C EDUCATION RESULT READINESS",
    paste0("Registry: ", normalizePath(registry_path, winslash = "/")),
    paste0("Registry SHA-256: ", file_sha256(registry_path)),
    paste0("Reviewed groups: ", nrow(result_readiness), "/8"),
    paste0("Reviewed edges: ", nrow(edge_accuracy_review), "/288"),
    paste0(
      "Retained edges available for cautious within-network description: ",
      sum(
        edge_accuracy_review$
          Retained_under_EBICglasso
      ),
      "/288"
    ),
    paste0(
      paste(
        "Retained edges with percentile interval",
        "including zero (audit only):"
      ),
      " ",
      sum(
        edge_accuracy_review$
          Retained_under_EBICglasso &
          edge_accuracy_review$
            Percentile_CI_includes_zero
      )
    ),
    paste0("Figures created: ", nrow(figure_manifest), "/16"),
    paste0("Final validation checks passed: ", sum(final_validation$Passed),
      "/", nrow(final_validation)),
    "Network estimation or re-estimation: NOT RUN",
    "Bootstrap execution or re-execution: NOT RUN",
    "Centrality: NOT RUN",
    "NCT: NOT RUN",
    "Formal education-group inference: NOT RUN",
    "Political interpretation: NOT RUN"
  ),
  log_file
)

output_paths <- c(
  edge_accuracy_review_file,
  widest_intervals_file,
  result_readiness_file,
  figure_manifest_file,
  file.path(formal_table_dir, "08C_package_gate.csv"),
  file.path(formal_table_dir, "08C_decision_gate.csv"),
  file.path(formal_table_dir, "08C_source_hash_gate.csv"),
  file.path(formal_table_dir, "08C_input_validation.csv"),
  file.path(formal_table_dir, "08C_prohibited_call_scan.csv"),
  file.path(formal_table_dir, "08C_edge_review_validation.csv"),
  file.path(formal_table_dir, "08C_visual_specification.csv"),
  file.path(formal_table_dir, "08C_final_validation.csv"),
  network_figure_files,
  accuracy_figure_files,
  log_file,
  session_file
)

output_paths <- normalizePath(
  output_paths,
  winslash = "/",
  mustWork = TRUE
)

output_inventory <- data.frame(
  File = basename(output_paths),
  Path = output_paths,
  Bytes = file.info(output_paths)$size,
  SHA256 = vapply(output_paths, file_sha256, FUN.VALUE = character(1)),
  stringsAsFactors = FALSE
)

write.csv(
  output_inventory,
  file.path(formal_table_dir, "08C_output_hash_inventory.csv"),
  row.names = FALSE
)

stopifnot(
  all(output_inventory$Bytes > 0),
  all(nzchar(output_inventory$SHA256))
)

output_inventory
```

# Final bounded status

The output permits cautious within-network description of edges retained under the frozen EBICglasso procedure when their point estimates, bootstrap intervals and interval widths are reported together. Interval inclusion of zero is retained for audit only and is not a second edge-selection or reporting rule.The output does not establish any statistically significant difference between education groups.
