---
title: "09 Sensitivity Analyses"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
params:
  execute_engine: false
---

## Purpose

This notebook reports the authorised minimum Step 09 sensitivity set. It
contains D11 pairwise-available polychoric point networks, the D20A gamma
0.25 and no-VCF0839 point networks, reuse-only D15 weighted point-network
summaries, and D19 all-internet diagnostics and point networks for
2012--2024.

These prespecified analyses provide descriptive robustness evidence. They
do not select or replace the primary specification, and no significance
tests are performed.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  error = FALSE,
  warning = TRUE,
  message = TRUE
)
```

## Stage 1: locate the project

```{r project-root}
current_dir <- normalizePath(".", winslash = "/")

if (file.exists(file.path(current_dir, "ANES.Rproj"))) {
  project_root <- current_dir
} else if (
  basename(current_dir) == "notebooks" &&
    file.exists(file.path(dirname(current_dir), "ANES.Rproj"))
) {
  project_root <- dirname(current_dir)
} else {
  stop("Project root could not be identified.")
}

knitr::opts_knit$set(root.dir = project_root)
table_dir <- file.path(project_root, "outputs", "tables", "sensitivity")
model_dir <- file.path(project_root, "outputs", "models", "sensitivity")
log_dir <- file.path(project_root, "outputs", "logs")
engine_file <- file.path(project_root, "R", "09_sensitivity_analysis_engine.R")
```

## Stage 2: optional controlled execution

The engine is disabled during ordinary rendering so a report refresh does not
re-estimate the point networks. Set `execute_engine: true` only when the
authorised inputs and decision registry have been verified and a new execution
is intended.

```{r controlled-execution}
if (isTRUE(params$execute_engine)) {
  stopifnot(file.exists(engine_file))
  source(engine_file, local = new.env(parent = globalenv()))
}
```

## Stage 3: decision and exclusion gates

```{r decision-gates}
decision_gate <- read.csv(
  file.path(table_dir, "09_decision_gate.csv"),
  check.names = FALSE
)
run_controls <- read.csv(
  file.path(table_dir, "09_run_controls.csv"),
  check.names = FALSE
)
prohibited_gate <- read.csv(
  file.path(table_dir, "09_prohibited_call_gate.csv"),
  check.names = FALSE
)

stopifnot(
  all(decision_gate$Passed),
  !any(prohibited_gate$Detected),
  !any(run_controls$Bootstrap),
  !any(run_controls$Centrality),
  !any(run_controls$NCT),
  !any(run_controls$Significance_testing)
)

knitr::kable(decision_gate)
knitr::kable(run_controls)
```

## Stage 4: input, hash and package gates

```{r input-gates}
input_hashes <- read.csv(
  file.path(table_dir, "09_input_hash_validation.csv"),
  check.names = FALSE
)
package_gate <- read.csv(
  file.path(table_dir, "09_package_gate.csv"),
  check.names = FALSE
)
d11_archive_gate <- read.csv(
  file.path(table_dir, "09_d11_archive_hash_gate.csv"),
  check.names = FALSE
)

stopifnot(
  all(input_hashes$Hash_matches),
  all(package_gate$Passed),
  all(d11_archive_gate$Hash_matches)
)

knitr::kable(package_gate)
```

## Stage 5: primary-claim registry validation

```{r claim-registry-gate}
claim_gate <- read.csv(
  file.path(table_dir, "09_primary_claim_registry_validation.csv"),
  check.names = FALSE
)
stopifnot(all(claim_gate$Passed))
knitr::kable(claim_gate)
```

## Stage 6: D11 pairwise-available polychoric networks

```{r d11-results}
d11_validation <- read.csv(
  file.path(table_dir, "09_d11_validation.csv"),
  check.names = FALSE
)
d11_comparison <- read.csv(
  file.path(table_dir, "09_d11_primary_comparison.csv"),
  check.names = FALSE
)
knitr::kable(d11_validation)
knitr::kable(d11_comparison)
```

## Stage 7: D20A gamma 0.25 networks

```{r gamma-results}
gamma_validation <- read.csv(
  file.path(table_dir, "09_d20a_gamma025_validation.csv"),
  check.names = FALSE
)
gamma_comparison <- read.csv(
  file.path(table_dir, "09_d20a_gamma025_primary_comparison.csv"),
  check.names = FALSE
)
knitr::kable(gamma_validation)
knitr::kable(gamma_comparison)
```

## Stage 8: D20A no-VCF0839 networks

```{r node-results}
node_validation <- read.csv(
  file.path(table_dir, "09_d20a_no_vcf0839_identity_validation.csv"),
  check.names = FALSE
)
node_comparison <- read.csv(
  file.path(table_dir, "09_d20a_no_vcf0839_primary_comparison.csv"),
  check.names = FALSE
)
knitr::kable(node_validation)
knitr::kable(node_comparison)
```

## Stage 9: D15 weighted point-network reuse

```{r d15-results}
d15_validation <- read.csv(
  file.path(table_dir, "09_d15_reuse_validation.csv"),
  check.names = FALSE
)
d15_summary <- read.csv(
  file.path(table_dir, "09_d15_weighted_network_summary.csv"),
  check.names = FALSE
)
knitr::kable(d15_validation)
knitr::kable(d15_summary)
```

## Stage 10: D19 all-internet diagnostics and networks

```{r d19-results}
d19_validation <- read.csv(
  file.path(table_dir, "09_d19_validation.csv"),
  check.names = FALSE
)
d19_comparison <- read.csv(
  file.path(table_dir, "09_d19_primary_comparison.csv"),
  check.names = FALSE
)
knitr::kable(d19_validation)
knitr::kable(d19_comparison)
```

## Stage 11: consolidated robustness and completion audit

```{r consolidated-results}
completion <- read.csv(
  file.path(table_dir, "09_completion_status.csv"),
  check.names = FALSE
)
summary_by_wave <- read.csv(
  file.path(table_dir, "09_sensitivity_summary_by_wave.csv"),
  check.names = FALSE
)
claim_summary <- read.csv(
  file.path(table_dir, "09_primary_claim_robustness_summary.csv"),
  check.names = FALSE
)

stopifnot(
  all(completion$Complete),
  !any(completion$Prohibited_analysis_run)
)

knitr::kable(completion)
knitr::kable(summary_by_wave)
knitr::kable(as.data.frame(table(
  claim_summary$Final_robustness_status,
  claim_summary$Evaluation_completeness
)))
```

## Interpretation boundary

`UNCHANGED`, `CHANGED` and `REVERSED` are mechanical sign/retention audit
labels. They are not significance or equivalence tests. Magnitude changes are
reported continuously. `NOT_ESTIMABLE` denotes a failed mathematical or model
gate; `NOT_EVALUABLE` denotes an applicable registered claim that could not be
assessed. D19 does not isolate a causal survey-mode effect, and D15 does not
provide full complex-survey population-network inference.

## Reproducibility record

```{r reproducibility-record}
inventory <- read.csv(
  file.path(table_dir, "09_output_hash_inventory.csv"),
  check.names = FALSE
)
knitr::kable(inventory)

cat(readLines(file.path(log_dir, "09_sessionInfo.txt")), sep = "\n")
```
