---
title: "09B Specificity and Sample-Size Sensitivity"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
params:
  execute_engine: false
---

## Purpose and boundary

Step 09B audits whether selected descriptive patterns in the five primary
overall networks depend strongly on unequal wave sample sizes or on the
optional `qgraph` high-specificity threshold. It does not replace the primary
models and does not select a preferred model from the observed results.

The analysis does not run bootstrap, centrality, NCT, education-group models,
survey-weight re-estimation, significance tests, matrix repair, category
merging or model substitution.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  error = FALSE,
  warning = TRUE,
  message = TRUE
)
```

## Locate the project

```{r project-root}
current_dir <- normalizePath(".", winslash = "/")
if (file.exists(file.path(current_dir, "ANES.Rproj"))) {
  project_root <- current_dir
} else if (
  basename(current_dir) == "notebooks" &&
    file.exists(file.path(dirname(current_dir), "ANES.Rproj"))
) {
  project_root <- dirname(current_dir)
} else {
  stop("Project root could not be identified.")
}
knitr::opts_knit$set(root.dir = project_root)
table_dir <- file.path(project_root, "outputs", "tables", "sensitivity")
log_dir <- file.path(project_root, "outputs", "logs")
engine_file <- file.path(
  project_root,
  "R",
  "09B_specificity_sample_size_sensitivity_engine.R"
)
```

## Optional controlled execution

Ordinary rendering reuses validated outputs and does not rerun the experiment.
The engine runs only when `execute_engine` is explicitly set to `true` and the
required environment variables identify both projects.

```{r controlled-execution}
if (isTRUE(params$execute_engine)) {
  stopifnot(file.exists(engine_file))
  source(engine_file, local = new.env(parent = globalenv()))
}
```

## Governance and input gates

```{r gates}
decision_gate <- read.csv(
  file.path(table_dir, "09B_decision_gate.csv"),
  check.names = FALSE
)
input_gate <- read.csv(
  file.path(table_dir, "09B_primary_input_validation.csv"),
  check.names = FALSE
)
identity_gate <- read.csv(
  file.path(table_dir, "09B_primary_identity_validation.csv"),
  check.names = FALSE
)
run_controls <- read.csv(
  file.path(table_dir, "09B_run_controls.csv"),
  check.names = FALSE
)

stopifnot(
  all(decision_gate$Passed),
  all(input_gate$Passed),
  all(identity_gate$Passed),
  !any(run_controls$Executed)
)

knitr::kable(decision_gate)
knitr::kable(input_gate)
```

## Technical pilot

```{r pilot}
pilot <- read.csv(
  file.path(table_dir, "09B_pilot_validation.csv"),
  check.names = FALSE
)
stopifnot(all(pilot$Pilot_passed))
knitr::kable(pilot)
```

## A1 full-N high-specificity networks

```{r full-n-threshold}
a1 <- read.csv(
  file.path(table_dir, "09B_fullN_threshold_diagnostics.csv"),
  check.names = FALSE
)
stopifnot(nrow(a1) == 5L, all(a1$Success))
knitr::kable(a1)
```

## A2 and A3 equal-N validation

```{r equal-n-validation}
equaln_validation <- read.csv(
  file.path(table_dir, "09B_equalN_validation.csv"),
  check.names = FALSE
)
stopifnot(all(equaln_validation$Evaluable))
knitr::kable(equaln_validation)
```

## Equal-N network summaries

```{r equal-n-summary}
network_summary <- read.csv(
  file.path(table_dir, "09B_equalN_network_summary.csv"),
  check.names = FALSE
)
knitr::kable(network_summary)
```

## Registered contrast robustness

```{r contrast-summary}
contrast_summary <- read.csv(
  file.path(table_dir, "09B_contrast_robustness.csv"),
  check.names = FALSE
)
knitr::kable(contrast_summary)
```

## Pre-result claim audit

```{r claim-audit}
claim_summary <- read.csv(
  file.path(table_dir, "09B_claim_robustness_summary.csv"),
  check.names = FALSE
)
knitr::kable(claim_summary)
```

`ROBUST`, `MIXED` and `NOT_ROBUST` are pre-specified sensitivity labels.
They are not p-values, equivalence tests or evidence that deleted edges are
false positives. Wave-level values remain descriptive estimates.

## Completion and reproducibility

```{r completion}
completion <- read.csv(
  file.path(table_dir, "09B_completion_status.csv"),
  check.names = FALSE
)
hashes <- read.csv(
  file.path(table_dir, "09B_input_hash_validation.csv"),
  check.names = FALSE
)
stopifnot(all(completion$Complete), all(hashes$Hash_matches))
knitr::kable(completion)
```

## Independent correction and audit

The first independent audit identified a vectorisation error in one derived
edge-state matching proportion. The saved sample indices, polychoric inputs,
network matrices, edge-weight distributions and network summaries were not
affected. The affected derived summaries were recomputed from the saved RDS
arrays without re-estimating a model, and the pre-correction versions were
retained in the audit archive.

```{r independent-audit}
correction <- read.csv(
  file.path(table_dir, "09B_correction_validation.csv"),
  check.names = FALSE
)
diagnostic_correction <- read.csv(
  file.path(table_dir, "09B_diagnostic_label_correction_validation.csv"),
  check.names = FALSE
)
independent_audit <- read.csv(
  file.path(table_dir, "09B_independent_audit_validation.csv"),
  check.names = FALSE
)
stopifnot(
  all(correction$Passed),
  all(diagnostic_correction$Passed),
  all(independent_audit$Passed)
)
knitr::kable(correction)
knitr::kable(diagnostic_correction)
knitr::kable(independent_audit)
```

## Interpretation boundary

The equal-N results isolate sensitivity to available sample size; they do not
make the five election-year samples exchangeable. The thresholded results
audit a specificity-oriented estimator option; they do not prove that removed
edges are false. Neither arm repairs cross-wave measurement differences,
complete-case selection or complex-survey limitations.

## Session information

```{r session-information}
cat(readLines(file.path(log_dir, "09B_sessionInfo.txt")), sep = "\n")
```
