---
title: "10 Results and Reproducibility Assembly"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
params:
  execute_engine: false
  execute_independent_audit: false
---

## Purpose

Step 10 assembles verified outputs from Steps 00--09B into a Results and
reproducibility package. It reuses saved estimates under the existing
specification without further model estimation, resampling or NCT.

All comparisons remain descriptive unless an explicitly authorised formal test
exists. No such cross-wave or education-group significance test is present in
the current design. A regularised zero means that an edge was not retained by
the selected procedure; it does not prove absence. Retained-edge count and
global strength are not treated as ideological constraint.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  error = FALSE,
  warning = TRUE,
  message = TRUE,
  echo = FALSE
)
```

## Stage 1: locate the project

```{r project-root}
current_dir <- normalizePath(".", winslash = "/")

if (file.exists(file.path(current_dir, "ANES.Rproj"))) {
  project_root <- current_dir
} else if (
  basename(current_dir) == "notebooks" &&
    file.exists(file.path(dirname(current_dir), "ANES.Rproj"))
) {
  project_root <- dirname(current_dir)
} else {
  stop("Project root could not be identified.")
}

knitr::opts_knit$set(root.dir = project_root)
final_table_dir <- file.path(project_root, "outputs", "tables", "final")
final_figure_dir <- file.path(project_root, "outputs", "figures", "final")
final_audit_dir <- file.path(project_root, "outputs", "audits", "final")
engine_file <- file.path(project_root, "R", "10_results_and_reproducibility_engine.R")
audit_file <- file.path(project_root, "R", "10_results_and_reproducibility_independent_audit.R")
```

## Stage 2: optional controlled execution

Ordinary rendering is read-only. The engine and independent audit may be run
from a clean R session before rendering, or enabled explicitly through the
parameters below.

```{r controlled-execution}
if (isTRUE(params$execute_engine)) {
  source(engine_file, local = new.env(parent = globalenv()))
}

if (isTRUE(params$execute_independent_audit)) {
  source(audit_file, local = new.env(parent = globalenv()))
}
```

## Stage 3: completion and validation gates

```{r completion-gates}
completion <- read.csv(
  file.path(final_table_dir, "10_completion_status.csv"),
  check.names = FALSE
)
validation <- read.csv(
  file.path(final_table_dir, "10_validation.csv"),
  check.names = FALSE
)
independent_audit <- read.csv(
  file.path(final_table_dir, "10_independent_audit_validation.csv"),
  check.names = FALSE
)

stopifnot(all(validation$Passed))
stopifnot(all(independent_audit$Passed))
stopifnot(
  completion$Status[completion$Component == "Independent Step 10 audit"] ==
    "COMPLETE_AND_VALIDATED"
)

knitr::kable(completion, caption = "Step 10 completion status")
```

## Stage 4: current analysis scope

The current researcher-selected working title is:

> Structural Differences in Selected Policy-Attitude Networks across U.S.
> Election Years: Evidence from ANES Samples, 2004--2024

The title and exact research-question wording still need to be entered into the
formal decision log; supervisor confirmation remains advised. The completed
outputs support:

1. descriptive comparison of selected nine-node policy-attitude conditional-
   association networks in the 2004, 2012, 2016, 2020 and 2024 ANES analysis
   samples; and
2. descriptive examination of corresponding edges and relational
   configurations within the two education groups in 2012, 2016, 2020 and
   2024.

Expected Influence and Strength are exploratory, not separate confirmatory
research questions.

## Stage 5: primary sample and five-wave network descriptors

```{r primary-summary}
primary <- read.csv(
  file.path(final_table_dir, "10_sample_primary_summary.csv"),
  check.names = FALSE
)

knitr::kable(
  primary[, c(
    "Wave",
    "Analysis_ready_n",
    "Retained_edges",
    "Global_strength_descriptive",
    "Median_percentile_interval_width",
    "Maximum_percentile_interval_width"
  )],
  digits = 4,
  caption = "Verified primary sample and descriptive network summary"
)
```

All five primary networks produced the saved dense-network warning. Small
retained edges therefore require particular caution. Edge count and global
strength are reported as descriptive network quantities and are not equated
with ideological constraint. The repeated cross-sectional design cannot
identify individual change or causal ordering.

```{r overall-network-figure, out.width='100%', fig.cap='Five primary ANES analysis-sample networks using a common node order, layout and edge scale. Edges are descriptive conditional associations. No NCT or population-design-consistent comparison is implied.'}
knitr::include_graphics(
  normalizePath(
    file.path(final_figure_dir, "10_08_overall_networks_five_wave_panel.png"),
    winslash = "/",
    mustWork = TRUE
  ),
  rel_path = FALSE
)
```

## Stage 6: within-wave edge accuracy

```{r edge-accuracy-summary}
knitr::kable(
  primary[, c(
    "Wave",
    "Retained_edges",
    "Retained_edges_percentile_interval_excluding_zero_audit_only",
    "Median_percentile_interval_width",
    "Maximum_percentile_interval_width"
  )],
  digits = 4,
  caption = "Within-wave edge-accuracy descriptors"
)
```

The percentile-interval columns are retained as precision diagnostics. They are
not used as a second rule for accepting or rejecting regularised edges.

## Stage 7: exploratory centrality

```{r centrality-summary}
knitr::kable(
  primary[, c(
    "Wave",
    "Highest_Expected_Influence_node",
    "Highest_Expected_Influence",
    "Expected_Influence_CS",
    "Strength_CS"
  )],
  digits = 4,
  caption = "Exploratory within-wave centrality and case-dropping stability"
)
```

The CS values reached the highest case-drop proportion examined (approximately
0.75). This supports cautious exploratory within-wave interpretation; it does
not establish cross-wave centrality differences, political importance, public
salience, causal influence or intervention value.

```{r centrality-figures, out.width='49%', fig.show='hold', fig.cap='Exploratory Expected Influence and auxiliary Strength. Lines connect repeated cross-sectional wave estimates for display only; no cross-wave centrality significance test was performed.'}
knitr::include_graphics(
  normalizePath(
    c(
      file.path(final_figure_dir, "10_08_expected_influence_over_time.png"),
      file.path(final_figure_dir, "10_08_strength_over_time.png")
    ),
    winslash = "/",
    mustWork = TRUE
  ),
  rel_path = FALSE
)
```

## Stage 8: education-group result readiness

```{r education-summary}
education <- read.csv(
  file.path(final_table_dir, "10_education_summary.csv"),
  check.names = FALSE
)

knitr::kable(
  education[, c(
    "Wave",
    "Education_group",
    "N",
    "Retained_edges",
    "Median_percentile_CI_width",
    "Maximum_percentile_CI_width",
    "Sparse_cell_status",
    "Result_readiness"
  )],
  digits = 4,
  caption = "Education-group within-network result readiness"
)
```

The eight networks may be described separately and cautiously. The analysis did
not test whether the two education groups differ statistically. Educational
attainment is not interpreted as elite status or as a direct measure of
political sophistication.

## Stage 9: sensitivity and joint-robustness evidence

```{r sensitivity-summary}
sensitivity <- read.csv(
  file.path(final_table_dir, "10_sensitivity_summary.csv"),
  check.names = FALSE
)

knitr::kable(
  sensitivity,
  digits = 4,
  caption = "Pre-specified sensitivity and reproducibility evidence"
)
```

Sensitivity labels summarise whether descriptive results reproduced under the
pre-specified alternatives. `ROBUST`, `MIXED`, `NOT_ROBUST`, `UNCHANGED` and
`CHANGED` are not p-values, significance tests or equivalence tests.

## Stage 10: Results claim-to-output control

```{r claim-registry}
claims <- read.csv(
  file.path(final_table_dir, "10_result_claim_registry.csv"),
  check.names = FALSE
)

knitr::kable(
  head(claims, 20),
  caption = "First 20 rows of the claim-to-output registry"
)
```

The full registry remains a machine-readable audit file. Dissertation prose
must be drafted later from these verified records rather than from memory or a
chat summary.

## Stage 11: method citations and Zotero gaps

```{r citations}
citations <- read.csv(
  file.path(final_table_dir, "10_method_citation_register.csv"),
  check.names = FALSE
)
not_in_zotero <- read.csv(
  file.path(final_table_dir, "10_references_not_in_zotero.csv"),
  check.names = FALSE
)

knitr::kable(
  citations[, c(
    "Reference_ID", "Authors", "Year", "Title", "Role_in_Step10",
    "Full_text_or_official_source_status", "Zotero_status"
  )],
  caption = "Step 10 method and data-source citation register"
)

knitr::kable(
  not_in_zotero[, c(
    "Reference_ID", "Authors", "Year", "Title", "Zotero_status",
    "Verification_note"
  )],
  caption = "Sources used in Step 10 that are not currently recorded in Zotero"
)
```

The core network-method claims are supported by Epskamp and Fried (2018),
Epskamp, Borsboom and Fried (2018), Burger et al. (2023), and Bringmann et al.
(2019), all of which have inspected local full text. The original Expected
Influence definition is traced to Robinaugh, Millner and McNally (2016); the
full text was inspected through an accessible PMC copy, but no local Zotero PDF
is present. The qgraph software citation and official ANES documentation are
also explicitly marked as absent from Zotero.

## Stage 12: reproducibility package

```{r reproducibility}
packages <- read.csv(
  file.path(final_table_dir, "10_package_manifest.csv"),
  check.names = FALSE
)
reproducibility <- read.csv(
  file.path(final_table_dir, "10_reproducibility_manifest.csv"),
  check.names = FALSE
)
figures <- read.csv(
  file.path(final_table_dir, "10_figure_manifest.csv"),
  check.names = FALSE
)

knitr::kable(packages, caption = "Software and package-version manifest")
knitr::kable(reproducibility, caption = "Reproducibility-material status")
knitr::kable(
  figures[, c(
    "Figure_ID", "Final_path", "Figure_role", "Bytes", "Byte_identical",
    "Required_caption_boundary"
  )],
  caption = "Final figure manifest"
)
```

The final figure copies are byte-identical to the validated Step 07A/08/08C
sources. Step 10 records a package manifest, `sessionInfo()`, `renv` status,
input hashes, output hashes and an independent recalculation audit. Raw ANES
data are not redistributed.

```{r renv-status-statement, results='asis'}
renv_snapshot_status <- reproducibility$Status[
  reproducibility$Component == "renv status snapshot"
]

if (length(renv_snapshot_status) == 1L && identical(renv_snapshot_status, "COMPLETED")) {
  cat(
    "The recorded `renv::status()` snapshot reports that the project library ",
    "and lockfile are synchronized. The saved numerical outputs also passed ",
    "independent recalculation.\n",
    sep = ""
  )
} else {
  cat(
    "The recorded `renv::status()` snapshot does not establish a synchronized ",
    "project environment. The final reproducibility archive must retain the ",
    "corresponding limitation until reconciliation is completed.\n",
    sep = ""
  )
}
```

## Stage 13: independent audit result

```{r independent-audit}
knitr::kable(
  independent_audit,
  caption = "Independent Step 10 audit"
)
```

Every check above must remain `TRUE` before this report is treated as current.
Re-rendering the report ordinarily reads the saved audit package and performs
no new model computation.
